---
title: "Eelgrass GO Fitness Comparisons"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/camillerumberger/Documents/GitHub/BalticEelgrass")
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# packages
library(car)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
library(knitr)
library(kableExtra)
library(emmeans)
library(multcomp)
library(multcompView)

# data
# # exp fit
# fit_full <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
# surv <- read.csv("data/experiment/survival_dat_trt_pop.csv")[,-1]
# grow <- read.csv("data/experiment/fit_growth_dat.csv")[,-1]

# fit offsets
fit_off_clim <- read.csv("results/fit_results/climate_change_fitness_offsets_2026-05-28.csv")[,-1]

# go
lea <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_2025-12-03.csv")[,-1]
rda <- read.csv("results/GO_results/RDA_offsets_exptrt_wide_2026-01-02.csv")[,-1]
gf <- read.csv("results/GO_results/GF_offsets_exptrt_wide_2025-12-03.csv")[,-1]
env_dist <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]
```

## Genomic Offset - Fitness Offset Comparisons

This script compares genomic offset against fitness metrics in a "common garden" (across genotypes) and "current future" (across environments) comparison. 

### Data Prep

#### GO Data Processing

These data are already in a matrix format, but I will still make a long form combined df for plotting. 

```{r}
head(lea)
head(rda)
head(gf)

# long versions
# lea
lea_long <- lea %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

# rda
rda_long <- rda %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "rda_offset")

# gf
gf_long <- gf %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "gf_offset")

# env dist
env_dist_trt <- env_dist %>% filter(pops1 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu") & !(pops2 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu"))) %>% dplyr::select(pops1, pops2, distances) %>% rename(pop = pops2, trt = pops1, env_dist = distances)

# put everything together
go1 <- merge(lea_long, rda_long, by = c("trt", "cc", "pop"))
go2 <- merge(go1, gf_long[gf_long$mod == "geno", c("trt","pop","cc","gf_offset")], by = c("trt", "cc", "pop"))
go_wide <- merge(go2, env_dist_trt, by = c("pop", "trt"))

# long version of all
go_long <- go_wide %>% pivot_longer(
  cols = c("lea_offset","rda_offset","gf_offset", "env_dist"),
  names_to = "method",
  values_to = "offset")
```

#### Experimental Data

This data is prepped in the fitness offset climate change script.

### GO v Fitness

To assess whether genomic offset predictions align with experimental outcomes, I conduct a correlation analysis between genomic offset and different fitness and fitness offset metrics, namely absolute and relative fitness, as well as both absolute and relative current-future fitness offsets.

#### Data Merge

```{r}
# isolate relevant trt/pop combos
go_clim_w <- go_long %>% filter(pop %in% c("VAT", "VIK", "HOG", "BAR") & grepl("TempWarm", trt)) %>% mutate(coast = "West")
go_clim_e <- go_long %>% filter(pop %in% c("KUR", "KAL", "HOR", "BJO") & grepl("TempWarm", trt)) %>% mutate(coast = "East")
go_clim <- rbind(go_clim_w, go_clim_e)

# combine fitness and offset data
fit_go <- merge(go_clim, fit_off_clim, by = c("pop", "coast"))

# ordered list of pops
pops <- unique(fit_go[,c("pop", "pop_ord")]) %>% arrange(pop_ord)
```

#### Correlations

```{r}
# function for correlations current future offset
#################################################
# df is dataframe, go is lea_offset/rda_offset/gf_offset/env_dist, clone is mll/full, fit_met is surv/len/weigh
fitcorrCF <- function(df, go, clone, fit_met) {
  # subset data
  df_subset <- fit_go[fit_go$method == go & fit_go$cc == clone & fit_go$metric == fit_met,]
  
  # set up df
  corrs_cf <- data.frame(fit_metric = NA, method = NA, cc = NA, abs_corr = NA, abs_pval = NA, rel_corr = NA, rel_pval = NA, cf_abs_corr = NA, cf_abs_pval = NA, cf_rel_corr = NA, cf_rel_pval = NA)
  
  # metadata
  corrs_cf$fit_metric <- fit_met
  corrs_cf$method <- go
  corrs_cf$cc <- clone
  
  # correlation testing
  test_abs = cor.test(df_subset$abs_future, df_subset$offset, method = "kendall", na.action = "na.exclude")
  corrs_cf$abs_corr = test_abs$estimate
  corrs_cf$abs_pval = test_abs$p.value

  test_rel = cor.test(df_subset$CF_relative, df_subset$offset, method = "kendall", na.action = "na.exclude")
  corrs_cf$rel_corr = test_rel$estimate
  corrs_cf$rel_pval = test_rel$p.value

  test_cf_abs = cor.test(df_subset$CF_abs_offset, df_subset$offset, method = "kendall", na.action = "na.exclude")
  corrs_cf$cf_abs_corr = test_cf_abs$estimate
  corrs_cf$cf_abs_pval = test_cf_abs$p.value

  test_cf_rel = cor.test(df_subset$CF_rel_offset, df_subset$offset, method = "kendall", na.action = "na.exclude")
  corrs_cf$cf_rel_corr = test_cf_rel$estimate
  corrs_cf$cf_rel_pval = test_cf_rel$p.value

  # for (i in 1:length(pops$pop)) {
  #   pop_temp = pops[i,]$pop
  #   temp = df_subset[df_subset$pop == pop_temp,]
  #   
  #   # metadata
  #   corrs_cf[i,]$fit_metric <- fit_met
  #   corrs_cf[i,]$method <- go
  #   corrs_cf[i,]$cc <- clone
  #   
  #   # correlation analysis
  #   test_abs = cor.test(temp$abs_future, temp$offset, method = "kendall", na.action = "na.exclude")
  #   corrs_cf[i,]$abs_corr = test_abs$estimate
  #   corrs_cf[i,]$abs_pval = test_abs$p.value
  #   
  #   test_rel = cor.test(temp$CF_relative, temp$offset, method = "kendall", na.action = "na.exclude")
  #   corrs_cf[i,]$rel_corr = test_rel$estimate
  #   corrs_cf[i,]$rel_corr_pval = test_rel$p.value
  #   
  #   test_cf_abs = cor.test(temp$CF_abs_offset, temp$offset, method = "kendall", na.action = "na.exclude")
  #   corrs_cf[i,]$cf_abs_corr = test_cf_abs$estimate
  #   corrs_cf[i,]$cf_abs_pval = test_cf_abs$p.value
  #   
  #   test_cf_rel = cor.test(temp$CF_rel_offset, temp$offset, method = "kendall", na.action = "na.exclude")
  #   corrs_cf[i,]$cf_rel_corr = test_cf_rel$estimate
  #   corrs_cf[i,]$cf_rel_pval = test_cf_rel$p.value
  # }
  
  return(corrs_cf)
}
#################################################

# test correlations
###################
# lea
lea_cf <- rbind(fitcorrCF(fit_go, go = "lea_offset", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "lea_offset", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "lea_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "lea_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "lea_offset", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "lea_offset", clone = "mll", fit_met = "weigh"))

# rda
rda_cf <- rbind(fitcorrCF(fit_go, go = "rda_offset", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "rda_offset", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "rda_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "rda_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "rda_offset", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "rda_offset", clone = "mll", fit_met = "weigh"))

# gf
gf_cf <- rbind(fitcorrCF(fit_go, go = "gf_offset", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "gf_offset", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "gf_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "gf_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "gf_offset", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "gf_offset", clone = "mll", fit_met = "weigh"))

# env dist - note we use both full and mll to match with other methods when subsetting
envdist_cf <- rbind(fitcorrCF(fit_go, go = "env_dist", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "env_dist", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "env_dist", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "env_dist", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "env_dist", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "env_dist", clone = "mll", fit_met = "weigh"))

# combine
cf_clim_summary <- rbind(lea_cf, rda_cf, gf_cf, envdist_cf)

# pivot correlations and pvals to long format
cf_clim_long <- cf_clim_summary %>% pivot_longer(cols = c(abs_corr, rel_corr, cf_abs_corr, cf_rel_corr, abs_pval, rel_pval, cf_abs_pval, cf_rel_pval),
    names_to = c("fit_type", ".value"),
    names_pattern = "(abs|rel|cf_abs|cf_rel)_(corr|pval)")

# signif & adjustments
cf_clim_long$signif <- ifelse(cf_clim_long$pval < 0.05, T, F)
###################

# and by coast
##############
# data subset by coast
west_fitgo <- fit_go[fit_go$coast == "West",]
east_fitgo <- fit_go[fit_go$coast == "East",]

# west coast
# lea
lea_cf_w <- rbind(fitcorrCF(west_fitgo, go = "lea_offset", clone = "full", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "lea_offset", clone = "full", fit_met = "len"),
fitcorrCF(west_fitgo, go = "lea_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(west_fitgo, go = "lea_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "lea_offset", clone = "mll", fit_met = "len"),
fitcorrCF(west_fitgo, go = "lea_offset", clone = "mll", fit_met = "weigh"))

# rda
rda_cf_w <- rbind(fitcorrCF(west_fitgo, go = "rda_offset", clone = "full", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "rda_offset", clone = "full", fit_met = "len"),
fitcorrCF(west_fitgo, go = "rda_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(west_fitgo, go = "rda_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "rda_offset", clone = "mll", fit_met = "len"),
fitcorrCF(west_fitgo, go = "rda_offset", clone = "mll", fit_met = "weigh"))

# gf
gf_cf_w <- rbind(fitcorrCF(west_fitgo, go = "gf_offset", clone = "full", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "gf_offset", clone = "full", fit_met = "len"),
fitcorrCF(west_fitgo, go = "gf_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(west_fitgo, go = "gf_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "gf_offset", clone = "mll", fit_met = "len"),
fitcorrCF(west_fitgo, go = "gf_offset", clone = "mll", fit_met = "weigh"))

# env dist - note we use both full and mll to match with other methods when subsetting
envdist_cf_w <- rbind(fitcorrCF(west_fitgo, go = "env_dist", clone = "full", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "env_dist", clone = "full", fit_met = "len"),
fitcorrCF(west_fitgo, go = "env_dist", clone = "full", fit_met = "weigh"),
fitcorrCF(west_fitgo, go = "env_dist", clone = "mll", fit_met = "surv"),
fitcorrCF(west_fitgo, go = "env_dist", clone = "mll", fit_met = "len"),
fitcorrCF(west_fitgo, go = "env_dist", clone = "mll", fit_met = "weigh"))

cf_clim_w <- rbind(lea_cf_w, rda_cf_w, gf_cf_w, envdist_cf_w)

# pivot correlations and pvals to long format
cf_clim_w_long <- cf_clim_w %>% pivot_longer(cols = c(abs_corr, rel_corr, cf_abs_corr, cf_rel_corr, abs_pval, rel_pval, cf_abs_pval, cf_rel_pval),
    names_to = c("fit_type", ".value"),
    names_pattern = "(abs|rel|cf_abs|cf_rel)_(corr|pval)")

# signif & adjustments
cf_clim_w_long$signif <- ifelse(cf_clim_w_long$pval < 0.05, T, F)

# east coast
# only survival is informative for east coast, since only one pop has any growth data, but we will try all combos
# lea
lea_cf_e <- rbind(fitcorrCF(east_fitgo, go = "lea_offset", clone = "full", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "lea_offset", clone = "full", fit_met = "len"),
fitcorrCF(east_fitgo, go = "lea_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(east_fitgo, go = "lea_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "lea_offset", clone = "mll", fit_met = "len"),
fitcorrCF(east_fitgo, go = "lea_offset", clone = "mll", fit_met = "weigh"))

# rda
rda_cf_e <- rbind(fitcorrCF(east_fitgo, go = "rda_offset", clone = "full", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "rda_offset", clone = "full", fit_met = "len"),
fitcorrCF(east_fitgo, go = "rda_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(east_fitgo, go = "rda_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "rda_offset", clone = "mll", fit_met = "len"),
fitcorrCF(east_fitgo, go = "rda_offset", clone = "mll", fit_met = "weigh"))

# gf
gf_cf_e <- rbind(fitcorrCF(east_fitgo, go = "gf_offset", clone = "full", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "gf_offset", clone = "full", fit_met = "len"),
fitcorrCF(east_fitgo, go = "gf_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(east_fitgo, go = "gf_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "gf_offset", clone = "mll", fit_met = "len"),
fitcorrCF(east_fitgo, go = "gf_offset", clone = "mll", fit_met = "weigh"))

# env dist - note we use both full and mll to match with other methods when subsetting
envdist_cf_e <- rbind(fitcorrCF(east_fitgo, go = "env_dist", clone = "full", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "env_dist", clone = "full", fit_met = "len"),
fitcorrCF(east_fitgo, go = "env_dist", clone = "full", fit_met = "weigh"),
fitcorrCF(east_fitgo, go = "env_dist", clone = "mll", fit_met = "surv"),
fitcorrCF(east_fitgo, go = "env_dist", clone = "mll", fit_met = "len"),
fitcorrCF(east_fitgo, go = "env_dist", clone = "mll", fit_met = "weigh"))

cf_clim_e <- rbind(lea_cf_e, rda_cf_e, gf_cf_e, envdist_cf_e)

# pivot correlations and pvals to long format
cf_clim_e_long <- cf_clim_e %>% pivot_longer(cols = c(abs_corr, rel_corr, cf_abs_corr, cf_rel_corr, abs_pval, rel_pval, cf_abs_pval, cf_rel_pval),
    names_to = c("fit_type", ".value"),
    names_pattern = "(abs|rel|cf_abs|cf_rel)_(corr|pval)")

# signif & adjustments
cf_clim_e_long$signif <- ifelse(cf_clim_e_long$pval < 0.05, T, F)
##############

# save outputs
##############
write.csv(cf_clim_long, paste0("climchg_gofit_corrs_", Sys.Date(), ".csv"))
write.csv(cf_clim_w_long, paste0("climchg_gofit_corrs_west_", Sys.Date(), ".csv"))
write.csv(cf_clim_e_long, paste0("climchg_gofit_corrs_east_", Sys.Date(), ".csv"))
##############
```

### Visualization

Visualize strength of correlations. Try a heatmap with genomic offset as columns and fitness/fitness offset as rows


