---
title: "Eelgrass GO Fitness Comparisons"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/camillerumberger/Documents/GitHub/BalticEelgrass")
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# packages
library(car)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
library(knitr)
library(kableExtra)
library(emmeans)
library(multcomp)
library(multcompView)

# data
# exp fit
fit_full <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
surv <- read.csv("data/experiment/survival_dat_trt_pop.csv")[,-1]
grow <- read.csv("data/experiment/fit_growth_dat.csv")[,-1]

# fit offsets
fit_off_clim <- read.csv("results/fit_results/fit_offsets_clim_2026-05-21.csv")

# go
lea <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_2025-12-03.csv")[,-1]
rda <- read.csv("results/GO_results/RDA_offsets_exptrt_wide_2026-01-02.csv")[,-1]
gf <- read.csv("results/GO_results/GF_offsets_exptrt_wide_2025-12-03.csv")[,-1]
env_dist <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]
```

## Genomic Offset - Fitness Offset Comparisons

This script compares genomic offset against fitness metrics in a "common garden" (across genotypes) and "current future" (across environments) comparison. 


### GO Data Processing

These data are already in a matrix format, but I will still make a long form combined df for plotting. 

```{r}
head(lea)
head(rda)
head(gf)

# long versions
# lea
lea_long <- lea %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

# rda
rda_long <- rda %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "rda_offset")

# gf
gf_long <- gf %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "gf_offset")

# env dist
env_dist_trt <- env_dist %>% filter(pops1 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu") & !(pops2 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu"))) %>% dplyr::select(pops1, pops2, distances) %>% rename(pop = pops2, trt = pops1, env_dist = distances)

# put everything together
go1 <- merge(lea_long, rda_long, by = c("trt", "cc", "pop"))
go2 <- merge(go1, gf_long[gf_long$mod == "geno", c("trt","pop","cc","gf_offset")], by = c("trt", "cc", "pop"))
go_wide <- merge(go2, env_dist_trt, by = c("pop", "trt"))

# long version of all
go_long <- go_wide %>% pivot_longer(
  cols = c("lea_offset","rda_offset","gf_offset", "env_dist"),
  names_to = "method",
  values_to = "offset")
```

### Experimental Data

```{r}
# survival
head(surv)
dim(surv)
surv$trt <- gsub("_", "-", surv$trt)
fit_full$isAlive <- ifelse(fit_full$isAlive == TRUE, 1, 0)

# surv only fitness, all plants
surv_all <- fit_full[,c("pop", "trt", "tank_ID", "isAlive")]
surv_all$trt <- gsub("_", "-", surv_all$trt)
  
# also a format with survival in each tank/pop/trt
surv_tanks <- fit_full %>% 
  dplyr::select(pop, trt, tank_ID, isAlive) %>%
  group_by(pop, tank_ID, trt) %>%
  summarize(surv = mean(isAlive),
            surv_se = sd(isAlive)/length(isAlive))

# growth
head(grow)
grow_trim <- grow[,c("trt", "pop", "pop_trt", "tank_ID", "length_growth", "length_std_change", "weight_change", "weight_std_change", "above_below_weight_ratio")]
grow_trim$trt <- gsub("_","-",grow_trim$trt)

# some pop/trt level data
grow_pop_trt <- grow_trim %>% 
  group_by(pop, trt) %>% 
  summarize(avg_length_growth = mean(length_growth, na.rm = T),
    avg_length_std_change = mean(length_std_change, na.rm = T),
    avg_weight_change = mean(weight_change, na.rm = T),
    avg_weight_std_change = mean(weight_std_change, na.rm = T),
    avg_above_below_weight_ratio = mean(above_below_weight_ratio, na.rm = T))
```

### GO v Fitness

#### Data Merge

```{r}
# combine fitness and offset data
surv_go <- merge(go_wide, surv, by = c("pop", "trt"))
grow_go <- merge(go_wide, grow_trim, by = c("pop", "trt"))
surv_all_go <- merge(go_long, surv_all, by = c("pop", "trt"))

# add surv col to surv_all_go from isAlive for consistency in corr analysis
surv_all_go$surv <- surv_all_go$isAlive

# recode method
surv_all_go$method <- str_split_fixed(surv_all_go$method, "_", 2)[,1]

# numeric tank coding
tanks_df <- data.frame(tank_ID = levels(as.factor(surv_all_go$tank_ID)),
                       tank_num = 1:length(levels(as.factor(surv_all_go$tank_ID))))
surv_all_go <- merge(surv_all_go, tanks_df, by = "tank_ID")
surv_all_go$tank_num <- as.numeric(surv_all_go$tank_num)

# long versions
surv_go_long <- surv_go %>% pivot_longer(names_to = "method", cols = c("lea_offset", "rda_offset", "gf_offset", "env_dist"), values_to = "offset") %>% as.data.frame()
surv_go_long$method <- str_split_fixed(surv_go_long$method, "_", 2)[,1]
grow_go_long <- grow_go %>% pivot_longer(names_to = "method", cols = c("lea_offset", "rda_offset", "gf_offset", "env_dist"), values_to = "offset") %>% as.data.frame()
grow_go_long$method <- str_split_fixed(grow_go_long$method, "_", 2)[,1]
# surv all is already long format

# for correlation/signif testing 
trts <- levels(as.factor(surv_go$trt))
pops <- levels(as.factor(surv_go$pop))

# some checks
levels(as.factor(fit_off_clim$fit)) # only relative fitness here, no abs

# add fitness offsets
surv_go_long_off_clim <- fit_off_clim %>%
  filter(metric == "surv" & method == "cf") %>%
  rename(method_fit = method) %>%
  pivot_wider(id_cols = c(trt, pop, trt_ord, pop_ord, metric, coast),
              names_from = c(method_fit, fit),
              names_sep = "_",
              values_from = fit_off,
              values_fn = first) %>%
  mutate(trt = gsub("_", "-", trt))
surv_go_long_offset <- merge(surv_go_long, surv_go_long_off_clim, by = c("trt", "pop"))

grow_go_long_off_clim <- fit_off_clim_un %>% 
  filter(metric %in% c("len", "weigh")) %>% rename(method_fit = method) %>% pivot_wider(id_cols = c(trt, pop, trt_ord, pop_ord, metric), names_from = c(method_fit, fit), names_sep = "_", values_from = fit_off) %>% mutate(trt = gsub("_", "-", trt), metric = metric)
grow_go_long_offset <- merge(grow_go_long, grow_go_long_off_clim, by = c("trt", "pop"))

# save combined dfs
#write.csv(surv_go, "results/GO_compare/survival_offset_avgs.csv", row.names = F)
#write.csv(surv_all_go, "results/GO_compare/survival_offset_binary.csv", row.names = F)
#write.csv(grow_go, "results/GO_compare/grow_offset.csv", row.names = F)
#write.csv(surv_go_long_offset, "results/GO_compare/surv_fitoffset.csv", row.names = F)
#write.csv(grow_go_long_offset, "results/GO_compare/grow_fitoffset.csv", row.names = F)
```

#### Correlations

##### Correlations CF Offsets

```{r}
# function for correlations current future offset
 # df is dataframe, go is lea/rda/gf/env, clone is mll/full, coast is W/E, fit_met is surv/len/weigh, type is abs/rel
fitcorrCF <- function(df, go, clone, coast, fit_met, type) {
  # subset data
  df_subset <- df[df$method == go & df$cc == clone & df$coast == coast & df$metric == fit_met,]
  
  # set up df
  corrs_cf <- data.frame(fit_metric = NA, go_method = NA, coast = NA, pop = NA, cc = NA, kendall = NA, pval = NA, signif = NA)

    # calc fit offset
  for (i in 1:length(pops)){
    # subset
    pop_temp <- pops[i]
    temp <- df_subset[df_subset$pop == pop_temp,]
    
    # metadata
    corrs_cf[i,]$fit_metric <- fit_met
    corrs_cf[i,]$go_method <- go
    corrs_cf[i,]$cc <- clone
    corrs_cf[i,]$coast <- coast
    corrs_cf[i,]$pop <- pop_temp  
    
    # na
    prep <- temp[complete.cases(temp[, as.character(paste0("cf_", type))], temp$offset),]

    # extract data
    fit_offset <- as.numeric(prep[,paste0("cf_", type)])
    gen_offset <- as.numeric(prep$offset)
      
    # correlation
    if (nrow(temp) > 0){
      test <- cor.test(gen_offset, fit_offset, method = "kendall", )
      corrs_cf[i,]$kendall <- test$estimate
      corrs_cf[i,]$pval <- test$p.value
      corrs_cf[i,]$signif <- ifelse(test$p.value < 0.05, T, F)
    }elseif (nrow(temp) == 0){
      test <- NA
      corrs_cf[i,]$kendall_surv <- NA
      corrs_cf[i,]$pval_surv <- NA
      corrs_cf[i,]$signif <- NA
    }
  }
  return(corrs_cf)
}

# I get a ton of NAs here in kendall testing surv from populations with no individuals in the final treatment (i.e. full mortality) - trying where I set 0 values to 1e-10
surv_go_long_offset$surv_adj <- ifelse(surv_go_long_offset$surv == 0, 1e-10, surv_go_long_offset$surv)

# and calc
combos_surv_clim <- expand.grid(go = c("lea", "rda", "gf", "env"),
                      clone = c("mll", "full"),
                      fit_type = c("abs", "rel"),
                      fit_met = "surv_adj",
                      coast = c("West", "East"),
                      stringsAsFactors = F)
combos_grow_clim <- expand.grid(go = c("lea", "rda", "gf", "env"),
                      clone = c("mll", "full"),
                      fit_type = c("abs", "rel"),
                      fit_met = c("len", "weigh"),
                      coast = c("West", "East"),
                      stringsAsFactors = F)

surv_corrs_clim <- do.call(rbind, mapply(fitcorrCF, go = combos_surv_clim$go, clone = combos_surv_clim$clone, fit_met = combos_surv_clim$fit_met, type = combos_surv_clim$fit_type, coast = combos_surv_clim$coast,
                              MoreArgs = list(df = surv_go_long_offset), 
                              SIMPLIFY = F)) 
grow_corrs_clim <- do.call(rbind, mapply(fitcorrCF, go = combos_grow_clim$go, clone = combos_grow_clim$clone, fit_met = combos_grow_clim$fit_met, type = combos_grow_clim$fit_type, coast = combos_grow_clim$coast,
                              MoreArgs = list(df = grow_go_long_offset), 
                              SIMPLIFY = F))

# signif T/F
surv_corrs_clim$pval_adj <- p.adjust(surv_corrs_clim$pval)
surv_corrs_clim$signif_adj <- ifelse(surv_corrs_cg$pval_surv < 0.05, T, F)
###########

# save outputs
##############
#write.csv(surv_corrs_cf, "results/GO_compare/surv_corrs_cf.csv", row.names = F)
#write.csv(surv_corrs_cg, "results/GO_compare/surv_corrs_cg.csv", row.names = F)
#write.csv(grow_corr_cf, "results/GO_compare/grow_corr_cf.csv", row.names = F)
#write.csv(grow_corr_cg, "results/GO_compare/grow_corr_cg.csv", row.names = F)
##############
```

##### Fitness Offset Correlations CG & CF

```{r}
## SURVIVAL
###########
# relevant datasets
head(surv_go_long_offset)
head(grow_go_long_offset)

# function for correlations both common garden/current future abs/rel fitness offsets
fit_corr_clim <- function(df, go, clone, coast, fit_typ, fit_met) { # go can be lea, rda, gf, or env; clone can be mll or full; fit_off can be cf or cg, fit_typ can be abs or rel; fit_met can be surv, weigh, or len
  # setup empty df to populate
  cor_df <- data.frame(go_method = NA, fit_metric, fit_method = NA, fit_type = NA, cc = NA, pearson = NA, pearson_pval = NA, spearman = NA, spearman_pval = NA, kendall = NA, kendall_pval = NA, r2 = NA, signif = NA)
  
  # subset data
  df_subset <- df[df$method == go & df$cc == clone & df$metric == fit_met,]
  
  # na
  df_prep <- df_subset[complete.cases(df_subset[, as.character(paste0(fit_off, "_", fit_typ))], df_subset$offset),]

  # extract data
  fit_offset <- as.numeric(df_prep[,paste0(fit_off, "_", fit_typ)])
#    na.omit(as.numeric(df_subset[, lookup_metric[[fit_method]][[fit_type]]]))
  gen_offset <- as.numeric(df_prep$offset)

  # metadata
  cor_df$go_method <- as.character(go)
  cor_df$fit_metric <- as.character(fit_met)
  cor_df$fit_method <- as.character(fit_off)
  cor_df$fit_type <- as.character(fit_typ)
  cor_df$cc <- as.character(clone)
    
  # correlation
  cor_df$pearson <- cor.test(fit_offset, gen_offset, method = "pearson")$estimate
  cor_df$pearson_pval <- cor.test(fit_offset, gen_offset, method = "pearson")$p.value
  cor_df$spearman <- cor.test(fit_offset, gen_offset, method = "spearman")$estimate
  cor_df$spearman_pval <- cor.test(fit_offset, gen_offset, method = "spearman")$p.value
  cor_df$kendall <- cor.test(fit_offset, gen_offset, method = "kendall")$estimate
  cor_df$kendall_pval <- cor.test(fit_offset, gen_offset, method = "kendall")$p.value
  cor_df$r2 <- summary(lm(gen_offset ~ fit_offset))$r.squared
  cor_df$signif <- ifelse(cor_df$kendall_pval < 0.05, T, F)
  
  return(cor_df)
}

combos_surv <- expand.grid(go = c("lea", "rda", "gf", "env"),
                      clone = c("mll", "full"),
                      fit_type = c("abs", "rel"),
                      fit_met = "surv",
                      coast = c("West", "East"),
                      stringsAsFactors = F)

combos_grow <- expand.grid(go = c("lea", "rda", "gf", "env"),
                      clone = c("mll", "full"),
                      fit_type = c("abs", "rel"),
                      fit_met = c("len", "weigh"),
                      coast = c("West", "East"),
                      stringsAsFactors = F)

surv_corrs <- do.call(rbind, mapply(fit_corr_clim, go = combos_surv$go, clone = combos_surv$clone, fit_met = combos_surv$fit_met, fit_type = combos_surv$fit_type, 
                              MoreArgs = list(df = surv_go_long_offset), 
                              SIMPLIFY = F))
grow_corrs <- do.call(rbind, mapply(fit_corr_clim, go = combos_grow$go, clone = combos_grow$clone, fit_met = combos_grow$fit_met, fit_type = combos_grow$fit_typ, 
                              MoreArgs = list(df = grow_go_long_offset), 
                              SIMPLIFY = F))

row.names(surv_corrs) <- row.names(grow_corrs) <- NULL

# correction for multiple testing
surv_corrs$pval_adj <- p.adjust(surv_corrs$kendall_pval, method = "holm")
surv_corrs$signif_adj <- ifelse(surv_corrs$pval_adj < 0.05, T, F)
grow_corrs$pval_adj <- p.adjust(grow_corrs$kendall_pval, method = "holm")
grow_corrs$signif_adj <- ifelse(grow_corrs$pval_adj < 0.05, T, F)

# combine two dfs
fitoff_corrs <- rbind(surv_corrs, grow_corrs)

# save outputs
write.csv(fitoff_corrs, "results/GO_compare/fitoff_corrs.csv", row.names = F)
###########
```


#### Modeling

##### GO only

Survival 

```{r}
# quick checks + fixes
hist(surv_all_go$surv) # we will use binary outcomes first

## lea
######
# model full (no cc)
surv_mod_lea_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "lea",],
               na.action = na.exclude)
summary(surv_mod_lea_full) 
#plot(surv_mod_lea_full) # these look pretty bad
Anova(surv_mod_lea_full) # signif

# model mll (cc)
surv_mod_lea_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_mod_lea_mll) 
#plot(surv_mod_lea_mll) # these look pretty bad
Anova(surv_mod_lea_mll) # signif

# now the tank level pop/trt
surv_mod_lea_poptrt_full <- glm(surv ~ lea_offset,
               family = binomial(link = "logit"),
               data = surv_go)
summary(surv_mod_lea_poptrt_full) 
#plot(surv_mod_lea_full)
Anova(surv_mod_lea_poptrt_full) # signif

surv_mod_lea_poptrt_mll <- glm(surv ~ lea_offset,
               family = binomial(link = "logit"),
               data = surv_go[surv_go$cc == "mll",])
summary(surv_mod_lea_poptrt_mll) 
#plot(surv_mod_lea_poptrt_mll)
Anova(surv_mod_lea_poptrt_mll) # signif

# which pops diff?
emm_surv_lea_full <- emmeans(surv_mod_lea_full, "offset")
pairs(emm_surv_lea_full)
######

## rda
######
# model full (no cc)
surv_mod_rda_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_mod_rda_full) 
#plot(surv_mod_rda_full) # these look pretty bad
Anova(surv_mod_rda_full) # signif

# model mll (cc)
surv_mod_rda_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_mod_rda_mll) 
#plot(surv_mod_rda_mll) # these look pretty bad
Anova(surv_mod_rda_mll) # signif

# now the tank level pop/trt
surv_mod_rda_poptrt_full <- glm(surv ~ rda_offset,
               family = binomial(link = "logit"),
               data = surv_go[surv_go$cc == "full",])
summary(surv_mod_rda_poptrt_full) 
#plot(surv_mod_rda_full)
Anova(surv_mod_rda_poptrt_full) # not signif

surv_mod_rda_poptrt_mll <- glm(surv ~ rda_offset,
               family = binomial(link = "logit"),
               data = surv_go[surv_go$cc == "mll",])
summary(surv_mod_rda_poptrt_mll) 
#plot(surv_mod_rda_poptrt_mll)
Anova(surv_mod_rda_poptrt_mll) # not signif
######

## gf
#####
# model full (no cc)
surv_mod_gf_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_mod_gf_full) 
#plot(surv_mod_gf_full) # these look pretty bad
Anova(surv_mod_gf_full) # signif

# model mll (cc)
surv_mod_gf_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_mod_gf_mll) 
#plot(surv_mod_gf_mll) # these look pretty bad
Anova(surv_mod_gf_mll) # signif

# now the tank level pop/trt
surv_mod_gf_poptrt_full <- glm(surv ~ gf_offset,
               family = binomial(link = "logit"),
               data = surv_go_full)
summary(surv_mod_gf_poptrt_full) 
#plot(surv_mod_gf_poptrt_full) # better than binary model
Anova(surv_mod_gf_poptrt_full) # not signif

surv_mod_gf_poptrt_mll <- glm(surv ~ gf_offset,
               family = binomial(link = "logit"),
               data = surv_go_mll)
summary(surv_mod_gf_poptrt_mll) 
#plot(surv_mod_gf_poptrt_mll)
Anova(surv_mod_gf_poptrt_mll) # not signif
######
```

Length

```{r}
# length
########
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
# model full (no cc)
len_mod_lea_full <- lm(length_std_change ~ log(lea_offset), data = grow_go_full)
summary(len_mod_lea_full) 
#plot(len_mod_lea_full) # diverges pretty significantly on right tail
Anova(len_mod_lea_full) # signif

# model mll (cc)
len_mod_lea_mll <- lm(length_std_change ~ log(lea_offset), data = grow_go_mll)
summary(len_mod_lea_mll) 
#plot(len_mod_lea_mll) # diverges pretty significantly on right tail
Anova(len_mod_lea_mll) # signif

## rda
# model full (no cc)
len_mod_rda_full <- lm(length_std_change ~ log(rda_offset), data = grow_go_full)
summary(len_mod_rda_full) 
#plot(len_mod_rda_full) # diverges pretty significantly on right tail
Anova(len_mod_rda_full) # signif

# model mll (cc)
len_mod_rda_mll <- lm(length_std_change ~ log(rda_offset), data = grow_go_mll)
summary(len_mod_rda_mll) 
#plot(len_mod_rda_mll) # diverges pretty significantly on right tail
Anova(len_mod_rda_mll) # signif

## gf
# model full (no cc)
len_mod_gf_full <- lm(length_std_change ~ log(gf_offset), data = grow_go_full)
summary(len_mod_gf_full) 
#plot(len_mod_gf_full) # diverges pretty significantly on right tail
Anova(len_mod_gf_full) # signif

# model mll (cc)
len_mod_gf_mll <- lm(length_std_change ~ log(gf_offset), data = grow_go_mll)
summary(len_mod_gf_mll) 
#plot(len_mod_gf_mll) # diverges pretty significantly on right tail
Anova(len_mod_gf_mll) # signif
########
```

Growth

```{r}
# weight
########
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
# model full (no cc)
weigh_mod_lea_full <- lm(weight_std_change ~ lea_offset, data = grow_go_full)
summary(weigh_mod_lea_full) 
#plot(weigh_mod_lea_full) # diverges pretty significantly on right tail
Anova(weigh_mod_lea_full) # signif

# model mll (cc)
weigh_mod_lea_mll <- lm(weight_std_change ~ lea_offset, data = grow_go_mll)
summary(weigh_mod_lea_mll) 
#plot(weigh_mod_lea_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_lea_mll) # signif

## rda
# model full (no cc)
weigh_mod_rda_full <- lm(weight_std_change ~ (rda_offset), data = grow_go_full)
summary(weigh_mod_rda_full) 
#plot(weigh_mod_rda_full) # diverges pretty significantly on right tail
Anova(weigh_mod_rda_full) # signif

# model mll (cc)
weigh_mod_rda_mll <- lm(weight_std_change ~ (rda_offset), data = grow_go_mll)
summary(weigh_mod_rda_mll) 
#plot(weigh_mod_rda_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_rda_mll) # signif

## gf
# model full (no cc)
weigh_mod_gf_full <- lm(weight_std_change ~ (gf_offset), data = grow_go_full)
summary(weigh_mod_gf_full) 
#plot(weigh_mod_gf_full) # diverges pretty significantly on right tail
Anova(weigh_mod_gf_full) # signif

# model mll (cc)
weigh_mod_gf_mll <- lm(weight_std_change ~ (gf_offset), data = grow_go_mll)
summary(weigh_mod_gf_mll) 
#plot(weigh_mod_gf_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_gf_mll) # signif

# save these outputs
sink("results/fit_results/survival_modeling.txt")
"Survival\n"
"LEA MLL\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"LEA Full\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"RDA MLL\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"RDA Full\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"GF MLL\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"GF Full\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

sink()
########
```

##### With population as fixed effect

Survival 

```{r}
# quick checks + fixes
hist(surv_all_go$surv) # we will use binary outcomes first

## lea
######
# model full (no cc)
surv_pop_lea_full <- lm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",])
summary(surv_pop_lea_full) 
#plot(surv_pop_lea_full)
Anova(surv_pop_lea_full) # signif
emm_survpop_lea_full <- emmeans(surv_pop_lea_full, "pop")
letters_survpop_lea_full <- cld(emm_survpop_lea_full, Letters = letters)

# model mll (cc)
surv_pop_lea_mll <- lm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_pop_lea_mll) 
#plot(surv_pop_lea_mll)
Anova(surv_pop_lea_mll) # signif
emm_survpop_lea_mll <- emmeans(surv_pop_lea_mll, "pop")
letters_survpop_lea_mll <- cld(emm_survpop_lea_mll, Letters = letters)

# model full (no cc) log
surv_pop_lea_full_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",],
               family = binomial)
summary(surv_pop_lea_full_log) 
#plot(surv_pop_lea_full_log)
Anova(surv_pop_lea_full_log) # signif
emm_survpop_lea_full_log <- emmeans(surv_pop_lea_full_log, "pop")
letters_survpop_lea_full_log <- cld(emm_survpop_lea_full_log, Letters = letters)

# model mll (cc) log
surv_pop_lea_mll_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",],
               family = binomial)
summary(surv_pop_lea_mll_log) 
#plot(surv_pop_lea_mll_log)
Anova(surv_pop_lea_mll_log) # signif
emm_survpop_lea_mll_log <- emmeans(surv_pop_lea_mll_log, "pop")
letters_survpop_lea_mll_log <- cld(emm_survpop_lea_mll_log, Letters = letters)
######

## rda
######
# model full (no cc)
surv_pop_rda_full <- lm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_pop_rda_full) 
#plot(surv_pop_rda_full)
Anova(surv_pop_rda_full) # signif
emm_survpop_rda_full <- emmeans(surv_pop_rda_full, "pop")
letters_survpop_rda_full <- cld(emm_survpop_rda_full, Letters = letters)

# model mll (cc)
surv_pop_rda_mll <- lm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_pop_rda_mll) 
#plot(surv_pop_rda_mll)
Anova(surv_pop_rda_mll) # signif
emm_survpop_rda_mll <- emmeans(surv_pop_rda_mll, "pop")
letters_survpop_rda_mll <- cld(emm_survpop_rda_mll, Letters = letters)

# model full (no cc) log
surv_pop_rda_full_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",],
               family = binomial)
summary(surv_pop_rda_full_log) 
#plot(surv_pop_rda_full_log)
Anova(surv_pop_rda_full_log) # signif
emm_survpop_rda_full_log <- emmeans(surv_pop_rda_full_log, "pop")
letters_survpop_rda_full_log <- cld(emm_survpop_rda_full_log, Letters = letters)

# model mll (cc) log
surv_pop_rda_mll_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",],
               family = binomial)
summary(surv_pop_rda_mll_log) 
#plot(surv_pop_rda_mll_log)
Anova(surv_pop_rda_mll_log) # signif
emm_survpop_rda_mll_log <- emmeans(surv_pop_rda_mll_log, "pop")
letters_survpop_rda_mll_log <- cld(emm_survpop_rda_mll_log, Letters = letters)
######

## gf
#####
# model full (no cc)
surv_pop_gf_full <- lm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_pop_gf_full) 
#plot(surv_pop_gf_full)
Anova(surv_pop_gf_full) # signif
emm_survpop_gf_full <- emmeans(surv_pop_gf_full, "pop")
letters_survpop_gf_full <- cld(emm_survpop_gf_full, Letters = letters)

# model mll (cc)
surv_pop_gf_mll <- lm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_pop_gf_mll) 
#plot(surv_pop_gf_mll)
Anova(surv_pop_gf_mll) # signif
emm_survpop_gf_mll <- emmeans(surv_pop_gf_mll, "pop")
letters_survpop_gf_mll <- cld(emm_survpop_gf_mll, Letters = letters)

# model full (no cc) log
surv_pop_gf_full_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",],
               family = binomial)
summary(surv_pop_gf_full_log) 
#plot(surv_pop_gf_full_log)
Anova(surv_pop_gf_full_log) # signif
emm_survpop_gf_full_log <- emmeans(surv_pop_gf_full_log, "pop")
letters_survpop_gf_full_log <- cld(emm_survpop_gf_full_log, Letters = letters)

# model mll (cc) log
surv_pop_gf_mll_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",],
               family = binomial)
summary(surv_pop_gf_mll_log) 
#plot(surv_pop_gf_mll_log)
Anova(surv_pop_gf_mll_log) # signif
emm_survpop_gf_mll_log <- emmeans(surv_pop_gf_mll_log, "pop")
letters_survpop_gf_mll_log <- cld(emm_survpop_gf_mll_log, Letters = letters)
#####
```

Length

```{r}
# length/pop
############
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
######
# model full (no cc)
len_pop_lea_full <- lm(log(length_std_change) ~ pop + lea_offset, data = grow_go_full)
summary(len_pop_lea_full) 
#plot(len_pop_lea_full) # diverges pretty significantly on left tail
Anova(len_pop_lea_full) # signif
emm_lenpop_lea_full <- emmeans(len_pop_lea_full, "pop")
letters_lenpop_lea_full <- cld(emm_lenpop_lea_full, Letters = letters)

# model mll (cc)
len_pop_lea_mll <- lm(log(length_std_change) ~ pop + lea_offset, data = grow_go_mll)
summary(len_pop_lea_mll) 
#plot(len_pop_lea_mll) # diverges pretty significantly on right tail
Anova(len_pop_lea_mll) # signif
emm_lenpop_lea_mll <- emmeans(len_pop_lea_mll, "pop")
letters_lenpop_lea_mll <- cld(emm_lenpop_lea_mll, Letters = letters)
######

## rda
######
# model full (no cc)
len_pop_rda_full <- lm(log(length_std_change) ~ pop + rda_offset, data = grow_go_full)
summary(len_pop_rda_full) 
#plot(len_pop_rda_full)
Anova(len_pop_rda_full) # rda offset not signif
emm_lenpop_rda_full <- emmeans(len_pop_rda_full, "pop")
letters_lenpop_rda_full <- cld(emm_lenpop_rda_full, Letters = letters)

# model mll (cc)
len_pop_rda_mll <- lm(log(length_std_change) ~ pop + rda_offset, data = grow_go_mll)
summary(len_pop_rda_mll) 
#plot(len_pop_rda_mll)
Anova(len_pop_rda_mll) # signif
emm_lenpop_rda_mll <- emmeans(len_pop_rda_mll, "pop")
letters_lenpop_rda_mll <- cld(emm_lenpop_rda_mll, Letters = letters)
######

## gf
#####
# model full (no cc)
len_pop_gf_full <- lm(log(length_std_change) ~ pop + gf_offset, data = grow_go_full)
summary(len_pop_gf_full) 
#plot(len_pop_gf_full)
Anova(len_pop_gf_full) # signif
emm_lenpop_gf_full <- emmeans(len_pop_gf_full, "pop")
letters_lenpop_gf_full <- cld(emm_lenpop_gf_full, Letters = letters)

# model mll (cc)
len_pop_gf_mll <- lm(log(length_std_change) ~ pop + gf_offset, data = grow_go_mll)
summary(len_pop_gf_mll) 
#plot(len_pop_gf_mll)
Anova(len_pop_gf_mll) # signif
emm_lenpop_gf_mll <- emmeans(len_pop_gf_mll, "pop")
letters_lenpop_gf_mll <- cld(emm_lenpop_gf_mll, Letters = letters)
#####
```

Weight

```{r}
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
######
# model full (no cc)
weigh_pop_lea_full <- lm(weight_std_change ~ pop + lea_offset, data = grow_go_full)
summary(weigh_pop_lea_full) 
#plot(weigh_pop_lea_full)
Anova(weigh_pop_lea_full) # signif
emm_weighpop_lea_full <- emmeans(weigh_pop_lea_full, "pop")
letters_weighpop_lea_full <- cld(emm_weighpop_lea_full, Letters = letters)

# model mll (cc)
weigh_pop_lea_mll <- lm(weight_std_change ~ pop + lea_offset, data = grow_go_mll)
summary(weigh_pop_lea_mll) 
#plot(weigh_pop_lea_mll)
Anova(weigh_pop_lea_mll) # signif
######

## rda
# model full (no cc)
weigh_pop_rda_full <- lm(weight_std_change ~ pop + rda_offset, data = grow_go_full)
summary(weigh_pop_rda_full) 
#plot(weigh_pop_rda_full) 
Anova(weigh_pop_rda_full) # signif

# model mll (cc)
weigh_pop_rda_mll <- lm(weight_std_change ~ pop + rda_offset, data = grow_go_mll)
summary(weigh_pop_rda_mll) 
#plot(weigh_pop_rda_mll) 
Anova(weigh_pop_rda_mll) # signif
######

## gf
#####
# model full (no cc)
weigh_pop_gf_full <- lm(weight_std_change ~ pop + gf_offset, data = grow_go_full)
summary(weigh_pop_gf_full) 
#plot(weigh_pop_gf_full) 
Anova(weigh_pop_gf_full) # signif

# model mll (cc)
weigh_pop_gf_mll <- lm(weight_std_change ~ pop + gf_offset, data = grow_go_mll)
summary(weigh_pop_gf_mll) 
#plot(weigh_pop_gf_mll)
Anova(weigh_pop_gf_mll) # signif
#####
```


##### With treatment as fixed effect

Survival 

```{r}
# quick checks + fixes
#hist(surv_all_go$surv) # we will use binary outcomes first

## lea
######
# model full (no cc)
surv_trt_lea_full <- lm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",])
summary(surv_trt_lea_full) 
#plot(surv_trt_lea_full)
Anova(surv_trt_lea_full) # signif
emm_survtrt_lea_full <- emmeans(surv_trt_lea_full, "trt")
letters_survtrt_lea_full <- cld(emm_survtrt_lea_full, Letters = letters)

# model mll (cc)
surv_trt_lea_mll <- lm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_trt_lea_mll) 
#plot(surv_trt_lea_mll)
Anova(surv_trt_lea_mll) # signif
emm_survtrt_lea_mll <- emmeans(surv_trt_lea_mll, "trt")
letters_survtrt_lea_mll <- cld(emm_survtrt_lea_mll, Letters = letters)

# model full (no cc) log
surv_trt_lea_full_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",],
               family = binomial)
summary(surv_trt_lea_full_log) 
#plot(surv_trt_lea_full_log)
Anova(surv_trt_lea_full_log) # signif
emm_survtrt_lea_full_log <- emmeans(surv_trt_lea_full_log, "trt")
letters_survtrt_lea_full_log <- cld(emm_survtrt_lea_full_log, Letters = letters)

# model mll (cc) log
surv_trt_lea_mll_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",],
               family = binomial)
summary(surv_trt_lea_mll_log) 
#plot(surv_trt_lea_mll_log)
Anova(surv_trt_lea_mll_log) # signif
emm_survtrt_lea_mll_log <- emmeans(surv_trt_lea_mll_log, "trt")
letters_survtrt_lea_mll_log <- cld(emm_survtrt_lea_mll_log, Letters = letters)
######

## rda
######
# model full (no cc)
surv_trt_rda_full <- lm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_trt_rda_full) 
#plot(surv_trt_rda_full)
Anova(surv_trt_rda_full) # signif trt, not offset
emm_survtrt_rda_full <- emmeans(surv_trt_rda_full, "trt")
letters_survtrt_rda_full <- cld(emm_survtrt_rda_full, Letters = letters)

# model mll (cc)
surv_trt_rda_mll <- lm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_trt_rda_mll) 
#plot(surv_trt_rda_mll)
Anova(surv_trt_rda_mll) # signif trt, not offset
emm_survtrt_rda_mll <- emmeans(surv_trt_rda_mll, "trt")
letters_survtrt_rda_mll <- cld(emm_survtrt_rda_mll, Letters = letters)

# model full (no cc) log
surv_trt_rda_full_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",],
               family = binomial)
summary(surv_trt_rda_full_log) 
#plot(surv_trt_rda_full_log)
Anova(surv_trt_rda_full_log) # signif trt, not offset
emm_survtrt_rda_full_log <- emmeans(surv_trt_rda_full_log, "trt")
letters_survtrt_rda_full_log <- cld(emm_survtrt_rda_full_log, Letters = letters)

# model mll (cc) log
surv_trt_rda_mll_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",],
               family = binomial)
summary(surv_trt_rda_mll_log) 
#plot(surv_trt_rda_mll_log)
Anova(surv_trt_rda_mll_log) # signif trt, not offset
emm_survtrt_rda_mll_log <- emmeans(surv_trt_rda_mll_log, "trt")
letters_survtrt_rda_mll_log <- cld(emm_survtrt_rda_mll_log, Letters = letters)
######

## gf
#####
# model full (no cc)
surv_trt_gf_full <- lm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_trt_gf_full) 
#plot(surv_trt_gf_full)
Anova(surv_trt_gf_full) # signif
emm_survtrt_gf_full <- emmeans(surv_trt_gf_full, "trt")
letters_survtrt_gf_full <- cld(emm_survtrt_gf_full, Letters = letters)

# model mll (cc)
surv_trt_gf_mll <- lm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_trt_gf_mll) 
#plot(surv_trt_gf_mll)
Anova(surv_trt_gf_mll) # signif
emm_survtrt_gf_mll <- emmeans(surv_trt_gf_mll, "trt")
letters_survtrt_gf_mll <- cld(emm_survtrt_gf_mll, Letters = letters)

# model full (no cc)
surv_trt_gf_full_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",],
               family = binomial)
summary(surv_trt_gf_full_log) 
#plot(surv_trt_gf_full_log)
Anova(surv_trt_gf_full_log) # signif
emm_survtrt_gf_full_log <- emmeans(surv_trt_gf_full_log, "trt")
letters_survtrt_gf_ful_log <- cld(emm_survtrt_gf_full_log, Letters = letters)

# model mll (cc)
surv_trt_gf_mll_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",],
               family = binomial)
summary(surv_trt_gf_mll_log) 
#plot(surv_trt_gf_mll_log)
Anova(surv_trt_gf_mll_log) # signif
emm_survtrt_gf_mll_log <- emmeans(surv_trt_gf_mll_log, "trt")
letters_survtrt_gf_mll_log <- cld(emm_survtrt_gf_mll_log, Letters = letters)
#####
```

Length

```{r}
# length/trt
############
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
######
# model full (no cc)
len_trt_lea_full <- lm(log(length_std_change) ~ trt + lea_offset, data = grow_go_full)
summary(len_trt_lea_full) 
#plot(len_trt_lea_full) # diverges pretty significantly on left tail
Anova(len_trt_lea_full) # signif
emm_lentrt_lea_full <- emmeans(len_trt_lea_full, "trt")
letters_lentrt_lea_full <- cld(emm_lentrt_lea_full, Letters = letters)

# model mll (cc)
len_trt_lea_mll <- lm(log(length_std_change) ~ trt + lea_offset, data = grow_go_mll)
summary(len_trt_lea_mll) 
#plot(len_trt_lea_mll) # diverges pretty significantly on right tail
Anova(len_trt_lea_mll) # not signif
emm_lentrt_lea_mll <- emmeans(len_trt_lea_mll, "trt")
letters_lentrt_lea_mll <- cld(emm_lentrt_lea_mll, Letters = letters)
######

## rda
######
# model full (no cc)
len_trt_rda_full <- lm(log(length_std_change) ~ trt + rda_offset, data = grow_go_full)
summary(len_trt_rda_full) 
#plot(len_trt_rda_full)
Anova(len_trt_rda_full) # signif
emm_lentrt_rda_full <- emmeans(len_trt_rda_full, "trt")
letters_lentrt_rda_full <- cld(emm_lentrt_rda_full, Letters = letters)

# model mll (cc)
len_trt_rda_mll <- lm(log(length_std_change) ~ trt + rda_offset, data = grow_go_mll)
summary(len_trt_rda_mll) 
#plot(len_trt_rda_mll)
Anova(len_trt_rda_mll) # signif
emm_lentrt_rda_mll <- emmeans(len_trt_rda_mll, "trt")
letters_lentrt_rda_mll <- cld(emm_lentrt_rda_mll, Letters = letters)
######

## gf
#####
# model full (no cc)
len_trt_gf_full <- lm(log(length_std_change) ~ trt + gf_offset, data = grow_go_full)
summary(len_trt_gf_full) 
#plot(len_trt_gf_full)
Anova(len_trt_gf_full) # signif
emm_lentrt_gf_full <- emmeans(len_trt_rda_mll, "trt")
letters_lentrt_rda_mll <- cld(emm_lentrt_gf_full, Letters = letters)

# model mll (cc)
len_trt_gf_mll <- lm(log(length_std_change) ~ trt + gf_offset, data = grow_go_mll)
summary(len_trt_gf_mll) 
#plot(len_trt_gf_mll)
Anova(len_trt_gf_mll) # signif
emm_lentrt_rda_mll <- emmeans(len_trt_rda_mll, "trt")
letters_lentrt_rda_mll <- cld(emm_lentrt_rda_mll, Letters = letters)
#####
```

Weight

```{r}
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
######
# model full (no cc)
weigh_trt_lea_full <- lm(weight_std_change ~ trt + lea_offset, data = grow_go_full)
summary(weigh_trt_lea_full) 
#plot(weigh_trt_lea_full)
Anova(weigh_trt_lea_full) # signif
emm_weightrt_lea_full <- emmeans(weigh_trt_lea_full, "trt")
letters_weightrt_lea_full <- cld(emm_weightrt_lea_full, Letters = letters)

# model mll (cc)
weigh_trt_lea_mll <- lm(weight_std_change ~ trt + lea_offset, data = grow_go_mll)
summary(weigh_trt_lea_mll) 
#plot(weigh_trt_lea_mll)
Anova(weigh_trt_lea_mll) # signif
emm_weightrt_lea_mll <- emmeans(weigh_trt_lea_mll, "trt")
letters_weightrt_lea_mll <- cld(emm_weightrt_lea_mll, Letters = letters)
######

## rda
# model full (no cc)
weigh_trt_rda_full <- lm(weight_std_change ~ trt + rda_offset, data = grow_go_full)
summary(weigh_trt_rda_full) 
#plot(weigh_trt_rda_full) 
Anova(weigh_trt_rda_full) # signif
emm_weightrt_rda_full <- emmeans(weigh_trt_rda_full, "trt")
letters_weightrt_rda_full <- cld(emm_weightrt_rda_full, Letters = letters)

# model mll (cc)
weigh_trt_rda_mll <- lm(weight_std_change ~ trt + rda_offset, data = grow_go_mll)
summary(weigh_trt_rda_mll) 
#plot(weigh_trt_rda_mll) 
Anova(weigh_trt_rda_mll) # signif
emm_weightrt_rda_mll <- emmeans(weigh_trt_rda_mll, "trt")
letters_weightrt_rda_mll <- cld(emm_weightrt_rda_mll, Letters = letters)
######

## gf
#####
# model full (no cc)
weigh_trt_gf_full <- lm(weight_std_change ~ trt + gf_offset, data = grow_go_full)
summary(weigh_trt_gf_full) 
#plot(weigh_trt_gf_full) 
Anova(weigh_trt_gf_full) # signif
emm_weightrt_gf_full <- emmeans(weigh_trt_gf_full, "trt")
letters_weightrt_gf_full <- cld(emm_weightrt_gf_full, Letters = letters)

# model mll (cc)
weigh_trt_gf_mll <- lm(weight_std_change ~ trt + gf_offset, data = grow_go_mll)
summary(weigh_trt_gf_mll) 
#plot(weigh_trt_gf_mll)
Anova(weigh_trt_gf_mll) # signif
emm_weightrt_gf_mll <- emmeans(weigh_trt_gf_mll, "trt")
letters_weightrt_gf_mll <- cld(emm_weightrt_gf_mll, Letters = letters)
#####
```

##### Model Outputs

Survival

```{r}
# list of info for all survival ancova tables
surv_mod_info <- list(
  list(model = Anova(surv_pop_lea_full), title = "ANCOVA: Survival, LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_full), title = "ANCOVA: Survival, LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_full), title = "ANCOVA: Survival, RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_full), title = "ANCOVA: Survival, RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_full), title = "ANCOVA: Survival, GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_full), title = "ANCOVA: Survival, GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_lea_mll), title = "ANCOVA: Survival, LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_mll), title = "ANCOVA: Survival, LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_mll), title = "ANCOVA: Survival, RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_mll), title = "ANCOVA: Survival, RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_mll), title = "ANCOVA: Survival, GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_mll), title = "ANCOVA: Survival, GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
surv_tables <- lapply(surv_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
surv_output <- paste(
  "<!DOCTYPE html><html><head><title>ANCOVA Results</title></head><body>",
  paste(surv_tables, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(surv_output, file = "results/GO_compare/survival_ancovas.html")

# for logistic regression as well
surv_log_mod_info <- list(
  list(model = Anova(surv_pop_lea_full_log), title = "Log Regression: Survival, LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_full_log), title = "Log Regression: Survival, LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_full_log), title = "Log Regression: Survival, RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_full_log), title = "Log Regression: Survival, RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_full_log), title = "Log Regression: Survival, GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_full_log), title = "Log Regression: Survival, GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_lea_mll_log), title = "Log Regression: Survival, LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_mll_log), title = "Log Regression: Survival, LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_mll_log), title = "Log Regression: Survival, RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_mll_log), title = "Log Regression: Survival, RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_mll_log), title = "Log Regression: Survival, GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_mll_log), title = "Log Regression: Survival, GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
surv_tables_log <- lapply(surv_log_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
surv_output_log <- paste(
  "<!DOCTYPE html><html><head><title>Log Regression Results</title></head><body>",
  paste(surv_tables_log, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(surv_output_log, file = "results/GO_compare/survival_log_regression.html")
```

Length

```{r}
# list of info for all survival ancova tables
len_mod_info <- list(
  list(model = Anova(len_pop_lea_full), title = "ANCOVA: Length Change (Standardized), LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_lea_full), title = "ANCOVA: Length Change (Standardized), LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_rda_full), title = "ANCOVA: Length Change (Standardized), RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_rda_full), title = "ANCOVA: Length Change (Standardized), RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_gf_full), title = "ANCOVA: Length Change (Standardized), GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_gf_full), title = "ANCOVA: Length Change (Standardized), GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_lea_mll), title = "ANCOVA: Length Change (Standardized), LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_lea_mll), title = "ANCOVA: Length Change (Standardized), LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_rda_mll), title = "ANCOVA: Length Change (Standardized), RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_rda_mll), title = "ANCOVA: Length Change (Standardized), RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_gf_mll), title = "ANCOVA: Length Change (Standardized), GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_gf_mll), title = "ANCOVA: Length Change (Standardized), GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
len_tables <- lapply(len_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
len_output <- paste(
  "<!DOCTYPE html><html><head><title>ANCOVA Results</title></head><body>",
  paste(len_tables, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(len_output, file = "results/GO_compare/length_ancovas.html")
```

Weight

```{r}
# list of info for all weighival ancova tables
weigh_mod_info <- list(
  list(model = Anova(weigh_pop_lea_full), title = "ANCOVA: Weight Change (Standardized), LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_lea_full), title = "ANCOVA: Weight Change (Standardized), LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_rda_full), title = "ANCOVA: Weight Change (Standardized), RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_rda_full), title = "ANCOVA: Weight Change (Standardized), RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_gf_full), title = "ANCOVA: Weight Change (Standardized), GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_gf_full), title = "ANCOVA: Weight Change (Standardized), GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_lea_mll), title = "ANCOVA: Weight Change (Standardized), LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_lea_mll), title = "ANCOVA: Weight Change (Standardized), LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_rda_mll), title = "ANCOVA: Weight Change (Standardized), RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_rda_mll), title = "ANCOVA: Weight Change (Standardized), RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_gf_mll), title = "ANCOVA: Weight Change (Standardized), GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_gf_mll), title = "ANCOVA: Weight Change (Standardized), GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
weigh_tables <- lapply(weigh_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
weigh_output <- paste(
  "<!DOCTYPE html><html><head><title>ANCOVA Results</title></head><body>",
  paste(weigh_tables, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(weigh_output, file = "results/GO_compare/weight_ancovas.html")
```
