---
title: "Eelgrass GO Fitness Comparisons"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# packages
library(car)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
library(knitr)
library(kableExtra)

# data
# exp fit
fit_full <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
surv <- read.csv("data/experiment/survival_dat_trt_pop.csv")[,-1]
grow <- read.csv("data/experiment/fit_growth_dat.csv")[,-1]

# go
lea <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_2025-12-03.csv")[,-1]
rda <- read.csv("results/GO_results/RDA_offsets_exptrt_wide_2026-01-02.csv")[,-1]
gf <- read.csv("results/GO_results/GF_offsets_exptrt_wide_2025-12-03.csv")[,-1]
```

## Genomic Offset - Fitness Offset Comparisons

This script compares genomic offset against fitness metrics in a "common garden" (across genotypes) and "current future" (across environments) comparison. 


### GO Data Processing

These data are already in a matrix format, but I will still make a long form combined df for plotting. 

```{r}
head(lea)
head(rda)
head(gf)

# long versions
# lea
lea_long <- lea %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

# rda
rda_long <- rda %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "rda_offset")

# gf
gf_long <- gf %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "gf_offset")

# put everything together
go1 <- merge(lea_long, rda_long, by = c("trt", "cc", "pop"))
go_wide <- merge(go1, gf_long[gf_long$mod == "geno", c("trt","pop","cc","gf_offset")], by = c("trt", "cc", "pop"))

# long version of all
go_long <- go_wide %>% pivot_longer(
  cols = c("lea_offset","rda_offset","gf_offset"),
  names_to = "method",
  values_to = "offset")
```

### Experimental Data

```{r}
# survival
head(surv)
dim(surv)
surv$trt <- gsub("_", "-", surv$trt)
fit_full$isAlive <- ifelse(fit_full$isAlive == TRUE, 1, 0)

# surv only fitness, all plants
surv_all <- fit_full[,c("pop", "trt", "tank_ID", "isAlive")]
surv_all$trt <- gsub("_", "-", surv_all$trt)
  
# also a format with survival in each tank/pop/trt
surv_tanks <- fit_full %>% 
  dplyr::select(pop, trt, tank_ID, isAlive) %>%
  group_by(pop, tank_ID, trt) %>%
  summarize(surv = mean(isAlive),
            surv_se = sd(isAlive)/length(isAlive))

# growth
head(grow)
grow_trim <- grow[,c("trt", "pop", "pop_trt", "tank_ID", "length_growth", "length_std_change", "weight_change", "weight_std_change", "above_below_weight_ratio")]
grow_trim$trt <- gsub("_","-",grow_trim$trt)

# some pop/trt level data
grow_pop_trt <- grow_trim %>% 
  group_by(pop, trt) %>% 
  summarize(avg_length_growth = mean(length_growth, na.rm = T),
    avg_length_std_change = mean(length_std_change, na.rm = T),
    avg_weight_change = mean(weight_change, na.rm = T),
    avg_weight_std_change = mean(weight_std_change, na.rm = T),
    avg_above_below_weight_ratio = mean(above_below_weight_ratio, na.rm = T))
```

### GO v Fitness

#### Data Merge

```{r}
# combine fitness and offset data
surv_go <- merge(go_wide, surv, by = c("pop", "trt"))
grow_go <- merge(go_wide, grow_trim, by = c("pop", "trt"))
surv_all_go <- merge(go_long, surv_all, by = c("pop", "trt"))

# add surv col to surv_all_go from isAlive for consistency in corr analysis
surv_all_go$surv <- surv_all_go$isAlive

# recode method
surv_all_go$method <- str_split_fixed(surv_all_go$method, "_", 2)[,1]

# numeric tank coding
tanks_df <- data.frame(tank_ID = levels(as.factor(surv_all_go$tank_ID)),
                       tank_num = 1:length(levels(as.factor(surv_all_go$tank_ID))))
surv_all_go <- merge(surv_all_go, tanks_df, by = "tank_ID")
surv_all_go$tank_num <- as.numeric(surv_all_go$tank_num)

# long versions
surv_go_long <- surv_go %>% pivot_longer(names_to = "method", cols = c("lea_offset", "rda_offset", "gf_offset"), values_to = "offset") %>% as.data.frame()
surv_go_long$method <- str_split_fixed(surv_go_long$method, "_", 2)[,1]
grow_go_long <- grow_go %>% pivot_longer(names_to = "method", cols = c("lea_offset", "rda_offset", "gf_offset"), values_to = "offset") %>% as.data.frame()
grow_go_long$method <- str_split_fixed(grow_go_long$method, "_", 2)[,1]
# surv all is already long format

# for correlation/signif testing 
trts <- levels(as.factor(surv_go$trt))
pops <- levels(as.factor(surv_go$pop))

# save combined dfs
#write.csv(surv_go, "results/GO_compare/survival_offset_avgs.csv", row.names = F)
#write.csv(surv_all_go, "results/GO_compare/survival_offset_binary.csv", row.names = F)
#write.csv(grow_go, "results/GO_compare/grow_offset.csv", row.names = F)
```

#### Correlations

```{r}
## SURVIVAL
###########
# function for correlations both common garden/current future offset
surv_corr <- function(df, method, gen_clone, fit_offset) {
  # subset data
  df_subset <- df[df$method == method & df$cc == gen_clone,]
  
  if(fit_offset == "cg") {
    # make df
    corrs_cg <- data.frame(method = NA, trt = NA, cc = NA, kendall_surv = NA, pval_surv = NA)
    
    # common garden
    for (i in 1:length(trts)){
      # subset
      trt_temp <- trts[i]
      temp <- df_subset[df_subset$trt == trt_temp,]
    
      # metadata
      corrs_cg[i,]$method <- method
      corrs_cg[i,]$cc <- gen_clone
      corrs_cg[i,]$trt <- trt_temp  
      
      # correlation
      test <- cor.test(temp$surv, temp$offset, method = "kendall")
      corrs_cg[i,]$kendall_surv <- test$estimate
      corrs_cg[i,]$pval_surv <- test$p.value
    }
    return(corrs_cg)
  } else if (fit_offset == "cf") {
    # make df
    corrs_cf <- data.frame(method = NA, pop = NA, cc = NA, kendall_surv = NA, pval_surv = NA)
    
    # current future
    for (i in 1:length(pops)){
      # subset
      pop_temp <- pops[i]
      temp <- df_subset[df_subset$pop == pop_temp,]
    
      # metadata
      corrs_cf[i,]$method <- method
      corrs_cf[i,]$cc <- gen_clone
      corrs_cf[i,]$pop <- pop_temp  
      
      # correlation
      test <- cor.test(temp$surv, temp$offset, method = "kendall")
      corrs_cf[i,]$kendall_surv <- test$estimate
      corrs_cf[i,]$pval_surv <- test$p.value
    }
    return(corrs_cf)
  } else {
    print("Please choose cg for common garden or cf for current-future")
  }
}

surv_corr_cg <- rbind(surv_corr(surv_go_long, "lea", "full", "cg"),
                      surv_corr(surv_go_long, "rda", "full", "cg"),
                      surv_corr(surv_go_long, "gf", "full", "cg"),
                      surv_corr(surv_go_long, "lea", "mll", "cg"),
                      surv_corr(surv_go_long, "rda", "mll", "cg"),
                      surv_corr(surv_go_long, "gf", "mll", "cg"))

surv_corr_all_cg <- rbind(surv_corr(surv_all_go, "lea", "full", "cg"),
                      surv_corr(surv_all_go, "rda", "full", "cg"),
                      surv_corr(surv_all_go, "gf", "full", "cg"),
                      surv_corr(surv_all_go, "lea", "mll", "cg"),
                      surv_corr(surv_all_go, "rda", "mll", "cg"),
                      surv_corr(surv_all_go, "gf", "mll", "cg"))

surv_corr_cf <- rbind(surv_corr(surv_go_long, "lea", "full", "cf"),
                      surv_corr(surv_go_long, "rda", "full", "cf"),
                      surv_corr(surv_go_long, "gf", "full", "cf"),
                      surv_corr(surv_go_long, "lea", "mll", "cf"),
                      surv_corr(surv_go_long, "rda", "mll", "cf"),
                      surv_corr(surv_go_long, "gf", "mll", "cf"))

surv_corr_all_cf <- rbind(surv_corr(surv_all_go, "lea", "full", "cf"),
                      surv_corr(surv_all_go, "rda", "full", "cf"),
                      surv_corr(surv_all_go, "gf", "full", "cf"),
                      surv_corr(surv_all_go, "lea", "mll", "cf"),
                      surv_corr(surv_all_go, "rda", "mll", "cf"),
                      surv_corr(surv_all_go, "gf", "mll", "cf"))

# add info about surv
surv_corr_cg$type_surv <- rep("proportion", nrow(surv_corr_cg))
surv_corr_cf$type_surv <- rep("proportion", nrow(surv_corr_cf))
surv_corr_all_cg$type_surv <- rep("binary", nrow(surv_corr_all_cg))
surv_corr_all_cf$type_surv <- rep("binary", nrow(surv_corr_all_cf))

# combine into single df
surv_corrs_cf <- rbind(surv_corr_cf, surv_corr_all_cf)
surv_corrs_cg <- rbind(surv_corr_cg, surv_corr_all_cg)

# signif T/F
surv_corrs_cf$signif <- ifelse(surv_corrs_cf$pval_surv < 0.05, T, F)
surv_corrs_cg$signif <- ifelse(surv_corrs_cg$pval_surv < 0.05, T, F)
###########

## GROWTH
#########
grow_corr <- function(df, method, gen_clone, fit_offset) {
  # subset data
  df_subset <- df[df$method == method & df$cc == gen_clone,]
  
  if(fit_offset == "cg") {
    # make df
    corrs_cg <- data.frame(method = NA, trt = NA, cc = NA, kendall_len = NA, pval_len = NA, kendall_weigh = NA, pval_weigh = NA)
    
    # common garden
    for (i in 1:length(trts)){
      # subset
      trt_temp <- trts[i]
      temp <- df_subset[df_subset$trt == trt_temp,]
    
      # metadata
      corrs_cg[i,]$method <- method
      corrs_cg[i,]$cc <- gen_clone
      corrs_cg[i,]$trt <- trt_temp  
      
      # correlation
      test_len <- cor.test(temp$length_std_change, temp$offset, method = "kendall")
      corrs_cg[i,]$kendall_len <- test_len$estimate
      corrs_cg[i,]$pval_len <- test_len$p.value
      
      test_weigh <- cor.test(temp$weight_std_change, temp$offset, method = "kendall")
      corrs_cg[i,]$kendall_weigh <- test_weigh$estimate
      corrs_cg[i,]$pval_weigh <- test_weigh$p.value
    }
    return(corrs_cg)
    
  } else if (fit_offset == "cf") {
    # make df
    corrs_cf <- data.frame(method = NA, pop = NA, cc = NA, kendall_len = NA, pval_len = NA, kendall_weigh = NA, pval_weigh = NA)
    
    # current future
    for (i in 1:length(pops)){
      # subset
      pop_temp <- pops[i]
      temp <- df_subset[df_subset$pop == pop_temp,]
    
      # metadata
      corrs_cf[i,]$method <- method
      corrs_cf[i,]$cc <- gen_clone
      corrs_cf[i,]$pop <- pop_temp  
      
      # correlation
      test_len <- cor.test(temp$length_std_change, temp$offset, method = "kendall")
      corrs_cf[i,]$kendall_len <- test_len$estimate
      corrs_cf[i,]$pval_len <- test_len$p.value
      
      test_weigh <- cor.test(temp$weight_std_change, temp$offset, method = "kendall")
      corrs_cf[i,]$kendall_weigh <- test_weigh$estimate
      corrs_cf[i,]$pval_weigh <- test_weigh$p.value
    }
    return(corrs_cf)
    
  } else {
    print("Please choose cg for common garden or cf for current-future")
  }
}

grow_corr_cg <- rbind(grow_corr(grow_go_long, "lea", "full", "cg"),
                      grow_corr(grow_go_long, "rda", "full", "cg"),
                      grow_corr(grow_go_long, "gf", "full", "cg"),
                      grow_corr(grow_go_long, "lea", "mll", "cg"),
                      grow_corr(grow_go_long, "rda", "mll", "cg"),
                      grow_corr(grow_go_long, "gf", "mll", "cg"))

grow_corr_cf <- rbind(grow_corr(grow_go_long, "lea", "full", "cf"),
                      grow_corr(grow_go_long, "rda", "full", "cf"),
                      grow_corr(grow_go_long, "gf", "full", "cf"),
                      grow_corr(grow_go_long, "lea", "mll", "cf"),
                      grow_corr(grow_go_long, "rda", "mll", "cf"),
                      grow_corr(grow_go_long, "gf", "mll", "cf"))

# signif T/F
grow_corr_cf$signif_len <- ifelse(grow_corr_cf$pval_len < 0.05, T, F)
grow_corr_cg$signif_len <- ifelse(grow_corr_cg$pval_len < 0.05, T, F)
grow_corr_cf$signif_weigh <- ifelse(grow_corr_cf$pval_weigh < 0.05, T, F)
grow_corr_cg$signif_weigh <- ifelse(grow_corr_cg$pval_weigh < 0.05, T, F)
#########

# save outputs
##############
#write.csv(surv_corrs_cf, "results/GO_compare/surv_corrs_cf.csv", row.names = F)
#write.csv(surv_corrs_cg, "results/GO_compare/surv_corrs_cg.csv", row.names = F)
#write.csv(grow_corr_cf, "results/GO_compare/grow_corr_cf.csv", row.names = F)
#write.csv(grow_corr_cg, "results/GO_compare/grow_corr_cg.csv", row.names = F)
##############
```

#### Modeling

##### GO only

Survival 

```{r}
# survival
##########
# quick checks + fixes
hist(surv_all_go$surv) # we will use binary outcomes first

## lea
# model full (no cc)
surv_mod_lea_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "lea",])
summary(surv_mod_lea_full) 
#plot(surv_mod_lea_full) # these look pretty bad
Anova(surv_mod_lea_full) # signif

# model mll (cc)
surv_mod_lea_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_mod_lea_mll) 
#plot(surv_mod_lea_mll) # these look pretty bad
Anova(surv_mod_lea_mll) # signif

# now the tank level pop/trt
surv_mod_lea_poptrt_full <- glm(surv ~ lea_offset,
               family = binomial(link = "logit"),
               data = surv_go)
summary(surv_mod_lea_poptrt_full) 
#plot(surv_mod_lea_full)
Anova(surv_mod_lea_poptrt_full) # signif

## rda
# model full (no cc)
surv_mod_rda_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_mod_rda_full) 
#plot(surv_mod_rda_full) # these look pretty bad
Anova(surv_mod_rda_full) # signif

# model mll (cc)
surv_mod_rda_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_mod_rda_mll) 
#plot(surv_mod_rda_mll) # these look pretty bad
Anova(surv_mod_rda_mll) # signif

# now the tank level pop/trt
surv_mod_rda_poptrt_full <- glm(surv ~ rda_offset,
               family = binomial(link = "logit"),
               data = surv_go[surv_go$cc == "full",])
summary(surv_mod_rda_poptrt_full) 
#plot(surv_mod_rda_full)
Anova(surv_mod_rda_poptrt_full) # signif

surv_mod_rda_poptrt_mll <- glm(surv ~ rda_offset,
               family = binomial(link = "logit"),
               data = surv_go[surv_go$cc == "mll",])
summary(surv_mod_rda_poptrt_mll) 
#plot(surv_mod_rda_poptrt_mll)
Anova(surv_mod_rda_poptrt_mll) # signif


## gf
# model full (no cc)
surv_mod_gf_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_mod_gf_full) 
#plot(surv_mod_gf_full) # these look pretty bad
Anova(surv_mod_gf_full) # signif

# model mll (cc)
surv_mod_gf_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_mod_gf_mll) 
#plot(surv_mod_gf_mll) # these look pretty bad
Anova(surv_mod_gf_mll) # signif

# now the tank level pop/trt
surv_mod_gf_poptrt_full <- glm(surv ~ gf_offset,
               family = binomial(link = "logit"),
               data = surv_go_full)
summary(surv_mod_gf_poptrt_full) 
#plot(surv_mod_gf_poptrt_full) # better than binary model
Anova(surv_mod_gf_poptrt_full) # signif

surv_mod_gf_poptrt_mll <- glm(surv ~ gf_offset,
               family = binomial(link = "logit"),
               data = surv_go_mll)
summary(surv_mod_gf_poptrt_mll) 
#plot(surv_mod_gf_poptrt_mll)
Anova(surv_mod_gf_poptrt_mll) # signif
##########
```

Length

```{r}
# length
########
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
# model full (no cc)
len_mod_lea_full <- lm(length_std_change ~ log(lea_offset), data = grow_go_full)
summary(len_mod_lea_full) 
#plot(len_mod_lea_full) # diverges pretty significantly on right tail
Anova(len_mod_lea_full) # signif

# model mll (cc)
len_mod_lea_mll <- lm(length_std_change ~ log(lea_offset), data = grow_go_mll)
summary(len_mod_lea_mll) 
#plot(len_mod_lea_mll) # diverges pretty significantly on right tail
Anova(len_mod_lea_mll) # signif

## rda
# model full (no cc)
len_mod_rda_full <- lm(length_std_change ~ log(rda_offset), data = grow_go_full)
summary(len_mod_rda_full) 
#plot(len_mod_rda_full) # diverges pretty significantly on right tail
Anova(len_mod_rda_full) # signif

# model mll (cc)
len_mod_rda_mll <- lm(length_std_change ~ log(rda_offset), data = grow_go_mll)
summary(len_mod_rda_mll) 
#plot(len_mod_rda_mll) # diverges pretty significantly on right tail
Anova(len_mod_rda_mll) # signif

## gf
# model full (no cc)
len_mod_gf_full <- lm(length_std_change ~ log(gf_offset), data = grow_go_full)
summary(len_mod_gf_full) 
#plot(len_mod_gf_full) # diverges pretty significantly on right tail
Anova(len_mod_gf_full) # signif

# model mll (cc)
len_mod_gf_mll <- lm(length_std_change ~ log(gf_offset), data = grow_go_mll)
summary(len_mod_gf_mll) 
#plot(len_mod_gf_mll) # diverges pretty significantly on right tail
Anova(len_mod_gf_mll) # signif
########
```

Growth

```{r}
# weight
########
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
# model full (no cc)
weigh_mod_lea_full <- lm(weight_std_change ~ lea_offset, data = grow_go_full)
summary(weigh_mod_lea_full) 
#plot(weigh_mod_lea_full) # diverges pretty significantly on right tail
Anova(weigh_mod_lea_full) # signif

# model mll (cc)
weigh_mod_lea_mll <- lm(weight_std_change ~ lea_offset, data = grow_go_mll)
summary(weigh_mod_lea_mll) 
#plot(weigh_mod_lea_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_lea_mll) # signif

## rda
# model full (no cc)
weigh_mod_rda_full <- lm(weight_std_change ~ (rda_offset), data = grow_go_full)
summary(weigh_mod_rda_full) 
#plot(weigh_mod_rda_full) # diverges pretty significantly on right tail
Anova(weigh_mod_rda_full) # signif

# model mll (cc)
weigh_mod_rda_mll <- lm(weight_std_change ~ (rda_offset), data = grow_go_mll)
summary(weigh_mod_rda_mll) 
#plot(weigh_mod_rda_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_rda_mll) # signif

## gf
# model full (no cc)
weigh_mod_gf_full <- lm(weight_std_change ~ (gf_offset), data = grow_go_full)
summary(weigh_mod_gf_full) 
#plot(weigh_mod_gf_full) # diverges pretty significantly on right tail
Anova(weigh_mod_gf_full) # signif

# model mll (cc)
weigh_mod_gf_mll <- lm(weight_std_change ~ (gf_offset), data = grow_go_mll)
summary(weigh_mod_gf_mll) 
#plot(weigh_mod_gf_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_gf_mll) # signif

# save these outputs
sink("results/fit_results/survival_modeling.txt")
"Survival\n"
"LEA MLL\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"LEA Full\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"RDA MLL\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"RDA Full\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"GF MLL\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

"GF Full\n"
summary(surv_mod_lea_mll)
Anova(surv_mod_lea_mll)

sink()
########
```

##### With population as fixed effect

Survival 

```{r}
# quick checks + fixes
hist(surv_all_go$surv) # we will use binary outcomes first

## lea
######
# model full (no cc)
surv_pop_lea_full <- lm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",])
summary(surv_pop_lea_full) 
#plot(surv_pop_lea_full)
Anova(surv_pop_lea_full) # signif

# model mll (cc)
surv_pop_lea_mll <- lm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_pop_lea_mll) 
#plot(surv_pop_lea_mll)
Anova(surv_pop_lea_mll) # signif

# model full (no cc) log
surv_pop_lea_full_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",],
               family = binomial)
summary(surv_pop_lea_full_log) 
#plot(surv_pop_lea_full_log)
Anova(surv_pop_lea_full_log) # signif

# model mll (cc) log
surv_pop_lea_mll_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",],
               family = binomial)
summary(surv_pop_lea_mll_log) 
#plot(surv_pop_lea_mll_log)
Anova(surv_pop_lea_mll_log) # signif
######

## rda
######
# model full (no cc)
surv_pop_rda_full <- lm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_pop_rda_full) 
#plot(surv_pop_rda_full)
Anova(surv_pop_rda_full) # signif

# model mll (cc)
surv_pop_rda_mll <- lm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_pop_rda_mll) 
#plot(surv_pop_rda_mll)
Anova(surv_pop_rda_mll) # signif

# model full (no cc) log
surv_pop_rda_full_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",],
               family = binomial)
summary(surv_pop_rda_full_log) 
#plot(surv_pop_rda_full_log)
Anova(surv_pop_rda_full_log) # signif

# model mll (cc) log
surv_pop_rda_mll_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",],
               family = binomial)
summary(surv_pop_rda_mll_log) 
#plot(surv_pop_rda_mll_log)
Anova(surv_pop_rda_mll_log) # signif
######

## gf
#####
# model full (no cc)
surv_pop_gf_full <- lm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_pop_gf_full) 
#plot(surv_pop_gf_full)
Anova(surv_pop_gf_full) # signif

# model mll (cc)
surv_pop_gf_mll <- lm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_pop_gf_mll) 
#plot(surv_pop_gf_mll)
Anova(surv_pop_gf_mll) # signif

# model full (no cc) log
surv_pop_gf_full_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",],
               family = binomial)
summary(surv_pop_gf_full_log) 
#plot(surv_pop_gf_full_log)
Anova(surv_pop_gf_full_log) # signif

# model mll (cc) log
surv_pop_gf_mll_log <- glm(isAlive ~ pop + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",],
               family = binomial)
summary(surv_pop_gf_mll_log) 
#plot(surv_pop_gf_mll_log)
Anova(surv_pop_gf_mll_log) # signif
#####
```

Length

```{r}
# length/pop
############
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
######
# model full (no cc)
len_pop_lea_full <- lm(log(length_std_change) ~ pop + lea_offset, data = grow_go_full)
summary(len_pop_lea_full) 
#plot(len_pop_lea_full) # diverges pretty significantly on left tail
Anova(len_pop_lea_full) # signif

# model mll (cc)
len_pop_lea_mll <- lm(log(length_std_change) ~ pop + lea_offset, data = grow_go_mll)
summary(len_pop_lea_mll) 
#plot(len_pop_lea_mll) # diverges pretty significantly on right tail
Anova(len_pop_lea_mll) # signif
######

## rda
######
# model full (no cc)
len_pop_rda_full <- lm(log(length_std_change) ~ pop + rda_offset, data = grow_go_full)
summary(len_pop_rda_full) 
#plot(len_pop_rda_full)
Anova(len_pop_rda_full) # rda offset not signif

# model mll (cc)
len_pop_rda_mll <- lm(log(length_std_change) ~ pop + rda_offset, data = grow_go_mll)
summary(len_pop_rda_mll) 
#plot(len_pop_rda_mll)
Anova(len_pop_rda_mll) # signif
######

## gf
#####
# model full (no cc)
len_pop_gf_full <- lm(log(length_std_change) ~ pop + gf_offset, data = grow_go_full)
summary(len_pop_gf_full) 
#plot(len_pop_gf_full)
Anova(len_pop_gf_full) # signif

# model mll (cc)
len_pop_gf_mll <- lm(log(length_std_change) ~ pop + gf_offset, data = grow_go_mll)
summary(len_pop_gf_mll) 
#plot(len_pop_gf_mll)
Anova(len_pop_gf_mll) # signif
#####
```

Weight

```{r}
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
######
# model full (no cc)
weigh_pop_lea_full <- lm(weight_std_change ~ pop + lea_offset, data = grow_go_full)
summary(weigh_pop_lea_full) 
#plot(weigh_pop_lea_full)
Anova(weigh_pop_lea_full) # signif

# model mll (cc)
weigh_pop_lea_mll <- lm(weight_std_change ~ pop + lea_offset, data = grow_go_mll)
summary(weigh_pop_lea_mll) 
#plot(weigh_pop_lea_mll)
Anova(weigh_pop_lea_mll) # signif
######

## rda
# model full (no cc)
weigh_pop_rda_full <- lm(weight_std_change ~ pop + rda_offset, data = grow_go_full)
summary(weigh_pop_rda_full) 
#plot(weigh_pop_rda_full) 
Anova(weigh_pop_rda_full) # signif

# model mll (cc)
weigh_pop_rda_mll <- lm(weight_std_change ~ pop + rda_offset, data = grow_go_mll)
summary(weigh_pop_rda_mll) 
#plot(weigh_pop_rda_mll) 
Anova(weigh_pop_rda_mll) # signif
######

## gf
#####
# model full (no cc)
weigh_pop_gf_full <- lm(weight_std_change ~ pop + gf_offset, data = grow_go_full)
summary(weigh_pop_gf_full) 
#plot(weigh_pop_gf_full) 
Anova(weigh_pop_gf_full) # signif

# model mll (cc)
weigh_pop_gf_mll <- lm(weight_std_change ~ pop + gf_offset, data = grow_go_mll)
summary(weigh_pop_gf_mll) 
#plot(weigh_pop_gf_mll)
Anova(weigh_pop_gf_mll) # signif
#####
```

##### With treatment as fixed effect

Survival 

```{r}
# quick checks + fixes
hist(surv_all_go$surv) # we will use binary outcomes first

## lea
######
# model full (no cc)
surv_trt_lea_full <- lm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",])
summary(surv_trt_lea_full) 
#plot(surv_trt_lea_full)
Anova(surv_trt_lea_full) # signif

# model mll (cc)
surv_trt_lea_mll <- lm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_trt_lea_mll) 
#plot(surv_trt_lea_mll)
Anova(surv_trt_lea_mll) # signif

# model full (no cc) log
surv_trt_lea_full_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "lea",],
               family = binomial)
summary(surv_trt_lea_full_log) 
#plot(surv_trt_lea_full_log)
Anova(surv_trt_lea_full_log) # signif

# model mll (cc) log
surv_trt_lea_mll_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",],
               family = binomial)
summary(surv_trt_lea_mll_log) 
#plot(surv_trt_lea_mll_log)
Anova(surv_trt_lea_mll_log) # signif
######

## rda
######
# model full (no cc)
surv_trt_rda_full <- lm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_trt_rda_full) 
#plot(surv_trt_rda_full)
Anova(surv_trt_rda_full) # signif trt, not offset

# model mll (cc)
surv_trt_rda_mll <- lm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_trt_rda_mll) 
#plot(surv_trt_rda_mll)
Anova(surv_trt_rda_mll) # signif trt, not offset

# model full (no cc) log
surv_trt_rda_full_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "rda",],
               family = binomial)
summary(surv_trt_rda_full_log) 
#plot(surv_trt_rda_full_log)
Anova(surv_trt_rda_full_log) # signif trt, not offset

# model mll (cc) log
surv_trt_rda_mll_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",],
               family = binomial)
summary(surv_trt_rda_mll_log) 
#plot(surv_trt_rda_mll_log)
Anova(surv_trt_rda_mll_log) # signif trt, not offset
######

## gf
#####
# model full (no cc)
surv_trt_gf_full <- lm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_trt_gf_full) 
#plot(surv_trt_gf_full)
Anova(surv_trt_gf_full) # signif

# model mll (cc)
surv_trt_gf_mll <- lm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_trt_gf_mll) 
#plot(surv_trt_gf_mll)
Anova(surv_trt_gf_mll) # signif

# model full (no cc)
surv_trt_gf_full_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_full[surv_all_go_full$method == "gf",],
               family = binomial)
summary(surv_trt_gf_full_log) 
#plot(surv_trt_gf_full_log)
Anova(surv_trt_gf_full_log) # signif

# model mll (cc)
surv_trt_gf_mll_log <- glm(isAlive ~ trt + offset,
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",],
               family = binomial)
summary(surv_trt_gf_mll_log) 
#plot(surv_trt_gf_mll_log)
Anova(surv_trt_gf_mll_log) # signif
#####
```

Length

```{r}
# length/trt
############
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
######
# model full (no cc)
len_trt_lea_full <- lm(log(length_std_change) ~ trt + lea_offset, data = grow_go_full)
summary(len_trt_lea_full) 
#plot(len_trt_lea_full) # diverges pretty significantly on left tail
Anova(len_trt_lea_full) # signif

# model mll (cc)
len_trt_lea_mll <- lm(log(length_std_change) ~ trt + lea_offset, data = grow_go_mll)
summary(len_trt_lea_mll) 
#plot(len_trt_lea_mll) # diverges pretty significantly on right tail
Anova(len_trt_lea_mll) # signif
######

## rda
######
# model full (no cc)
len_trt_rda_full <- lm(log(length_std_change) ~ trt + rda_offset, data = grow_go_full)
summary(len_trt_rda_full) 
#plot(len_trt_rda_full)
Anova(len_trt_rda_full) # rda offset not signif

# model mll (cc)
len_trt_rda_mll <- lm(log(length_std_change) ~ trt + rda_offset, data = grow_go_mll)
summary(len_trt_rda_mll) 
#plot(len_trt_rda_mll)
Anova(len_trt_rda_mll) # signif
######

## gf
#####
# model full (no cc)
len_trt_gf_full <- lm(log(length_std_change) ~ trt + gf_offset, data = grow_go_full)
summary(len_trt_gf_full) 
#plot(len_trt_gf_full)
Anova(len_trt_gf_full) # signif

# model mll (cc)
len_trt_gf_mll <- lm(log(length_std_change) ~ trt + gf_offset, data = grow_go_mll)
summary(len_trt_gf_mll) 
#plot(len_trt_gf_mll)
Anova(len_trt_gf_mll) # signif
#####
```

Weight

```{r}
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
######
# model full (no cc)
weigh_trt_lea_full <- lm(weight_std_change ~ trt + lea_offset, data = grow_go_full)
summary(weigh_trt_lea_full) 
#plot(weigh_trt_lea_full)
Anova(weigh_trt_lea_full) # signif

# model mll (cc)
weigh_trt_lea_mll <- lm(weight_std_change ~ trt + lea_offset, data = grow_go_mll)
summary(weigh_trt_lea_mll) 
#plot(weigh_trt_lea_mll)
Anova(weigh_trt_lea_mll) # signif
######

## rda
# model full (no cc)
weigh_trt_rda_full <- lm(weight_std_change ~ trt + rda_offset, data = grow_go_full)
summary(weigh_trt_rda_full) 
#plot(weigh_trt_rda_full) 
Anova(weigh_trt_rda_full) # signif

# model mll (cc)
weigh_trt_rda_mll <- lm(weight_std_change ~ trt + rda_offset, data = grow_go_mll)
summary(weigh_trt_rda_mll) 
#plot(weigh_trt_rda_mll) 
Anova(weigh_trt_rda_mll) # signif
######

## gf
#####
# model full (no cc)
weigh_trt_gf_full <- lm(weight_std_change ~ trt + gf_offset, data = grow_go_full)
summary(weigh_trt_gf_full) 
#plot(weigh_trt_gf_full) 
Anova(weigh_trt_gf_full) # signif

# model mll (cc)
weigh_trt_gf_mll <- lm(weight_std_change ~ trt + gf_offset, data = grow_go_mll)
summary(weigh_trt_gf_mll) 
#plot(weigh_trt_gf_mll)
Anova(weigh_trt_gf_mll) # signif
#####
```

##### Model Outputs

Survival

```{r}
# list of info for all survival ancova tables
surv_mod_info <- list(
  list(model = Anova(surv_pop_lea_full), title = "ANCOVA: Survival, LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_full), title = "ANCOVA: Survival, LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_full), title = "ANCOVA: Survival, RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_full), title = "ANCOVA: Survival, RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_full), title = "ANCOVA: Survival, GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_full), title = "ANCOVA: Survival, GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_lea_mll), title = "ANCOVA: Survival, LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_mll), title = "ANCOVA: Survival, LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_mll), title = "ANCOVA: Survival, RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_mll), title = "ANCOVA: Survival, RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_mll), title = "ANCOVA: Survival, GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_mll), title = "ANCOVA: Survival, GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
surv_tables <- lapply(surv_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
surv_output <- paste(
  "<!DOCTYPE html><html><head><title>ANCOVA Results</title></head><body>",
  paste(surv_tables, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(surv_output, file = "results/GO_compare/survival_ancovas.html")

# for logistic regression as well
surv_log_mod_info <- list(
  list(model = Anova(surv_pop_lea_full_log), title = "Log Regression: Survival, LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_full_log), title = "Log Regression: Survival, LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_full_log), title = "Log Regression: Survival, RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_full_log), title = "Log Regression: Survival, RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_full_log), title = "Log Regression: Survival, GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_full_log), title = "Log Regression: Survival, GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_lea_mll_log), title = "Log Regression: Survival, LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_lea_mll_log), title = "Log Regression: Survival, LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_rda_mll_log), title = "Log Regression: Survival, RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_rda_mll_log), title = "Log Regression: Survival, RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(surv_pop_gf_mll_log), title = "Log Regression: Survival, GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(surv_trt_gf_mll_log), title = "Log Regression: Survival, GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
surv_tables_log <- lapply(surv_log_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
surv_output_log <- paste(
  "<!DOCTYPE html><html><head><title>Log Regression Results</title></head><body>",
  paste(surv_tables_log, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(surv_output_log, file = "results/GO_compare/survival_log_regression.html")
```

Length

```{r}
# list of info for all survival ancova tables
len_mod_info <- list(
  list(model = Anova(len_pop_lea_full), title = "ANCOVA: Length Change (Standardized), LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_lea_full), title = "ANCOVA: Length Change (Standardized), LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_rda_full), title = "ANCOVA: Length Change (Standardized), RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_rda_full), title = "ANCOVA: Length Change (Standardized), RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_gf_full), title = "ANCOVA: Length Change (Standardized), GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_gf_full), title = "ANCOVA: Length Change (Standardized), GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_lea_mll), title = "ANCOVA: Length Change (Standardized), LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_lea_mll), title = "ANCOVA: Length Change (Standardized), LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_rda_mll), title = "ANCOVA: Length Change (Standardized), RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_rda_mll), title = "ANCOVA: Length Change (Standardized), RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(len_pop_gf_mll), title = "ANCOVA: Length Change (Standardized), GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(len_trt_gf_mll), title = "ANCOVA: Length Change (Standardized), GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
len_tables <- lapply(len_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
len_output <- paste(
  "<!DOCTYPE html><html><head><title>ANCOVA Results</title></head><body>",
  paste(len_tables, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(len_output, file = "results/GO_compare/length_ancovas.html")
```

Weight

```{r}
# list of info for all weighival ancova tables
weigh_mod_info <- list(
  list(model = Anova(weigh_pop_lea_full), title = "ANCOVA: Weight Change (Standardized), LEA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_lea_full), title = "ANCOVA: Weight Change (Standardized), LEA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_rda_full), title = "ANCOVA: Weight Change (Standardized), RDA, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_rda_full), title = "ANCOVA: Weight Change (Standardized), RDA, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_gf_full), title = "ANCOVA: Weight Change (Standardized), GF, Full, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_gf_full), title = "ANCOVA: Weight Change (Standardized), GF, Full, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_lea_mll), title = "ANCOVA: Weight Change (Standardized), LEA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_lea_mll), title = "ANCOVA: Weight Change (Standardized), LEA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_rda_mll), title = "ANCOVA: Weight Change (Standardized), RDA, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_rda_mll), title = "ANCOVA: Weight Change (Standardized), RDA, MLL, Treatment", labels = c("Treatment", "Offset")),
  list(model = Anova(weigh_pop_gf_mll), title = "ANCOVA: Weight Change (Standardized), GF, MLL, Population", labels = c("Population", "Offset")),
  list(model = Anova(weigh_trt_gf_mll), title = "ANCOVA: Weight Change (Standardized), GF, MLL, Treatment", labels = c("Treatment", "Offset")))

# format for html
weigh_tables <- lapply(weigh_mod_info, function(x) {
  df <- as.data.frame(x$model)
  rownames(df)[1:length(x$labels)] <- x$labels
  k_table <- kable(df, format = "html", caption = x$title, digits = 4) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "bordered"), 
                  full_width = FALSE, position = "left")
  return(as.character(k_table))
})

# combine + save
weigh_output <- paste(
  "<!DOCTYPE html><html><head><title>ANCOVA Results</title></head><body>",
  paste(weigh_tables, collapse = "<br><hr><br>"),
  "</body></html>"
)
cat(weigh_output, file = "results/GO_compare/weight_ancovas.html")
```
