---
title: "eelgrass_GOfit_cf.Rmd"
author: "Camille Rumberger"
date: '`r Sys.Date()`'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/camillerumberger/Documents/GitHub/BalticEelgrass")
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# packages
library(car)
library(lme4)
library(lmerTest)
library(tidyverse)
library(stringr)
library(emmeans)
library(multcomp)
library(multcompView)

# data
# exp fit
fit_full <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
# surv <- read.csv("data/experiment/survival_dat_trt_pop.csv")[,-1] # not using pop data
grow <- read.csv("data/experiment/fit_growth_dat.csv")[,-1]

# go
lea <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_2025-12-03.csv")[,-1]
# lea_cand <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_gea_2026-06-30.csv")[,-1]
rda <- read.csv("results/GO_results/RDA_offsets_exptrt_wide_2026-01-02.csv")[,-1]
gf <- read.csv("results/GO_results/GF_offsets_exptrt_wide_2025-12-03.csv")[,-1]
env_dist <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]

# individual
lea_ind <- read.csv("results/GO_results/LEA_offsets_exptrt_ind_long_2026-08-13.csv")[,-1]
rda_ind <- read.csv("results/GO_results/RDA_offsets_exptrt_ind_long_2026-08-13.csv")[,-1]
gf_ind <- read.csv("results/GO_results/GF_offsets_exptrt_ind_long_2026-08-13.csv")[,-1]
env_dist_ind <- read.csv("data/EnvDat/euclidian_trt_exp_ind.csv")[,-1]
```


## Genomic Offset-Fitness Comparisons

I will start by comparing genomic offset to fitness metrics, specifically survival. To do so, I will correlate survival and genomic offset using Kendall's Tau, and additionally model this relationship using a generalized linear modelling approach. I will use binary survival grouped by treatment and either population or coast.

### Pre-process data

#### Genomic Offsets

```{r}
## pop level
############
# lea
lea_long <- lea %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

# rda
rda_long <- rda %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "rda_offset")

# gf
gf_long <- gf %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "gf_offset")

# env dist
env_dist_trt <- env_dist %>% filter(pops1 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu") & !(pops2 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu"))) %>% dplyr::select(pops1, pops2, distances) %>% rename(pop = pops2, trt = pops1, env_dist = distances)

# put everything together
go1 <- merge(lea_long, rda_long, by = c("trt", "cc", "pop"))
go2 <- merge(go1, gf_long[gf_long$mod == "geno", c("trt","pop","cc","gf_offset")], by = c("trt", "cc", "pop"))
go_wide <- merge(go2, env_dist_trt, by = c("pop", "trt"))

# long version of all
go_long <- go_wide %>% pivot_longer(
  cols = c("lea_offset","rda_offset","gf_offset", "env_dist"),
  names_to = "method",
  values_to = "offset")
############

## ind level
############
# lea already in long format
lea_ind <- lea_ind %>% rename(bagnum = bagnum_text)

# rm extra info
lea_ind <- lea_ind_long %>% dplyr::select(pop, trt, bagnum, method, cc, offset)

# rda already in correct format
head(rda_ind)

# gradient forest
gf_ind$method <- "gf_offset"
gf_ind <- gf_ind[,c("pop", "trt", "bagnum", "method", "cc", "offset")]

# envdist
head(env_dist_ind)
env_dist_ind1 <- env_dist_ind %>% mutate(bagnum = as.numeric(bagnum_text), offset = dist, method = "env_dist", cc = "none") %>% dplyr::select(pop, trt, bagnum, method, cc, offset)

# put everything together
go_long_ind <- rbind(lea_ind, rda_ind, gf_ind, env_dist_ind1)
############
```

#### Fitness Data

I will use exclusively individual level survival and growth data for correlations, visualization, and analysis, as averaging data by population or tank gives very few data points to work with. 

```{r}
# pop/trt info
w_pops <- c("VAT", "VIK", "BAR", "HOG")
e_pops <- c("KUR", "KAL", "HOR", "BJO")
w_trts <- c("TempControl-21psu", "TempWarm-16psu")
e_trts <- c("TempControl-7psu", "TempWarm-5psu")
curr_trts <- c("TempControl-21psu", "TempControl-7psu")
fut_trts <- c("TempWarm-16psu", "TempWarm-5psu")

# add pop/trt metadata
fit_full$trt <- gsub("_", "-", fit_full$trt)
fit_full$coast_pop <- ifelse(fit_full$pop %in% w_pops, "west", "east")
fit_full$coast_trt <- ifelse(fit_full$trt %in% w_trts, "west", "east")
fit_full$curr_fut <- ifelse(fit_full$trt %in% curr_trts, "current", "future")

# surv
surv_all <- fit_full[,c("pop", "trt", "bagnum", "tank_ID", "isAlive", "coast_pop", "coast_trt", "curr_fut")]
surv_all$isAlive <- ifelse(surv_all$isAlive == TRUE, 1, 0)

# growth
grow_trim <- grow[,c("trt", "pop", "pop_trt", "bagnum", "tank_ID", "length_growth", "length_std_change", "weight_change", "weight_std_change", "above_below_weight_ratio")]
grow_trim$trt <- gsub("_", "-", grow_trim$trt)
grow_trim$coast_pop <- ifelse(grow_trim$pop %in% w_pops, "west", "east")
grow_trim$coast_trt <- ifelse(grow_trim$trt %in% w_trts, "west", "east")
grow_trim$curr_fut <- ifelse(grow_trim$trt %in% curr_trts, "current", "future")
```

#### Combine GOfit

```{r}
# isolate relevant trt/pop combos
go_clim_w <- go_long %>% filter(pop %in% w_pops & trt %in% w_trts)
go_clim_e <- go_long %>% filter(pop %in% e_pops & trt %in% e_trts)
go_clim <- rbind(go_clim_w, go_clim_e)

# and individual level data - just lea for now
go_clim_ind_w <- go_long_ind %>% filter(pop %in% w_pops & trt %in% w_trts)
go_clim_ind_e <- go_long_ind %>% filter(pop %in% e_pops & trt %in% e_trts)
go_clim_ind <- rbind(go_clim_ind_w, go_clim_ind_e)

# combine with survival and growth data
go_surv <- merge(go_clim, surv_all, by = c("pop", "trt"))
go_grow <- merge(go_clim, grow_trim, by = c("pop", "trt"))

# also combine individual level gos/fitness
go_surv_ind <- merge(go_clim_ind, surv_all, by = c("bagnum", "pop", "trt"))
go_grow_ind <- merge(go_clim_ind, grow_trim, by = c("bagnum", "pop", "trt"))
```


### Correlations

#### Functions

```{r}
# function for correlation testing
gofit_corr <- function(df, go, clone, fit_met) {
  # ensure proper naming of offsets and metrics
  if (!(fit_met %in% c("surv", "len", "weigh"))) {
    print("please pick from one of the following fitness metrics: surv, len, weigh")
    stop()
  }
  
  if (!(go %in% c("lea_offset", "rda_offset", "gf_offset", "env_dist"))) {
    print("please pick from one of the following offset metrics: lea_offset, rda_offset, gf_offset, env_dist")
    stop()
  }
  
  # subset data
  df_subset <- df[df$method == go & df$cc == clone,]

  # set up df
  corrs_cf <- data.frame(metric = NA, method = NA, cc = NA, corr = NA, pval = NA)
  
  # metadata
  corrs_cf$metric <- fit_met
  corrs_cf$method <- go
  corrs_cf$cc <- clone
  
  # testing
  if (fit_met == "surv") {
      test_cf = cor.test(df_subset$isAlive, df_subset$offset, method = "kendall", na.action = "na.exclude")
    corrs_cf$corr = test_cf$estimate
    corrs_cf$pval = test_cf$p.value
  } else if (fit_met == "len") {
    test_cf = cor.test(df_subset$length_std_change, df_subset$offset, method = "kendall", na.action = "na.exclude")
    corrs_cf$corr = test_cf$estimate
    corrs_cf$pval = test_cf$p.value
  } else {
    test_cf = cor.test(df_subset$weight_std_change, df_subset$offset, method = "kendall", na.action = "na.exclude")
    corrs_cf$corr = test_cf$estimate
    corrs_cf$pval = test_cf$p.value
  }
  return(corrs_cf)
}
```

#### Calculating Correlations - Populations

Overall

```{r}
## full
#######
# survival
surv_corr_lea_full <- gofit_corr(df = go_surv, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_full <- gofit_corr(df = go_surv, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_full <- gofit_corr(df = go_surv, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_full <- gofit_corr(df = go_surv, go = "env_dist", clone = "full", fit_met = "surv")
surv_corr_full <- rbind(surv_corr_lea_full, surv_corr_rda_full, surv_corr_gf_full, surv_corr_envdist_full)

# length
len_corr_lea_full <- gofit_corr(df = go_grow, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_full <- gofit_corr(df = go_grow, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_full <- gofit_corr(df = go_grow, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_full <- gofit_corr(df = go_grow, go = "env_dist", clone = "full", fit_met = "len")
len_corr_full <- rbind(len_corr_lea_full, len_corr_rda_full, len_corr_gf_full, len_corr_envdist_full)

# weight
weigh_corr_lea_full <- gofit_corr(df = go_grow, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_full <- gofit_corr(df = go_grow, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_full <- gofit_corr(df = go_grow, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_full <- gofit_corr(df = go_grow, go = "env_dist", clone = "full", fit_met = "weigh")
weigh_corr_full <- rbind(weigh_corr_lea_full, weigh_corr_rda_full, weigh_corr_gf_full, weigh_corr_envdist_full)
#######

## mll
######
# survival
surv_corr_lea_mll <- gofit_corr(df = go_surv, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_mll <- gofit_corr(df = go_surv, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_mll <- gofit_corr(df = go_surv, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_mll <- gofit_corr(df = go_surv, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_mll <- rbind(surv_corr_lea_mll, surv_corr_rda_mll, surv_corr_gf_mll, surv_corr_envdist_mll)

# length
len_corr_lea_mll <- gofit_corr(df = go_grow, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_mll <- gofit_corr(df = go_grow, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_mll <- gofit_corr(df = go_grow, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_mll <- gofit_corr(df = go_grow, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_mll <- rbind(len_corr_lea_mll, len_corr_rda_mll, len_corr_gf_mll, len_corr_envdist_mll)

# weight
weigh_corr_lea_mll <- gofit_corr(df = go_grow, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_mll <- gofit_corr(df = go_grow, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_mll <- gofit_corr(df = go_grow, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_mll <- gofit_corr(df = go_grow, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_mll <- rbind(weigh_corr_lea_mll, weigh_corr_rda_mll, weigh_corr_gf_mll, weigh_corr_envdist_mll)
######

# put together
fit_corr <- rbind(surv_corr_full, len_corr_full, weigh_corr_full, surv_corr_mll, len_corr_mll, weigh_corr_mll)
```

By coast

```{r}
## full
## east coast
#############
# subset to east coast
go_surv_e <- go_surv[go_surv$coast_pop == "east",]
go_grow_e <- go_grow[go_grow$coast_pop == "east",]

# survival
surv_corr_lea_e_full <- gofit_corr(df = go_surv_e, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_e_full <- gofit_corr(df = go_surv_e, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_e_full <- gofit_corr(df = go_surv_e, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_e_full <- gofit_corr(df = go_surv_e, go = "env_dist", clone = "full", fit_met = "surv")
surv_corr_e_full <- rbind(surv_corr_lea_e_full, surv_corr_rda_e_full, surv_corr_gf_e_full, surv_corr_envdist_e_full)

# length
len_corr_lea_e_full <- gofit_corr(df = go_grow_e, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_e_full <- gofit_corr(df = go_grow_e, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_e_full <- gofit_corr(df = go_grow_e, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_e_full <- gofit_corr(df = go_grow_e, go = "env_dist", clone = "full", fit_met = "len")
len_corr_e_full <- rbind(len_corr_lea_e_full, len_corr_rda_e_full, len_corr_gf_e_full, len_corr_envdist_e_full)

# weight
weigh_corr_lea_e_full <- gofit_corr(df = go_grow_e, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_e_full <- gofit_corr(df = go_grow_e, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_e_full <- gofit_corr(df = go_grow_e, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_e_full <- gofit_corr(df = go_grow_e, go = "env_dist", clone = "full", fit_met = "weigh")
weigh_corr_e_full <- rbind(weigh_corr_lea_e_full, weigh_corr_rda_e_full, weigh_corr_gf_e_full, weigh_corr_envdist_e_full)

# put together
fit_corr_e_full <- rbind(surv_corr_e_full, len_corr_e_full, weigh_corr_e_full) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_w <- go_surv[go_surv$coast_pop == "west",]
go_grow_w <- go_grow[go_grow$coast_pop == "west",]

# survival
surv_corr_lea_w_full <- gofit_corr(df = go_surv_w, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_w_full <- gofit_corr(df = go_surv_w, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_w_full <- gofit_corr(df = go_surv_w, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_w_full <- gofit_corr(df = go_surv_w, go = "env_dist", clone = "full", fit_met = "surv")
surv_corr_w_full <- rbind(surv_corr_lea_w_full, surv_corr_rda_w_full, surv_corr_gf_w_full, surv_corr_envdist_w_full)

# length
len_corr_lea_w_full <- gofit_corr(df = go_grow_w, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_w_full <- gofit_corr(df = go_grow_w, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_w_full <- gofit_corr(df = go_grow_w, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_w_full <- gofit_corr(df = go_grow_w, go = "env_dist", clone = "full", fit_met = "len")
len_corr_w_full <- rbind(len_corr_lea_w_full, len_corr_rda_w_full, len_corr_gf_w_full, len_corr_envdist_w_full)

# weight
weigh_corr_lea_w_full <- gofit_corr(df = go_grow_w, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_w_full <- gofit_corr(df = go_grow_w, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_w_full <- gofit_corr(df = go_grow_w, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_w_full <- gofit_corr(df = go_grow_w, go = "env_dist", clone = "full", fit_met = "weigh")
weigh_corr_w_full <- rbind(weigh_corr_lea_w_full, weigh_corr_rda_w_full, weigh_corr_gf_w_full, weigh_corr_envdist_w_full)

# put together
fit_corr_w_full <- rbind(surv_corr_w_full, len_corr_w_full, weigh_corr_w_full) %>% mutate(coast = "west")
#############

## mll
## east coast
#############
# subset to east coast
go_surv_e <- go_surv[go_surv$coast_pop == "east",]
go_grow_e <- go_grow[go_grow$coast_pop == "east",]

# survival
surv_corr_lea_e_mll <- gofit_corr(df = go_surv_e, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_e_mll <- gofit_corr(df = go_surv_e, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_e_mll <- gofit_corr(df = go_surv_e, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_e_mll <- gofit_corr(df = go_surv_e, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_e_mll <- rbind(surv_corr_lea_e_mll, surv_corr_rda_e_mll, surv_corr_gf_e_mll, surv_corr_envdist_e_mll)

# length
len_corr_lea_e_mll <- gofit_corr(df = go_grow_e, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_e_mll <- gofit_corr(df = go_grow_e, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_e_mll <- gofit_corr(df = go_grow_e, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_e_mll <- gofit_corr(df = go_grow_e, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_e_mll <- rbind(len_corr_lea_e_mll, len_corr_rda_e_mll, len_corr_gf_e_mll, len_corr_envdist_e_mll)

# weight
weigh_corr_lea_e_mll <- gofit_corr(df = go_grow_e, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_e_mll <- gofit_corr(df = go_grow_e, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_e_mll <- gofit_corr(df = go_grow_e, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_e_mll <- gofit_corr(df = go_grow_e, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_e_mll <- rbind(weigh_corr_lea_e_mll, weigh_corr_rda_e_mll, weigh_corr_gf_e_mll, weigh_corr_envdist_e_mll)

# put together
fit_corr_e_mll <- rbind(surv_corr_e_mll, len_corr_e_mll, weigh_corr_e_mll) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_w <- go_surv[go_surv$coast_pop == "west",]
go_grow_w <- go_grow[go_grow$coast_pop == "west",]

# survival
surv_corr_lea_w_mll <- gofit_corr(df = go_surv_w, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_w_mll <- gofit_corr(df = go_surv_w, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_w_mll <- gofit_corr(df = go_surv_w, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_w_mll <- gofit_corr(df = go_surv_w, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_w_mll <- rbind(surv_corr_lea_w_mll, surv_corr_rda_w_mll, surv_corr_gf_w_mll, surv_corr_envdist_w_mll)

# length
len_corr_lea_w_mll <- gofit_corr(df = go_grow_w, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_w_mll <- gofit_corr(df = go_grow_w, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_w_mll <- gofit_corr(df = go_grow_w, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_w_mll <- gofit_corr(df = go_grow_w, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_w_mll <- rbind(len_corr_lea_w_mll, len_corr_rda_w_mll, len_corr_gf_w_mll, len_corr_envdist_w_mll)

# weight
weigh_corr_lea_w_mll <- gofit_corr(df = go_grow_w, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_w_mll <- gofit_corr(df = go_grow_w, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_w_mll <- gofit_corr(df = go_grow_w, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_w_mll <- gofit_corr(df = go_grow_w, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_w_mll <- rbind(weigh_corr_lea_w_mll, weigh_corr_rda_w_mll, weigh_corr_gf_w_mll, weigh_corr_envdist_w_mll)

# put together
fit_corr_w_mll <- rbind(surv_corr_w_mll, len_corr_w_mll, weigh_corr_w_mll) %>% mutate(coast = "west")
#############

# together
fit_corr_coast <- rbind(fit_corr_w_full, fit_corr_e_full, fit_corr_w_mll, fit_corr_e_mll)
```

Future outcomes overall

```{r}
## subset to future
go_surv_fut <- go_surv[go_surv$curr_fut == "future",]
go_grow_fut <- go_grow[go_grow$curr_fut == "future",]

## full
#######
# survival
surv_corr_lea_full_fut <- gofit_corr(df = go_surv_fut, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_full_fut <- gofit_corr(df = go_surv_fut, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_full_fut <- gofit_corr(df = go_surv_fut, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_full_fut <- gofit_corr(df = go_surv_fut, go = "env_dist", clone = "full", fit_met = "surv")
surv_corr_full_fut <- rbind(surv_corr_lea_full_fut, surv_corr_rda_full_fut, surv_corr_gf_full_fut, surv_corr_envdist_full_fut)

# length
len_corr_lea_full_fut <- gofit_corr(df = go_grow_fut, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_full_fut <- gofit_corr(df = go_grow_fut, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_full_fut <- gofit_corr(df = go_grow_fut, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_full_fut <- gofit_corr(df = go_grow_fut, go = "env_dist", clone = "full", fit_met = "len")
len_corr_full_fut <- rbind(len_corr_lea_full_fut, len_corr_rda_full_fut, len_corr_gf_full_fut, len_corr_envdist_full_fut)

# weight
weigh_corr_lea_full_fut <- gofit_corr(df = go_grow_fut, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_full_fut <- gofit_corr(df = go_grow_fut, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_full_fut <- gofit_corr(df = go_grow_fut, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_full_fut <- gofit_corr(df = go_grow_fut, go = "env_dist", clone = "full", fit_met = "weigh")
weigh_corr_full_fut <- rbind(weigh_corr_lea_full_fut, weigh_corr_rda_full_fut, weigh_corr_gf_full_fut, weigh_corr_envdist_full_fut)
#######

## mll
######
# survival
surv_corr_lea_mll_fut <- gofit_corr(df = go_surv_fut, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_mll_fut <- gofit_corr(df = go_surv_fut, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_mll_fut <- gofit_corr(df = go_surv_fut, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_mll_fut <- gofit_corr(df = go_surv_fut, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_mll_fut <- rbind(surv_corr_lea_mll_fut, surv_corr_rda_mll_fut, surv_corr_gf_mll_fut, surv_corr_envdist_mll_fut)

# length
len_corr_lea_mll_fut <- gofit_corr(df = go_grow_fut, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_mll_fut <- gofit_corr(df = go_grow_fut, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_mll_fut <- gofit_corr(df = go_grow_fut, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_mll_fut <- gofit_corr(df = go_grow_fut, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_mll_fut <- rbind(len_corr_lea_mll_fut, len_corr_rda_mll_fut, len_corr_gf_mll_fut, len_corr_envdist_mll_fut)

# weight
weigh_corr_lea_mll_fut <- gofit_corr(df = go_grow_fut, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_mll_fut <- gofit_corr(df = go_grow_fut, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_mll_fut <- gofit_corr(df = go_grow_fut, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_mll_fut <- gofit_corr(df = go_grow_fut, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_mll_fut <- rbind(weigh_corr_lea_mll_fut, weigh_corr_rda_mll_fut, weigh_corr_gf_mll_fut, weigh_corr_envdist_mll_fut)
######

# put together
fit_corr_fut <- rbind(surv_corr_full_fut, len_corr_full_fut, weigh_corr_full_fut, surv_corr_mll_fut, len_corr_mll_fut, weigh_corr_mll_fut)
```

Future outcomes by coast (mll only)

```{r}
## east coast
#############
# subset to east coast
go_surv_e_fut <- go_surv[go_surv$coast_pop == "east" & go_surv$curr_fut == "future",]
go_grow_e_fut <- go_grow[go_grow$coast_pop == "east" & go_surv$curr_fut == "future",]

# survival
surv_corr_lea_e_fut <- gofit_corr(df = go_surv_e_fut, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_e_fut <- gofit_corr(df = go_surv_e_fut, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_e_fut <- gofit_corr(df = go_surv_e_fut, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_e_fut <- gofit_corr(df = go_surv_e_fut, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_e_fut <- rbind(surv_corr_lea_e_fut, surv_corr_rda_e_fut, surv_corr_gf_e_fut, surv_corr_envdist_e_fut)

# length
len_corr_lea_e_fut <- gofit_corr(df = go_grow_e_fut, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_e_fut <- gofit_corr(df = go_grow_e_fut, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_e_fut <- gofit_corr(df = go_grow_e_fut, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_e_fut <- gofit_corr(df = go_grow_e_fut, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_e_fut <- rbind(len_corr_lea_e_fut, len_corr_rda_e_fut, len_corr_gf_e_fut, len_corr_envdist_e_fut)

# weight
weigh_corr_lea_e_fut <- gofit_corr(df = go_grow_e_fut, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_e_fut <- gofit_corr(df = go_grow_e_fut, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_e_fut <- gofit_corr(df = go_grow_e_fut, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_e_fut <- gofit_corr(df = go_grow_e_fut, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_e_fut <- rbind(weigh_corr_lea_e_fut, weigh_corr_rda_e_fut, weigh_corr_gf_e_fut, weigh_corr_envdist_e_fut)

# put together
fit_corr_e_fut <- rbind(surv_corr_e_fut, len_corr_e_fut, weigh_corr_e_fut) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_w_fut <- go_surv[go_surv$coast_pop == "west" & go_surv$curr_fut == "future",]
go_grow_w_fut <- go_grow[go_grow$coast_pop == "west" & go_surv$curr_fut == "future",]

# survival
surv_corr_lea_w_fut <- gofit_corr(df = go_surv_w_fut, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_w_fut <- gofit_corr(df = go_surv_w_fut, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_w_fut <- gofit_corr(df = go_surv_w_fut, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_w_fut <- gofit_corr(df = go_surv_w_fut, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_w_fut <- rbind(surv_corr_lea_w_fut, surv_corr_rda_w_fut, surv_corr_gf_w_fut, surv_corr_envdist_w_fut)

# length
len_corr_lea_w_fut <- gofit_corr(df = go_grow_w_fut, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_w_fut <- gofit_corr(df = go_grow_w_fut, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_w_fut <- gofit_corr(df = go_grow_w_fut, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_w_fut <- gofit_corr(df = go_grow_w_fut, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_w_fut <- rbind(len_corr_lea_w_fut, len_corr_rda_w_fut, len_corr_gf_w_fut, len_corr_envdist_w_fut)

# weight
weigh_corr_lea_w_fut <- gofit_corr(df = go_grow_w_fut, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_w_fut <- gofit_corr(df = go_grow_w_fut, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_w_fut <- gofit_corr(df = go_grow_w_fut, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_w_fut <- gofit_corr(df = go_grow_w_fut, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_w_fut <- rbind(weigh_corr_lea_w_fut, weigh_corr_rda_w_fut, weigh_corr_gf_w_fut, weigh_corr_envdist_w_fut)

# put together
fit_corr_w_fut <- rbind(surv_corr_w_fut, len_corr_w_fut, weigh_corr_w_fut) %>% mutate(coast = "west")
#############

# put together
fit_corr_coast_fut <- rbind(fit_corr_w_fut, fit_corr_e_fut)
```

Current outcomes overall

```{r}
## subset to current
go_surv_curr <- go_surv[go_surv$curr_fut == "current",]
go_grow_curr <- go_grow[go_grow$curr_fut == "current",]

## full
#######
# survival
surv_corr_lea_full_curr <- gofit_corr(df = go_surv_curr, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_full_curr <- gofit_corr(df = go_surv_curr, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_full_curr <- gofit_corr(df = go_surv_curr, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_full_curr <- gofit_corr(df = go_surv_curr, go = "env_dist", clone = "full", fit_met = "surv")
surv_corr_full_curr <- rbind(surv_corr_lea_full_curr, surv_corr_rda_full_curr, surv_corr_gf_full_curr, surv_corr_envdist_full_curr)

# length
len_corr_lea_full_curr <- gofit_corr(df = go_grow_curr, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_full_curr <- gofit_corr(df = go_grow_curr, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_full_curr <- gofit_corr(df = go_grow_curr, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_full_curr <- gofit_corr(df = go_grow_curr, go = "env_dist", clone = "full", fit_met = "len")
len_corr_full_curr <- rbind(len_corr_lea_full_curr, len_corr_rda_full_curr, len_corr_gf_full_curr, len_corr_envdist_full_curr)

# weight
weigh_corr_lea_full_curr <- gofit_corr(df = go_grow_curr, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_full_curr <- gofit_corr(df = go_grow_curr, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_full_curr <- gofit_corr(df = go_grow_curr, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_full_curr <- gofit_corr(df = go_grow_curr, go = "env_dist", clone = "full", fit_met = "weigh")
weigh_corr_full_curr <- rbind(weigh_corr_lea_full_curr, weigh_corr_rda_full_curr, weigh_corr_gf_full_curr, weigh_corr_envdist_full_curr)
#######

## mll
######
# survival
surv_corr_lea_mll_curr <- gofit_corr(df = go_surv_curr, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_mll_curr <- gofit_corr(df = go_surv_curr, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_mll_curr <- gofit_corr(df = go_surv_curr, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_mll_curr <- gofit_corr(df = go_surv_curr, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_mll_curr <- rbind(surv_corr_lea_mll_curr, surv_corr_rda_mll_curr, surv_corr_gf_mll_curr, surv_corr_envdist_mll_curr)

# length
len_corr_lea_mll_curr <- gofit_corr(df = go_grow_curr, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_mll_curr <- gofit_corr(df = go_grow_curr, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_mll_curr <- gofit_corr(df = go_grow_curr, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_mll_curr <- gofit_corr(df = go_grow_curr, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_mll_curr <- rbind(len_corr_lea_mll_curr, len_corr_rda_mll_curr, len_corr_gf_mll_curr, len_corr_envdist_mll_curr)

# weight
weigh_corr_lea_mll_curr <- gofit_corr(df = go_grow_curr, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_mll_curr <- gofit_corr(df = go_grow_curr, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_mll_curr <- gofit_corr(df = go_grow_curr, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_mll_curr <- gofit_corr(df = go_grow_curr, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_mll_curr <- rbind(weigh_corr_lea_mll_curr, weigh_corr_rda_mll_curr, weigh_corr_gf_mll_curr, weigh_corr_envdist_mll_curr)
######

# put together
fit_corr_curr <- rbind(surv_corr_full_curr, len_corr_full_curr, weigh_corr_full_curr, surv_corr_mll_curr, len_corr_mll_curr, weigh_corr_mll_curr)
```

Current outcomes by coast (mll only)

```{r}
## east coast
#############
# subset to east coast
go_surv_e_curr <- go_surv[go_surv$coast_pop == "east" & go_surv$curr_fut == "current",]
go_grow_e_curr <- go_grow[go_grow$coast_pop == "east" & go_surv$curr_fut == "current",]

# survival
surv_corr_lea_e_curr <- gofit_corr(df = go_surv_e_curr, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_e_curr <- gofit_corr(df = go_surv_e_curr, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_e_curr <- gofit_corr(df = go_surv_e_curr, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_e_curr <- gofit_corr(df = go_surv_e_curr, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_e_curr <- rbind(surv_corr_lea_e_curr, surv_corr_rda_e_curr, surv_corr_gf_e_curr, surv_corr_envdist_e_curr)

# length
len_corr_lea_e_curr <- gofit_corr(df = go_grow_e_curr, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_e_curr <- gofit_corr(df = go_grow_e_curr, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_e_curr <- gofit_corr(df = go_grow_e_curr, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_e_curr <- gofit_corr(df = go_grow_e_curr, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_e_curr <- rbind(len_corr_lea_e_curr, len_corr_rda_e_curr, len_corr_gf_e_curr, len_corr_envdist_e_curr)

# weight
weigh_corr_lea_e_curr <- gofit_corr(df = go_grow_e_curr, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_e_curr <- gofit_corr(df = go_grow_e_curr, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_e_curr <- gofit_corr(df = go_grow_e_curr, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_e_curr <- gofit_corr(df = go_grow_e_curr, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_e_curr <- rbind(weigh_corr_lea_e_curr, weigh_corr_rda_e_curr, weigh_corr_gf_e_curr, weigh_corr_envdist_e_curr)

# put together
fit_corr_e_curr <- rbind(surv_corr_e_curr, len_corr_e_curr, weigh_corr_e_curr) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_w_curr <- go_surv[go_surv$coast_pop == "west" & go_surv$curr_fut == "current",]
go_grow_w_curr <- go_grow[go_grow$coast_pop == "west" & go_surv$curr_fut == "current",]

# survival
surv_corr_lea_w_curr <- gofit_corr(df = go_surv_w_curr, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_w_curr <- gofit_corr(df = go_surv_w_curr, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_w_curr <- gofit_corr(df = go_surv_w_curr, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_w_curr <- gofit_corr(df = go_surv_w_curr, go = "env_dist", clone = "mll", fit_met = "surv")
surv_corr_w_curr <- rbind(surv_corr_lea_w_curr, surv_corr_rda_w_curr, surv_corr_gf_w_curr, surv_corr_envdist_w_curr)

# length
len_corr_lea_w_curr <- gofit_corr(df = go_grow_w_curr, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_w_curr <- gofit_corr(df = go_grow_w_curr, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_w_curr <- gofit_corr(df = go_grow_w_curr, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_w_curr <- gofit_corr(df = go_grow_w_curr, go = "env_dist", clone = "mll", fit_met = "len")
len_corr_w_curr <- rbind(len_corr_lea_w_curr, len_corr_rda_w_curr, len_corr_gf_w_curr, len_corr_envdist_w_curr)

# weight
weigh_corr_lea_w_curr <- gofit_corr(df = go_grow_w_curr, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_w_curr <- gofit_corr(df = go_grow_w_curr, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_w_curr <- gofit_corr(df = go_grow_w_curr, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_w_curr <- gofit_corr(df = go_grow_w_curr, go = "env_dist", clone = "mll", fit_met = "weigh")
weigh_corr_w_curr <- rbind(weigh_corr_lea_w_curr, weigh_corr_rda_w_curr, weigh_corr_gf_w_curr, weigh_corr_envdist_w_curr)

# put together
fit_corr_w_curr <- rbind(surv_corr_w_curr, len_corr_w_curr, weigh_corr_w_curr) %>% mutate(coast = "west")
#############

# put together
fit_corr_coast_curr <- rbind(fit_corr_w_curr, fit_corr_e_curr)
```


#### Calculating Correlations - Individuals

Overall

```{r}
## full
#######
# survival
surv_corr_lea_full <- gofit_corr(df = go_surv_ind, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_full <- gofit_corr(df = go_surv_ind, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_full <- gofit_corr(df = go_surv_ind, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_full <- gofit_corr(df = go_surv_ind, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_full <- rbind(surv_corr_lea_full, surv_corr_rda_full, surv_corr_gf_full, surv_corr_envdist_full)

# length
len_corr_lea_full <- gofit_corr(df = go_grow_ind, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_full <- gofit_corr(df = go_grow_ind, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_full <- gofit_corr(df = go_grow_ind, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_full <- gofit_corr(df = go_grow_ind, go = "env_dist", clone = "none", fit_met = "len")
len_corr_full <- rbind(len_corr_lea_full, len_corr_rda_full, len_corr_gf_full, len_corr_envdist_full)

# weight
weigh_corr_lea_full <- gofit_corr(df = go_grow_ind, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_full <- gofit_corr(df = go_grow_ind, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_full <- gofit_corr(df = go_grow_ind, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_full <- gofit_corr(df = go_grow_ind, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_full <- rbind(weigh_corr_lea_full, weigh_corr_rda_full, weigh_corr_gf_full, weigh_corr_envdist_full)
#######

## mll
######
# survival
surv_corr_lea_mll <- gofit_corr(df = go_surv_ind, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_mll <- gofit_corr(df = go_surv_ind, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_mll <- gofit_corr(df = go_surv_ind, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_mll <- gofit_corr(df = go_surv_ind, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_mll <- rbind(surv_corr_lea_mll, surv_corr_rda_mll, surv_corr_gf_mll, surv_corr_envdist_mll)

# length
len_corr_lea_mll <- gofit_corr(df = go_grow_ind, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_mll <- gofit_corr(df = go_grow_ind, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_mll <- gofit_corr(df = go_grow_ind, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_mll <- gofit_corr(df = go_grow_ind, go = "env_dist", clone = "none", fit_met = "len")
len_corr_mll <- rbind(len_corr_lea_mll, len_corr_rda_mll, len_corr_gf_mll, len_corr_envdist_mll)

# weight
weigh_corr_lea_mll <- gofit_corr(df = go_grow_ind, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_mll <- gofit_corr(df = go_grow_ind, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_mll <- gofit_corr(df = go_grow_ind, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_mll <- gofit_corr(df = go_grow_ind, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_mll <- rbind(weigh_corr_lea_mll, weigh_corr_rda_mll, weigh_corr_gf_mll, weigh_corr_envdist_mll)
######

# put together
fit_corr <- rbind(surv_corr_full, len_corr_full, weigh_corr_full, surv_corr_mll, len_corr_mll, weigh_corr_mll)
```

By coast

```{r}
## full
## east coast
#############
# subset to east coast
go_surv_ind_e <- go_surv_ind[go_surv_ind$coast_pop == "east",]
go_grow_ind_e <- go_grow_ind[go_grow_ind$coast_pop == "east",]

# survival
surv_corr_lea_e_full <- gofit_corr(df = go_surv_ind_e, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_e_full <- gofit_corr(df = go_surv_ind_e, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_e_full <- gofit_corr(df = go_surv_ind_e, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_e_full <- gofit_corr(df = go_surv_ind_e, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_e_full <- rbind(surv_corr_lea_e_full, surv_corr_rda_e_full, surv_corr_gf_e_full, surv_corr_envdist_e_full)

# length
len_corr_lea_e_full <- gofit_corr(df = go_grow_ind_e, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_e_full <- gofit_corr(df = go_grow_ind_e, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_e_full <- gofit_corr(df = go_grow_ind_e, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_e_full <- gofit_corr(df = go_grow_ind_e, go = "env_dist", clone = "none", fit_met = "len")
len_corr_e_full <- rbind(len_corr_lea_e_full, len_corr_rda_e_full, len_corr_gf_e_full, len_corr_envdist_e_full)

# weight
weigh_corr_lea_e_full <- gofit_corr(df = go_grow_ind_e, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_e_full <- gofit_corr(df = go_grow_ind_e, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_e_full <- gofit_corr(df = go_grow_ind_e, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_e_full <- gofit_corr(df = go_grow_ind_e, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_e_full <- rbind(weigh_corr_lea_e_full, weigh_corr_rda_e_full, weigh_corr_gf_e_full, weigh_corr_envdist_e_full)

# put together
fit_corr_e_full <- rbind(surv_corr_e_full, len_corr_e_full, weigh_corr_e_full) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_ind_w <- go_surv_ind[go_surv_ind$coast_pop == "west",]
go_grow_ind_w <- go_grow_ind[go_grow_ind$coast_pop == "west",]

# survival
surv_corr_lea_w_full <- gofit_corr(df = go_surv_ind_w, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_w_full <- gofit_corr(df = go_surv_ind_w, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_w_full <- gofit_corr(df = go_surv_ind_w, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_w_full <- gofit_corr(df = go_surv_ind_w, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_w_full <- rbind(surv_corr_lea_w_full, surv_corr_rda_w_full, surv_corr_gf_w_full, surv_corr_envdist_w_full)

# length
len_corr_lea_w_full <- gofit_corr(df = go_grow_ind_w, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_w_full <- gofit_corr(df = go_grow_ind_w, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_w_full <- gofit_corr(df = go_grow_ind_w, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_w_full <- gofit_corr(df = go_grow_ind_w, go = "env_dist", clone = "none", fit_met = "len")
len_corr_w_full <- rbind(len_corr_lea_w_full, len_corr_rda_w_full, len_corr_gf_w_full, len_corr_envdist_w_full)

# weight
weigh_corr_lea_w_full <- gofit_corr(df = go_grow_ind_w, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_w_full <- gofit_corr(df = go_grow_ind_w, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_w_full <- gofit_corr(df = go_grow_ind_w, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_w_full <- gofit_corr(df = go_grow_ind_w, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_w_full <- rbind(weigh_corr_lea_w_full, weigh_corr_rda_w_full, weigh_corr_gf_w_full, weigh_corr_envdist_w_full)

# put together
fit_corr_w_full <- rbind(surv_corr_w_full, len_corr_w_full, weigh_corr_w_full) %>% mutate(coast = "west")
#############

## mll
## east coast
#############
# subset to east coast
go_surv_ind_e <- go_surv_ind[go_surv_ind$coast_pop == "east",]
go_grow_ind_e <- go_grow_ind[go_grow_ind$coast_pop == "east",]

# survival
surv_corr_lea_e_mll <- gofit_corr(df = go_surv_ind_e, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_e_mll <- gofit_corr(df = go_surv_ind_e, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_e_mll <- gofit_corr(df = go_surv_ind_e, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_e_mll <- gofit_corr(df = go_surv_ind_e, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_e_mll <- rbind(surv_corr_lea_e_mll, surv_corr_rda_e_mll, surv_corr_gf_e_mll, surv_corr_envdist_e_mll)

# length
len_corr_lea_e_mll <- gofit_corr(df = go_grow_ind_e, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_e_mll <- gofit_corr(df = go_grow_ind_e, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_e_mll <- gofit_corr(df = go_grow_ind_e, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_e_mll <- gofit_corr(df = go_grow_ind_e, go = "env_dist", clone = "none", fit_met = "len")
len_corr_e_mll <- rbind(len_corr_lea_e_mll, len_corr_rda_e_mll, len_corr_gf_e_mll, len_corr_envdist_e_mll)

# weight
weigh_corr_lea_e_mll <- gofit_corr(df = go_grow_ind_e, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_e_mll <- gofit_corr(df = go_grow_ind_e, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_e_mll <- gofit_corr(df = go_grow_ind_e, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_e_mll <- gofit_corr(df = go_grow_ind_e, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_e_mll <- rbind(weigh_corr_lea_e_mll, weigh_corr_rda_e_mll, weigh_corr_gf_e_mll, weigh_corr_envdist_e_mll)

# put together
fit_corr_e_mll <- rbind(surv_corr_e_mll, len_corr_e_mll, weigh_corr_e_mll) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_ind_w <- go_surv_ind[go_surv_ind$coast_pop == "west",]
go_grow_ind_w <- go_grow_ind[go_grow_ind$coast_pop == "west",]

# survival
surv_corr_lea_w_mll <- gofit_corr(df = go_surv_ind_w, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_w_mll <- gofit_corr(df = go_surv_ind_w, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_w_mll <- gofit_corr(df = go_surv_ind_w, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_w_mll <- gofit_corr(df = go_surv_ind_w, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_w_mll <- rbind(surv_corr_lea_w_mll, surv_corr_rda_w_mll, surv_corr_gf_w_mll, surv_corr_envdist_w_mll)

# length
len_corr_lea_w_mll <- gofit_corr(df = go_grow_ind_w, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_w_mll <- gofit_corr(df = go_grow_ind_w, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_w_mll <- gofit_corr(df = go_grow_ind_w, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_w_mll <- gofit_corr(df = go_grow_ind_w, go = "env_dist", clone = "none", fit_met = "len")
len_corr_w_mll <- rbind(len_corr_lea_w_mll, len_corr_rda_w_mll, len_corr_gf_w_mll, len_corr_envdist_w_mll)

# weight
weigh_corr_lea_w_mll <- gofit_corr(df = go_grow_ind_w, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_w_mll <- gofit_corr(df = go_grow_ind_w, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_w_mll <- gofit_corr(df = go_grow_ind_w, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_w_mll <- gofit_corr(df = go_grow_ind_w, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_w_mll <- rbind(weigh_corr_lea_w_mll, weigh_corr_rda_w_mll, weigh_corr_gf_w_mll, weigh_corr_envdist_w_mll)

# put together
fit_corr_w_mll <- rbind(surv_corr_w_mll, len_corr_w_mll, weigh_corr_w_mll) %>% mutate(coast = "west")
#############

# together
fit_corr_coast <- rbind(fit_corr_w_full, fit_corr_e_full, fit_corr_w_mll, fit_corr_e_mll)
```

Future outcomes overall

```{r}
## subset to future
go_surv_ind_fut <- go_surv_ind[go_surv_ind$curr_fut == "future",]
go_grow_ind_fut <- go_grow_ind[go_grow_ind$curr_fut == "future",]

## full
#######
# survival
surv_corr_lea_full_fut <- gofit_corr(df = go_surv_ind_fut, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_full_fut <- gofit_corr(df = go_surv_ind_fut, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_full_fut <- gofit_corr(df = go_surv_ind_fut, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_full_fut <- gofit_corr(df = go_surv_ind_fut, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_full_fut <- rbind(surv_corr_lea_full_fut, surv_corr_rda_full_fut, surv_corr_gf_full_fut, surv_corr_envdist_full_fut)

# length
len_corr_lea_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "env_dist", clone = "none", fit_met = "len")
len_corr_full_fut <- rbind(len_corr_lea_full_fut, len_corr_rda_full_fut, len_corr_gf_full_fut, len_corr_envdist_full_fut)

# weight
weigh_corr_lea_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_full_fut <- gofit_corr(df = go_grow_ind_fut, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_full_fut <- rbind(weigh_corr_lea_full_fut, weigh_corr_rda_full_fut, weigh_corr_gf_full_fut, weigh_corr_envdist_full_fut)
#######

## mll
######
# survival
surv_corr_lea_mll_fut <- gofit_corr(df = go_surv_ind_fut, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_mll_fut <- gofit_corr(df = go_surv_ind_fut, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_mll_fut <- gofit_corr(df = go_surv_ind_fut, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_mll_fut <- gofit_corr(df = go_surv_ind_fut, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_mll_fut <- rbind(surv_corr_lea_mll_fut, surv_corr_rda_mll_fut, surv_corr_gf_mll_fut, surv_corr_envdist_mll_fut)

# length
len_corr_lea_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "env_dist", clone = "none", fit_met = "len")
len_corr_mll_fut <- rbind(len_corr_lea_mll_fut, len_corr_rda_mll_fut, len_corr_gf_mll_fut, len_corr_envdist_mll_fut)

# weight
weigh_corr_lea_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_mll_fut <- gofit_corr(df = go_grow_ind_fut, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_mll_fut <- rbind(weigh_corr_lea_mll_fut, weigh_corr_rda_mll_fut, weigh_corr_gf_mll_fut, weigh_corr_envdist_mll_fut)
######

# put together
fit_corr_fut <- rbind(surv_corr_full_fut, len_corr_full_fut, weigh_corr_full_fut, surv_corr_mll_fut, len_corr_mll_fut, weigh_corr_mll_fut)
```

Future outcomes by coast (mll only)

```{r}
## east coast
#############
# subset to east coast
go_surv_ind_e_fut <- go_surv_ind[go_surv_ind$coast_pop == "east" & go_surv_ind$curr_fut == "future",]
go_grow_ind_e_fut <- go_grow_ind[go_grow_ind$coast_pop == "east" & go_surv_ind$curr_fut == "future",]

# survival
surv_corr_lea_e_fut <- gofit_corr(df = go_surv_ind_e_fut, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_e_fut <- gofit_corr(df = go_surv_ind_e_fut, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_e_fut <- gofit_corr(df = go_surv_ind_e_fut, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_e_fut <- gofit_corr(df = go_surv_ind_e_fut, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_e_fut <- rbind(surv_corr_lea_e_fut, surv_corr_rda_e_fut, surv_corr_gf_e_fut, surv_corr_envdist_e_fut)

# length
len_corr_lea_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "env_dist", clone = "none", fit_met = "len")
len_corr_e_fut <- rbind(len_corr_lea_e_fut, len_corr_rda_e_fut, len_corr_gf_e_fut, len_corr_envdist_e_fut)

# weight
weigh_corr_lea_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_e_fut <- gofit_corr(df = go_grow_ind_e_fut, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_e_fut <- rbind(weigh_corr_lea_e_fut, weigh_corr_rda_e_fut, weigh_corr_gf_e_fut, weigh_corr_envdist_e_fut)

# put together
fit_corr_e_fut <- rbind(surv_corr_e_fut, len_corr_e_fut, weigh_corr_e_fut) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_ind_w_fut <- go_surv_ind[go_surv_ind$coast_pop == "west" & go_surv_ind$curr_fut == "future",]
go_grow_ind_w_fut <- go_grow_ind[go_grow_ind$coast_pop == "west" & go_surv_ind$curr_fut == "future",]

# survival
surv_corr_lea_w_fut <- gofit_corr(df = go_surv_ind_w_fut, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_w_fut <- gofit_corr(df = go_surv_ind_w_fut, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_w_fut <- gofit_corr(df = go_surv_ind_w_fut, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_w_fut <- gofit_corr(df = go_surv_ind_w_fut, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_w_fut <- rbind(surv_corr_lea_w_fut, surv_corr_rda_w_fut, surv_corr_gf_w_fut, surv_corr_envdist_w_fut)

# length
len_corr_lea_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "env_dist", clone = "none", fit_met = "len")
len_corr_w_fut <- rbind(len_corr_lea_w_fut, len_corr_rda_w_fut, len_corr_gf_w_fut, len_corr_envdist_w_fut)

# weight
weigh_corr_lea_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_w_fut <- gofit_corr(df = go_grow_ind_w_fut, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_w_fut <- rbind(weigh_corr_lea_w_fut, weigh_corr_rda_w_fut, weigh_corr_gf_w_fut, weigh_corr_envdist_w_fut)

# put together
fit_corr_w_fut <- rbind(surv_corr_w_fut, len_corr_w_fut, weigh_corr_w_fut) %>% mutate(coast = "west")
#############

# put together
fit_corr_coast_fut <- rbind(fit_corr_w_fut, fit_corr_e_fut)
```

Current outcomes overall

```{r}
## subset to current
go_surv_ind_curr <- go_surv_ind[go_surv_ind$curr_fut == "current",]
go_grow_ind_curr <- go_grow_ind[go_grow_ind$curr_fut == "current",]

## full
#######
# survival
surv_corr_lea_full_curr <- gofit_corr(df = go_surv_ind_curr, go = "lea_offset", clone = "full", fit_met = "surv")
surv_corr_rda_full_curr <- gofit_corr(df = go_surv_ind_curr, go = "rda_offset", clone = "full", fit_met = "surv")
surv_corr_gf_full_curr <- gofit_corr(df = go_surv_ind_curr, go = "gf_offset", clone = "full", fit_met = "surv")
surv_corr_envdist_full_curr <- gofit_corr(df = go_surv_ind_curr, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_full_curr <- rbind(surv_corr_lea_full_curr, surv_corr_rda_full_curr, surv_corr_gf_full_curr, surv_corr_envdist_full_curr)

# length
len_corr_lea_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "lea_offset", clone = "full", fit_met = "len")
len_corr_rda_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "rda_offset", clone = "full", fit_met = "len")
len_corr_gf_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "gf_offset", clone = "full", fit_met = "len")
len_corr_envdist_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "env_dist", clone = "none", fit_met = "len")
len_corr_full_curr <- rbind(len_corr_lea_full_curr, len_corr_rda_full_curr, len_corr_gf_full_curr, len_corr_envdist_full_curr)

# weight
weigh_corr_lea_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_full_curr <- gofit_corr(df = go_grow_ind_curr, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_full_curr <- rbind(weigh_corr_lea_full_curr, weigh_corr_rda_full_curr, weigh_corr_gf_full_curr, weigh_corr_envdist_full_curr)
#######

## mll
######
# survival
surv_corr_lea_mll_curr <- gofit_corr(df = go_surv_ind_curr, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_mll_curr <- gofit_corr(df = go_surv_ind_curr, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_mll_curr <- gofit_corr(df = go_surv_ind_curr, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_mll_curr <- gofit_corr(df = go_surv_ind_curr, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_mll_curr <- rbind(surv_corr_lea_mll_curr, surv_corr_rda_mll_curr, surv_corr_gf_mll_curr, surv_corr_envdist_mll_curr)

# length
len_corr_lea_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "env_dist", clone = "none", fit_met = "len")
len_corr_mll_curr <- rbind(len_corr_lea_mll_curr, len_corr_rda_mll_curr, len_corr_gf_mll_curr, len_corr_envdist_mll_curr)

# weight
weigh_corr_lea_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_mll_curr <- gofit_corr(df = go_grow_ind_curr, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_mll_curr <- rbind(weigh_corr_lea_mll_curr, weigh_corr_rda_mll_curr, weigh_corr_gf_mll_curr, weigh_corr_envdist_mll_curr)
######

# put together
fit_corr_curr <- rbind(surv_corr_full_curr, len_corr_full_curr, weigh_corr_full_curr, surv_corr_mll_curr, len_corr_mll_curr, weigh_corr_mll_curr)
```

Current outcomes by coast (mll only)

```{r}
## east coast
#############
# subset to east coast
go_surv_ind_e_curr <- go_surv_ind[go_surv_ind$coast_pop == "east" & go_surv_ind$curr_fut == "current",]
go_grow_ind_e_curr <- go_grow_ind[go_grow_ind$coast_pop == "east" & go_surv_ind$curr_fut == "current",]

# survival
surv_corr_lea_e_curr <- gofit_corr(df = go_surv_ind_e_curr, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_e_curr <- gofit_corr(df = go_surv_ind_e_curr, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_e_curr <- gofit_corr(df = go_surv_ind_e_curr, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_e_curr <- gofit_corr(df = go_surv_ind_e_curr, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_e_curr <- rbind(surv_corr_lea_e_curr, surv_corr_rda_e_curr, surv_corr_gf_e_curr, surv_corr_envdist_e_curr)

# length
len_corr_lea_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "env_dist", clone = "none", fit_met = "len")
len_corr_e_curr <- rbind(len_corr_lea_e_curr, len_corr_rda_e_curr, len_corr_gf_e_curr, len_corr_envdist_e_curr)

# weight
weigh_corr_lea_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "lea_offset", clone = "full", fit_met = "weigh")
weigh_corr_rda_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "rda_offset", clone = "full", fit_met = "weigh")
weigh_corr_gf_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "gf_offset", clone = "full", fit_met = "weigh")
weigh_corr_envdist_e_curr <- gofit_corr(df = go_grow_ind_e_curr, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_e_curr <- rbind(weigh_corr_lea_e_curr, weigh_corr_rda_e_curr, weigh_corr_gf_e_curr, weigh_corr_envdist_e_curr)

# put together
fit_corr_e_curr <- rbind(surv_corr_e_curr, len_corr_e_curr, weigh_corr_e_curr) %>% mutate(coast = "east")
#############

## west coast
#############
# subset to west coast
go_surv_ind_w_curr <- go_surv_ind[go_surv_ind$coast_pop == "west" & go_surv_ind$curr_fut == "current",]
go_grow_ind_w_curr <- go_grow_ind[go_grow_ind$coast_pop == "west" & go_surv_ind$curr_fut == "current",]

# survival
surv_corr_lea_w_curr <- gofit_corr(df = go_surv_ind_w_curr, go = "lea_offset", clone = "mll", fit_met = "surv")
surv_corr_rda_w_curr <- gofit_corr(df = go_surv_ind_w_curr, go = "rda_offset", clone = "mll", fit_met = "surv")
surv_corr_gf_w_curr <- gofit_corr(df = go_surv_ind_w_curr, go = "gf_offset", clone = "mll", fit_met = "surv")
surv_corr_envdist_w_curr <- gofit_corr(df = go_surv_ind_w_curr, go = "env_dist", clone = "none", fit_met = "surv")
surv_corr_w_curr <- rbind(surv_corr_lea_w_curr, surv_corr_rda_w_curr, surv_corr_gf_w_curr, surv_corr_envdist_w_curr)

# length
len_corr_lea_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "lea_offset", clone = "mll", fit_met = "len")
len_corr_rda_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "rda_offset", clone = "mll", fit_met = "len")
len_corr_gf_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "gf_offset", clone = "mll", fit_met = "len")
len_corr_envdist_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "env_dist", clone = "none", fit_met = "len")
len_corr_w_curr <- rbind(len_corr_lea_w_curr, len_corr_rda_w_curr, len_corr_gf_w_curr, len_corr_envdist_w_curr)

# weight
weigh_corr_lea_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "lea_offset", clone = "mll", fit_met = "weigh")
weigh_corr_rda_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "rda_offset", clone = "mll", fit_met = "weigh")
weigh_corr_gf_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "gf_offset", clone = "mll", fit_met = "weigh")
weigh_corr_envdist_w_curr <- gofit_corr(df = go_grow_ind_w_curr, go = "env_dist", clone = "none", fit_met = "weigh")
weigh_corr_w_curr <- rbind(weigh_corr_lea_w_curr, weigh_corr_rda_w_curr, weigh_corr_gf_w_curr, weigh_corr_envdist_w_curr)

# put together
fit_corr_w_curr <- rbind(surv_corr_w_curr, len_corr_w_curr, weigh_corr_w_curr) %>% mutate(coast = "west")
#############

# put together
fit_corr_coast_curr <- rbind(fit_corr_w_curr, fit_corr_e_curr)
```

### Modeling & ANCOVA

```{r}
# info for ordering, labelling, plotting
trt_labels <- c("TempControl-21psu" = "Current W Coast", "TempControl-7psu" = "Current E Coast",  "TempWarm-16psu" = "Future W Coast", "TempWarm-5psu" = "Future E Coast")
pop_colors <- c("VIK" = "#002e5a", "VAT" = "#0130b9", "HOG" = "#0078cf", "BAR" = "#82daff", "KUR" = "#d9ed92", "KAL" = "#4ad66d", "HOR" = "#25a244", "BJO" = "#155d27")
pop_order <- data.frame(pop = c("VIK", "VAT", "HOG", "BAR", "KUR", "KAL", "HOR", "BJO"), pop_order = 1:8)
trt_colors <- c("Current W Coast" = "#FFA203", "Current E Coast" = "#FE691E",  "Future W Coast" = "deeppink", "Future E Coast" = "deeppink4")
trt_order <- data.frame(trt_name = c("Current W Coast", "Current E Coast", "Future W Coast", "Future E Coast"), trt_order = 1:4)
```

By population: survival

```{r}
### LEA
## full
#######
# glm for binary survival
mod_go_surv_lea_full <- glm(isAlive ~ offset + pop, 
                            family = binomial(link = "logit"), 
                            data = go_surv[go_surv$method == "lea_offset" & go_surv$cc == "full",])
plot(mod_go_surv_lea_full)
summary(mod_go_surv_lea_full)
Anova(mod_go_surv_lea_full, type = "III") # significant!

# letters
emm_surv_lea_full <- emmeans(mod_go_surv_lea_full, "pop")
letters_surv_lea_full <- cld(emm_surv_lea_full, Letters = letters)
letters_surv_lea_full_mg <- merge(letters_surv_lea_full, pop_order, by = "pop") %>% arrange(pop_order)
#######

## mll
######
# glm for binary survival
mod_go_surv_lea_mll <- glm(isAlive ~ offset + pop, 
                            family = binomial(link = "logit"), 
                            data = go_surv[go_surv$method == "lea_offset" & go_surv$cc == "mll",])
plot(mod_go_surv_lea_mll)
summary(mod_go_surv_lea_mll)
Anova(mod_go_surv_lea_mll, type = "III") # significant!

# letters
emm_surv_lea_mll <- emmeans(mod_go_surv_lea_mll, "pop")
letters_surv_lea_mll <- cld(emm_surv_lea_mll, Letters = letters)
letters_surv_lea_mll_mg <- merge(letters_surv_lea_mll, pop_order, by = "pop") %>% arrange(pop_order)
######

### RDA
## full
#######
# glm for binary survival
mod_go_surv_rda_full <- glm(isAlive ~ offset + pop, 
                            family = binomial(link = "logit"), 
                            data = go_surv[go_surv$method == "rda_offset" & go_surv$cc == "full",])
plot(mod_go_surv_rda_full)
summary(mod_go_surv_rda_full)
Anova(mod_go_surv_rda_full, type = "III") # significant!

# letters
emm_surv_rda_full <- emmeans(mod_go_surv_rda_full, "pop")
letters_surv_rda_full <- cld(emm_surv_rda_full, Letters = letters)
letters_surv_rda_full_mg <- merge(letters_surv_rda_full, pop_order, by = "pop") %>% arrange(pop_order)
#######

## mll
#######
# glm for binary survival
mod_go_surv_rda_mll <- glm(isAlive ~ offset + pop, 
                            family = binomial(link = "logit"), 
                            data = go_surv[go_surv$method == "rda_offset" & go_surv$cc == "mll",])
plot(mod_go_surv_rda_mll)
summary(mod_go_surv_rda_mll)
Anova(mod_go_surv_rda_mll, type = "III") # significant!

# letters
emm_surv_rda_mll <- emmeans(mod_go_surv_rda_mll, "pop")
letters_surv_rda_mll <- cld(emm_surv_rda_mll, Letters = letters)
letters_surv_rda_mll_mg <- merge(letters_surv_rda_mll, pop_order, by = "pop") %>% arrange(pop_order)
#######

### GF
## full
#######
# glm for binary survival
mod_go_surv_gf_full <- glm(isAlive ~ offset + pop, 
                            family = binomial(link = "logit"), 
                            data = go_surv[go_surv$method == "gf_offset" & go_surv$cc == "full",])
plot(mod_go_surv_gf_full)
summary(mod_go_surv_gf_full)
Anova(mod_go_surv_gf_full, type = "III") # significant!

# letters
emm_surv_gf_full <- emmeans(mod_go_surv_gf_full, "pop")
letters_surv_gf_full <- cld(emm_surv_gf_full, Letters = letters)
letters_surv_gf_full_mg <- merge(letters_surv_gf_full, pop_order, by = "pop") %>% arrange(pop_order)
#######

## mll
#######
# glm for binary survival
mod_go_surv_gf_mll <- glm(isAlive ~ offset + pop, 
                            family = binomial(link = "logit"), 
                            data = go_surv[go_surv$method == "gf_offset" & go_surv$cc == "mll",])
plot(mod_go_surv_gf_mll)
summary(mod_go_surv_gf_mll)
Anova(mod_go_surv_gf_mll, type = "III") # significant!

# letters
emm_surv_gf_mll <- emmeans(mod_go_surv_gf_mll, "pop")
letters_surv_gf_mll <- cld(emm_surv_gf_mll, Letters = letters)
letters_surv_gf_mll_mg <- merge(letters_surv_gf_mll, pop_order, by = "pop") %>% arrange(pop_order)
#######
```

By coast: survival

```{r}
### LEA
## full
#######
# glm for binary survival
mod_go_surv_lea_full_coast <- glm(isAlive ~ offset + coast_pop, 
                                  family = binomial(link = "logit"), 
                                  data = go_surv[go_surv$method == "lea_offset" & go_surv$cc == "full",])
plot(mod_go_surv_lea_full_coast)
summary(mod_go_surv_lea_full_coast)
Anova(mod_go_surv_lea_full_coast, type = "III") # significant!

# letters
emm_surv_lea_full_coast <- emmeans(mod_go_surv_lea_full_coast, "coast_pop")
letters_surv_lea_full_coast <- cld(emm_surv_lea_full_coast, Letters = letters)
#######

## mll
######
# glm for binary survival
mod_go_surv_lea_mll_coast <- glm(isAlive ~ offset + coast_pop, 
                                 family = binomial(link = "logit"), 
                                 data = go_surv[go_surv$method == "lea_offset" & go_surv$cc == "mll",])
plot(mod_go_surv_lea_mll_coast)
summary(mod_go_surv_lea_mll_coast)
Anova(mod_go_surv_lea_mll_coast, type = "III") # significant!

# letters
emm_surv_lea_mll_coast <- emmeans(mod_go_surv_lea_mll_coast, "coast_pop")
letters_surv_lea_mll_coast <- cld(emm_surv_lea_mll_coast, Letters = letters)
######

### RDA
## full
#######
# glm for binary survival
mod_go_surv_rda_full_coast <- glm(isAlive ~ offset + coast_pop, 
                                  family = binomial(link = "logit"), 
                                  data = go_surv[go_surv$method == "rda_offset" & go_surv$cc == "full",])
plot(mod_go_surv_rda_full_coast)
summary(mod_go_surv_rda_full_coast)
Anova(mod_go_surv_rda_full_coast, type = "III") # significant!

# letters
emm_surv_rda_full_coast <- emmeans(mod_go_surv_rda_full_coast, "coast_pop")
letters_surv_rda_full_coast <- cld(emm_surv_rda_full_coast, Letters = letters)
#######

## mll
######
# glm for binary survival
mod_go_surv_rda_mll_coast <- glm(isAlive ~ offset + coast_pop, 
                                 family = binomial(link = "logit"), 
                                 data = go_surv[go_surv$method == "rda_offset" & go_surv$cc == "mll",])
plot(mod_go_surv_rda_mll_coast)
summary(mod_go_surv_rda_mll_coast)
Anova(mod_go_surv_rda_mll_coast, type = "III") # significant!

# letters
emm_surv_rda_mll_coast <- emmeans(mod_go_surv_rda_mll_coast, "coast_pop")
letters_surv_rda_mll_coast <- cld(emm_surv_rda_mll_coast, Letters = letters)
######

### GF
## full
#######
# glm for binary survival
mod_go_surv_gf_full_coast <- glm(isAlive ~ offset + coast_pop, 
                                  family = binomial(link = "logit"), 
                                  data = go_surv[go_surv$method == "gf_offset" & go_surv$cc == "full",])
plot(mod_go_surv_gf_full_coast)
summary(mod_go_surv_gf_full_coast)
Anova(mod_go_surv_gf_full_coast, type = "III") # significant!

# letters
emm_surv_gf_full_coast <- emmeans(mod_go_surv_gf_full_coast, "coast_pop")
letters_surv_gf_full_coast <- cld(emm_surv_gf_full_coast, Letters = letters)
#######

## mll
######
# glm for binary survival
mod_go_surv_gf_mll_coast <- glm(isAlive ~ offset + coast_pop, 
                                 family = binomial(link = "logit"), 
                                 data = go_surv[go_surv$method == "gf_offset" & go_surv$cc == "mll",])
plot(mod_go_surv_gf_mll_coast)
summary(mod_go_surv_gf_mll_coast)
Anova(mod_go_surv_gf_mll_coast, type = "III") # significant!

# letters
emm_surv_gf_mll_coast <- emmeans(mod_go_surv_gf_mll_coast, "coast_pop")
letters_surv_gf_mll_coast <- cld(emm_surv_gf_mll_coast, Letters = letters)
######
```

### Visualizations

#### Plotting Info

```{r}
# plotting info
trt_labels <- c("TempControl-21psu" = "Current W Coast", "TempControl-7psu" = "Current E Coast",  "TempWarm-16psu" = "Future W Coast", "TempWarm-5psu" = "Future E Coast")
pop_colors <- c("VIK" = "#002e5a", "VAT" = "#0130b9", "HOG" = "#0078cf", "BAR" = "#82daff", "KUR" = "#d9ed92", "KAL" = "#4ad66d", "HOR" = "#25a244", "BJO" = "#155d27")
pop_order <- data.frame(pop = c("VIK", "VAT", "HOG", "BAR", "KUR", "KAL", "HOR", "BJO"), pop_order = 1:8)
trt_colors <- c("Current W Coast" = "#FFA203", "Current E Coast" = "#FE691E",  "Future W Coast" = "deeppink", "Future E Coast" = "deeppink4")
trt_order <- data.frame(trt_name = c("Current W Coast", "Current E Coast", "Future W Coast", "Future E Coast"), trt_order = 1:4)
coast_cols <- c("east" = "#2b991f", "west" = "#1f4699")
coast_shapes <- c("east" = 21, "west" = 23)
we_cols_fill <- c(pop_colors, coast_cols)

# combine with data
go_surv_plot <- merge(go_surv, pop_order, by = "pop")
```

By individual: mll only

```{r}
## lea
######
# mll
surv_mll_binlog_lea <- go_surv_plot %>%
  filter(method == "lea_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order), shape = coast_pop)) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 3) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Survival in Current and Future Treatments vs \nLEA Offset", 
       x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 23),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_lea

# mll, single coast line
surv_mll_coast_binlog_lea <- go_surv_plot %>%
  filter(method == "lea_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = coast_pop)) + 
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = coast_pop), method = "glm", method.args = list(family = "binomial"), se = T) + 
  labs(title = "Survival in Current and Future Treatments vs \nLEA Offset", 
       x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Coast",
                    values = coast_cols) +
  scale_colour_manual(name = "Coast",
                    values = coast_cols) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_coast_binlog_lea

# lea future
surv_mll_fut_all_lea <- go_surv_plot %>%
  filter(method == "lea_offset" & cc == "mll" & curr_fut == "future") %>%
  ggplot(aes(x = offset, y = isAlive)) +
  geom_point(aes(fill = fct_reorder(pop, pop_order)), size = 2, shape = 21) + 
  geom_smooth(method = "lm", colour = "black", se = T) + 
  labs(title = "Future Survival vs LEA Offset", 
       x = "LEA Offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = we_cols_fill,
                    breaks = names(pop_colors)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_fut_all_lea
######

## rda
######
# mll
surv_mll_binlog_rda <- go_surv_plot %>%
  filter(method == "rda_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Survival in Current and Future Treatments vs \nRDA Offset", 
       x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_rda

# mll, single coast line
surv_mll_coast_binlog_rda <- go_surv_plot %>%
  filter(method == "rda_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = coast_pop)) + 
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = coast_pop), method = "glm", method.args = list(family = "binomial"), se = T) + 
  labs(title = "Survival in Current and Future Treatments vs \nRDA Offset", 
       x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Coast",
                    values = coast_cols) +
  scale_colour_manual(name = "Coast",
                    values = coast_cols) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_coast_binlog_rda

# rda future
surv_mll_fut_all_rda <- go_surv_plot %>%
  filter(method == "rda_offset" & cc == "mll" & curr_fut == "future") %>%
  ggplot(aes(x = offset, y = isAlive)) +
  geom_point(aes(fill = fct_reorder(pop, pop_order)), size = 2, shape = 21) + 
  geom_smooth(method = "lm", colour = "black", se = T) + 
  labs(title = "Future Survival vs RDA Offset", 
       x = "RDA Offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = we_cols_fill,
                    breaks = names(pop_colors)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_fut_all_rda
######

## gf
######
# mll
surv_mll_binlog_gf <- go_surv_plot %>%
  filter(method == "gf_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Survival in Current and Future Treatments vs \nGF Offset", 
       x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_gf

# mll, single coast line
surv_mll_coast_binlog_gf <- go_surv_plot %>%
  filter(method == "gf_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = coast_pop)) + 
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = coast_pop), method = "glm", method.args = list(family = "binomial"), se = T) + 
  labs(title = "Survival in Current and Future Treatments vs \nGF Offset", 
       x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Coast",
                    values = coast_cols) +
  scale_colour_manual(name = "Coast",
                    values = coast_cols) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_coast_binlog_gf

# gf future
surv_mll_fut_all_gf <- go_surv_plot %>%
  filter(method == "gf_offset" & cc == "mll" & curr_fut == "future") %>%
  ggplot(aes(x = offset, y = isAlive)) +
  geom_point(aes(fill = fct_reorder(pop, pop_order)), size = 2, shape = 21) + 
  geom_smooth(method = "lm", colour = "black", se = T) + 
  labs(title = "Future Survival vs GF Offset", 
       x = "GF Offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = we_cols_fill,
                    breaks = names(pop_colors)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_fut_all_gf
######

## env dist
###########
# mll
surv_mll_binlog_envdist <- go_surv_plot %>%
  filter(method == "env_dist" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Survival in Current and Future Treatments vs \nEnvironmental Distance", 
       x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_envdist

# mll, single coast line
surv_mll_coast_binlog_envdist <- go_surv_plot %>%
  filter(method == "env_dist" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = coast_pop)) + 
  geom_point(size = 2, shape = 21) + 
  geom_smooth(aes(col = coast_pop), method = "glm", method.args = list(family = "binomial"), se = T) + 
  labs(title = "Survival in Current and Future Treatments vs \nEnvironmental Distance", 
       x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Coast",
                    values = coast_cols) +
  scale_colour_manual(name = "Coast",
                    values = coast_cols) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_coast_binlog_envdist

# env dist future
surv_mll_fut_all_envdist <- go_surv_plot %>%
  filter(method == "env_dist" & cc == "mll" & curr_fut == "future") %>%
  ggplot(aes(x = offset, y = isAlive)) +
  geom_point(aes(fill = fct_reorder(pop, pop_order)), size = 2, shape = 21) + 
  geom_smooth(method = "lm", colour = "black", se = T) + 
  labs(title = "Future Survival vs Environmental Distance", 
       x = "Environmental Distance", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = we_cols_fill,
                    breaks = names(pop_colors)) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_fut_all_envdist
###########
```

By coast

```{r}

```

Future outcomes only

```{r}

```

### Exporting Results

```{r}

```
