---
title: "Eelgrass GO Fitness Comparisons"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "/Users/camillerumberger/Documents/GitHub/BalticEelgrass")
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# packages
library(car)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)
library(knitr)
library(kableExtra)
library(emmeans)
library(multcomp)
library(multcompView)

# data
# exp fit
fit_full <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
surv <- read.csv("data/experiment/survival_dat_trt_pop.csv")[,-1]
grow <- read.csv("data/experiment/fit_growth_dat.csv")[,-1]

# fit offsets
fit_off_clim <- read.csv("results/fit_results/climate_change_fitness_offsets_2026-05-28.csv")[,-1]
fit_off_clim_ind <- read.csv("results/fit_results/climate_change_fitness_offsets_ind_2026-07-10.csv")[,-1]

# go
lea <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_2025-12-03.csv")[,-1]
lea_cand <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_gea_2026-06-30.csv")[,-1]
rda <- read.csv("results/GO_results/RDA_offsets_exptrt_wide_2026-01-02.csv")[,-1]
gf <- read.csv("results/GO_results/GF_offsets_exptrt_wide_2025-12-03.csv")[,-1]
env_dist <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]
```

## Genomic Offset - Fitness Offset Comparisons

This script compares genomic offset against fitness metrics in a "common garden" (across genotypes) and "current future" (across environments) comparison. 

### Data Prep

#### GO Data Processing

These data are already in a matrix format, but I will still make a long form combined df for plotting. 

```{r}
head(lea)
head(rda)
head(gf)

# long versions
# lea
lea_long <- lea %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

lea_cand_long <- lea_cand %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

# rda
rda_long <- rda %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "rda_offset")

# gf
gf_long <- gf %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "gf_offset")

# env dist
env_dist_trt <- env_dist %>% filter(pops1 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu") & !(pops2 %in% c("TempWarm-5psu", "TempWarm-16psu", "TempControl-7psu", "TempControl-21psu"))) %>% dplyr::select(pops1, pops2, distances) %>% rename(pop = pops2, trt = pops1, env_dist = distances)

# put everything together
go1 <- merge(lea_long, rda_long, by = c("trt", "cc", "pop"))
go2 <- merge(go1, gf_long[gf_long$mod == "geno", c("trt","pop","cc","gf_offset")], by = c("trt", "cc", "pop"))
go_wide <- merge(go2, env_dist_trt, by = c("pop", "trt"))

# long version of all
go_long <- go_wide %>% pivot_longer(
  cols = c("lea_offset","rda_offset","gf_offset", "env_dist"),
  names_to = "method",
  values_to = "offset")
```

#### Experimental Data

```{r}
# survival
head(surv)
dim(surv)
surv$trt <- gsub("_", "-", surv$trt)
fit_full$isAlive <- ifelse(fit_full$isAlive == TRUE, 1, 0)

# surv only fitness, all plants
surv_all <- fit_full[,c("pop", "trt", "tank_ID", "isAlive")]
surv_all$trt <- gsub("_", "-", surv_all$trt)
  
# also a format with survival in each tank/pop/trt
surv_tanks <- fit_full %>% 
  dplyr::select(pop, trt, tank_ID, isAlive) %>%
  group_by(pop, tank_ID, trt) %>%
  summarize(surv = mean(isAlive),
            surv_se = sd(isAlive)/length(isAlive))

# growth
head(grow)
grow_trim <- grow[,c("trt", "pop", "pop_trt", "tank_ID", "length_growth", "length_std_change", "weight_change", "weight_std_change", "above_below_weight_ratio")]
grow_trim$trt <- gsub("_","-",grow_trim$trt)

# some pop/trt level data
grow_pop_trt <- grow_trim %>% 
  group_by(pop, trt) %>% 
  summarize(avg_length_growth = mean(length_growth, na.rm = T),
    avg_length_std_change = mean(length_std_change, na.rm = T),
    avg_weight_change = mean(weight_change, na.rm = T),
    avg_weight_std_change = mean(weight_std_change, na.rm = T),
    avg_above_below_weight_ratio = mean(above_below_weight_ratio, na.rm = T))

# fitness offsets
head(fit_off_clim) # only goes with the future climate stuff
```


### GO v Fitness

To assess whether genomic offset predictions align with experimental outcomes, I conduct a correlation analysis between genomic offset and different fitness and fitness offset metrics, namely absolute and relative fitness, as well as both absolute and relative current-future fitness offsets.

#### Data Merge

```{r}
# isolate relevant trt/pop combos
go_clim_w <- go_long %>% filter(pop %in% c("VAT", "VIK", "HOG", "BAR") & trt %in% c("TempControl-21psu", "TempWarm-16psu")) %>% mutate(coast = "West", cf_trt = ifelse(grepl("TempWarm", trt), "future", "current"))
go_clim_e <- go_long %>% filter(pop %in% c("KUR", "KAL", "HOR", "BJO") & trt %in% c("TempControl-7psu", "TempWarm-5psu")) %>% mutate(coast = "East", cf_trt = ifelse(grepl("TempWarm", trt), "future", "current"))
go_clim <- rbind(go_clim_w, go_clim_e)
go_clim_fut <- rbind(go_clim_w[go_clim_w$cf_trt == "future",], go_clim_e[go_clim_e$cf_trt == "future",])

# combine fitness and offset data
fit_go <- merge(go_clim_fut, fit_off_clim, by = c("pop", "coast"))

# ordered list of pops
pops <- unique(fit_go[,c("pop", "pop_ord")]) %>% arrange(pop_ord)

# also for the raw fitness data
# combine fitness and offset data
surv_all_go <- merge(go_clim, surv_all, by = c("pop", "trt"))
grow_go <- merge(go_clim, grow_trim, by = c("pop", "trt"))

surv_w <- surv %>% filter(pop %in% c("VIK", "VAT", "HOG", "BAR") & trt %in% c("TempControl-21psu", "TempWarm-16psu"))
surv_e <- surv %>% filter(pop %in% c("KUR", "KAL", "HOR", "BJO") & trt %in% c("TempControl-7psu", "TempWarm-5psu"))
surv_prop_clim <- rbind(surv_w, surv_e)
surv_prop_go <- merge(surv_prop_clim[,-6], go_clim[,-6], by = c("pop", "trt"))
```

#### Correlations

```{r}
# function for correlations current future offset
#################################################
# df is dataframe, go is lea_offset/rda_offset/gf_offset/env_dist, clone is mll/full, fit_met is surv/len/weigh
fitcorrCF <- function(df, go, clone, fit_met) {
  # subset data
  df_subset <- df[df$method == go & df$cc == clone & df$metric == fit_met & df$cf_trt == "future",]
  
  # set up df
  corrs_cf <- data.frame(fit_metric = NA, method = NA, cc = NA, abs_corr = NA, abs_pval = NA, rel_corr = NA, rel_pval = NA, cf_abs_corr = NA, cf_abs_pval = NA, cf_rel_corr = NA, cf_rel_pval = NA)
  
  # metadata
  corrs_cf$fit_metric <- fit_met
  corrs_cf$method <- go
  corrs_cf$cc <- clone
  
  # correlation testing
  test_abs = cor.test(df_subset$abs_future, df_subset$offset, method = "kendall", na.action = "na.omit")
  corrs_cf$abs_corr = test_abs$estimate
  corrs_cf$abs_pval = test_abs$p.value

  test_rel = cor.test(df_subset$CF_relative, df_subset$offset, method = "kendall", na.action = "na.omit")
  corrs_cf$rel_corr = test_rel$estimate
  corrs_cf$rel_pval = test_rel$p.value

  test_cf_abs = cor.test(df_subset$CF_abs_offset, df_subset$offset, method = "kendall", na.action = "na.omit")
  corrs_cf$cf_abs_corr = test_cf_abs$estimate
  corrs_cf$cf_abs_pval = test_cf_abs$p.value

  test_cf_rel = cor.test(df_subset$CF_rel_offset, df_subset$offset, method = "kendall", na.action = "na.omit")
  corrs_cf$cf_rel_corr = test_cf_rel$estimate
  corrs_cf$cf_rel_pval = test_cf_rel$p.value
  return(corrs_cf)
}

# function for survival proportions
fitcorrSurv <- function(df, go, clone) {
  # subset data
  df_subset <- df[df$method == go & df$cc == clone,]
  
  # set up df
  corrs_cf <- data.frame(fit_metric = NA, method = NA, cc = NA, surv_corr = NA, surv_pval = NA)
  
  # metadata
  corrs_cf$fit_metric <- "surv"
  corrs_cf$method <- go
  corrs_cf$cc <- clone
  
  # correlation testing
  test_surv = cor.test(df_subset$surv, df_subset$offset, method = "kendall", na.action = "na.exclude")
  corrs_cf$surv_corr = test_surv$estimate
  corrs_cf$surv_pval = test_surv$p.value
  return(corrs_cf)
}
#################################################

# test correlations
###################
# lea
lea_cf <- rbind(fitcorrCF(fit_go, go = "lea_offset", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "lea_offset", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "lea_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "lea_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "lea_offset", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "lea_offset", clone = "mll", fit_met = "weigh"))

lea_cf_ew <- rbind(cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "lea_offset", clone = "full", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "lea_offset", clone = "mll", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "lea_offset", clone = "full", fit_met = "surv"), coast = "East"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "lea_offset", clone = "mll", fit_met = "surv"), coast = "East"))

# rda
rda_cf <- rbind(fitcorrCF(fit_go, go = "rda_offset", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "rda_offset", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "rda_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "rda_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "rda_offset", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "rda_offset", clone = "mll", fit_met = "weigh"))

rda_cf_ew <- rbind(cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "rda_offset", clone = "full", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "rda_offset", clone = "mll", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "rda_offset", clone = "full", fit_met = "surv"), coast = "East"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "rda_offset", clone = "mll", fit_met = "surv"), coast = "East"))

# gf
gf_cf <- rbind(fitcorrCF(fit_go, go = "gf_offset", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "gf_offset", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "gf_offset", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "gf_offset", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "gf_offset", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "gf_offset", clone = "mll", fit_met = "weigh"))

gf_cf_ew <- rbind(cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "gf_offset", clone = "full", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "gf_offset", clone = "mll", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "gf_offset", clone = "full", fit_met = "surv"), coast = "East"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "gf_offset", clone = "mll", fit_met = "surv"), coast = "East"))

# env dist - note we use both full and mll to match with other methods when subsetting
envdist_cf <- rbind(fitcorrCF(fit_go, go = "env_dist", clone = "full", fit_met = "surv"),
fitcorrCF(fit_go, go = "env_dist", clone = "full", fit_met = "len"),
fitcorrCF(fit_go, go = "env_dist", clone = "full", fit_met = "weigh"),
fitcorrCF(fit_go, go = "env_dist", clone = "mll", fit_met = "surv"),
fitcorrCF(fit_go, go = "env_dist", clone = "mll", fit_met = "len"),
fitcorrCF(fit_go, go = "env_dist", clone = "mll", fit_met = "weigh"))

envdist_cf_ew <- rbind(cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "env_dist", clone = "full", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "West",], go = "env_dist", clone = "mll", fit_met = "surv"), coast = "West"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "env_dist", clone = "full", fit_met = "surv"), coast = "East"),
cbind(fitcorrCF(fit_go[fit_go$coast == "East",], go = "env_dist", clone = "mll", fit_met = "surv"), coast = "East"))

# combine
cf_clim_summary <- rbind(lea_cf, rda_cf, gf_cf, envdist_cf)
cf_clim_coast_summary <- rbind(lea_cf_ew, rda_cf_ew, gf_cf_ew, envdist_cf_ew)

# pivot correlations and pvals to long format
cf_clim_long <- cf_clim_summary %>% pivot_longer(cols = c(abs_corr, rel_corr, cf_abs_corr, cf_rel_corr, abs_pval, rel_pval, cf_abs_pval, cf_rel_pval),
    names_to = c("fit_type", ".value"),
    names_pattern = "(abs|rel|cf_abs|cf_rel)_(corr|pval)")

cf_clim_coast_long <- cf_clim_coast_summary %>% pivot_longer(cols = c(abs_corr, rel_corr, cf_abs_corr, cf_rel_corr, abs_pval, rel_pval, cf_abs_pval, cf_rel_pval),
    names_to = c("fit_type", ".value"),
    names_pattern = "(abs|rel|cf_abs|cf_rel)_(corr|pval)")

# signif & adjustments - keep in mind these correlations are between literally four data points total for the coast and eight for the overall, so significance is largely meaningless here - should I try per tank averages instead of per population? Then I can't report a single value though
cf_clim_long$signif <- ifelse(cf_clim_long$pval < 0.05, T, F)
cf_clim_coast_long$signif <- ifelse(cf_clim_coast_long$pval < 0.05, T, F)
###################

# save outputs
##############
write.csv(fit_go, paste0("results/GO_compare/climchg_gofit_df_", Sys.Date(), ".csv"))
write.csv(cf_clim_long, paste0("results/GO_compare/climchg_gofit_corrs_", Sys.Date(), ".csv"))
write.csv(cf_clim_coast_long, paste0("results/GO_compare/climchg_gofit_corrs_coast_", Sys.Date(), ".csv"))
##############
```


Katie's chunk
```{r}
fit_goMini <- fit_go[fit_go$cc == "mll" & fit_go$method == "lea_offset" & fit_go$metric == "surv",]
dim(fit_goMini)
cor.plot(fit_goMini[,11:14])
cor(fit_goMini[,11:14])
fit_goMini
fit_goMini$abs_future-fit_goMini$abs_current
fit_goMini$abs_future/fit_goMini$abs_current
(fit_goMini$abs_future-fit_goMini$abs_current)/fit_goMini$abs_current
```

### Visualization

Visualize strength of correlations. Try a heatmap with genomic offset as columns and fitness/fitness offset as rows

```{r}
# quick checks
head(cf_clim_long)

# ordering for plotting
fit_ord <- c("abs", "rel", "cf_abs", "cf_rel")
method_ord <- c("lea_offset", "rda_offset", "gf_offset", "env_dist")

# surv
surv_corplot_full <- cf_clim_long %>% 
  filter(cc == "full" & fit_metric == "surv") %>% 
  mutate(fit_type = factor(fit_type, levels = rev(fit_ord)), method = factor(method, levels = method_ord), corr = round(corr, 3)) %>%
  ggplot(aes(x = method, y = (fit_type), fill = corr)) +
  geom_tile() +
  geom_text(aes(label = corr), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  scale_y_discrete(labels = c("abs" = "Absolute\nFitness", "rel" = "Relative\nFitness", "cf_abs" = "Absolute Current\nFuture Offset", "cf_rel" = "Relative Current\nFuture Offset")) +
  labs(title = "Survival", subtitle = "Correlation between Fitness and Genomic Offsets", x = "Genomic Offset", y = "Fitness") +
  theme_classic()
surv_corplot_full

# length
len_corplot_full <- cf_clim_long %>% 
  filter(cc == "full" & fit_metric == "len") %>% 
  mutate(fit_type = factor(fit_type, levels = rev(fit_ord)), method = factor(method, levels = method_ord), corr = round(corr, 3)) %>%
  ggplot(aes(x = method, y = (fit_type), fill = corr)) +
  geom_tile() +
  geom_text(aes(label = corr), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  scale_y_discrete(labels = c("abs" = "Absolute\nFitness", "rel" = "Relative\nFitness", "cf_abs" = "Absolute Current\nFuture Offset", "cf_rel" = "Relative Current\nFuture Offset")) +
  labs(title = "Standardized Length Change", subtitle = "Correlation between Fitness and Genomic Offsets", x = "Genomic Offset", y = "Fitness") +
  theme_classic()
len_corplot_full

# weight
weigh_corplot_full <- cf_clim_long %>% 
  filter(cc == "full" & fit_metric == "weigh") %>% 
  mutate(fit_type = factor(fit_type, levels = rev(fit_ord)), method = factor(method, levels = method_ord), corr = round(corr, 3)) %>%
  ggplot(aes(x = method, y = (fit_type), fill = corr)) +
  geom_tile() +
  geom_text(aes(label = corr), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  scale_y_discrete(labels = c("abs" = "Absolute\nFitness", "rel" = "Relative\nFitness", "cf_abs" = "Absolute Current\nFuture Offset", "cf_rel" = "Relative Current\nFuture Offset")) +
  labs(title = "Standardized Weight Change", subtitle = "Correlation between Fitness and Genomic Offsets", x = "Genomic Offset", y = "Fitness") +
  theme_classic()
weigh_corplot_full

## coast
########
surv_corplot_w <- cf_clim_coast_long %>% 
  filter(cc == "mll" & coast == "West") %>% 
  mutate(fit_type = factor(fit_type, levels = rev(fit_ord)), method = factor(method, levels = method_ord), corr = round(corr, 3)) %>%
  ggplot(aes(x = method, y = (fit_type), fill = corr)) +
  geom_tile() +
  geom_text(aes(label = corr), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  scale_y_discrete(labels = c("abs" = "Absolute\nFitness", "rel" = "Relative\nFitness", "cf_abs" = "Absolute Current\nFuture Offset", "cf_rel" = "Relative Current\nFuture Offset")) +
  labs(title = "Survival", subtitle = "Correlation between Fitness and Genomic Offsets, \nWest Coast", x = "Genomic Offset", y = "Fitness") +
  theme_classic()
surv_corplot_w

surv_corplot_e <- cf_clim_coast_long %>% 
  filter(cc == "mll" & coast == "East") %>% 
  mutate(fit_type = factor(fit_type, levels = rev(fit_ord)), method = factor(method, levels = method_ord), corr = round(corr, 3)) %>%
  ggplot(aes(x = method, y = (fit_type), fill = corr)) +
  geom_tile() +
  geom_text(aes(label = corr), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  scale_y_discrete(labels = c("abs" = "Absolute\nFitness", "rel" = "Relative\nFitness", "cf_abs" = "Absolute Current\nFuture Offset", "cf_rel" = "Relative Current\nFuture Offset")) +
  labs(title = "Survival", subtitle = "Correlation between Fitness and Genomic Offsets, \nEast Coast", x = "Genomic Offset", y = "Fitness") +
  theme_classic()
surv_corplot_e
```


Also try plotting genomic offset against different fitness offsets.

#### Prep plotting data

```{r}
# plotting info
trt_labels <- c("TempControl-21psu" = "Current W Coast", "TempControl-7psu" = "Current E Coast",  "TempWarm-16psu" = "Future W Coast", "TempWarm-5psu" = "Future E Coast")
pop_colors <- c("VIK" = "#002e5a", "VAT" = "#0130b9", "HOG" = "#0078cf", "BAR" = "#82daff", "KUR" = "#d9ed92", "KAL" = "#4ad66d", "HOR" = "#25a244", "BJO" = "#155d27")
pop_order <- data.frame(pop = c("VIK", "VAT", "HOG", "BAR", "KUR", "KAL", "HOR", "BJO"), pop_order = 1:8)
trt_colors <- c("Current W Coast" = "#FFA203", "Current E Coast" = "#FE691E",  "Future W Coast" = "deeppink", "Future E Coast" = "deeppink4")
trt_order <- data.frame(trt_name = c("Current W Coast", "Current E Coast", "Future W Coast", "Future E Coast"), trt_order = 1:4)

# survival
# binary
surv_all_go <- surv_all_go %>% mutate(trt_name = case_match(trt,
    "TempControl-21psu" ~ "Current W Coast",
    "TempControl-7psu" ~ "Current E Coast",
    "TempWarm-16psu" ~ "Future W Coast",
    "TempWarm-5psu" ~ "Future E Coast",
    .default = NA))
surv_all_go1 <- merge(surv_all_go, pop_order, by = "pop")
surv_all_go2 <- merge(surv_all_go1, trt_order, by = "trt_name")

# proportions
surv_prop_go1 <- surv_prop_go %>% mutate(trt_name = case_match(trt,
    "TempControl-21psu" ~ "Current W Coast",
    "TempControl-7psu" ~ "Current E Coast",
    "TempWarm-16psu" ~ "Future W Coast",
    "TempWarm-5psu" ~ "Future E Coast",
    .default = NA))
surv_prop_go2 <- merge(surv_prop_go1, pop_order, by = "pop")
surv_prop_go3 <- merge(surv_prop_go2, trt_order, by = "trt_name")
```

#### Survival 

Analysis of go/fit relationship

```{r}
## lea
######
# model full log
surv_lea_full_log <- surv_all_go %>% filter(method == "lea_offset" & cc == "full")
mod_surv_lea_full_log1 <- glm(isAlive ~ pop * offset, family = binomial(link = "logit"), data = surv_lea_full_log)
step(mod_surv_lea_full_log1) # remove interaction - slopes are not meaningfully different
mod_surv_lea_full_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_lea_full_log)
summary(mod_surv_lea_full_log) 
#plot(mod_surv_lea_full_log)
Anova(mod_surv_lea_full_log) # both pop and offset signif
emm_surv_lea_full_log <- emmeans(mod_surv_lea_full_log, "pop")
letters_surv_lea_full_log <- cld(emm_surv_lea_full_log, Letters = letters)
letters_surv_lea_full_log_mg <- merge(letters_surv_lea_full_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_lea_full_log_mg1 <- letters_surv_lea_full_log_mg %>% mutate(group_lea = .group, cc = "full") %>% dplyr::select(pop, group_lea, cc)

# model mll log
surv_lea_mll_log <- surv_all_go %>% filter(method == "lea_offset" & cc == "mll")
mod_surv_lea_mll_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_lea_mll_log)
summary(mod_surv_lea_mll_log) 
#plot(mod_surv_lea_mll_log)
Anova(mod_surv_lea_mll_log) # both pop and offset signif
emm_surv_lea_mll_log <- emmeans(mod_surv_lea_mll_log, "pop")
letters_surv_lea_mll_log <- cld(emm_surv_lea_mll_log, Letters = letters)
letters_surv_lea_mll_log_mg <- merge(letters_surv_lea_mll_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_lea_mll_log_mg1 <- letters_surv_lea_mll_log_mg %>% mutate(group_lea = .group, cc = "mll") %>% dplyr::select(pop, group_lea, cc)

# more specific grouping
# letters_surv_lea_mll_log_mg$group_a <- grepl("a", letters_surv_lea_mll_log_mg$.group)
# letters_surv_lea_mll_log_mg$group_b <- grepl("b", letters_surv_lea_mll_log_mg$.group)
# letters_surv_lea_mll_log_mg$group_c <- grepl("c", letters_surv_lea_mll_log_mg$.group)
######

## rda
######
# model full log
surv_rda_full_log <- surv_all_go %>% filter(method == "rda_offset" & cc == "full")
mod_surv_rda_full_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_rda_full_log)
summary(mod_surv_rda_full_log) 
#plot(mod_surv_rda_full_log)
Anova(mod_surv_rda_full_log) # both pop and offset signif
emm_surv_rda_full_log <- emmeans(mod_surv_rda_full_log, "pop")
letters_surv_rda_full_log <- cld(emm_surv_rda_full_log, Letters = letters)
letters_surv_rda_full_log_mg <- merge(letters_surv_rda_full_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_rda_full_log_mg1 <- letters_surv_rda_full_log_mg %>% mutate(group_rda = .group, cc = "full") %>% dplyr::select(pop, group_rda, cc)

# model mll log
surv_rda_mll_log <- surv_all_go %>% filter(method == "rda_offset" & cc == "mll")
mod_surv_rda_mll_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_rda_mll_log)
summary(mod_surv_rda_mll_log) 
#plot(mod_surv_rda_mll_log)
Anova(mod_surv_rda_mll_log) # both pop and offset signif
emm_surv_rda_mll_log <- emmeans(mod_surv_rda_mll_log, "pop")
letters_surv_rda_mll_log <- cld(emm_surv_rda_mll_log, Letters = letters)
letters_surv_rda_mll_log_mg <- merge(letters_surv_rda_mll_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_rda_mll_log_mg$group_a <- grepl("a", letters_surv_rda_mll_log_mg$.group)
letters_surv_rda_mll_log_mg$group_b <- grepl("b", letters_surv_rda_mll_log_mg$.group)
letters_surv_rda_mll_log_mg1 <- letters_surv_rda_mll_log_mg %>% mutate(group_rda = .group, cc = "full") %>% dplyr::select(pop, group_rda, cc)
######

## gf
######
# model full log
surv_gf_full_log <- surv_all_go %>% filter(method == "gf_offset" & cc == "full")
mod_surv_gf_full_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_gf_full_log)
summary(mod_surv_gf_full_log) 
#plot(mod_surv_gf_full_log)
Anova(mod_surv_gf_full_log) # both pop and offset signif
emm_surv_gf_full_log <- emmeans(mod_surv_gf_full_log, "pop")
letters_surv_gf_full_log <- cld(emm_surv_gf_full_log, Letters = letters)
letters_surv_gf_full_log_mg <- merge(letters_surv_gf_full_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_gf_full_log_mg1 <- letters_surv_gf_full_log_mg %>% mutate(group_gf = .group, cc = "full") %>% dplyr::select(pop, group_gf, cc)

# model mll log
surv_gf_mll_log <- surv_all_go %>% filter(method == "gf_offset" & cc == "mll")
mod_surv_gf_mll_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_gf_mll_log)
summary(mod_surv_gf_mll_log) 
#plot(mod_surv_gf_mll_log)
Anova(mod_surv_gf_mll_log) # both pop and offset signif
emm_surv_gf_mll_log <- emmeans(mod_surv_gf_mll_log, "pop")
letters_surv_gf_mll_log <- cld(emm_surv_gf_mll_log, Letters = letters)
letters_surv_gf_mll_log_mg <- merge(letters_surv_gf_mll_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_gf_mll_log_mg1 <- letters_surv_gf_mll_log_mg %>% mutate(group_gf = .group, cc = "mll") %>% dplyr::select(pop, group_gf, cc)

letters_surv_gf_mll_log_mg$group_a <- grepl("a", letters_surv_gf_mll_log_mg$.group)
letters_surv_gf_mll_log_mg$group_b <- grepl("b", letters_surv_gf_mll_log_mg$.group)
letters_surv_gf_mll_log_mg$group_c <- grepl("c", letters_surv_gf_mll_log_mg$.group)
######

## env dist
######
# model full log
surv_envdist_full_log <- surv_all_go %>% filter(method == "env_dist" & cc == "full")
mod_surv_envdist_full_log <- glm(isAlive ~ pop + offset, family = binomial(link = "logit"), data = surv_envdist_full_log)
summary(mod_surv_envdist_full_log) 
#plot(mod_surv_envdist_full_log)
Anova(mod_surv_envdist_full_log) # both pop and offset signif
emm_surv_envdist_full_log <- emmeans(mod_surv_envdist_full_log, "pop")
letters_surv_envdist_full_log <- cld(emm_surv_envdist_full_log, Letters = letters)
letters_surv_envdist_full_log_mg <- merge(letters_surv_envdist_full_log, pop_order, by = "pop") %>% arrange(pop_order)
letters_surv_envdist_mll_log_mg1 <- letters_surv_envdist_full_log_mg %>% mutate(group_envdist = .group, cc = "mll") %>% dplyr::select(pop, group_envdist, cc)
letters_surv_envdist_full_log_mg1 <- letters_surv_envdist_full_log_mg %>% mutate(group_envdist = .group, cc = "full") %>% dplyr::select(pop, group_envdist, cc) # just incase i want to plot full
######
```


```{r}
surv_all_go_test <- merge(go_long, surv_all, by = c("pop", "trt"))
surv_envdist_full_log_test <- surv_all_go_test %>% filter(method == "env_dist" & cc == "full")
mod_surv_envdist_full_log_test <- glmer(isAlive ~ offset + pop + (1|tank_ID), family = binomial(link = "logit"), data = surv_envdist_full_log_test)
summary(mod_surv_envdist_full_log_test) 
#plot(mod_surv_envdist_full_log_test)
Anova(mod_surv_envdist_full_log_test) # pop signif, not env dist
emm_surv_envdist_full_log_test <- emmeans(mod_surv_envdist_full_log_test, "pop")
letters_surv_envdist_full_log_test <- cld(emm_surv_envdist_full_log_test, Letters = letters)
letters_surv_envdist_full_log_mg_test <- merge(letters_surv_envdist_full_log_test, pop_order, by = "pop") %>% arrange(pop_order) # all group together except vat (a) with vik/kal (b) - others all (ab)
```

Plotting

```{r}
# survival comparison
# add letters
letters_surv_bin <- list(rbind(letters_surv_lea_full_log_mg1, letters_surv_lea_mll_log_mg1), rbind(letters_surv_rda_full_log_mg1, letters_surv_rda_mll_log_mg1), rbind(letters_surv_gf_full_log_mg1, letters_surv_gf_mll_log_mg1), rbind(letters_surv_envdist_full_log_mg1, letters_surv_envdist_mll_log_mg1)) %>% reduce(full_join, by = c("pop", "cc"))
plotting_survgo_all <- merge(surv_all_go2, letters_surv_bin, by = c("pop", "cc"))

## lea
######
# full
surv_full_binlog_lea <- plotting_survgo_all %>%
  filter(method == "lea_offset" & cc == "full") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, Full): \nBinary Survival, Logistic Regression", 
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_full_binlog_lea

# mll
surv_mll_binlog_lea <- plotting_survgo_all %>%
  filter(method == "lea_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, MLL): \nBinary Survival, Logistic Regression", 
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_lea

# correlation
lea_temp <- plotting_survgo_all %>% filter(cc == "mll", method == "lea_offset")
lea_temp_w <- lea_temp %>% filter(trt %in% c("TempControl-21psu", "TempWarm-16psu"))
lea_temp_e <- lea_temp %>% filter(trt %in% c("TempControl-7psu", "TempWarm-5psu"))

cor_lea <- cor.test(lea_temp$offset, lea_temp$surv, method = "kendall") # cor -0.229, p = 0.0001829 < 0.001 ***
cor_lea_w <- cor.test(lea_temp_w$offset, lea_temp_w$surv, method = "kendall") # cor -0.442, p = 8.647e-07 < 0.001 ***
cor_lea_e <- cor.test(lea_temp_e$offset, lea_temp_e$surv, method = "kendall") # cor -0.434, p = 1.03e-06 < 0.001 ***
lea_cor_df <- data.frame(method = "lea_offset", cc = "mll", coast = c("both", "west", "east"), kendall = c(cor_lea$estimate, cor_lea_w$estimate, cor_lea_e$estimate), pval = c(cor_lea$p.value, cor_lea_w$p.value, cor_lea_e$p.value))
######

## rda
######
# full
surv_full_binlog_rda <- surv_all_go2 %>%
  filter(method == "rda_offset" & cc == "full") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, Full): \nBinary Survival, Logistic Regression", 
       x = "rda offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_full_binlog_rda

# mll
surv_mll_binlog_rda <- surv_all_go2 %>%
  filter(method == "rda_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, MLL): \nBinary Survival, Logistic Regression", 
       x = "rda offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_rda

# correlation
rda_temp <- plotting_survgo_all %>% filter(cc == "mll", method == "rda_offset")
rda_temp_w <- rda_temp %>% filter(trt %in% c("TempControl-21psu", "TempWarm-16psu"))
rda_temp_e <- rda_temp %>% filter(trt %in% c("TempControl-7psu", "TempWarm-5psu"))

cor_rda <- cor.test(rda_temp$offset, rda_temp$surv, method = "kendall") # cor 0.174, p = 0.005 < 0.01 **
cor_rda_w <- cor.test(rda_temp_w$offset, rda_temp_w$surv, method = "kendall") # cor 0.237, p = 0.008 < 0.01 **
cor_rda_e <- cor.test(rda_temp_e$offset, rda_temp_e$surv, method = "kendall") # cor -0.416, p = 2.78e-06 < 0.001 ***

rda_cor_df <- data.frame(method = "rda_offset", cc = "mll", coast = c("both", "west", "east"), kendall = c(cor_rda$estimate, cor_rda_w$estimate, cor_rda_e$estimate), pval = c(cor_rda$p.value, cor_rda_w$p.value, cor_rda_e$p.value))
######

## gf
#####
# full
surv_full_binlog_gf <- surv_all_go2 %>%
  filter(method == "gf_offset" & cc == "full") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, Full): \nBinary Survival, Logistic Regression", 
       x = "gf offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_full_binlog_gf

# mll
surv_mll_binlog_gf <- surv_all_go2 %>%
  filter(method == "gf_offset" & cc == "mll") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, MLL): \nBinary Survival, Logistic Regression", 
       x = "gf offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_mll_binlog_gf

# correlation
gf_temp <- plotting_survgo_all %>% filter(cc == "mll", method == "gf_offset")
gf_temp_w <- gf_temp %>% filter(trt %in% c("TempControl-21psu", "TempWarm-16psu"))
gf_temp_e <- gf_temp %>% filter(trt %in% c("TempControl-7psu", "TempWarm-5psu"))

cor_gf <- cor.test(gf_temp$offset, gf_temp$surv, method = "kendall") # cor -0.183, p = 0.003 < 0.01 **
cor_gf_w <- cor.test(gf_temp_w$offset, gf_temp_w$surv, method = "kendall") # cor -0.432, p = 1.466e-06 < 0.001 ***
cor_gf_e <- cor.test(gf_temp_e$offset, gf_temp_e$surv, method = "kendall") # cor -0.452, p = 3.67e-07 < 0.001 ***

gf_cor_df <- data.frame(method = "gf_offset", cc = "mll", coast = c("both", "west", "east"), kendall = c(cor_gf$estimate, cor_gf_w$estimate, cor_gf_e$estimate), pval = c(cor_gf$p.value, cor_gf_w$p.value, cor_gf_e$p.value))
######

## env_dist
###########
# full
surv_binlog_env_dist <- surv_all_go2 %>%
  filter(method == "env_dist" & cc == "full") %>%
  ggplot(aes(x = offset, y = isAlive, fill = fct_reorder(pop, pop_order))) + # shape = fct_reorder(pop_coast, pop_coast_ord))
  geom_point(size = 4, shape = 21) + 
  geom_smooth(aes(col = fct_reorder(pop, pop_order)), method = "glm", method.args = list(family = "binomial"), se = F) + 
  labs(title = "Current-Future Fitness vs Environmental Distance: \nBinary Survival, Logistic Regression", 
       x = "Env. Dist.", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = pop_colors) +
  scale_colour_manual(name = "Population",
                    values = pop_colors) +
  scale_shape_manual(name = "Coast",
                    values = c(21, 24),
                    labels = c("West Coast", "East Coast")) +
  guides(fill = guide_legend(override.aes = list(shape = 21))) +
  theme_minimal()
surv_binlog_env_dist

# correlation
envdist_temp <- plotting_survgo_all %>% filter(cc == "mll", method == "env_dist")
envdist_temp_w <- envdist_temp %>% filter(trt %in% c("TempControl-21psu", "TempWarm-16psu"))
envdist_temp_e <- envdist_temp %>% filter(trt %in% c("TempControl-7psu", "TempWarm-5psu"))

cor_envdist <- cor.test(envdist_temp$offset, envdist_temp$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
cor_envdist_w <- cor.test(envdist_temp_w$offset, envdist_temp_w$surv, method = "kendall") # cor -0.442, p = 8.647e-07 < 0.001 ***
cor_envdist_e <- cor.test(envdist_temp_e$offset, envdist_temp_e$surv, method = "kendall") # cor -0.434, p = 1.03e-06 < 0.001 ***

envdist_cor_df <- data.frame(method = "env_dist", cc = "mll", coast = c("both", "west", "east"), kendall = c(cor_gf$estimate, cor_gf_w$estimate, cor_gf_e$estimate), pval = c(cor_gf$p.value, cor_gf_w$p.value, cor_gf_e$p.value))
###########

# just future
#############
lea_temp2 <- plotting_survgo_all %>% filter(cc == "mll", method == "lea_offset" & trt %in% c("TempWarm-16psu", "TempWarm-5psu"))
lea_temp3 <- lea_temp2 %>% filter(trt == "TempWarm-16psu")
lea_temp4 <- lea_temp2 %>% filter(trt == "TempWarm-5psu")
lea_cor_fut <- cor.test(lea_temp2$offset, lea_temp2$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
lea_cor_fut_w <- cor.test(lea_temp3$offset, lea_temp3$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
lea_cor_fut_e <- cor.test(lea_temp4$offset, lea_temp4$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
lea_cor_fut_df <- data.frame(method = "lea_offset", cc = "mll", coast = c("both", "west", "east"), kendall = c(lea_cor_fut$estimate, lea_cor_fut_w$estimate, lea_cor_fut_e$estimate), pval = c(lea_cor_fut$p.value, lea_cor_fut_w$p.value, lea_cor_fut_e$p.value))

rda_temp2 <- plotting_survgo_all %>% filter(cc == "mll", method == "rda_offset" & trt %in% c("TempWarm-16psu", "TempWarm-5psu"))
rda_temp3 <- rda_temp2 %>% filter(trt == "TempWarm-16psu")
rda_temp4 <- rda_temp2 %>% filter(trt == "TempWarm-5psu")
rda_cor_fut <- cor.test(rda_temp2$offset, rda_temp2$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
rda_cor_fut_w <- cor.test(rda_temp3$offset, rda_temp3$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
rda_cor_fut_e <- cor.test(rda_temp4$offset, rda_temp4$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
rda_cor_fut_df <- data.frame(method = "rda_offset", cc = "mll", coast = c("both", "west", "east"), kendall = c(rda_cor_fut$estimate, rda_cor_fut_w$estimate, rda_cor_fut_e$estimate), pval = c(rda_cor_fut$p.value, rda_cor_fut_w$p.value, rda_cor_fut_e$p.value))

gf_temp2 <- plotting_survgo_all %>% filter(cc == "mll", method == "gf_offset" & trt %in% c("TempWarm-16psu", "TempWarm-5psu"))
gf_temp3 <- gf_temp2 %>% filter(trt == "TempWarm-16psu")
gf_temp4 <- gf_temp2 %>% filter(trt == "TempWarm-5psu")
gf_cor_fut <- cor.test(gf_temp2$offset, gf_temp2$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
gf_cor_fut_w <- cor.test(gf_temp3$offset, gf_temp3$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
gf_cor_fut_e <- cor.test(gf_temp4$offset, gf_temp4$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
gf_cor_fut_df <- data.frame(method = "gf_offset", cc = "mll", coast = c("both", "west", "east"), kendall = c(gf_cor_fut$estimate, gf_cor_fut_w$estimate, gf_cor_fut_e$estimate), pval = c(gf_cor_fut$p.value, gf_cor_fut_w$p.value, gf_cor_fut_e$p.value))

envdist_temp2 <- plotting_survgo_all %>% filter(cc == "mll", method == "env_dist" & trt %in% c("TempWarm-16psu", "TempWarm-5psu"))
envdist_temp3 <- plotting_survgo_all %>% filter(cc == "mll", method == "env_dist" & trt == "TempWarm-16psu")
envdist_temp4 <- plotting_survgo_all %>% filter(cc == "mll", method == "env_dist" & trt == "TempWarm-5psu")
envdist_cor_fut <- cor.test(envdist_temp2$offset, envdist_temp2$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
envdist_cor_fut_w <- cor.test(envdist_temp3$offset, envdist_temp3$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
envdist_cor_fut_e <- cor.test(envdist_temp4$offset, envdist_temp4$surv, method = "kendall") # cor -0.216, p = 0.0004079 < 0.001 ***
envdist_cor_fut_df <- data.frame(method = "env_dist", cc = "mll", coast = c("both", "west", "east"), kendall = c(envdist_cor_fut$estimate, envdist_cor_fut_w$estimate, envdist_cor_fut_e$estimate), pval = c(envdist_cor_fut$p.value, envdist_cor_fut_w$p.value, envdist_cor_fut_e$p.value))
#############

# corrs
#######
corrs_binsurv <- rbind(lea_cor_df, rda_cor_df, gf_cor_df, envdist_cor_df) %>% mutate(surv_coast = ifelse(coast == "both", "Survival", ifelse(coast == "west", "Survival (West)", "Survival (East)"))) # %>% pivot_wider(names_from = coast, values_from = c(kendall, pval))

corrs_binsurv_fut <- rbind(lea_cor_fut_df, rda_cor_fut_df, gf_cor_fut_df, envdist_cor_fut_df) %>% mutate(surv_coast = ifelse(coast == "both", "Survival", ifelse(coast == "west", "Survival (West)", "Survival (East)")))

surv_corplot_bin <- corrs_binsurv %>% 
  ggplot(aes(x = method, y = surv_coast, fill = kendall)) +
  geom_tile(color = "black") +
  geom_text(aes(label = round(kendall, 3)), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  labs(title = "(A) Correlation between Survival and \nGenomic Offsets", x = NULL, y = NULL, fill = "Kendall's\nTau") +
  theme_classic()
surv_corplot_bin

surv_corplot_bin_fut <- corrs_binsurv_fut %>% 
  ggplot(aes(x = method, y = surv_coast, fill = kendall)) +
  geom_tile(color = "black") +
  geom_text(aes(label = round(kendall, 3)), color = "black", size = 4) +
  scale_fill_gradient2(low = "#4c9220", mid = "white", high = "#c5247d", midpoint = 0, limits = c(-1, 1), na.value = "lightgray", guide = guide_colourbar(reverse = TRUE)) +
  scale_x_discrete(labels = c("lea_offset" = "LEA", "rda_offset" = "RDA", "gf_offset" = "GF", "env_dist" = "Environmental \nDistance")) +
  labs(title = "(B) Correlation between Survival and \nGenomic Offsets (Future Treatments Only)", x = NULL, y = NULL, fill = "Kendall's\nTau") +
  theme_classic()
surv_corplot_bin_fut
#######

# export
########
library(officer)
library(rvg)

# survival regressions
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_mll_binlog_lea), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_mll_binlog_rda), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_mll_binlog_gf), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_binlog_env_dist), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))

print(doc, target = "results/figures/GO_surv_binary_climchg.pptx")

# survival correlations
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_corplot_bin), location = ph_location(left = 1, top = 1, width = 5.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_corplot_bin_fut), location = ph_location(left = 1, top = 1, width = 5.5, height = 4))

print(doc, target = "results/figures/GO_surv_corr_climchg.pptx")

# corrs
write.csv(corrs_binsurv, paste0("results/GO_compare/corrs_bin_surv_", Sys.Date(), ".csv"))
write.csv(corrs_binsurv_fut, paste0("results/GO_compare/corrs_bin_surv_fut_", Sys.Date(), ".csv"))
########
```
