---
title: "Eelgrass GO Fitness Comparisons"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Eelgrass/BalticEelgrass/")

# packages
library(car)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(stringr)

# data
# exp fit
fit_full <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
surv <- read.csv("data/experiment/survival_dat_trt_pop.csv")[,-1]
grow <- read.csv("data/experiment/fit_growth_dat.csv")[,-1]

# go
lea <- read.csv("results/GO_results/LEA_offsets_exptrt_wide_2025-12-03.csv")[,-1]
rda <- read.csv("results/GO_results/RDA_offsets_exptrt_wide_2026-01-02.csv")[,-1]
gf <- read.csv("results/GO_results/GF_offsets_exptrt_wide_2025-12-03.csv")[,-1]
```

## Genomic Offset - Fitness Offset Comparisons

This script compares genomic offset against fitness metrics in a "common garden" (across genotypes) and "current future" (across environments) comparison. 


### GO Data Processing

These data are already in a matrix format, but I will still make a long form combined df for plotting. 

```{r}
head(lea)
head(rda)
head(gf)

# long versions
# lea
lea_long <- lea %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "lea_offset")

# rda
rda_long <- rda %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "rda_offset")

# gf
gf_long <- gf %>% pivot_longer(
  cols = c(BAR, BJO, HOG, HOR, KAL, KUR, VAT, VIK),
  names_to = "pop",
  values_to = "gf_offset")

# put everything together
go1 <- merge(lea_long, rda_long, by = c("trt", "cc", "pop"))
go_wide <- merge(go1, gf_long[gf_long$mod == "geno", c("trt","pop","cc","gf_offset")], by = c("trt", "cc", "pop"))

# long version of all
go_long <- go_wide %>% pivot_longer(
  cols = c("lea_offset","rda_offset","gf_offset"),
  names_to = "method",
  values_to = "offset")
```

### Experimental Data

```{r}
# survival
head(surv)
dim(surv)
surv$trt <- gsub("_", "-", surv$trt)
fit_full$isAlive <- ifelse(fit_full$isAlive == TRUE, 1, 0)

# surv only fitness, all plants
surv_all <- fit_full[,c("pop", "trt", "tank_ID", "isAlive")]
surv_all$trt <- gsub("_", "-", surv_all$trt)
  
# also a format with survival in each tank/pop/trt
surv_tanks <- fit_full %>% 
  dplyr::select(pop, trt, tank_ID, isAlive) %>%
  group_by(pop, tank_ID, trt) %>%
  summarize(surv = mean(isAlive),
            surv_se = sd(isAlive)/length(isAlive))

# growth
head(grow)
grow_trim <- grow[,c("trt", "pop", "pop_trt", "tank_ID", "length_growth", "length_std_change", "weight_change", "weight_std_change", "above_below_weight_ratio")]
grow_trim$trt <- gsub("_","-",grow_trim$trt)

# some pop/trt level data
grow_pop_trt <- grow_trim %>% 
  group_by(pop, trt) %>% 
  summarize(avg_length_growth = mean(length_growth, na.rm = T),
    avg_length_std_change = mean(length_std_change, na.rm = T),
    avg_weight_change = mean(weight_change, na.rm = T),
    avg_weight_std_change = mean(weight_std_change, na.rm = T),
    avg_above_below_weight_ratio = mean(above_below_weight_ratio, na.rm = T))
```

### GO v Fitness

#### Data Merge

```{r}
# combine fitness and offset data
surv_go <- merge(go_wide, surv, by = c("pop", "trt"))
grow_go <- merge(go_wide, grow_trim, by = c("pop", "trt"))
surv_all_go <- merge(go_long, surv_all, by = c("pop", "trt"))

# add surv col to surv_all_go from isAlive for consistency in corr analysis
surv_all_go$surv <- surv_all_go$isAlive

# recode method
surv_all_go$method <- str_split_fixed(surv_all_go$method, "_", 2)[,1]

# numeric tank coding
tanks_df <- data.frame(tank_ID = levels(as.factor(surv_all_go$tank_ID)),
                       tank_num = 1:length(levels(as.factor(surv_all_go$tank_ID))))
surv_all_go <- merge(surv_all_go, tanks_df, by = "tank_ID")
surv_all_go$tank_num <- as.numeric(surv_all_go$tank_num)

# full vs mll (plotting)
surv_go_full <- surv_go[surv_go$cc == "full",]
surv_go_mll <- surv_go[surv_go$cc == "mll",]
surv_all_go_full <- surv_all_go[surv_all_go$cc == "full",]
surv_all_go_mll <- surv_all_go[surv_all_go$cc == "mll",]
grow_go_full <- grow_go[grow_go$cc == "full",]
grow_go_mll <- grow_go[grow_go$cc == "mll",]

# long versions
surv_go_long <- surv_go %>% pivot_longer(names_to = "method", cols = c("lea_offset", "rda_offset", "gf_offset"), values_to = "offset") %>% as.data.frame()
surv_go_long$method <- str_split_fixed(surv_go_long$method, "_", 2)[,1]
grow_go_long <- grow_go %>% pivot_longer(names_to = "method", cols = c("lea_offset", "rda_offset", "gf_offset"), values_to = "offset") %>% as.data.frame()
grow_go_long$method <- str_split_fixed(grow_go_long$method, "_", 2)[,1]
# surv all is already long format

# for correlation/signif testing 
trts <- levels(as.factor(surv_go$trt))
pops <- levels(as.factor(surv_go$pop))
```

#### Correlations

```{r}
## SURVIVAL
###########
# function for correlations both common garden/current future offset
surv_corr <- function(df, method, gen_clone, fit_offset) {
  # subset data
  df_subset <- df[df$method == method & df$cc == gen_clone,]
  
  if(fit_offset == "cg") {
    # make df
    corrs_cg <- data.frame(method = NA, trt = NA, cc = NA, kendall_surv = NA, pval_surv = NA)
    
    # common garden
    for (i in 1:length(trts)){
      # subset
      trt_temp <- trts[i]
      temp <- df_subset[df_subset$trt == trt_temp,]
    
      # metadata
      corrs_cg[i,]$method <- method
      corrs_cg[i,]$cc <- gen_clone
      corrs_cg[i,]$trt <- trt_temp  
      
      # correlation
      test <- cor.test(temp$surv, temp$offset, method = "kendall")
      corrs_cg[i,]$kendall_surv <- test$estimate
      corrs_cg[i,]$pval_surv <- test$p.value
    }
    return(corrs_cg)
  } else if (fit_offset == "cf") {
    # make df
    corrs_cf <- data.frame(method = NA, pop = NA, cc = NA, kendall_surv = NA, pval_surv = NA)
    
    # current future
    for (i in 1:length(pops)){
      # subset
      pop_temp <- pops[i]
      temp <- df_subset[df_subset$pop == pop_temp,]
    
      # metadata
      corrs_cf[i,]$method <- method
      corrs_cf[i,]$cc <- gen_clone
      corrs_cf[i,]$pop <- pop_temp  
      
      # correlation
      test <- cor.test(temp$surv, temp$offset, method = "kendall")
      corrs_cf[i,]$kendall_surv <- test$estimate
      corrs_cf[i,]$pval_surv <- test$p.value
    }
    return(corrs_cf)
  } else {
    print("Please choose cg for common garden or cf for current-future")
  }
}

surv_corr_cg <- rbind(surv_corr(surv_go_long, "lea", "full", "cg"),
                      surv_corr(surv_go_long, "rda", "full", "cg"),
                      surv_corr(surv_go_long, "gf", "full", "cg"),
                      surv_corr(surv_go_long, "lea", "mll", "cg"),
                      surv_corr(surv_go_long, "rda", "mll", "cg"),
                      surv_corr(surv_go_long, "gf", "mll", "cg"))

surv_corr_all_cf <- rbind(surv_corr(surv_all_go, "lea", "full", "cg"),
                      surv_corr(surv_all_go, "rda", "full", "cg"),
                      surv_corr(surv_all_go, "gf", "full", "cg"),
                      surv_corr(surv_all_go, "lea", "mll", "cg"),
                      surv_corr(surv_all_go, "rda", "mll", "cg"),
                      surv_corr(surv_all_go, "gf", "mll", "cg"))

surv_corr_cf <- rbind(surv_corr(surv_go_long, "lea", "full", "cf"),
                      surv_corr(surv_go_long, "rda", "full", "cf"),
                      surv_corr(surv_go_long, "gf", "full", "cf"),
                      surv_corr(surv_go_long, "lea", "mll", "cf"),
                      surv_corr(surv_go_long, "rda", "mll", "cf"),
                      surv_corr(surv_go_long, "gf", "mll", "cf"))

surv_corr_all_cf <- rbind(surv_corr(surv_all_go, "lea", "full", "cf"),
                      surv_corr(surv_all_go, "rda", "full", "cf"),
                      surv_corr(surv_all_go, "gf", "full", "cf"),
                      surv_corr(surv_all_go, "lea", "mll", "cf"),
                      surv_corr(surv_all_go, "rda", "mll", "cf"),
                      surv_corr(surv_all_go, "gf", "mll", "cf"))
###########

## GROWTH
#########
grow_corr <- function(df, method, gen_clone, fit_offset) {
  # subset data
  df_subset <- df[df$method == method & df$cc == gen_clone,]
  
  if(fit_offset == "cg") {
    # make df
    corrs_cg <- data.frame(method = NA, trt = NA, cc = NA, kendall_len = NA, pval_len = NA, kendall_weigh = NA, pval_weigh = NA)
    
    # common garden
    for (i in 1:length(trts)){
      # subset
      trt_temp <- trts[i]
      temp <- df_subset[df_subset$trt == trt_temp,]
    
      # metadata
      corrs_cg[i,]$method <- method
      corrs_cg[i,]$cc <- gen_clone
      corrs_cg[i,]$trt <- trt_temp  
      
      # correlation
      test_len <- cor.test(temp$length_std_change, temp$offset, method = "kendall")
      corrs_cg[i,]$kendall_len <- test_len$estimate
      corrs_cg[i,]$pval_len <- test_len$p.value
      
      test_weigh <- cor.test(temp$weight_std_change, temp$offset, method = "kendall")
      corrs_cg[i,]$kendall_weigh <- test_weigh$estimate
      corrs_cg[i,]$pval_weigh <- test_weigh$p.value
    }
    return(corrs_cg)
    
  } else if (fit_offset == "cf") {
    # make df
    corrs_cf <- data.frame(method = NA, pop = NA, cc = NA, kendall_len = NA, pval_len = NA, kendall_weigh = NA, pval_weigh = NA)
    
    # current future
    for (i in 1:length(pops)){
      # subset
      pop_temp <- pops[i]
      temp <- df_subset[df_subset$pop == pop_temp,]
    
      # metadata
      corrs_cf[i,]$method <- method
      corrs_cf[i,]$cc <- gen_clone
      corrs_cf[i,]$pop <- pop_temp  
      
      # correlation
      test_len <- cor.test(temp$length_std_change, temp$offset, method = "kendall")
      corrs_cf[i,]$kendall_len <- test_len$estimate
      corrs_cf[i,]$pval_len <- test_len$p.value
      
      test_weigh <- cor.test(temp$weight_std_change, temp$offset, method = "kendall")
      corrs_cf[i,]$kendall_weigh <- test_weigh$estimate
      corrs_cf[i,]$pval_weigh <- test_weigh$p.value
    }
    return(corrs_cf)
    
  } else {
    print("Please choose cg for common garden or cf for current-future")
  }
}

grow_corr_cg <- rbind(grow_corr(grow_go_long, "lea", "full", "cg"),
                      grow_corr(grow_go_long, "rda", "full", "cg"),
                      grow_corr(grow_go_long, "gf", "full", "cg"),
                      grow_corr(grow_go_long, "lea", "mll", "cg"),
                      grow_corr(grow_go_long, "rda", "mll", "cg"),
                      grow_corr(grow_go_long, "gf", "mll", "cg"))

grow_corr_cf <- rbind(grow_corr(grow_go_long, "lea", "full", "cf"),
                      grow_corr(grow_go_long, "rda", "full", "cf"),
                      grow_corr(grow_go_long, "gf", "full", "cf"),
                      grow_corr(grow_go_long, "lea", "mll", "cf"),
                      grow_corr(grow_go_long, "rda", "mll", "cf"),
                      grow_corr(grow_go_long, "gf", "mll", "cf"))
#########
```

#### Modeling

```{r}
# survival
##########
# quick checks + fixes
hist(surv_all_go$surv) # we will use binary outcomes first

## lea
# model full (no cc)
surv_mod_lea_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "lea",])
summary(surv_mod_lea_full) 
#plot(surv_mod_lea_full) # these look pretty bad
Anova(surv_mod_lea_full) # signif

# model mll (cc)
surv_mod_lea_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "lea",])
summary(surv_mod_lea_mll) 
#plot(surv_mod_lea_mll) # these look pretty bad
Anova(surv_mod_lea_mll) # signif

# now the tank level pop/trt
surv_mod_lea_poptrt_full <- glm(surv ~ lea_offset,
               family = binomial(link = "logit"),
               data = surv_go)
summary(surv_mod_lea_poptrt_full) 
#plot(surv_mod_lea_full)
Anova(surv_mod_lea_poptrt_full) # signif

## rda
# model full (no cc)
surv_mod_rda_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "rda",])
summary(surv_mod_rda_full) 
#plot(surv_mod_rda_full) # these look pretty bad
Anova(surv_mod_rda_full) # signif

# model mll (cc)
surv_mod_rda_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "rda",])
summary(surv_mod_rda_mll) 
#plot(surv_mod_rda_mll) # these look pretty bad
Anova(surv_mod_rda_mll) # signif

# now the tank level pop/trt
surv_mod_rda_poptrt_full <- glm(surv ~ rda_offset,
               family = binomial(link = "logit"),
               data = surv_go)
summary(surv_mod_rda_poptrt_full) 
#plot(surv_mod_rda_full)
Anova(surv_mod_rda_poptrt_full) # signif

## gf
# model full (no cc)
surv_mod_gf_full <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_full[surv_all_go_full$method == "gf",])
summary(surv_mod_gf_full) 
#plot(surv_mod_gf_full) # these look pretty bad
Anova(surv_mod_gf_full) # signif

# model mll (cc)
surv_mod_gf_mll <- glm(isAlive ~ offset,
               family = binomial(link = "logit"),
               data = surv_all_go_mll[surv_all_go_mll$method == "gf",])
summary(surv_mod_gf_mll) 
#plot(surv_mod_gf_mll) # these look pretty bad
Anova(surv_mod_gf_mll) # signif

# now the tank level pop/trt
surv_mod_gf_poptrt_full <- glm(surv ~ gf_offset,
               family = binomial(link = "logit"),
               data = surv_go_full)
summary(surv_mod_gf_poptrt_full) 
#plot(surv_mod_gf_poptrt_full) # better than binary model
Anova(surv_mod_gf_poptrt_full) # signif

surv_mod_gf_poptrt_mll <- glm(surv ~ gf_offset,
               family = binomial(link = "logit"),
               data = surv_go_mll)
summary(surv_mod_gf_poptrt_mll) 
#plot(surv_mod_gf_poptrt_mll)
Anova(surv_mod_gf_poptrt_mll) # signif
##########

# length
########
# quick checks + fixes
hist(grow_go$length_std_change) 
hist(log(grow_go$length_std_change)) # log-normal distribution

## lea
# model full (no cc)
len_mod_lea_full <- lm(length_std_change ~ log(lea_offset), data = grow_go_full)
summary(len_mod_lea_full) 
#plot(len_mod_lea_full) # diverges pretty significantly on right tail
Anova(len_mod_lea_full) # signif

# model mll (cc)
len_mod_lea_mll <- lm(length_std_change ~ log(lea_offset), data = grow_go_mll)
summary(len_mod_lea_mll) 
#plot(len_mod_lea_mll) # diverges pretty significantly on right tail
Anova(len_mod_lea_mll) # signif

## rda
# model full (no cc)
len_mod_rda_full <- lm(length_std_change ~ log(rda_offset), data = grow_go_full)
summary(len_mod_rda_full) 
#plot(len_mod_rda_full) # diverges pretty significantly on right tail
Anova(len_mod_rda_full) # signif

# model mll (cc)
len_mod_rda_mll <- lm(length_std_change ~ log(rda_offset), data = grow_go_mll)
summary(len_mod_rda_mll) 
#plot(len_mod_rda_mll) # diverges pretty significantly on right tail
Anova(len_mod_rda_mll) # signif

## gf
# model full (no cc)
len_mod_gf_full <- lm(length_std_change ~ log(gf_offset), data = grow_go_full)
summary(len_mod_gf_full) 
#plot(len_mod_gf_full) # diverges pretty significantly on right tail
Anova(len_mod_gf_full) # signif

# model mll (cc)
len_mod_gf_mll <- lm(length_std_change ~ log(gf_offset), data = grow_go_mll)
summary(len_mod_gf_mll) 
#plot(len_mod_gf_mll) # diverges pretty significantly on right tail
Anova(len_mod_gf_mll) # signif
########

# weight
########
# quick checks + fixes
hist(grow_go$weight_std_change) # normal distribution :D

## lea
# model full (no cc)
weigh_mod_lea_full <- lm(weight_std_change ~ lea_offset, data = grow_go_full)
summary(weigh_mod_lea_full) 
#plot(weigh_mod_lea_full) # diverges pretty significantly on right tail
Anova(weigh_mod_lea_full) # signif

# model mll (cc)
weigh_mod_lea_mll <- lm(weight_std_change ~ lea_offset, data = grow_go_mll)
summary(weigh_mod_lea_mll) 
#plot(weigh_mod_lea_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_lea_mll) # signif

## rda
# model full (no cc)
weigh_mod_rda_full <- lm(weight_std_change ~ (rda_offset), data = grow_go_full)
summary(weigh_mod_rda_full) 
#plot(weigh_mod_rda_full) # diverges pretty significantly on right tail
Anova(weigh_mod_rda_full) # signif

# model mll (cc)
weigh_mod_rda_mll <- lm(weight_std_change ~ (rda_offset), data = grow_go_mll)
summary(weigh_mod_rda_mll) 
#plot(weigh_mod_rda_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_rda_mll) # signif

## gf
# model full (no cc)
weigh_mod_gf_full <- lm(weight_std_change ~ (gf_offset), data = grow_go_full)
summary(weigh_mod_gf_full) 
#plot(weigh_mod_gf_full) # diverges pretty significantly on right tail
Anova(weigh_mod_gf_full) # signif

# model mll (cc)
weigh_mod_gf_mll <- lm(weight_std_change ~ (gf_offset), data = grow_go_mll)
summary(weigh_mod_gf_mll) 
#plot(weigh_mod_gf_mll) # diverges pretty significantly on right tail
Anova(weigh_mod_gf_mll) # signif
########
```

### Plotting

#### Survival

```{r}
## lea
######
# means by pop/treatment
# common garden: LEA
surv_go_full %>% ggplot(aes(x = lea_offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, Full): Survival",
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
surv_go_mll %>% ggplot(aes(x = lea_offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, MLL): Survival",
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: LEA
surv_go_full %>% ggplot(aes(x = lea_offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, Full): Survival", 
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
surv_go_mll %>% ggplot(aes(x = lea_offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, MLL): Survival", 
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()

# binary outcome, full data
# common garden: LEA
surv_all_go_full %>% 
  filter(method == "lea") %>%
  ggplot(aes(x = offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, Full): \nBinary Survival",
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
surv_all_go_mll %>% 
  filter(method == "lea") %>%
  ggplot(aes(x = offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, MLL): \nBinary Survival",
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: LEA
surv_all_go_full %>%
  filter(method == "lea") %>%
  ggplot(aes(x = offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, Full): \nBinary Survival", 
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
surv_all_go_mll %>% 
  filter(method == "lea") %>%
  ggplot(aes(x = offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, MLL): \nBinary Survival", 
       x = "LEA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######

## rda
######
# common garden: RDA
surv_go_full %>% 
  ggplot(aes(x = rda_offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, Full): Survival",
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
surv_go_mll %>% ggplot(aes(x = rda_offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, MLL): Survival",
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: RDA
surv_go_full %>% ggplot(aes(x = rda_offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, Full): Survival", 
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
surv_go_mll %>% ggplot(aes(x = rda_offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, MLL): Survival", 
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()

# binary
# common garden: RDA
surv_all_go_full %>% 
  filter(method == "rda") %>% 
  ggplot(aes(x = offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, Full): \nBinary Survival",
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
surv_all_go_mll %>% 
  filter(method == "rda") %>% 
  ggplot(aes(x = offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, MLL): \nBinary Survival",
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: RDA
surv_all_go_full %>% 
  filter(method == "rda") %>% 
  ggplot(aes(x = offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, Full): \nBinary Survival", 
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
surv_all_go_mll %>% 
  filter(method == "rda") %>% 
  ggplot(aes(x = offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, MLL): \nBinary Survival", 
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######

## gf
######
# proportions
# common garden: GF
surv_go_full %>% ggplot(aes(x = gf_offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, Full): Survival",
       x = "GF offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
surv_go_mll %>% ggplot(aes(x = gf_offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, MLL): Survival",
       x = "GF offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: GF
surv_go_full %>% ggplot(aes(x = gf_offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, Full): Survival", 
       x = "GF offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
surv_go_mll %>% ggplot(aes(x = gf_offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, MLL): Survival", 
       x = "GF offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()

# binary
# common garden: GF
surv_all_go_full %>% 
  filter(method == "gf") %>% 
  ggplot(aes(x = offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, Full): \nBinary Survival",
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
surv_all_go_mll %>% 
  filter(method == "gf") %>% 
  ggplot(aes(x = offset, y = surv, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, MLL): \nBinary Survival",
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: GF
surv_all_go_full %>% 
  filter(method == "gf") %>% 
  ggplot(aes(x = offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, Full): \nBinary Survival", 
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
surv_all_go_mll %>% 
  filter(method == "gf") %>% 
  ggplot(aes(x = offset, y = surv, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, MLL): \nBinary Survival", 
       x = "RDA offset", y = "Survival") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######
```

#### Length

Length

```{r}
## lea
######
# common garden: LEA
grow_go_full %>% ggplot(aes(x = lea_offset, y = length_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, Full): Length",
       x = "LEA offset", y = "Length") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = lea_offset, y = length_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, MLL): Length",
       x = "LEA offset", y = "Length") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: LEA
grow_go_full %>% ggplot(aes(x = lea_offset, y = length_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, Full): Length", 
       x = "LEA offset", y = "Length") +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = lea_offset, y = length_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, MLL): Length", 
       x = "LEA offset", y = "Length") +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######

## rda
######
# common garden: RDA
grow_go_full %>% ggplot(aes(x = rda_offset, y = length_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, Full): Length",
       x = "RDA offset", y = "Length") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = rda_offset, y = length_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, MLL): Length",
       x = "RDA offset", y = "Length") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: RDA
grow_go_full %>% ggplot(aes(x = rda_offset, y = length_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, Full): Length", 
       x = "RDA offset", y = "Length") +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = rda_offset, y = length_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, MLL): Length", 
       x = "RDA offset", y = "Length") +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######

## gf
######
# common garden: GF
grow_go_full %>% ggplot(aes(x = gf_offset, y = length_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, Full): Length",
       x = "GF offset", y = "Length") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = gf_offset, y = length_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, MLL): Length",
       x = "GF offset", y = "Length") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: GF
grow_go_full %>% ggplot(aes(x = gf_offset, y = length_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, Full): Length", 
       x = "GF offset", y = "Length") +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = gf_offset, y = length_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, MLL): Length", 
       x = "GF offset", y = "Length") +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######
```

#### Weight

```{r}
## lea
######
# common garden: LEA
grow_go_full %>% ggplot(aes(x = lea_offset, y = weight_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, Full): Weight",
       x = "LEA offset", y = "Weight") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = lea_offset, y = weight_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (LEA, MLL): Weight",
       x = "LEA offset", y = "Weight") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: LEA
grow_go_full %>% ggplot(aes(x = lea_offset, y = weight_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, Full): Weight", 
       x = "LEA offset", y = "Weight") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = lea_offset, y = weight_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (LEA, MLL): Weight", 
       x = "LEA offset", y = "Weight") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######

## rda
######
# common garden: RDA
grow_go_full %>% ggplot(aes(x = rda_offset, y = weight_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, Full): Weight",
       x = "RDA offset", y = "Weight") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = rda_offset, y = weight_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (RDA, MLL): Weight",
       x = "RDA offset", y = "Weight") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: RDA
grow_go_full %>% ggplot(aes(x = rda_offset, y = weight_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, Full): Weight", 
       x = "RDA offset", y = "Weight") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = rda_offset, y = weight_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (RDA, MLL): Weight", 
       x = "RDA offset", y = "Weight") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######

## gf
######
# common garden: GF
grow_go_full %>% ggplot(aes(x = gf_offset, y = weight_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, Full): Weight",
       x = "GF offset", y = "Weight") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = gf_offset, y = weight_std_change, fill = trt)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = trt), method = "lm", se = T) + 
  labs(title = "Common Garden Fitness vs Genomic Offset (GF, MLL): Weight",
       x = "GF offset", y = "Weight") +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  scale_colour_manual(name = "Treatment",
                    values = c("#FFA203","#FE691E","deeppink","deeppink4"),
                    labels = c("Current W Coast", "Current E Coast",
                               "Future W Coast", "Future E Coast")) +
  theme_minimal()

# current future: GF
grow_go_full %>% ggplot(aes(x = gf_offset, y = weight_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, Full): Weight", 
       x = "GF offset", y = "Weight") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
grow_go_mll %>% ggplot(aes(x = gf_offset, y = weight_std_change, fill = pop)) +
  geom_point(shape = 21, size = 4) + 
  geom_smooth(aes(col = pop), method = "lm", se = F) + 
  labs(title = "Current-Future Fitness vs Genomic Offset (GF, MLL): Weight", 
       x = "GF offset", y = "Weight") +
  scale_fill_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_colour_manual(name = "Population",
                    values = c("#023e8a","#0096c7","#90e0ef","#caf0f8", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  theme_minimal()
######
```

