---
title: "eelgrass_env_process.Rmd"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Environmental Data Processing

This script is for processing environmental data for genomic offset calculation of the project "VR_Zostera_genomic_offset." It includes processing of the seascape data used to train offset models, as well as processing of the mesocosm environmental data used to make final offset predictions. The copernicus data is processed already in the "ncDat_process.R" script. Hobo logger data is also pre-processed in the "exp_hobo_loggers.R" script. Scripts for environmental distance calculation and plotting of environmental trends are named "env_dist.R" and "env_plot.R" respectively. 

### Setup

```{r}
# packages
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(LEA) # landscape genomics
library(psych) # environmental data cleaning
library(vegan) # environmental data cleaning
library(tidyr) # data cleaning
library(dplyr) # data cleaning
library(stringr) # data cleaning
library(vcfR) # for vcf
library(ggplot2) # plotting
library(viridis) # plotting
library(forcats) # order data
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ncdf4) # for nc data
library(raster) # for nc data
library(pdist) # for env dists

# setwd for data
setwd("~/Documents/GitHub/BalticEelgrass")

# data
seascape_dat <- read.csv("data/EnvDat/summary_env_data_meadows_GEA.csv") # seascape env data
exp_dat_cop <- read.csv("data/EnvDat/complete_expSites_copernicus_2025-12-01.csv")[,-1]
exp_dat <- read_excel("data/EnvDat/Camille_env_data_experiment.xlsx")
trt_dat <- read.csv("data/EnvDat/trt_env_dat_2025-12-31.csv")
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx") # seascape individuals data
eelgrass_vcf_MLL <- read.vcfR("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.vcf") # read in vcf eelgrass mll
indiv_trt_env_dat <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-12-31.csv")[,-1]
```

### Pre-processing 

Start with formatting

```{r}
# seascape env data reads in funky, change column names and remove first row
colnames(seascape_dat) <- seascape_dat[1,]
seascape_dat <- seascape_dat[-1,]

# the column names have spaces in them, change names to be more code friendly
colnames(seascape_dat) <- c("region","site_full","pop","long","lat","median_temp","median_sal","temp_8.5","sal_8.5","max_temp","min_sal")

# also deal with lat/long
seascape_dat$long <- as.numeric(gsub(",", ".", gsub("\\.", "", seascape_dat$long)))
seascape_dat$lat <- as.numeric(gsub(",", ".", gsub("\\.", "", seascape_dat$lat)))

# and ensure consistency in site abbreviations by getting rid of funky characters
seascape_dat$pop[seascape_dat$pop == "ÅLA"] <- "ALA"
seascape_dat$pop[seascape_dat$pop == "BÅD"] <- "BAD"

# exp dat also has these column name issues
exp_dat2 <- exp_dat[,1:10]
colnames(exp_dat2) <- c("site_full","pop","long","lat","median_temp","temp_4.5","temp_8.5","median_sal","sal_4.5","sal_8.5")

# trim experimental individual data to just relevant columns
exp_ind <- exp_inds[,c("bagnum_num", "tank_ID", "DNA_label", "new_label", "temp_full", "sal_full")]
```

Now that I have environmental data for all my sites in a coding-friendly format, I need to figure out which of these variables should be included in my model. To do so, I will make sure that I don't have any variables that are too correlated. I should also separate out the "future" environmental data from the dataset I want to use for training offsets. Here I will start with the seascape data and the experimental data (i.e. everything from various environmental models)

```{r}
# start by splitting current and future data
curr_sea <- seascape_dat[,c("region","site_full","pop","long","lat","median_temp","max_temp","median_sal","min_sal")]
fut_sea <- seascape_dat[,c("region","site_full","pop","long","lat","temp_8.5","sal_8.5")]

# let's also keep a df with the population data only
pop_sea <- curr_sea[,c("region","site_full","pop","long","lat")]

# now I need to look at variable correlation
pairs.panels(curr_sea[,6:9]) # perfect, nothing is too correlated! probably Steffi already did this step :)

# merge the experimental pop data
(ex_pop_dat <- merge(exp_dat2[,c("site_full","pop","long","lat","median_temp","median_sal")], exp_dat_cop[,c("pop","temp_max","sal_min")], by = c("pop")))

colnames(ex_pop_dat)[colnames(ex_pop_dat) == "temp_max"] <- "max_temp"
colnames(ex_pop_dat)[colnames(ex_pop_dat) == "sal_min"] <- "min_sal"
```

I need to adjust the environmental data for the site-of-origin for experimental pops according to the procedure outlined [here](docs.google.com/document/d/1J54wsnxtk5UFdv5PXPYA2CnWi6fN4DvJ/view). Specifically, I will adjust to use data from symphony +4C for all mean temps and -5psu for W coast mean salinities.

```{r}
# adjust temp
ex_pop_dat$temp_median_adjust <- (ex_pop_dat$median_temp + 4)

# add a coast identifier
ex_pop_dat$coast <- ifelse(ex_pop_dat$pop %in% c("VIK","VAT","HOG","BAR"), "West", "East")

# adjust salinity
ex_pop_dat$sal_median_adjust <- ifelse(ex_pop_dat$coast == "West", ex_pop_dat$median_sal - 5, ex_pop_dat$median_sal)

# trim
exp_pop_env <- ex_pop_dat[,c("pop","coast","long","lat","temp_median_adjust","max_temp","sal_median_adjust","min_sal")]

# colnames
colnames(exp_pop_env)[colnames(exp_pop_env) == "temp_median_adjust"] <- "median_temp"
colnames(exp_pop_env)[colnames(exp_pop_env) == "sal_median_adjust"] <- "median_sal"
```

Now I will prep the treatment data to be comparable to the seascape and experimental source data

```{r}
# reduce treatment data to comparable dataset
trt_dat_red <- trt_dat[,c("trt","median_temp","max_temp","median_sal","min_sal")]

# line up all the mins/maxes/averages for both temp and salinity
trt_env <- trt_dat_red[,c("median_temp","max_temp","median_sal","min_sal")]
rownames(trt_env) <- trt_dat_red$trt
trt_env$df <- "treatment"
exp_env <- exp_pop_env[,c("median_temp","max_temp","median_sal","min_sal")]
rownames(exp_env) <- exp_pop_env$pop
exp_env$df <- "experiment"
sea_env <- seascape_dat2[,c("median_temp","max_temp","median_sal","min_sal")]
rownames(sea_env) <- seascape_dat2$pop
sea_env$df <- "seascape"
colnames(trt_env) <- colnames(exp_env) <- colnames(sea_env) <- c("median_temp","max_temp","median_sal","min_sal","df")
all_env <- rbind(trt_env, exp_env, sea_env) # this is not what will actually be used for scaling, just for single df to look at correlations

# # ensure proper data types
all_env$median_temp <- as.numeric(all_env$median_temp)
all_env$max_temp <- as.numeric(all_env$max_temp)
all_env$median_sal <- as.numeric(all_env$median_sal)
all_env$min_sal <- as.numeric(all_env$min_sal)

# look at correlations
pairs.panels(all_env[,1:4]) # median and min sal are at 0.7 corr, so right on the brink
```

For the actual genomic offset predictions, I will use individual level data such that each row in my seascape and experimental datasets will line up with either the genomic data (seascape) or the fitness data (experiment)

```{r}
# deal with individual data
indiv_list <- as.data.frame(indiv_list)
pops_split <- stringr::str_split_fixed(indiv_list$`Indv #`, "-", 2) # split Indv # to have population in a seperate column
indiv_list_full <- cbind(indiv_list, pops_split) # add this info back to original file
colnames(indiv_list_full) <- c("mll_no","indv_no","pop","pop_no")
table(indiv_list_full$pop) # between 4 and 20 individuals in each pop - most 17+ inds

# also put together an mll dataset
mll_split <- stringr::str_split_fixed(colnames(eelgrass_vcf_MLL@gt), "-", 3)
mll_inds <- mll_split[-1,3]
indiv_list_mll <- indiv_list_full[indiv_list_full$indv_no %in% mll_inds,]

# subset just relevant env dat from trts, rename to clarify origin
trt_env_ind <- indiv_trt_env_dat %>%
  dplyr::select(pop, trt, tank_ID, bagnum_text, median_temp, max_temp, median_sal, min_sal) %>%
  mutate(df = "treatment")

# expand relevant data to individual level
seascape_dat_expand <- merge(indiv_list_full, seascape_dat, by = "pop") %>%
  dplyr::select(pop, mll_no, median_temp, max_temp, median_sal, min_sal) %>%
  arrange(mll_no)

exp_env$pop <- rownames(exp_env)
exp_dat_expand <- merge(trt_env_ind[,c("pop","bagnum_text")], exp_env, by = "pop") %>%
  arrange(bagnum_text)

# combine
sea_env_prep <- seascape_dat_expand %>%
  dplyr::select(median_temp, max_temp, median_sal, min_sal) %>%
  mutate_all(as.numeric) %>%
  mutate(df = "seascape", order = 1:nrow(seascape_dat_expand)) %>%
  as.data.frame()

exp_env_prep <- exp_dat_expand %>%
  as.data.frame() %>%
  dplyr::select(median_temp, max_temp, median_sal, min_sal, df) %>%
  mutate(order = exp_dat_expand$bagnum_text) %>%
  arrange(order) %>%
  as.data.frame()

trt_env_prep <- trt_env_ind %>%
  dplyr::select(median_temp, max_temp, median_sal, min_sal, bagnum_text, df) %>%
  rename(order = bagnum_text) %>%
  as.data.frame()
```

### Data Scaling

Now I will scale the data for the seascape training data, save those scaling factors, and also apply them to the experimental source and mesocosm treatment data. 

```{r}
# use expanded dat
head(seascape_dat_expand)
seascape_dat_expand$mll_no <- as.factor(seascape_dat_expand$mll_no)
seascape_dat_expand[,c("median_temp","max_temp","median_sal","min_sal")] <- apply(seascape_dat_expand[,c("median_temp","max_temp","median_sal","min_sal")], 2, as.numeric)
summary(seascape_dat_expand)

# scale and save scaling factors
seascape_scaled_ind <- scale(seascape_dat_expand[,c("median_temp","max_temp","median_sal","min_sal")])
center_sea <- attr(seascape_scaled_ind,"scaled:center")
scale_sea <- attr(seascape_scaled_ind,"scaled:scale")

seascape_scaled_ind1 <- cbind(seascape_dat_expand, seascape_scaled_ind)

# apply to ind lvl exp dat and trt dat
exp_scaled_ind <- scale(exp_dat_expand[,c("median_temp","max_temp","median_sal","min_sal")], center = center_sea, scale = scale_sea)
trt_scaled_ind <- scale(trt_env_ind[,c("median_temp","max_temp","median_sal","min_sal")], center = center_sea, scale = scale_sea)

# reattach ind dat
colnames(seascape_scaled_ind) <- colnames(exp_scaled_ind) <- colnames(trt_scaled_ind) <- c("median_temp_std","max_temp_std","median_sal_std","min_sal_std")
seascape_scaled_ind1 <- cbind(seascape_dat_expand, seascape_scaled_ind)
seascape_scaled_ind1$df <- "seascape"
exp_scaled_ind1 <- cbind(exp_dat_expand, exp_scaled_ind)
trt_scaled_ind1 <- cbind(trt_env_ind, trt_scaled_ind)

# mll data for seascape
seascape_scaled_ind1_mll <- seascape_scaled_ind1[!duplicated(seascape_scaled_ind1$mll_no),]

# output dat
write.csv(seascape_scaled_ind1, paste0("data/EnvDat/full_seascape_scaled_ind_only", Sys.Date(), ".csv"))
write.csv(seascape_scaled_ind1_mll, paste0("data/EnvDat/mll_seascape_scaled_ind_only", Sys.Date(), ".csv"))
write.csv(exp_scaled_ind1, paste0("data/EnvDat/exp_scaled_ind_only", Sys.Date(), ".csv"))
write.csv(trt_scaled_ind1, paste0("data/EnvDat/trt_scaled_ind_only", Sys.Date(), ".csv"))
```

Now I can format this data for the different GO methods...

### Formatting for LEA

LEA requires environmental data be formatted as a .env file type.

```{r}
mat1 <- as.matrix(seascape_scaled_ind1[,c("median_temp_std", "max_temp_std", "median_sal_std", "min_sal_std")])
mat2 <- as.matrix(seascape_scaled_ind1_mll[,c("median_temp_std", "max_temp_std", "median_sal_std", "min_sal_std")])
mat3 <- as.matrix(exp_scaled_ind1[,c("median_temp_std", "max_temp_std", "median_sal_std", "min_sal_std")])
mat4 <- as.matrix(trt_scaled_ind1[,c("median_temp_std", "max_temp_std", "median_sal_std", "min_sal_std")])

colnames(mat1) <- colnames(mat2) <- colnames(mat3) <- colnames(mat4) <- c("median_temp", "max_temp", "median_sal", "min_sal")

write.env(mat1, paste0("data/EnvDat/full_seascapeEnv_indlvl_only_", Sys.Date(), ".env"))
write.env(mat2, paste0("data/EnvDat/mll_seascapeEnv_indlvl_only_", Sys.Date(), ".env"))
write.env(mat3, paste0("data/EnvDat/full_expEnv_indlvl_only_", Sys.Date(), ".env"))
write.env(mat4, paste0("data/EnvDat/full_trtEnv_indlvl_only_", Sys.Date(), ".env"))
```

### Formatting for RDA & GF

RDA needs information in the exact format I already had it in, so no need to do any further processing - just know that I will use the csv for RDA. When I go to run RDA, I will likely also want to correct for population structure by including PCs from a PCA, but this will come when the analysis is actually run (since that requires genetic information). GF requires environmental data in the same format as RDA, just a regular table format. 
