---
title: "EelgrassEnvDataProcessing"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Eelgrass/BalticEelgrass/data")
```

# Environmental Data Processing

This script is for processing environmental data for genomic offset calculation of the project "VR_Zostera_genomic_offset." It includes processing of the seascape data used to train offset models, as well as processing of the mesocosm environmental data used to make final offset predictions. 

## Seascape Data

This data was already compiled by Steffi Ries & Marlene Jahnke to run GEAs. I am just processing it to be in formats compatible with LEA, RDA, and GF. More details about how this data was processed to find these averages can be found [here](https://drive.google.com/drive/u/0/folders/1CNLeIIg3IUlEkCgfIC6P8k9tesJ-BNtq).

### Setup & Formatting

I have already set my working directory in the setup chunk to be the "data" folder in the larger eelgrass directory. Now I need to include certain libraries and load relevant data files. 

```{r}
# packages
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(LEA) # landscape genomics
library(psych) # environmental data cleaning
library(vegan) # environmental data cleaning
library(tidyr) # data cleaning
library(dplyr) # data cleaning
library(stringr) # data cleaning
library(vcfR) # for vcf
library(ggplot2) # plotting
library(viridis) # plotting
library(forcats) # order data
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ncdf4) # for nc data
library(raster) # for nc data

# data
seascape_dat <- read.csv("data/EnvDat/summary_env_data_meadows_GEA.csv") # seascape env data
exp_dat <- read.csv("data/EnvDat/complete_expSites_copernicus_2025-08-25.csv")[,-1]
exp_dat_cop <- read_excel("data/EnvDat/Camille_env_data_experiment.xlsx")
trt_dat <- read.csv("data/EnvDat/pop_trt_env_dat_2025-09-12.csv")
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx") # seascape individuals data
eelgrass_vcf_MLL <- read.vcfR("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.vcf")# read in vcf eelgrass mll
```

Data is now loaded, but I need make some formatting changes so that it's more usable.

```{r}
# seascape env data reads in funky, change column names and remove first row
colnames(seascape_dat) <- seascape_dat[1,]
seascape_dat <- seascape_dat[-1,]

# the column names have spaces in them, change names to be more code friendly
colnames(seascape_dat) <- c("region","site_full","Pop","long","lat","temp_curr","sal_curr","temp_8.5","sal_8.5","max_temp_curr","min_sal_curr")

# also deal with lat/long
seascape_dat$long <- as.numeric(gsub(",", ".", gsub("\\.", "", seascape_dat$long)))
seascape_dat$lat <- as.numeric(gsub(",", ".", gsub("\\.", "", seascape_dat$lat)))

# and ensure consistency in site abbreviations by getting rid of funky characters
seascape_dat$Pop[seascape_dat$Pop == "ÅLA"] <- "ALA"
seascape_dat$Pop[seascape_dat$Pop == "BÅD"] <- "BAD"
```


### Pre-processing

Now that I have environmental data for all my sites in a coding-friendly format, I need to figure out which of these variables should be included in my model. To do so, I will make sure that I don't have any variables that are too correlated. I should also separate out the "future" environmental data from the dataset I want to use for training offsets. Here I will start with the seascape data and the experimental data (i.e. everything from various environmental models)

```{r}
# start by splitting current and future data
curr_sea <- seascape_dat[,c("region","site_full","Pop","long","lat","temp_curr","max_temp_curr","sal_curr","min_sal_curr")]
fut_sea <- seascape_dat[,c("region","site_full","Pop","long","lat","temp_8.5","sal_8.5")]

# let's also keep a df with the population data only
pop_sea <- curr_sea[,c("region","site_full","Pop","long","lat")]
#write.csv(pop,"data/seascape_data/population_data.csv")

# now I need to look at variable correlation
pairs.panels(curr_sea[,6:9]) # perfect, nothing is too correlated! probably Steffi already did this step :)

# now look at the sites for experimental pops
exp_dat
names(exp_dat)[names(exp_dat) == "site_abbrev"] <- "Pop"
exp_dat_cop
colnames(exp_dat_cop) <- c("site_full","Pop","long","lat","temp_mean_cop","temp_4.5","temp_8.5","sal_mean_cop","sal_4.5","sal_8.5","temp_4.5_med_chg","temp_8.5_med_chg","sal_4.5_med_chg","sal_8.5_med_chg","temp_4.5_max_chg","temp_8.5_max_chg","sal_4.5_max_chg","sal_8.5_max_chg")

# merge the experimental pop data
(ex_pop_dat <- merge(exp_dat_cop, exp_dat, by = c("Pop")))
```

I need to adjust the environmental data for the site-of-origin for experimental pops according to the procedure outlined [here](docs.google.com/document/d/1J54wsnxtk5UFdv5PXPYA2CnWi6fN4DvJ/view). Specifically, I will adjust to use data from symphony +4C for all mean temps and -5psu for W coast mean salinities.

```{r}
# first trim to just relevant variables
exp_pop_dat_trim <- ex_pop_dat[,c("Pop","long","lat","temp_mean_cop","temp_min","temp_max","sal_mean_cop","sal_min","sal_max")]

# adjust temp
exp_pop_dat_trim$temp_mean_adjust <- (exp_pop_dat_trim$temp_mean_cop + 4)

# add a coast identifier
exp_pop_dat_trim$coast <- ifelse(exp_pop_dat_trim$Pop %in% c("VIK","VAT","HOG","BAR"), "West", "East")

# adjust salinity
exp_pop_dat_trim$sal_mean_adjust <- ifelse(exp_pop_dat_trim$coast == "West", exp_pop_dat_trim$sal_mean_cop - 5, exp_pop_dat_trim$sal_mean_cop)

# look at the correlations between variables
pairs.panels(exp_pop_dat_trim[,c("temp_min", "temp_max", "temp_mean_adjust", "sal_min", "sal_max", "sal_mean_adjust")])

# remove overly correlated variables
exp_pop_env <- exp_pop_dat_trim[,c("Pop","coast","lat","long","temp_mean_adjust","temp_max","sal_min","sal_mean_adjust")]

# rename variables to match seascape
colnames(exp_pop_env) <- c("Pop","coast","lat","long","tempcurr","max_temp_curr","sal_curr","min_sal_curr")
```


Before I can scale any environmental data, I need to also add in the environments I want to predict to (i.e. experimental site of origin, experimental treatments). I will thus process all of these data as well, then scale and standardize everything. 

Pre-processing of these data has happened in other scripts - the experimental treatment data is prepped in "exp_hobo_loggers.R", and the experimental site of origin data is prepped in "ncDat_explore.R".

```{r}
# inspect our datasets
trt_dat
exp_pop_env
seascape_dat

# make sure values look consistent across dfs
exp_pop_env
seascape_dat[seascape_dat$Pop %in% c("BAR", "BJO", "HOG", "HOR", "KAL", "KUR", "VAT", "VIK"),]

# remove duplicates in exp and sea dfs
seascape_dat2 <- seascape_dat[!(seascape_dat$Pop %in% c("BAR", "BJO", "HOG", "HOR", "KAL", "KUR", "VAT", "VIK")),]

# trt dat doesn't need to be on pop level yet
trt_dat_red <- trt_dat[!duplicated(trt_dat$trt),c("trt","avg_temp","max_temp","avg_sal","min_sal")]

# line up all the mins/maxes/averages for both temp and salinity
trt_env <- trt_dat_red[,c("avg_temp","max_temp","avg_sal","min_sal")]
trt_env$df <- "treatment"
exp_env <- exp_pop_env[,c("tempcurr","max_temp_curr","sal_curr","min_sal_curr")]
exp_env$df <- "experiment"
sea_env <- seascape_dat2[,c("temp_curr","max_temp_curr","sal_curr","min_sal_curr")]
sea_env$df <- "seascape"
colnames(trt_env) <- colnames(exp_env) <- colnames(sea_env) <- c("temp","max_temp","sal","min_sal","df")
all_env <- rbind(trt_env, exp_env, sea_env)

# ensure proper data types
all_env$temp <- as.numeric(all_env$temp)
all_env$max_temp <- as.numeric(all_env$max_temp)
all_env$sal <- as.numeric(all_env$sal)
all_env$min_sal <- as.numeric(all_env$min_sal)

# now scale and standardize
std <- scale(all_env[,c("temp","max_temp","sal","min_sal")])
colnames(std) <-  c("temp_scaled","max_temp_scaled","sal_scaled","min_sal_scaled")
all_env2 <- cbind(all_env, std)

# put together with original dfs
trt_scaled <- cbind(trt_dat_red, all_env2[all_env2$df == "treatment", c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])
exp_scaled <- cbind(exp_pop_env, all_env2[all_env2$df == "experiment", c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])

# to get seascape need to add back in the exp sites
exp_scaled0 <- exp_scaled[exp_scaled$Pop %in% c(seascape_dat$Pop),]
sea_exp <- seascape_dat[seascape_dat$Pop %in% exp_dat$Pop, ]
sea_scaled0 <- merge(sea_exp, exp_scaled0[,c("Pop", "temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")], by = c("Pop"))
sea_scaled1 <- cbind(seascape_dat2, all_env2[all_env2$df == "seascape", c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])
sea_scaled <- rbind(sea_scaled1, sea_scaled0)

# save all these datasets
write.csv(trt_scaled, paste0("data/EnvDat/trt_scaled_envdat_", Sys.Date(), ".csv"), row.names = F)
write.csv(exp_scaled, paste0("data/EnvDat/exp_scaled_envdat_", Sys.Date(), ".csv"), row.names = F)
write.csv(sea_scaled, paste0("data/EnvDat/sea_scaled_envdat_", Sys.Date(), ".csv"), row.names = F)
```


Everything is looking good! I want to make sure now that my environmental data will match my genomic data by having one row in my seascape environmental dataset for each individual in my genomic data set. To do so, I will need to process the individual data and pair this with my environmental data.

```{r}
# deal with individual data
indiv_list <- as.data.frame(indiv_list)
pops_split <- stringr::str_split_fixed(indiv_list$`Indv #`, "-", 2) # split Indv # to have population in a seperate column
indiv_list_full <- cbind(indiv_list, pops_split) # add this info back to original file
colnames(indiv_list_full) <- c("MLL_no","Indv_no","Pop","Pop_no")
table(indiv_list_full$Pop) # between 4 and 20 individuals in each pop - most 17+ inds

# also put together an mll dataset
mll_split <- stringr::str_split_fixed(colnames(eelgrass_vcf_MLL@gt), "-", 3)
mll_inds <- mll_split[-1,3]
indiv_list_mll <- indiv_list_full[indiv_list_full$Indv_no %in% mll_inds,]

# now put this together with the environmental dataset
head(sea_scaled)
head(indiv_list_full)
head(indiv_list_mll)

# oops! some of our environmental data is listed as character when I want it numeric
sea_scaled$temp_curr <- as.numeric(curr_sea$temp_curr)
sea_scaled$max_temp_curr <- as.numeric(curr_sea$max_temp_curr)
sea_scaled$sal_curr <- as.numeric(curr_sea$sal_curr)
sea_scaled$min_sal_curr <- as.numeric(curr_sea$min_sal_curr)
sea_scaled$temp_8.5 <- as.numeric(fut_sea$temp_8.5)
sea_scaled$sal_8.5 <- as.numeric(fut_sea$sal_8.5)

# I can put it together using "Pop"
full_sea_scaled <- merge(indiv_list_full, sea_scaled, by = c("Pop"))
mll_sea_scaled <- merge(indiv_list_mll, sea_scaled, by = c("Pop"))

# output for rda
write.csv(full_sea_scaled, paste0("data/EnvDat/sea_scaled_envdat_full_", Sys.Date(), ".csv"), row.names = F)
write.csv(mll_sea_scaled, paste0("data/EnvDat/sea_scaled_envdat_mll_", Sys.Date(), ".csv"), row.names = F)
```


### Formatting for LEA

LEA requires environmental data be formatted as a .env file type. 

```{r}
# prep as matrix
mat1 <- as.matrix(full_sea_scaled[,c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])
mat2 <- as.matrix(mll_sea_scaled[,c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])
mat3 <- as.matrix(exp_scaled[,c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])
mat4 <- as.matrix(trt_scaled[,c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")])

# and convert to env format
write.env(mat1, paste0("data/EnvDat/full_seascapeEnv_", Sys.Date(), ".env"))
write.env(mat2, paste0("data/EnvDat/mll_seascapeEnv_", Sys.Date(), ".env"))
write.env(mat3, paste0("data/EnvDat/expEnv_", Sys.Date(), ".env"))
write.env(mat4, paste0("data/EnvDat/trtEnv_", Sys.Date(), ".env"))
```

### Formatting for RDA & GF

RDA needs information in the exact format I already had it in, so no need to do any further processing - just know that I will use the csv for RDA. When I go to run RDA, I will likely also want to correct for population structure by including PCs from a PCA, but this will come when the analysis is actually run (since that requires genetic information). GF requires environmental data in the same format as RDA, just a regular table format. 

### Environmental Distance

I need to calculate environmental distance between the tanks and the site-of-origin for the experimental populations.

```{r}
# assign relevant data to new objs
(trt_scaled_fordist <- trt_scaled)
(exp_scaled_fordist <- exp_scaled)

# reorder populations by coast
exp_scaled_fordist <- exp_scaled_fordist[order(match(exp_scaled_fordist$Pop, c("VAT", "VIK", "HOG", "BAR", "KUR", "KAL", "HOR", "BJO"))),]

# change column names for compatibility
names(trt_scaled_fordist)[names(trt_scaled_fordist) == "trt"] <- "trtpop"
names(exp_scaled_fordist)[names(exp_scaled_fordist) == "Pop"] <- "trtpop"

# put together the data
fordist <- rbind(trt_scaled_fordist[,c("trtpop","temp_scaled","max_temp_scaled","sal_scaled","min_sal_scaled")], exp_scaled_fordist[,c("trtpop","temp_scaled","max_temp_scaled","sal_scaled","min_sal_scaled")])

# calc dists
euc_dists <- round(vegdist(fordist[,-1], method = "euclidian",
                           upper = F, diag = T), 4)
euc_dists_df <- as.data.frame(as.matrix(euc_dists))

# add pops back as colnames & rownames
rownames(euc_dists_df) <- colnames(euc_dists_df) <- fordist[,1]

# keep this df
write.csv(euc_dists_df, "data/EnvDat/euclidian_trt_exp.csv")

# distances df for plotting
euc_dists_df2 <- euc_dists_df
euc_dists_df2$pops1 <- rownames(euc_dists_df)
euc_dists_long <- euc_dists_df2 %>% 
  pivot_longer(-pops1, names_to = "pops2", values_to = "distances") %>%
  as.data.frame()
euc_dists_long

popsdf1 <- data.frame(pops = fordist$trtpop, order = 1:length(fordist$trtpop))
popsdf2 <- data.frame(pops = fordist$trtpop, order = 1:length(fordist$trtpop))
colnames(popsdf1) <- c("pops1","order1")
colnames(popsdf2) <- c("pops2","order2")
euc_dists_long_mg <- merge(popsdf1[,c("pops1","order1")], euc_dists_long, by = "pops1")
euc_dists_long_mg2 <- merge(popsdf2[,c("pops2","order2")], euc_dists_long_mg, by = "pops2")
euc_dist_trtpop <- ggplot(euc_dists_long_mg2, aes(x = fct_reorder(pops1, order1), 
                              y = fct_reorder(pops2, order2), 
                              fill = distances))+
  theme_classic()+
  geom_tile(color = "black") +
#  geom_text(aes(label = distances), color = "white", size = 4)+
  scale_fill_viridis_c(name = "Euclidean 
Distance", limits = c(0, 7.5))+
  theme(plot.title = element_text(size = 20), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top") +
  ylab("")+
  xlab("")+
  ggtitle("Environmental Distance between Sites and Treatments") +
  theme(plot.title = element_text(size = 18), 
        legend.title = element_text(size = 14),
        legend.position = "right", legend.justification = "top",
        axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))
euc_dist_trtpop
```


Also plot treatment data against experimental site of origin data

```{r}
# pop info
order_pops <- c("VAT","VIK","HOG","BAR","KUR","KAL","HOR","BJO","TempControl-21psu","TempControl-7psu","TempWarm-16psu","TempWarm-5psu")

# order pops_in_exp & set explicit colors
pops_in_df <- data.frame(poptrt = order_pops, order = 1:12, cols = c("#155d27","#25a244","#4ad66d","#b7efc5","#caf0f8","#90e0ef","#0096c7","#023e8a","#FFA203","#FE691E","deeppink","deeppink4"), coast = c(rep("West",4), rep("East",4), rep("trt",4)))

pops_in_df1 <- data.frame(poptrt = order_pops, order = 1:12, cols = c("#002800","#155d27","#25a244","#6ede8a","#90e0ef","#00b4d8","#0077b6","#03045e","#FFA203","#FE691E","deeppink","deeppink4"), coast = c(rep("West",4), rep("East",4), rep("trt",4)))

# keep this order df for other purposes
write.csv(pops_in_df, "data/experiment/poptrts_df.csv", row.names = F)

# read in copernicus time series data
copernicus_plot <- read.csv("data/EnvDat/copernicus_plotting_data.csv")
copernicus_plot$date <- as.Date(copernicus_plot$date)

copernicus_plot$temp_adjust <- copernicus_plot$temp + 4
copernicus_plot$sal_adjust <- ifelse(copernicus_plot$coast == "West", copernicus_plot$sal - 5, copernicus_plot$sal)

ggplot(copernicus_plot, aes(x = date, y = temp_adjust, 
                            fill = fct_reorder(site_name, order), 
                            col = fct_reorder(site_name, order))) +
  geom_line()+
#  geom_line(na.rm = T, aes(group = time_block)) +
  facet_grid(~ year, scales = "free_x") +
  scale_fill_manual(name = "Site",
                    values = pops_in_df1$cols,
                    labels = pops_in_df1$poptrt) +
  scale_color_manual(name = "Site",
                     values = pops_in_df1$cols,
                     labels = pops_in_df1$poptrt) +
  geom_hline(yintercept = 20.07134, col = "#FFA202", linewidth = 1) +
  geom_hline(yintercept = 25.80038, col = "#FF1393", linewidth = 1) +
  labs(title = "Temperature at Experimental Sites 2011-2021",
       x = "Date", y = "Temperature (ºC)", color = "Site", fill = "Site") +
  theme_bw()

ggplot(copernicus_plot, aes(x = date, y = sal, 
                            fill = fct_reorder(site_name, order), 
                            col = fct_reorder(site_name, order))) +
  geom_line()+
#  geom_line(na.rm = T, aes(group = time_block)) +
  facet_grid(~ year, scales = "free_x") +
  scale_fill_manual(name = "Site",
                    values = pops_in_df1$cols,
                    labels = pops_in_df1$poptrt) +
  scale_color_manual(name = "Site",
                     values = pops_in_df1$cols,
                     labels = pops_in_df1$poptrt) +
  geom_hline(yintercept = 20.07134, col = "#FFA202", linewidth = 1) +
  geom_hline(yintercept = 25.80038, col = "#FF1393", linewidth = 1) +
  labs(title = "Salinity at Experimental Sites 2011-2021",
       x = "Date", y = "Salinity (PSU)", color = "Site", fill = "Site") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1)) +
  theme_bw()

ggplot(copernicus_plot, 
       aes(x = date, y = sal, 
           fill = fct_reorder(site_name, order), 
           col = fct_reorder(site_name, order))) +
  geom_smooth() +
  scale_fill_manual(name = "Site",
                    values = pops_in_df1$cols,
                    labels = pops_in_df1$poptrt) +
  scale_color_manual(name = "Site",
                     values = pops_in_df1$cols,
                     labels = pops_in_df1$poptrt) +
  geom_hline(yintercept = 21, col = "#FFA202", linewidth = 1) +
  geom_hline(yintercept = 7, col = "#FF1393", linewidth = 1) +
  geom_hline(yintercept = 16, col = "#FE691F", linewidth = 1) +
  geom_hline(yintercept = 5, col = "#8B0A50", linewidth = 1) +
  labs(title = "Salinity at Experimental Sites 2011-2021",
       x = "Date", y = "Salinity (PSU)", color = "Site", fill = "Site") +
  theme_bw()
```


I also want violin plots, which require putting together the different raw datasets for both treatment and site of origin 

```{r}
# read in raw temperature data from tanks
tanks_dat <- read.csv("data/EnvDat/all_tanks_logger_dat.csv")[,-1]

# expand out the date time info from tanks
tanks_dat$date <- str_split_fixed(tanks_dat$datetime, " ", 2)[,1]
tanks_dat$date <- as.Date(tanks_dat$date)
tanks_dat$time <- str_split_fixed(tanks_dat$datetime, " ", 2)[,2]
tanks_dat$poptrt <- tanks_dat$trt

# look at just these data
ggplot(tanks_dat) +
  geom_violin(aes(x = trt, y = temp, fill = trt)) +
#  geom_boxplot(aes(x = trt, y = temp, fill = trt), width = 0.1) +
  scale_fill_manual(name = "Treatment",
                    values = c("#FFA202","#FF1393"),
                    labels = c("Control", "Heated")) +
  theme_bw()

# now put it with the other sites temp data
pops_in_df2 <- data.frame(poptrt = c("VAT","VIK","HOG","BAR","KUR","KAL","HOR","BJO","ctrl","heat"), order = 1:10, cols = c("#10451d","#155d27","#25a244","#6ede8a","#90e0ef","#00b4d8","#0077b6","#03045e","#FFA203","#FF1393"), coast = c(rep("West",4), rep("East",4), rep("trt", 2)))

temps_all <- rbind(tanks_dat[,c("date","poptrt","temp")], 
                   copernicus_plotting[,c("date","poptrt","temp")])

temps_plotting <- merge(temps_all, pops_in_df2, by = c("poptrt"))

# plot
ggplot(temps_plotting) +
  geom_violin(aes(x = as.factor(fct_reorder(poptrt, order)), y = temp, fill = fct_reorder(poptrt, order))) +
#  geom_boxplot(aes(x = trt, y = temp, fill = trt), width = 0.1) +
  scale_fill_manual(name = "Site",
                    values = pops_in_df2$cols,
                    labels = pops_in_df2$poptrt) +
    labs(title = "Temperature at Experimental Sites 2011-2021",
       x = "Date", y = "Temperature (ºC)", color = "Site", fill = "Site") +
  theme_bw()

# we use daily checks data for salinity
bags <- read.csv("data/EnvDat/GOEEL-eelgrass-Sweden-exp - Bags.csv")
daily_checks <- read.csv("data/EnvDat/GOEEL-eelgrass-Sweden-exp - DailyChecks.csv")

# put together the salinity data so we can look at trends by tank/treatment
sal_dat <- merge(bags, daily_checks, by = c("bagKey"))

# expand out the date time info from daily checks
sal_dat$date <- str_split_fixed(sal_dat$dailyTimestamp, " ", 2)[,1]
sal_dat$date <- as.Date(sal_dat$date)
sal_dat$time <- str_split_fixed(sal_dat$dailyTimestamp, " ", 2)[,2]
sal_dat$poptrt <- paste0(sal_dat$PopID, "_", sal_dat$trt)

# 

# plot
ggplot(sal_plotting) +
  geom_violin(aes(x = as.factor(fct_reorder(poptrt, order)), y = (sal), fill = fct_reorder(poptrt, order))) +
  scale_fill_manual(name = "Site",
                    values = pops_in_df$cols,
                    labels = pops_in_df$poptrt) +
  labs(title = "Temperature at Experimental Sites 2011-2021",
       x = "Site", y = "Salinity (PSU)", color = "Site", fill = "Site") +
  theme_bw()

# plot just sites
ggplot(copernicus_plot) +
  geom_violin(aes(x = as.factor(fct_reorder(poptrt, order)), y = as.numeric(sal), fill = fct_reorder(poptrt, order))) +
#  geom_boxplot(aes(x = fct_reorder(poptrt, order), y = sal, fill = fct_reorder(poptrt, order)), width = 0.1) +
  scale_fill_manual(name = "Site",
                    values = pops_in_df2$cols,
                    labels = pops_in_df2$poptrt) +
  labs(title = "Salinity at Experimental Sites 2011-2021",
       x = "Date", y = "Salinity (PSU)", color = "Site", fill = "Site") +
  theme_bw()
```
