---
title: "eelgrass_fitness_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Github/BalticEelgrass")

# setwd
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse)
library(psych)
library(stringr)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(performance)
library(corrplot)

# data
fit <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
env <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-10-02.csv")[,-1]
```

# Fitness Modeling

This script is to compare fitness across populations/treatments in the GO eelgrass mesocosm experiment. I may also use some site of origin environmental data to see if environmental factors explain performance differences, but this will come later (if at all).

## Setup

I need first to put together my data so that I have all the fitness and growth metrics possible.

```{r data pre-process}
# data cleaning to just relevant columns
fit_trim <- fit[,c("DNA_label", "new_label", "pop", "genet", "bagnum", "tank_ID", "sal_full", "temp_full", "trt", "nbr_leaves_start", "sideShoot", "nbr_sideShootLeaves_start", "length_longest_leaf_start", "weight_start", "IsBlackLesions_start", "nbr_leaves_end", "nbr_new_leaves", "width_leaf", "Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Nbr_new_shoots", "Nbr_leaves_new_shoot1", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Nbr_leaves_new_shoot2", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5", "below_ground_weight", "above_ground_weight", "weight_new_shoot1", "weight_new_shoot2", "total_weight", "lesions.", "isAlive", "Length_flower1", "Length_bract1")]

# binary survival column
fit_trim$isAlive <- ifelse(fit_trim$isAlive == TRUE, 1, 0)

# pop_trt factor for later significance testing
fit_trim$pop_trt <- as.factor(paste0(fit_trim$pop, "_", fit_trim$trt))

# add treatment block to both data frames for testing potential random effects
# block changes every 8 IDs - note though that salinity was independently controlled in each bag, so block may not actually appropriate descriptor of design
# note two plants were removed
table(fit_trim$genet) # gen-13 & gen-35
table(fit_trim$pop) # VAT & HOG
fit_trim_ord <- fit_trim[order(fit_trim$bagnum),]
fit_trim_ord$block <- ifelse(fit_trim_ord$bagnum <= 384, ((fit_trim_ord$bagnum - 1) %/% 8) %% 24 + 1, NA)

# ensure correct coding of relevant variables
fit_trim_ord$pop <- as.factor(fit_trim_ord$pop)
fit_trim_ord$tank_ID <- as.factor(fit_trim_ord$tank_ID)
fit_trim_ord$trt <- as.factor(fit_trim_ord$trt)
fit_trim_ord$block <- as.factor(fit_trim_ord$block)

# combine with the exp env dat
fit_trim_env <- merge(fit_trim_ord, env[,c("DNA_label", "mean_sal", "median_sal", "min_sal", "max_sal", "sd_sal", "mean_temp", "median_temp", "min_temp", "max_temp", "sd_temp", "x_coord", "y_coord", "position")], by = c("DNA_label"))

# also add in the actual bag level salinity


# length and weight growth
# for the growth analyses, we want a df with no flowering individuals and no dead individuals
fit_growth <- fit_trim_env[is.na(fit_trim_env$Length_flower1),]
fit_growth <- fit_growth[fit_growth$isAlive == 1,]

# length
fit_growth$length_growth <- rowSums(fit_growth[,c("Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5")], na.rm = T)

# percent length growth
fit_growth$length_std_change <- fit_growth$length_growth / fit_growth$length_longest_leaf_start

# weight
fit_growth$weight_change <- fit_growth$total_weight - fit_growth$weight_start

# percent weight change
fit_growth$weight_std_change <- fit_growth$weight_change / fit_growth$weight_start

# ratio of above vs below ground weight
fit_growth$above_below_weight_ratio <- fit_growth$above_ground_weight / fit_growth$below_ground_weight


# correlations between fitness metrics
# survival
fit_surv_cordf <- fit_trim[,c("IsBlackLesions_start","isAlive","nbr_leaves_start","weight_start","length_longest_leaf_start")]
pairs.panels(fit_surv_cordf)
fit_surv_cors <- cor(fit_surv_cordf)
corrplot(fit_surv_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)

# growth
fit_growth_cordf <- fit_growth[,c("IsBlackLesions_start","nbr_leaves_start","nbr_leaves_end","length_longest_leaf_start","length_growth","length_std_change","weight_change","weight_std_change","weight_start","total_weight","above_ground_weight","below_ground_weight","above_below_weight_ratio")]
pairs.panels(fit_growth_cordf)
fit_growth_cors <- cor(fit_growth_cordf, use = "complete.obs")
corrplot(fit_growth_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)
```

## Fitness Difference Modeling

Here I will compare differences among groups. I will use tank, treatment, and population as explanatory factors in my analysis.

### Survival Modeling

I will start by modeling differences in survival. To do so, I will use a binomial GLMM with binary survival data as response variable, fixed explanatory factors treatment & population, and random factor tank.

```{r survival modeling trt}
# analysis of fixed effects
anova(lm(isAlive ~ pop + trt, data = fit_trim_env)) # both significant

# start by looking for evidence of random effects
# model with no random effect
surv_glm <- glm(isAlive ~ pop * trt,
               family = binomial(link = "logit"),
               data = fit_trim_env)
plot(surv_glm)
Anova(surv_glm, type = "III")
drop1(surv_glm) # keep everything
summary(surv_glm)

# now look at residuals
resid_surv <- data.frame(residuals = residuals(surv_glm), tank_ID = fit_trim_env$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_surv, main = "Residuals of Survival GLM by Tank") # tank doesn't seem to have any huge systematic effects, some differences in variance but not mean

# # try with random effects
# surv_glmm <- glmer(isAlive ~ pop * trt + (1 | tank_ID),
#                family = binomial(link = "logit"),
#                data = fit_trim_env)
# plot(surv_glmm)
# Anova(surv_glmm, type = "III")
# drop1(surv_glmm) # keep all
# summary(surv_glmm)

# check dispersion & multicollinearity
check_overdispersion(surv_glm) # looks good

# significance testing with tukey
surv_glm_poptrt <- glm(isAlive ~ pop_trt,
                       family = binomial,
                       data = fit_trim_env)
tuk_surv <- TukeyHSD(aov(surv_glm_poptrt), conf.level = 0.95)
cld_surv <- multcompLetters4(aov(surv_glm_poptrt), tuk_surv)

# get significance letters
letters_surv_df <- data.frame(letters = as.data.frame.list(cld_surv$pop_trt), popsite = names((cld_surv$pop_trt)$Letters))
letters_surv_df_red <- letters_surv_df[,c("letters.Letters","popsite")]
colnames(letters_surv_df_red) <- c("letters","popsite")
rownames(letters_surv_df_red) <- NULL
```

Rather than using treatment directly, I will try using salinity and temperature to represent temperature. Additionally, because tank_ID is confounded with temperature, I will try using x/y coordinates as a fixed effect and avoid having random effects altogether. If I do this, I can use a GLM. 

```{r survival modeling sal/temp}
# anova
anova(lm(isAlive ~ pop + median_sal + median_temp, data = fit_trim_env)) # pop, sal, temp all significant

# start with a glm
surv_glm_env <- glm(isAlive ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp,
                family = binomial(link = "logit"),
                data = fit_trim_env) # warning: glm.fit: fitted probabilities numerically 0 or 1 occurred

# check dispersion & multicollinearity
check_overdispersion(surv_glm_env) # looks good
vif(surv_glm_env) # pop is 21.827777, pop:median_temp is 26.376966

# inspect model
summary(surv_glm_env)
plot(surv_glm_env)
Anova(surv_glm_env, type = "III") # pop, sal, temp all signif
step(surv_glm_env)

# check for random effects
resid_surv_env <- data.frame(residuals = residuals(surv_glmm_env), tank_ID = fit_trim_env$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_surv_env, main = "Residuals of Survival GLM by Tank") # tank effects look minimal, but some amount of randomness - try a glmm

# try a glmm
surv_glmm_env <- glmer(isAlive ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp + (1|tank_ID),
                family = binomial(link = "logit"),
                data = fit_trim_env)
# Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge with max|grad| = 0.0119148 (tol = 0.002, component 1)
# Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model is nearly unidentifiable: large eigenvalue ratio
#  - Rescale variables?

# check dispersion & multicollinearity
check_overdispersion(surv_glmm_env) # looks good
vif(surv_glmm_env) # similar numbers to the glm

# inspect model
summary(surv_glmm_env)
plot(surv_glmm_env)
Anova(surv_glmm_env, type = "III") # sal, temp signif - pop and any pop interactions not significant
drop1(surv_glmm_env) # remove pop:median_temp

# removing pop temp interaction
surv_glmm_env.1 <- glmer(isAlive ~ pop + median_sal + median_temp + pop:median_sal + (1|tank_ID),
                family = binomial(link = "logit"),
                data = fit_trim_env) 
# boundary (singular) fit: see help('isSingular')

# check dispersion & multicollinearity
check_overdispersion(surv_glmm_env.1) # looks good
vif(surv_glmm_env.1) # looks good

# inspect model
summary(surv_glmm_env.1)
plot(surv_glmm_env.1)
Anova(surv_glmm_env.1, type = "III") # intercept, sal, temp significant
drop1(surv_glmm_env.1) # remove pop:median_sal

# removing pop sal interaction, this is solely an additive model
surv_glmm_env.2 <- glmer(isAlive ~ pop + median_sal + median_temp + (1|tank_ID),
                family = binomial(link = "logit"),
                data = fit_trim_env) 
# boundary (singular) fit: see help('isSingular')

# check dispersion & multicollinearity
check_overdispersion(surv_glmm_env.2) # looks good
vif(surv_glmm_env.2) # looks good

# inspect model
summary(surv_glmm_env.2)
plot(surv_glmm_env.2)
Anova(surv_glmm_env.2, type = "III") # everything significant
drop1(surv_glmm_env.2) # final model
```

### Growth Modeling

I will now model differences in growth, namely changes in length and changes in weight. I no longer need to use a GLM, I can use an LMM since I am not working with a binary response variable. 

Notes on this dataset: The dataset used in this modeling approach excludes individuals that died. The reason for this is that there is no biologically relevant way to code this into the model - if I code the dead individuals as having no length or weight change, they will appear more fit than many other plants that lost length/biomass, and thus had negative length/weight changes. 

I will start by building models for length change (standardized) using treatment, then environmental variables, to explain differences in length change. 

```{r length modeling standardized}
# analysis of fixed effects
anova(lm(length_std_change ~ pop + trt, data = fit_growth)) # both significant

# build a model
growlenstd_lmm <- lmer(length_std_change ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(growlenstd_lmm)
Anova(growlenstd_lmm, type = "III") # all signif except trt * pop
drop1(growlenstd_lmm) # drop interaction
summary(growlenstd_lmm)

growlenstd_lmm.1 <- lmer(length_std_change ~ pop + trt + (1 | tank_ID), data = fit_growth)
plot(growlenstd_lmm.1)
Anova(growlenstd_lmm.1, type = "III") # all signif
drop1(growlenstd_lmm.1) # keep all
summary(growlenstd_lmm.1)

# look for random effects in residuals
resid_growlenstd <- data.frame(residuals = residuals(growlenstd_lmm.1), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_growlenstd, main = "Residuals of Length Growth LMM by Tank") # more variance by tank too, but nothing significantly different

# check dispersion & multicollinearity
check_overdispersion(growlenstd_lmm.1) # looks good
vif(growlenstd_lmm.1) # suggests no collinearity, everything is around 1
```

I will look at separating out salinity and temperature for these models as well. I will only do this for the standardized data.

```{r length modeling sal/temp}
# analysis of fixed effects
anova(lm(length_std_change ~ pop + median_sal + median_temp, data = fit_growth)) # pop and temp signif, not sal

# glm without random effects, but with interactions
len_saltemp_lmm <- lmer(length_std_change ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp, data = fit_growth)

# now try building an additive model
growlen_saltemp_lmm.1 <- lmer(length_growth ~ pop + median_sal + median_temp + (1 | tank_ID), data = fit_growth) # boundary (singular) fit: see help('isSingular')
plot(growlen_saltemp_lmm.1)
Anova(growlen_saltemp_lmm.1, type = "III") # all significant
drop1(growlen_saltemp_lmm.1) # keep everything
summary(growlen_saltemp_lmm.1)

# residuals
resid_saltemp_lmm.1_growlen <- data.frame(residuals = residuals(growlen_saltemp_lmm.1), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_saltemp_lmm.1_growlen, main = "Residuals of Length Growth LMM by Tank") # more variance by tank too, but nothing significantly different

# check dispersion & multicollinearity
check_overdispersion(growlen_saltemp_lmm.1) # looks good
vif(growlen_saltemp_lmm.1) # suggests no collinearity, everything is around 1
```

Now look into weight. I am testing only standardized weight change.

```{r}
# std weight model
growweightstd_lmm <- lmer(weight_std_change ~ pop + trt + (1 | tank_ID), data = fit_growth)
plot(growweightstd_lmm)
Anova(growweightstd_lmm, type = "III")
drop1(growweightstd_lmm)
summary(growweightstd_lmm)

# check residuals
resid_growweightstd <- data.frame(residuals = residuals(growweightstd_lmm), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_growweightstd, main = "Residuals of Weight Change (std) LMM by Tank") # some tank variance in mean, more in variance of residuals themselves

# check dispersion & multicollinearity
check_overdispersion(growweightstd_lmm) # looks good
vif(growweightstd_lmm) # suggests no collinearity, everything is around 1
```

I will now look at this same data using salinity/temp in the place of treatment. Again, I will only do this for standardized data.

```{r weight modeling sal/temp}
# analysis of fixed effects
anova(lm(weight_std_change ~ pop + median_sal + median_temp, data = fit_growth)) # pop and temp signif, not sal

# now try building an additive model
growweight_saltemp_lmm <- lmer(weight_std_change ~ pop + median_sal + median_temp + (1 | tank_ID), data = fit_growth) # boundary (singular) fit: see help('isSingular')
plot(growweight_saltemp_lmm)
Anova(growweight_saltemp_lmm, type = "III") # median sal not significant
drop1(growweight_saltemp_lmm) # keep everything
summary(growweight_saltemp_lmm)

# residuals
resid_saltemp_lmm_growweight <- data.frame(residuals = residuals(growweight_saltemp_lmm), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_saltemp_lmm_growweight, main = "Residuals of Weight Change (std) LMM by Tank") # more variance by tank too, but nothing significantly different

# check dispersion & multicollinearity
check_overdispersion(resid_saltemp_lmm_growweight) # looks good
vif(resid_saltemp_lmm_growweight) # suggests no collinearity, everything is around 1
```


I will also try these growth models without random factors, using a basic linear model. Note that a linear model is inappropriate for survival analyses because the response data is binary.

```{r growth models lm}
# # growth models
# lm_len <- lm(length_growth ~ pop + median_sal + median_temp + block, data = fit_growth)
# lm_weight <- lm(weight_change ~ pop + median_sal + median_temp + block, data = fit_growth)
# 
# # anova of these models
# anova(lm_len) # pop, sal, temp all significant
# anova(lm_weight) # temp significant
# 
# # check multicollinearity
# vif(lm_len) # suggests no collinearity
# vif(lm_weight) # suggests no collinearity
# 
# # inspect models
# # length
# summary(lm_len)
# plot(lm_len)
# Anova(lm_len, type = "III") # intercept, pop, sal, temp all signif
# 
# # weight
# summary(lm_weight)
# plot(lm_weight)
# Anova(lm_weight, type = "III") # intercept and temp signif
# 
# # model selection
# # length
# step(lm_len) # keeps pop, median sal, median temp
# 
# # weight
# step(lm_weight) # keeps just median temp
```

