---
title: "eelgrass_fitness_offsets_clim"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Github/BalticEelgrass")

# setwd
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse) # for data cleaning
library(stringr) # for data cleaning
library(dplyr) # for data cleaning
library(ggplot2) # for plotting
library(raster) # for environmental pixels
library(viridis) # plotting
library(sf) # mapping
library(s2) # mapping
library(rnaturalearth) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(stars) # mapping
library(ggpubr) # plotting

# data
cf_off <- read.csv("results/fit_results/climate_change_fitness_offsets_2026-05-28.csv")[,-1]
sites <- read.csv("data/experiment/eelgrass_exp_sites.csv")
poptrts_df <- read.csv("data/experiment/poptrts_df.csv")
```

# Plotting Fitness Offsets

This script plots current-future fitness offsets from W & E coast conditions to future W & E coast conditions, respectively. Fitness offsets are calculated in the script "eelgrass_fitness_offsets.Rmd"

## Data Prep

Combine data

```{r}
# check data
head(cf_off)
head(sites)

# rename
sites$pop <- sites$site_abbrev

# combine
clim_offs <- merge(sites, cf_off, by = "pop")
```

## Plotting

### Maps

Start by visualizing these fitness offsets as maps

I'm going to start off making "climate change" maps, which visualize the offset between current/future W and current/future E on the same plot.

First prep the plotting range.

```{r}
# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))
```

#### Survival Maps

```{r}
# relevant data
surv_cf <- clim_offs %>% filter(metric == "surv")

# mapping
# absolute survival
###################
surv_abs_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = abs_future), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Survival in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Survival")
surv_abs_map
###################

# relative survival
###################
surv_rel_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = CF_relative), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Common Future Relative Survival \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nSurvival")
surv_rel_map
###################

# abs surv offset
#################
surv_abs_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = CF_abs_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Common Future Survival Offset \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Fitness Offset")
surv_abs_off_map
#################

# abs surv offset
#################
surv_rel_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = CF_rel_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Common Future Relative Survival \nOffset in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nFitness Offset")
surv_rel_off_map
#################

# combine plots
###############
surv_cf_maps <- ggarrange(plotlist = c(surv_abs_map, surv_rel_map, surv_abs_off_map, surv_rel_off_map))
###############
```

#### Length Plots

```{r}
# relevant data
len_cf <- clim_offs %>% filter(metric == "len")

# mapping
# absolute length
#################
len_abs_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = abs_future), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,2)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,2)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Length Change in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Length Change")
len_abs_map
###################

# relative length
#################
len_rel_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = CF_relative), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Common Future Relative Length Change \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nLength Change")
len_rel_map
#################

# abs len offset
################
len_abs_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = CF_abs_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-2,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-2,0)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Common Future Length Offset \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Fitness Offset")
len_abs_off_map
################

# rel len offset
################
len_rel_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = CF_rel_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Common Future Relative Length \nOffset in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nFitness Offset")
len_rel_off_map
################

# combine plots
###############
len_cf_maps <- ggarrange(plotlist = c(len_abs_map, len_rel_map, len_abs_off_map, len_rel_off_map))
###############
```

#### Weight Plots

```{r}
# relevant data
weigh_cf <- clim_offs %>% filter(metric == "weigh")

# mapping
# absolute weight chg
###################
weigh_abs_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = abs_future), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Weight Change in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Weight Change")
weigh_abs_map
###################

# relative weight chg
###################
weigh_rel_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = CF_relative), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-12,6)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-12,6)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Common Future Relative Weight Change \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nWeight Change")
weigh_rel_map
###################

# abs weigh offset
#################
weigh_abs_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = CF_abs_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Common Future Weight Change Offset \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Fitness Offset")
weigh_abs_off_map
#################

# abs weigh offset
#################
weigh_rel_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = CF_rel_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-13,5)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-13,5)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Common Future Relative Weight Change \nOffset in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nFitness Offset")
weigh_rel_off_map
#################

# combine plots
###############
weigh_cf_maps <- ggarrange(plotlist = c(weigh_abs_map, weigh_rel_map, weigh_abs_off_map, weigh_rel_off_map))
###############
```

#### Export

```{r}
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_rel_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_rel_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_abs_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_rel_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_abs_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_rel_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_abs_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_rel_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_abs_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_rel_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))

print(doc, target = "results/figures/climchg_fit_offset_maps.pptx")

doc2 <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_cf_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_cf_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_cf_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) 

print(doc2, target = "results/figures/climchg_fit_offset_maps_combined.pptx")
```
