---
title: "eelgrass_fitness_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Github/BalticEelgrass")

# setwd
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse)
library(psych)
library(stringr)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(performance)
library(corrplot)

# data
fit <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
env <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-10-02.csv")[,-1]
```

# Fitness Modeling

This script is to compare fitness across populations/treatments in the GO eelgrass mesocosm experiment. I may also use some site of origin environmental data to see if environmental factors explain performance differences, but this will come later (if at all).

## Setup

I need first to put together my data so that I have all the fitness and growth metrics possible.

```{r data pre-process}
# data cleaning to just relevant columns
fit_trim <- fit[,c("DNA_label", "new_label", "pop", "genet", "bagnum", "tank_ID", "sal_full", "temp_full", "trt", "nbr_leaves_start", "sideShoot", "nbr_sideShootLeaves_start", "length_longest_leaf_start", "weight_start", "IsBlackLesions_start", "nbr_leaves_end", "nbr_new_leaves", "width_leaf", "Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Nbr_new_shoots", "Nbr_leaves_new_shoot1", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Nbr_leaves_new_shoot2", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5", "below_ground_weight", "above_ground_weight", "weight_new_shoot1", "weight_new_shoot2", "total_weight", "lesions.", "isAlive", "Length_flower1", "Length_bract1")]

# binary survival column
fit_trim$isAlive <- ifelse(fit_trim$isAlive == TRUE, 1, 0)

# pop_trt factor for later significance testing
fit_trim$pop_trt <- as.factor(paste0(fit_trim$pop, "_", fit_trim$trt))

# add treatment block to both data frames for testing potential random effects
# block changes every 8 IDs - note though that salinity was independently controlled in each bag, so block may not actually appropriate descriptor of design
# note two plants were removed
table(fit_trim$genet) # gen-13 & gen-35
table(fit_trim$pop) # VAT & HOG
fit_trim_ord <- fit_trim[order(fit_trim$bagnum),]
fit_trim_ord$block <- ifelse(fit_trim_ord$bagnum <= 384, ((fit_trim_ord$bagnum - 1) %/% 8) %% 24 + 1, NA)

# ensure correct coding of relevant variables
fit_trim_ord$pop <- as.factor(fit_trim_ord$pop)
fit_trim_ord$tank_ID <- as.factor(fit_trim_ord$tank_ID)
fit_trim_ord$trt <- as.factor(fit_trim_ord$trt)
fit_trim_ord$block <- as.factor(fit_trim_ord$block)

# combine with the exp env dat
fit_trim_env <- merge(fit_trim_ord, env[,c("DNA_label", "mean_sal", "median_sal", "min_sal", "max_sal", "sd_sal", "mean_temp", "median_temp", "min_temp", "max_temp", "sd_temp", "x_coord", "y_coord", "position")], by = c("DNA_label"))

# length and weight growth
# for the growth analyses, we want a df with no flowering individuals and no dead individuals
fit_growth <- fit_trim_env[is.na(fit_trim$Length_flower1),]
fit_growth <- fit_growth[fit_growth$isAlive == 1,]

# length
fit_growth$length_growth <- rowSums(fit_growth[,c("Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5")], na.rm = T)

# percent length growth
fit_growth$length_std_change <- fit_growth$length_growth / fit_growth$length_longest_leaf_start

# weight
fit_growth$weight_change <- fit_growth$total_weight - fit_growth$weight_start

# percent weight change
fit_growth$weight_std_change <- fit_growth$weight_change / fit_growth$weight_start

# ratio of above vs below ground weight
fit_growth$above_below_weight_ratio <- fit_growth$above_ground_weight / fit_growth$below_ground_weight


# correlations between fitness metrics
# survival
fit_surv_cordf <- fit_trim[,c("IsBlackLesions_start","isAlive","nbr_leaves_start","weight_start","length_longest_leaf_start")]
pairs.panels(fit_surv_cordf)
fit_surv_cors <- cor(fit_surv_cordf)
corrplot(fit_surv_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)

# growth
fit_growth_cordf <- fit_growth[,c("IsBlackLesions_start","nbr_leaves_start","nbr_leaves_end","length_longest_leaf_start","length_growth","length_std_change","weight_change","weight_std_change","weight_start","total_weight","above_ground_weight","below_ground_weight","above_below_weight_ratio")]
pairs.panels(fit_growth_cordf)
fit_growth_cors <- cor(fit_growth_cordf, use = "complete.obs")
corrplot(fit_growth_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)
```

## Fitness Difference Modeling

Here I will compare differences among groups. I will use tank, treatment, and population as explanatory factors in my analysis.

### Survival Modeling

I will start by modeling differences in survival. To do so, I will use a binomial GLMM with binary survival data as response variable, fixed explanatory factors treatment & population, and random factor tank.

```{r survival modeling trt}
# analysis of fixed effects
anova(lm(isAlive ~ pop + trt, data = fit_trim_env)) # both significant

# now try building a model
surv_glmm <- glmer(isAlive ~ pop * trt + (1 | tank_ID), # may need to add more components of treatment after tank ID
               family = binomial(link = "logit"),
               data = fit_trim_env)
plot(surv_glmm)
Anova(surv_glmm, type = "III")
drop1(surv_glmm) # keep all
summary(surv_glmm)

# look for random effects in residuals
resid_surv <- data.frame(residuals = residuals(surv_glmm), block = fit_trim_env$block, tank_ID = fit_trim_env$tank_ID)
boxplot(residuals ~ block, data = resid_surv, main = "Residuals of Survival GLMM by Block") # boxplot bc block is a factor - doesn't look like any systematic effect of block
boxplot(residuals ~ tank_ID, data = resid_surv, main = "Residuals of Survival GLMM by Tank") # tank also doesn't seem to have any systematic effects

# significance testing with tukey
surv_glmm.1 <- glmer(isAlive ~ pop_trt + (1 | tank_ID), # getting boundary (singular) fit: see help('isSingular')
               family = binomial(link = "logit"),
               data = fit_trim)
surv_glm_poptrt <- glm(isAlive ~ pop_trt,
                       family = binomial,
                       data = fit_trim_env)

tuk_surv <- TukeyHSD(aov(surv_glm_poptrt), conf.level = 0.95)
cld_surv <- multcompLetters4(aov(surv_glm_poptrt), tuk_surv)

# multcomp letters to get letters
letters_surv_df <- data.frame(letters = as.data.frame.list(cld_surv$pop_trt), popsite = names((cld_surv$pop_trt)$Letters))
letters_surv_df_red <- letters_surv_df[,c("letters.Letters","popsite")]
colnames(letters_surv_df_red) <- c("letters","popsite")
rownames(letters_surv_df_red) <- NULL
```

Rather than using treatment directly, I will try using salinity and temperature to represent temperature. Additionally, because tank_ID is confounded with temperature, I will try using x/y coordinates as a fixed effect and avoid having random effects altogether. If I do this, I can use a GLM. 

```{r survival modeling sal/temp}
# anova
anova(lm(isAlive ~ pop + median_sal + median_temp, data = fit_trim_env)) # pop, sal, temp all significant

# try a glm
surv_glm <- glm(isAlive ~ pop + median_sal + median_temp,
                family = binomial(link = "logit"),
                data = fit_trim_env)

# check dispersion & multicollinearity
check_overdispersion(surv_glm) # looks good
vif(surv_glm) # suggests no collinearity, everything is around 1

# inspect model
summary(surv_glm)
plot(surv_glm)
Anova(surv_glm, type = "III") # pop, sal, temp all signif
step(surv_glm)

# what if we use block?
anova(lm(isAlive ~ pop + median_sal + median_temp + block, data = fit_trim_env)) # pop, sal, temp all significant

# try a glm
surv_glm_block <- glm(isAlive ~ pop + median_sal + median_temp + block,
                family = binomial(link = "logit"),
                data = fit_trim_env)

# check dispersion & multicollinearity
check_overdispersion(surv_glm_block) # looks good
vif(surv_glm_block) # suggests median sal has collinearity - this makes sense because the blocks align with salinity

# inspect model
summary(surv_glm_block)
plot(surv_glm_block)
Anova(surv_glm_block, type = "III") # pop, temp signif

# model selection
step(surv_glm_block) # gets rid of block, so back to the original (surv_glm)
```

### Growth Modeling

I will now model differences in growth, namely changes in length and changes in weight. I no longer need to use a GLMM, I can use an LMM since I am not working with a binary response variable. The following is for length growth (unstandardized)

```{r length modeling unstandardized}
# start with length
# analysis of fixed effects
anova(lm(length_growth ~ pop + trt, data = fit_growth)) # both significant

# now try building a model
growth_lmm <- lmer(length_growth ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(growth_lmm)
Anova(growth_lmm, type = "III") # all signif except trt * pop
drop1(growth_lmm) # drop interaction
summary(growth_lmm)

growth_lmm.1 <- lmer(length_growth ~ pop + trt + (1 | tank_ID), data = fit_growth)
plot(growth_lmm.1)
Anova(growth_lmm.1, type = "III") # all signif
drop1(growth_lmm.1) # keep all
summary(growth_lmm.1)

# look for random effects in residuals
resid_growlen <- data.frame(residuals = residuals(growth_lmm.1), block = fit_growth$block, tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ block, data = resid_growlen, main = "Residuals of Length Growth GLMM by Block") # here we do seem to have more variance in the data by block
boxplot(residuals ~ tank_ID, data = resid_growlen, main = "Residuals of Length Growth GLMM by Tank") # more variance by tank too, but nothing significantly different

# significance testing with tukey - throws an error
#growth_lmm.2 <- lmer(length_growth ~ pop_trt + (1 | tank_ID), data = fit_growth) # singular, throws error
#tuk_len <- TukeyHSD(aov(growth_lmm.2), conf.level = 0.95) # invalid x type in x && y
#cld_len <- multcompLetters4(aov(growth_lmm.2), tuk_len)
```

I also want to try this with my standardized length data.

```{r length modeling standardized}
# analysis of fixed effects
anova(lm(length_std_change ~ pop + trt, data = fit_growth)) # both significant

# build a model
growlen_lmm <- lmer(length_growth ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(growlen_lmm)
Anova(growlen_lmm, type = "III") # all signif except trt * pop
drop1(growlen_lmm) # drop interaction
summary(growlen_lmm)

growlen_lmm.1 <- lmer(length_growth ~ pop + trt + (1 | tank_ID), data = fit_growth)
plot(growlen_lmm.1)
Anova(growlen_lmm.1, type = "III") # all signif
drop1(growlen_lmm.1) # keep all
summary(growlen_lmm.1)

# look for random effects in residuals
resid_growlen <- data.frame(residuals = residuals(growlen_lmm.1), block = fit_growth$block, tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ block, data = resid_growlen, main = "Residuals of Length Growth LMM by Block") # here we do seem to have more variance in the data by block
boxplot(residuals ~ tank_ID, data = resid_growlen, main = "Residuals of Length Growth LMM by Tank") # more variance by tank too, but nothing significantly different

# significance testing with tukey
#growlen_lmm.2 <- lmer(length_growth ~ pop_trt + (1 | tank_ID), data = fit_growth) # singular, throws error
#tuk_len <- TukeyHSD(aov(growth_lmm.2), conf.level = 0.95)
#cld_len <- multcompLetters4(aov(growth_lmm.2), tuk_len)
```

I will look at separating out salinity and temperature for these models as well. I will only do this for the standardized data for now.

```{r length modeling sal/temp}
# analysis of fixed effects
anova(lm(length_std_change ~ pop + median_sal + median_temp, data = fit_growth)) # pop and temp signif, not sal

# now try building a model
growlen_saltemp_lmm <- lmer(length_growth ~ pop * median_sal * median_temp + (1 | tank_ID), data = fit_growth)
plot(growlen_saltemp_lmm)
Anova(growlen_saltemp_lmm, type = "III") # nothing is significant
drop1(growlen_saltemp_lmm) # don't need to drop anything if nothing matters anyways
summary(growlen_saltemp_lmm)

# additive model
growlen_saltemp_lmm.1 <- lmer(length_growth ~ pop + median_sal + median_temp + (1 | tank_ID), data = fit_growth) # boundary (singular) fit: see help('isSingular')
plot(growlen_saltemp_lmm.1)
Anova(growlen_saltemp_lmm.1, type = "III") # all significant
drop1(growlen_saltemp_lmm.1) # keep everything
summary(growlen_saltemp_lmm.1)

# residuals
resid_saltemp_lmm.1_growlen <- data.frame(residuals = residuals(growlen_saltemp_lmm.1), block = fit_growth$block, tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ block, data = resid_saltemp_lmm.1_growlen, main = "Residuals of Length Growth LMM by Block") # here we do seem to have more variance in the data by block
boxplot(residuals ~ tank_ID, data = resid_saltemp_lmm.1_growlen, main = "Residuals of Length Growth LMM by Tank") # more variance by tank too, but nothing significantly different
```

Now look into weight. I will first test un-standardized, then standardized weight change.

```{r weight modeling}
# analysis of fixed effects
anova(lm(weight_change ~ pop + trt, data = fit_growth)) # only trt significant

# build a model
growweight_lmm <- lmer(weight_change ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(growweight_lmm)
Anova(growweight_lmm, type = "III") # only interaction is significant
drop1(growweight_lmm) # can't drop anything
summary(growweight_lmm)

# check residuals
resid_growweight <- data.frame(residuals = residuals(growweight_lmm), block = fit_growth$block, tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ block, data = resid_growweight, main = "Residuals of Weight Change LMM by Block") # some variance in the data by block
boxplot(residuals ~ tank_ID, data = resid_growweight, main = "Residuals of Weight Change LMM by Tank") # some tank variance in mean but nothing that insane, more in variance of residuals themselves

# now for the standardized weight change
# fixed effects
anova(lm(weight_std_change ~ pop + trt, data = fit_growth)) # both significant

# std weight model
growweightstd_lmm <- lmer(weight_std_change ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(growweightstd_lmm)
Anova(growweightstd_lmm, type = "III") # nothing is significant
drop1(growweightstd_lmm) # model isn't significant
summary(growweightstd_lmm)

# check residuals
resid_growweightstd <- data.frame(residuals = residuals(growweightstd_lmm), block = fit_growth$block, tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ block, data = resid_growweightstd, main = "Residuals of Weight Change (std) LMM by Block") # some variance in the data by block
boxplot(residuals ~ tank_ID, data = resid_growweightstd, main = "Residuals of Weight Change (std) LMM by Tank") # some tank variance in mean, more in variance of residuals themselves
```

I will also try these growth models without random factors, using a basic linear model. Note that a linear model is inappropriate for survival analyses because the response data is binary.

```{r growth models lm}
# growth models
lm_len <- lm(length_growth ~ pop + median_sal + median_temp + x_coord + y_coord, data = fit_growth)
lm_weight <- lm(weight_change ~ pop + median_sal + median_temp + x_coord + y_coord, data = fit_growth)

# anova of these models
anova(lm_len) # pop, sal, temp all significant
anova(lm_weight) # temp significant

# check multicollinearity
vif(lm_len) # suggests no collinearity
vif(lm_weight) # suggests no collinearity

# inspect models
# length
summary(lm_len)
plot(lm_len)
Anova(lm_len, type = "III") # intercept, pop, sal, temp all signif

# weight
summary(lm_weight)
plot(lm_weight)
Anova(lm_weight, type = "III") # intercept and temp signif

# model selection
# length
step(lm_len) # keeps pop, median sal, median temp

# weight
step(lm_weight) # keeps just median temp
```

