---
title: "eelgrass_fitness_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Github/BalticEelgrass")

# setwd
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse)
library(psych)
library(stringr)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(performance)
library(corrplot)
library(blme)

# data
fit <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
env <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-12-31.csv")[,-1]
env_pop <- read.csv("data/EnvDat/exp_scaled_envdat_2025-12-03.csv")
fit_trim <- read.csv("data/experiment/expfit_trim.csv")
fit_growth <- read.csv("data/experiment/expfit_grow_trim.csv")
mlg <- read.csv("data/experiment/mlg_assignments.csv")
```

# Fitness Modeling

This script is to compare fitness across populations/treatments in the GO eelgrass mesocosm experiment. I may also use some site of origin environmental data to see if environmental factors explain performance differences, but this will come later (if at all).

## Setup

I need first to put together my data so that I have all the fitness and growth metrics possible. Note that all salinity stats are summarized through the experimental period at the individual bag level, while temperature stats are summarized through the experimental period at the tank level.

```{r data pre-process}
# add treatment block to both data frames for testing potential random effects
# block changes every 8 IDs - note though that salinity was independently controlled in each bag, so block may not actually appropriate descriptor of design
# note two plants were removed
table(fit_trim$genet) # gen-13 & gen-35
table(fit_trim$pop) # VAT & HOG
fit_trim_ord <- fit_trim[order(fit_trim$bagnum),]
fit_trim_ord$block <- ifelse(fit_trim_ord$bagnum <= 384, ((fit_trim_ord$bagnum - 1) %/% 8) %% 24 + 1, NA)

# ensure correct coding of relevant variables
fit_trim_ord$pop <- as.factor(fit_trim_ord$pop)
fit_trim_ord$tank_ID <- as.factor(fit_trim_ord$tank_ID)
fit_trim_ord$trt <- as.factor(fit_trim_ord$trt)
fit_trim_ord$block <- as.factor(fit_trim_ord$block)

# combine with the exp env dat
fit_trim_env <- merge(fit_trim_ord, env[,c("DNA_label", "mean_sal", "median_sal", "min_sal", "max_sal", "sd_sal", "mean_temp", "median_temp", "min_temp", "max_temp", "sd_temp", "x_coord", "y_coord", "position")], by = c("DNA_label"))
colnames(env_pop) <- c("pop", "coast", "long", "lat", "median_temp_home", "max_temp_home", "median_sal_home", "min_sal_home", "median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")
fit_trim_env1 <- merge(env_pop, fit_trim_env, by = c("pop"))
fit_growth1 <- merge(env_pop, fit_growth, by = c("pop"))

# correlations between fitness metrics
# survival
fit_surv_cordf <- fit_trim[,c("IsBlackLesions_start","isAlive","nbr_leaves_start","weight_start","length_longest_leaf_start")]
pairs.panels(fit_surv_cordf)
fit_surv_cors <- cor(fit_surv_cordf)
corrplot(fit_surv_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)

# growth
fit_growth_cordf <- fit_growth1[,c("IsBlackLesions_start","nbr_leaves_start","nbr_leaves_end","length_longest_leaf_start","length_growth","length_std_change","weight_change","weight_std_change","weight_start","total_weight","above_ground_weight","below_ground_weight","above_below_weight_ratio")]
pairs.panels(fit_growth_cordf)
fit_growth_cors <- cor(fit_growth_cordf, use = "complete.obs")
corrplot(fit_growth_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)

# add mlgs to both dfs - mlg samples are identified by pop & genet
# compatible formatting
mlg$pop <- str_split_fixed(mlg$sample, "-Zm", 2)[,1]
mlg$genet <- paste0("gen-",str_split_fixed(mlg$sample, "-Zm", 2)[,2])
mlg$MLG <- as.factor(mlg$MLG)

dim(fit_trim_env) # 382
dim(fit_growth) # 198 (surviving plants)
dim(mlg) # only 372, not all plants were able to be sequenced

# merge
fit_trim_env_mlg <- merge(fit_trim_env1, mlg, by = c("pop","genet")) # 370 plants, 2 in mll analysis were tossed in experiment
fit_growth_mlg <- merge(fit_growth1, mlg, by = c("pop","genet")) # 191 plants

# now we subset so only plants in home coast treatments are considered, i.e. west plants in curr and fut w coast
w_coast <- c("VAT", "VIK", "HOG", "BAR")
e_coast <- c("KUR", "KAL", "HOR", "BJO")
coast_fit_trim_env_mlg <- fit_trim_env_mlg %>% 
  filter(case_match(trt,
    c("TempControl_21psu", "TempWarm_16psu") ~ pop %in% w_coast,
    c("TempControl_7psu", "TempWarm_5psu") ~ pop %in% e_coast)) %>% 
  mutate(coast = ifelse(pop %in% w_coast, "West", "East"))
coast_fit_growth_mlg <- fit_growth_mlg %>% 
  filter(case_match(trt,
    c("TempControl_21psu", "TempWarm_16psu") ~ pop %in% w_coast,
    c("TempControl_7psu", "TempWarm_5psu") ~ pop %in% e_coast)) %>% 
  mutate(coast = ifelse(pop %in% w_coast, "West", "East"))

# code as current or future treatment
coast_fit_trim_env_mlg$cf_trt <- ifelse(coast_fit_trim_env_mlg$trt %in% c("TempWarm_16psu", "TempWarm_5psu"), "future", "current")
coast_fit_growth_mlg$cf_trt <- ifelse(coast_fit_growth_mlg$trt %in% c("TempWarm_16psu", "TempWarm_5psu"), "future", "current")
```

## Fitness Difference Modeling

Here I will compare differences among groups. I will use tank, treatment, and population as explanatory factors in my analysis.

### Survival Modeling

I will start by modeling differences in survival between current and future treatments. 

```{r}
# anova fixed effects
anova(lm(isAlive ~ pop + cf_trt, data = coast_fit_trim_env_mlg)) # pop, cf

# binomial regression (glm)
surv_glm_env <- glm(isAlive ~ pop * cf_trt,
                 family = binomial(link = "logit"),
                 data = coast_fit_trim_env_mlg)
plot(surv_glm_env)
summary(surv_glm_env)
Anova(surv_glm_env, type = "III") # main effects are significant, not interaction

# check dispersion & multicollinearity
check_overdispersion(surv_glm_env) # looks good
vif(surv_glm_env) # super bad, cf_trt is 7K VIF

# without interaction
surv_glm_env1 <- glm(isAlive ~ pop + cf_trt,
                 family = binomial(link = "logit"),
                 data = coast_fit_trim_env_mlg)
plot(surv_glm_env1)
summary(surv_glm_env1)
Anova(surv_glm_env1, type = "III") # both significant

# check for random effects
resid_surv_env <- data.frame(residuals = residuals(surv_glm_env1), tank_ID = coast_fit_trim_env_mlg$tank_ID, MLG = coast_fit_trim_env_mlg$MLG)
boxplot(residuals ~ tank_ID, data = resid_surv_env, main = "Residuals of Survival GLM by Tank") # yes, tank effects
boxplot(residuals ~ MLG, data = resid_surv_env, main = "Residuals of Survival GLM by MLG") # yes, MLG effects

# glmm
surv_glmm_env <- glmer(isAlive ~ pop + cf_trt + (1|tank_ID),
                 family = binomial(link = "logit"),
                 data = coast_fit_trim_env_mlg) # Model is nearly unidentifiable: large eigenvalue ratio -- seems to be because of VIK near 100% survival in both current and future
plot(surv_glmm_env)
summary(surv_glmm_env)
Anova(surv_glmm_env, type = "III") # all signif
drop1(surv_glmm_env) # none

# overdispersion & multicollinearity
check_overdispersion(surv_glmm_env) # looks good
vif(surv_glmm_env) # good

# run with blme? doesn't solve the issue
surv_glmm_env_b <- bglmer(isAlive ~ pop + cf_trt + (1|tank_ID),
                          family = binomial(link = "logit"),
                          control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 1e5)),
                          data = coast_fit_trim_env_mlg)
summary(surv_glmm_env_b)
Anova(surv_glmm_env_b, type = "III")

# run without VIK? the underlying issue I'm having is lack of variance in VIK
surv_glmm_env_v <- glmer(isAlive ~ pop + cf_trt + (1|tank_ID),
                         family = binomial(link = "logit"),
                         data = coast_fit_trim_env_mlg[!(coast_fit_trim_env_mlg$pop == "VIK"),])
summary(surv_glmm_env_v)
Anova(surv_glmm_env_v, type = "III") # all signif

# extract residuals
df.residual(surv_glmm_env_v)

# KW estimation of degrees of freedom
```


Quick look at survival against each environmental factor

```{r}
surv_env_df <- fit_trim_env_mlg %>% group_by(pop, trt) %>%
  summarise(
    mean_surv = mean(isAlive),
    .groups = 'drop'
  ) %>% as.data.frame()
fit_env_trim <- fit_trim_env_mlg %>% dplyr::select(pop, trt, median_sal, median_temp, max_temp, min_sal, median_sal_home, median_temp_home, max_temp_home, min_sal_home)
env_surv_mg <- merge(fit_env_trim, surv_env_df, by = c("pop", "trt"))

# clim chg only
env_surv_mg_w <- env_surv_mg %>% filter(pop %in% c("VAT", "VIK", "BAR", "HOG") & trt %in% c("TempControl_21psu", "TempWarm_16psu"))
env_surv_mg_e <- env_surv_mg %>% filter(!(pop %in% c("VAT", "VIK", "BAR", "HOG")) & !(trt %in% c("TempControl_21psu", "TempWarm_16psu")))
env_surv_mg_clim <- rbind(env_surv_mg_w, env_surv_mg_e)

# quick and dirty correlation
cor(env_surv_mg$mean_surv, env_surv_mg$median_temp)
cor(env_surv_mg$mean_surv, env_surv_mg$max_temp)
cor(env_surv_mg$mean_surv, env_surv_mg$median_sal)
cor(env_surv_mg$mean_surv, env_surv_mg$min_sal)

cor(env_surv_mg_clim$mean_surv, env_surv_mg_clim$median_temp_home)
cor(env_surv_mg_clim$mean_surv, env_surv_mg_clim$max_temp_home)
cor(env_surv_mg_clim$mean_surv, env_surv_mg_clim$median_sal_home)
cor(env_surv_mg_clim$mean_surv, env_surv_mg_clim$min_sal_home)

# add pop info
pop_coast <- data.frame(pop = c("VIK","VAT","HOG","BAR",
                                "KUR","KAL","HOR","BJO"),
                        pop_ord = 1:8,
                        pop_coast = c("West", "West", "West", "West", 
                                      "East", "East", "East", "East"),
                        pop_coast_ord = c(1,1,1,1,2,2,2,2))
trt_coast <- data.frame(#trt = c("TempControl-21psu", "TempControl-7psu", 
                        #        "TempWarm-16psu", "TempWarm-5psu"),
                        trt = c("TempControl_21psu", "TempControl_7psu",
                                "TempWarm_16psu", "TempWarm_5psu"),
                        trt_ord = 1:4,
                        trt_coast = c("West", "East", "West", "East"),
                        trt_coast_ord = c(1,2,1,2))
fit_env_mlg_surv <- merge(pop_coast, fit_trim_env_mlg, by = c("pop"))
fit_env_mlg_surv_clim <- merge(pop_coast, env_surv_mg_clim, by = c("pop"))

# plots - experimental env
fit_env_mlg_surv %>% ggplot(aes(x = median_temp, y = isAlive)) + 
  geom_point(shape = 21, fill = "#fbd1a2", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "goldenrod1", fill = "#fbd1a2") + 
  labs(title = "Survival as a Function of Median Temperature", x = "Median Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% ggplot(aes(x = max_temp, y = isAlive)) + 
  geom_point(shape = 21, fill = "#f79256", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "darkorange2", fill = "#f79256") + 
  labs(title = "Survival as a Function of Maximum Temperature", x = "Maximum Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% ggplot(aes(x = median_sal, y = isAlive)) + 
  geom_point(shape = 21, fill = "#7dcfb6", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "cyan4", fill = "#7dcfb6") + 
  labs(title = "Survival as a Function of Median Salinity", x = "Median Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% ggplot(aes(x = min_sal, y = isAlive)) + 
  geom_point(shape = 21, fill = "#00b2ca", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "royalblue3", fill = "#00b2ca") + 
  labs(title = "Survival as a Function of Minimum Salinity", x = "Minimum Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

# with populations as individual regressions
fit_env_mlg_surv %>% ggplot(aes(x = median_temp, y = isAlive, fill = fct_reorder(pop, pop_ord))) + 
  geom_point(shape = 21, colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), aes(colour = fct_reorder(pop, pop_ord)), se = F) + 
  scale_fill_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_color_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  labs(title = "Survival as a Function of Median Temperature", x = "Median Temperature (ºC)", y = "Alive or Dead", fill = "Population") +
  theme_classic()

fit_env_mlg_surv %>% ggplot(aes(x = max_temp, y = isAlive, fill = fct_reorder(pop, pop_ord))) + 
  geom_point(shape = 21, colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), aes(colour = fct_reorder(pop, pop_ord)), se = F) + 
  scale_fill_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_color_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  labs(title = "Survival as a Function of Maximum Temperature", x = "Maximum Temperature (ºC)", y = "Alive or Dead", fill = "Population") +
  theme_classic()

fit_env_mlg_surv %>% ggplot(aes(x = median_sal, y = isAlive, fill = fct_reorder(pop, pop_ord))) + 
  geom_point(shape = 21, colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), aes(colour = fct_reorder(pop, pop_ord)), se = F) + 
  scale_fill_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_color_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  labs(title = "Survival as a Function of Median Salinity", x = "Salinity (psu)", y = "Alive or Dead", fill = "Population") +
  theme_classic()

fit_env_mlg_surv %>% ggplot(aes(x = min_sal, y = isAlive, fill = fct_reorder(pop, pop_ord))) + 
  geom_point(shape = 21, colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), aes(colour = fct_reorder(pop, pop_ord)), se = F) + 
  scale_fill_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  scale_color_manual(name = "Population",
                    values = c("#002e5a","#0130b9","#0078cf","#82daff", 
                               "#d9ed92","#4ad66d","#25a244","#155d27"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  labs(title = "Survival as a Function of Minimum Salinity", x = "Salinity (psu)", y = "Alive or Dead", fill = "Population") +
  theme_classic()

# pops - env of origin
fit_env_mlg_surv %>% filter(trt == "TempWarm_16psu") %>% ggplot(aes(x = median_temp_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#fbd1a2", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "goldenrod1", fill = "#fbd1a2") + 
  labs(title = "Survival as a Function of Median Temperature (West)", x = "Median Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt == "TempWarm_5psu") %>% ggplot(aes(x = median_temp_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#fbd1a2", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "goldenrod1", fill = "#fbd1a2") + 
  labs(title = "Survival as a Function of Median Temperature (East)", x = "Median Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu", "TempWarm_5psu")) %>% ggplot(aes(x = median_temp_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#fbd1a2", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "goldenrod1", fill = "#fbd1a2") + 
  labs(title = "Survival (Future) as a Function of Median Temperature", x = "Median Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu")) %>% ggplot(aes(x = max_temp_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#f79256", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "darkorange2", fill = "#f79256") + 
  labs(title = "Survival as a Function of Maximum Temperature (West)", x = "Maximum Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_5psu")) %>% ggplot(aes(x = max_temp_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#f79256", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "darkorange2", fill = "#f79256") + 
  labs(title = "Survival as a Function of Maximum Temperature (East)", x = "Maximum Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu", "TempWarm_5psu")) %>% ggplot(aes(x = max_temp_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#f79256", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "darkorange2", fill = "#f79256") + 
  labs(title = "Survival (Future) as a Function of Maximum Temperature", x = "Maximum Temperature (ºC)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu")) %>% ggplot(aes(x = median_sal_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#7dcfb6", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "cyan4", fill = "#7dcfb6") + 
  labs(title = "Survival as a Function of Median Salinity (West)", x = "Median Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_5psu")) %>% ggplot(aes(x = median_sal_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#7dcfb6", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "cyan4", fill = "#7dcfb6") + 
  labs(title = "Survival as a Function of Median Salinity (East)", x = "Median Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu", "TempWarm_5psu")) %>% ggplot(aes(x = median_sal_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#7dcfb6", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "cyan4", fill = "#7dcfb6") + 
  labs(title = "Survival (Future) as a Function of Median Salinity", x = "Median Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu")) %>% ggplot(aes(x = min_sal_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#00b2ca", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "royalblue3", fill = "#00b2ca") + 
  labs(title = "Survival as a Function of Minimum Salinity (West)", x = "Minimum Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_5psu")) %>% ggplot(aes(x = min_sal_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#00b2ca", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "royalblue3", fill = "#00b2ca") + 
  labs(title = "Survival as a Function of Minimum Salinity (East)", x = "Minimum Salinity (psu)", y = "Alive or Dead") +
  theme_classic()

fit_env_mlg_surv %>% filter(trt %in% c("TempWarm_16psu", "TempWarm_5psu")) %>% ggplot(aes(x = min_sal_home, y = isAlive)) + 
  geom_point(shape = 21, fill = "#00b2ca", colour = "black", size = 4) +
  geom_smooth(method = "glm", method.args = list(family = "binomial"), se = T, color = "royalblue3", fill = "#00b2ca") + 
  labs(title = "Survival (Future) as a Function of Minimum Salinity", x = "Minimum Salinity (psu)", y = "Alive or Dead") +
  theme_classic()
```


And now more in depth modeling using environmental site of origin data. Will specifically look at this within each treatment

```{r}
# subsets
w_fut <- coast_fit_trim_env_mlg[coast_fit_trim_env_mlg$trt == "TempWarm_16psu",]
e_fut <- coast_fit_trim_env_mlg[coast_fit_trim_env_mlg$trt == "TempWarm_5psu",]
w_cur <- coast_fit_trim_env_mlg[coast_fit_trim_env_mlg$trt == "TempControl_21psu",]
e_cur <- coast_fit_trim_env_mlg[coast_fit_trim_env_mlg$trt == "TempControl_7psu",]
w_sub <- coast_fit_trim_env_mlg[coast_fit_trim_env_mlg$coast == "West",]
e_sub <- coast_fit_trim_env_mlg[coast_fit_trim_env_mlg$coast == "East",]

surv_env_w_fut <- glm(isAlive ~ max_temp_home * min_sal_home,
  #glm(isAlive ~ median_temp_home + max_temp_home + median_sal_home + min_sal_home,
                      family = binomial(link = "logit"), 
                      data = w_fut,
                      na.action = na.exclude)
#plot(surv_env_w_fut) # not great
summary(surv_env_w_fut)
Anova(surv_env_w_fut, type = "III") # sal, temp, and interaction all significant
drop1(surv_env_w_fut) # keep all

# plot residuals to look for random effects
surv_env_w_fut_check <- data.frame(resids = surv_env_w_fut$residuals, tank_ID = w_fut$tank_ID, mlg = w_fut$MLG)
plot(surv_env_w_fut_check$tank_ID, surv_env_w_fut_check$resids) # yeah some big effects

library(brms)
brms_w_fut <- brm(formula = isAlive ~ median_temp_scaled + median_sal_scaled + (1|tank_ID),
  family = bernoulli(link = "logit"),
  data = w_fut,
  prior = c(prior(normal(0, 2.5), class = "b"),
            prior(student_t(3, 0, 2.5), class = "sd")),
  chains = 4, cores = 2, iter = 2000)
summary(brms_w_fut)

# east coast
brms_e_fut <- brm(formula = isAlive ~ median_temp_scaled + median_sal_scaled + (1|tank_ID),
  family = bernoulli(link = "logit"),
  data = e_fut,
  prior = c(prior(normal(0, 2.5), class = "b"),
            prior(student_t(3, 0, 2.5), class = "sd")),
  chains = 4, cores = 2, iter = 2000)
summary(brms_e_fut)

brms_w_cur <- brm(formula = isAlive ~ median_temp_scaled + median_sal_scaled + (1|tank_ID),
  family = bernoulli(link = "logit"),
  data = w_cur,
  prior = c(prior(normal(0, 2.5), class = "b"),
            prior(student_t(3, 0, 2.5), class = "sd")),
  chains = 4, cores = 2, iter = 2000)
summary(brms_w_cur)

brms_e_cur <- brm(formula = isAlive ~ median_temp_scaled + median_sal_scaled + (1|tank_ID),
  family = bernoulli(link = "logit"),
  data = e_cur,
  prior = c(prior(normal(0, 2.5), class = "b"),
            prior(student_t(3, 0, 2.5), class = "sd")),
  chains = 4, cores = 2, iter = 2000)
summary(brms_e_cur)

brms_all <- brm(formula = isAlive ~ median_temp_scaled + median_sal_scaled + (1|tank_ID),
  family = bernoulli(link = "logit"),
  data = coast_fit_trim_env_mlg,
  prior = c(prior(normal(0, 2.5), class = "b"),
            prior(student_t(3, 0, 2.5), class = "sd")),
  chains = 4, cores = 2, iter = 2000)
summary(brms_all)
```

### Growth Modeling

I will now model differences in growth, namely changes in length and changes in weight. I no longer need to use a GLM, I can use an LMM since I am not working with a binary response variable. 

Notes on this dataset: The dataset used in this modeling approach excludes individuals that died. The reason for this is that there is no biologically relevant way to code this into the model - if I code the dead individuals as having no length or weight change, they will appear more fit than many other plants that lost length/biomass, and thus had negative length/weight changes. 

I will start by building models for length change (standardized) using treatment, then environmental variables, to explain differences in length change. 

```{r length modeling, standardized}
# quick check
anova(lm(length_std_change ~ pop + trt, data = fit_growth_mlg)) # both significant
hist(fit_growth$length_std_change) # left skewed, try log transformation
hist(log(fit_growth$length_std_change)) # yay it's normally distributed now

# start with lm since data is relatively normally distributed and not incorporating random effects yet
growlen_std_lm <- lm(log(length_std_change) ~ pop * trt, data = fit_growth_mlg)
plot(growlen_std_lm) # qq plot has some outliers deviating significantly from 1-1 line
anova(lm(length_std_change ~ pop * trt, data = fit_growth)) # both significant, interaction isn't
step(growlen_std_lm) # drop interaction
summary(growlen_std_lm)

# without interaction
growlen_std_lm.1 <- lm(log(length_std_change) ~ pop + trt, data = fit_growth_mlg)
plot(growlen_std_lm.1) # qq plot has some outliers deviating significantly from 1-1 line, but not as bad as before
anova(growlen_std_lm.1) # both significant
step(growlen_std_lm.1) # final model
summary(growlen_std_lm.1)

# check for random effects
resid_len <- data.frame(residuals = residuals(growlen_std_lm.1), tank_ID = fit_growth_mlg$tank_ID, MLG = fit_growth_mlg$MLG)
boxplot(residuals ~ tank_ID, data = resid_len, main = "Residuals of Length LM by Tank") # some variation, very slight, try an lmm
boxplot(residuals ~ MLG, data = resid_len, main = "Residuals of Length LM by MLG") # some variation, very slight, try an lmm

# lmm
growlen_std_lmm <- lmer(log(length_std_change) ~ pop * trt + (1|tank_ID), data = fit_growth)
plot(growlen_std_lmm)
Anova(growlen_std_lmm, type = "III") # all significant
step(growlen_std_lmm) # keep all
summary(growlen_std_lmm)
```

I will look at separating out salinity and temperature for these models as well. I will only do this for the standardized data.

```{r length modeling sal/temp, standardized}
# quick check
anova(lm(length_std_change ~ pop + median_sal + median_temp, data = fit_growth_mlg)) # pop and temp signif, not sal

# we know we have skew, so log transform
fit_growth_mlg$log_length_std_change <- log(fit_growth_mlg$length_std_change)

# now model
len_saltemp_lm <- lm(log_length_std_change ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp, data = fit_growth_mlg) # what family?
plot(len_saltemp_lm)
anova(len_saltemp_lm) # all signif except pop:median_temp interaction
step(len_saltemp_lm) # remove pop temp interaction
summary(len_saltemp_lm)

# without interaction
len_saltemp_lm.1 <- lm(log_length_std_change ~ pop + median_sal + median_temp + pop:median_sal, data = fit_growth_mlg) # what family?
plot(len_saltemp_lm.1)
anova(len_saltemp_lm.1) # all signif
step(len_saltemp_lm.1) # final model
summary(len_saltemp_lm.1)

# check residuals for random effects
resid_len_saltemp <- data.frame(residuals = residuals(len_saltemp_lm.1), tank_ID = fit_growth_mlg$tank_ID, MLG = fit_growth_mlg$MLG)
boxplot(residuals ~ tank_ID, data = resid_len_saltemp, main = "Residuals of Length LM by Tank") # some variation, very slight, could try an lmm
boxplot(residuals ~ MLG, data = resid_len_saltemp, main = "Residuals of Length LM by MLG") # lots of variation, try an lmm

# lmm
len_saltemp_lmm <- lmer(log_length_std_change ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp + (1|MLG), data = fit_growth_mlg)
plot(len_saltemp_lmm)
Anova(len_saltemp_lmm, type = "III") # intercept, temp, pop:sal significant
step(len_saltemp_lmm) # drop temp:pop, sal, and sal:pop, tankID as random effect
summary(len_saltemp_lmm)

len_saltemp_lmm.1 <- lmer(log_length_std_change ~ pop + median_temp + (1|MLG), data = fit_growth_mlg)
plot(len_saltemp_lmm.1)
Anova(len_saltemp_lmm.1, type = "III") # everything signif except sal & pop/sal
step(len_saltemp_lmm.1) # this is final model
summary(len_saltemp_lmm.1)

# check dispersion & multicollinearity
check_overdispersion(len_saltemp_lmm.1) # looks good
vif(len_saltemp_lmm.1) # suggests no collinearity

# # another option would be to do a glm with a gamma family, which is meant for skewed data
# len_saltemp_glm.1 <- glm(length_std_change ~ pop + median_temp, family = "gamma", data = fit_growth)
# plot(len_saltemp_glm.1)
# Anova(len_saltemp_glm.1, type = "III") # all signif
# step(len_saltemp_glm.1)
# summary(len_saltemp_glm.1)

# # check dispersion & multicollinearity
# check_overdispersion(len_saltemp_glm.1) # looks good
# vif(len_saltemp_glm.1) # suggests no collinearity, everything is around 1
```

Now look into weight. I am testing only standardized weight change.

```{r weight modeling, standardized}
# quick check
anova(lm(weight_std_change ~ pop + cf_trt, data = coast_fit_growth_mlg)) # both signif

# check distribution
hist(coast_fit_growth_mlg$weight_std_change) # beautifully normally distributed omg

# linear model to start
growweightstd_lm <- lm(weight_std_change ~ pop * cf_trt, data = coast_fit_growth_mlg)
plot(growweightstd_lm)
anova(growweightstd_lm) # pop & trt signif, not interaction
step(growweightstd_lm) # remove interaction

growweightstd_lm.1 <- lm(weight_std_change ~ pop + cf_trt, data = coast_fit_growth_mlg)
plot(growweightstd_lm.1)
anova(growweightstd_lm.1) # both signif
step(growweightstd_lm.1) # keep just cf_trt
summary(growweightstd_lm.1)

growweightstd_lm.2 <- lm(weight_std_change ~ cf_trt, data = coast_fit_growth_mlg)
plot(growweightstd_lm.2)
anova(growweightstd_lm.2) # signif
summary(growweightstd_lm.2)

# check for random effects
resid_weight <- data.frame(residuals = residuals(growweightstd_lm.2), tank_ID = coast_fit_growth_mlg$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_weight, main = "Residuals of Weight LM by Tank") # yes, tank effects

# lmm weight model
growweightstd_lmm <- lmer(weight_std_change ~ pop * cf_trt + (1|tank_ID), data = coast_fit_growth_mlg) # rank deficient (likely because some pop/trt combos empty)
plot(growweightstd_lmm)
Anova(growweightstd_lmm, type = "III") # pop is significant
step(growweightstd_lmm) # drop pop and interaction
summary(growweightstd_lmm)

# dropping interaction
growweightstd_lmm.1 <- lmer(weight_std_change ~ cf_trt + (1|tank_ID), data = coast_fit_growth_mlg)
plot(growweightstd_lmm.1)
Anova(growweightstd_lmm.1, type = "III") # significant
summary(growweightstd_lmm.1)

# check dispersion, no need for multicollinearity because only one variable
check_overdispersion(growweightstd_lmm.1) # looks good

# come back for estimation of variance/df
```

### Modeling Outputs

Now I have my beautiful models! I'll write relevant results out for future reference:

```{r}
sink(paste0("results/eelgrass_fitness_models_",Sys.Date(),".txt"))

"Survival Model Pop/trt\n"
summary(surv_glm)
Anova(surv_glm, type = "III")

"Survival Model Env\n"
summary(surv_glmm_env.2)
Anova(surv_glmm_env.2, type = "III")


"Length Model Pop/trt\n"
summary(growlen_std_lm.1)
anova(growlen_std_lm.1)

"Length Model Env\n"
summary(len_saltemp_lm.1)
anova(len_saltemp_lm.1)

"Weight Model Pop/trt\n"
summary(growweightstd_lm.1)
anova(growweightstd_lm.1)

"Weight Model Env\n"
summary(weight_saltemp_lm.1)
anova(weight_saltemp_lm.1)

sink()

# write out other fitness for later analysis
write.csv(fit_growth, "data/experiment/fit_growth_dat.csv")
```