---
title: "eelgrass_fitness_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Eelgrass/BalticEelgrass")
```

```{r}
# libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(performance)

# data
fit <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
env <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-10-02.csv")[,-1]
```

# Fitness Modeling

This script is to compare fitness across populations/treatments in the GO eelgrass mesocosm experiment. I may also use some site of origin environmental data to see if environmental factors explain performance differences, but this will come later (if at all).

## Setup

I need first to put together my data so that I have all the fitness and growth metrics possible.

```{r}
# data cleaning to just relevant columns
fit_trim <- fit[,c("DNA_label", "new_label", "pop", "genet", "bagnum", "tank_ID", "sal_full", "temp_full", "trt", "nbr_leaves_start", "sideShoot", "nbr_sideShootLeaves_start", "length_longest_leaf_start", "weight_start", "IsBlackLesions_start", "nbr_leaves_end", "nbr_new_leaves", "width_leaf", "Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Nbr_new_shoots", "Nbr_leaves_new_shoot1", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Nbr_leaves_new_shoot2", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5", "below_ground_weight", "above_ground_weight", "weight_new_shoot1", "weight_new_shoot2", "total_weight", "lesions.", "isAlive", "Length_flower1", "Length_bract1")]

# binary survival column
fit_trim$isAlive <- ifelse(fit_trim$isAlive == TRUE, 1, 0)

# pop_trt factor for later significance testing
fit_trim$pop_trt <- as.factor(paste0(fit_trim$pop, "_", fit_trim$trt))

# ensure correct coding of relevant variables
fit_trim$pop <- as.factor(fit_trim$pop)
fit_trim$tank_ID <- as.factor(fit_trim$tank_ID)
fit_trim$trt <- as.factor(fit_trim$trt)

# combine with the exp env dat
fit_trim_env <- merge(fit_trim, env[,c("DNA_label", "mean_sal", "median_sal", "min_sal", "max_sal", "sd_sal", "mean_temp", "median_temp", "min_temp", "max_temp", "sd_temp", "x_coord", "y_coord", "position")], by = c("DNA_label"))

# length and weight growth
# for the growth analyses, we want a df with no flowering individuals and no dead individuals
fit_growth <- fit_trim_env[is.na(fit_trim$Length_flower1),]
fit_growth <- fit_growth[fit_growth$isAlive == 1,]

# length
fit_growth$length_growth <- rowSums(fit_growth[,c("Growth_leaf1","Growth_leaf2","Growth_leaf3","Growth_leaf4","Growth_leaf5","Growth_leaf6","Growth_leaf7","Length_new_leaf1","Length_new_leaf2","Length_new_leaf3","Length_new_leaf4","Length_new_leaf5","Length_new_leaf6","Length_new_shoot1_leaf1","Length_new_shoot1_leaf2","Length_new_shoot1_leaf3","Length_new_shoot1_leaf4","Length_new_shoot1_leaf5","Length_new_shoot1_leaf6","Length_new_shoot2_leaf1","Length_new_shoot2_leaf2","Length_new_shoot2_leaf3","Length_new_shoot2_leaf4","Length_new_shoot2_leaf5")], na.rm = T)

# weight
fit_growth$weight_growth <- fit_growth$total_weight - fit_growth$weight_start

# ratio of above vs below ground weight
fit_growth$above_below_weight_ratio <- fit_growth$above_ground_weight / fit_growth$below_ground_weight
```

## Fitness Difference Modeling

Here I will compare differences among groups. I will use tank, treatment, and population as explanatory factors in my analysis.

### Survival Modeling

I will start by modeling differences in survival. To do so, I will use a binomial GLMM with binary survival data as response variable, fixed explanatory factors treatment & population, and random factor tank.

```{r}
# analysis of fixed effects
anova(lm(isAlive ~ pop + trt, data = fit_trim_env)) # both significant

# now try building a model
surv_glmm <- glmer(isAlive ~ pop * trt + (1 | tank_ID), # may need to add more components of treatment after tank ID
               family = binomial(link = "logit"),
               data = fit_trim_env)
plot(surv_glmm)
Anova(surv_glmm, type = "III")
drop1(surv_glmm) # keep all
summary(surv_glmm)

# significance testing with tukey
surv_glmm.1 <- glmer(isAlive ~ pop_trt + (1 | tank_ID), # getting boundary (singular) fit: see help('isSingular')
               family = binomial(link = "logit"),
               data = fit_trim)
surv_glm_poptrt <- glm(isAlive ~ pop_trt,
                       family = binomial,
                       data = fit_trim_env)

tuk_surv <- TukeyHSD(aov(surv_glm_poptrt), conf.level = 0.95)
cld_surv <- multcompLetters4(aov(surv_glm_poptrt), tuk_surv)

# multcomp letters to get letters
letters_surv_df <- data.frame(letters = as.data.frame.list(cld_surv$pop_trt), popsite = names((cld_surv$pop_trt)$Letters))
letters_surv_df_red <- letters_surv_df[,c("letters.Letters","popsite")]
colnames(letters_surv_df_red) <- c("letters","popsite")
rownames(letters_surv_df_red) <- NULL
```

Rather than using treatment directly, I will try using salinity and temperature to represent temperature. Additionally, because tank_ID is confounded with temperature, I will try using x/y coordinates as a fixed effect and avoid having random effects altogether. If I do this, I can use a GLM. 

```{r}
# anova
anova(lm(isAlive ~ pop + median_sal + median_temp + x_coord + y_coord, data = fit_trim_env)) # pop, sal, temp all significant

# try a glm
surv_glm <- glm(isAlive ~ pop + median_sal + median_temp + board_ID, # use board ID instead of position
                family = binomial(link = "logit"),
                data = fit_trim_env)

# check dispersion & multicollinearity
check_overdispersion(surv_glm) # looks good
vif(surv_glm) # suggests no collinearity, everything is around 1

# inspect model
summary(surv_glm)
plot(surv_glm)
Anova(surv_glm, type = "III") # pop, sal, temp all signif

# model selection
step(surv_glm)

# final model removes xy coords
surv_glm.1 <- glm(isAlive ~ pop + median_sal + median_temp, 
                  family = binomial(link = "logit"),
                  data = fit_trim_env)

# double check this model
check_overdispersion(surv_glm.1) # looks good
vif(surv_glm.1) # suggests no collinearity, everything is around 1
summary(surv_glm.1)
plot(surv_glm.1)
Anova(surv_glm.1, type = "III") # pop, sal, temp all signif
```

### Growth Modeling

I will now model differences in growth, namely changes in length and changes in weight. I no longer need to use a GLMM, I can use an LMM since I am not working with a binary response variable.

```{r}
# start with length
# analysis of fixed effects
anova(lm(length_growth ~ pop + trt, data = fit_growth)) # both significant

# now try building a model
mod2 <- lmer(length_growth ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(mod2)
Anova(mod2, type = "III") # all signif except trt * pop
drop1(mod2) # drop interaction
summary(mod2)

mod2.1 <- lmer(length_growth ~ pop + trt + (1 | tank_ID), data = fit_growth)
plot(mod2.1)
Anova(mod2.1, type = "III") # all signif
drop1(mod2.1) # keep all
summary(mod2.1)

# significance testing with tukey
mod2.2 <- lmer(length_growth ~ pop_trt + (1 | tank_ID), data = fit_growth) # singular, throws error
tuk_len <- TukeyHSD(aov(mod2.2), conf.level = 0.95)
cld_len <- multcompLetters4(aov(mod2.2), tuk_len)

# try a linear model
mod2.3 <- lm(length_growth ~ pop * trt, data = fit_growth)
mod2.4 <- lm(length_growth ~ pop_trt, data = fit_growth)
```

I will also try this one without random factors, using a basic linear model. 

```{r}
# growth models
lm_len <- lm(length_growth ~ pop + median_sal + median_temp + x_coord + y_coord, data = fit_growth)
lm_weight <- lm(weight_growth ~ pop + median_sal + median_temp + x_coord + y_coord, data = fit_growth)

# anova of these models
anova(lm_len) # pop, sal, temp all significant
anova(lm_weight) # temp significant

# check multicollinearity
vif(lm_len) # suggests no collinearity
vif(lm_weight) # suggests no collinearity

# inspect models
# length
summary(lm_len)
plot(lm_len)
Anova(lm_len, type = "III") # intercept, pop, sal, temp all signif

# weight
summary(lm_weight)
plot(lm_weight)
Anova(lm_weight, type = "III") # intercept and temp signif

# model selection
# length
step(lm_len) # keeps pop, median sal, median temp

# weight
step(lm_weight) # keeps just median temp
```

