---
title: "eelgrass_fitness_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Github/BalticEelgrass")

# setwd
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse)
library(psych)
library(stringr)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(performance)
library(corrplot)

# data
fit <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
env <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-10-02.csv")[,-1]
```

# Fitness Modeling

This script is to compare fitness across populations/treatments in the GO eelgrass mesocosm experiment. I may also use some site of origin environmental data to see if environmental factors explain performance differences, but this will come later (if at all).

## Setup

I need first to put together my data so that I have all the fitness and growth metrics possible. Note that all salinity stats are summarized through the experimental period at the individual bag level, while temperature stats are summarized through the experimental period at the tank level.

```{r data pre-process}
# data cleaning to just relevant columns
fit_trim <- fit[,c("DNA_label", "new_label", "pop", "genet", "bagnum", "tank_ID", "sal_full", "temp_full", "trt", "nbr_leaves_start", "sideShoot", "nbr_sideShootLeaves_start", "length_longest_leaf_start", "weight_start", "IsBlackLesions_start", "nbr_leaves_end", "nbr_new_leaves", "width_leaf", "Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Nbr_new_shoots", "Nbr_leaves_new_shoot1", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Nbr_leaves_new_shoot2", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5", "below_ground_weight", "above_ground_weight", "weight_new_shoot1", "weight_new_shoot2", "total_weight", "lesions.", "isAlive", "Length_flower1", "Length_bract1")]

# binary survival column
fit_trim$isAlive <- ifelse(fit_trim$isAlive == TRUE, 1, 0)

# pop_trt factor for later significance testing
fit_trim$pop_trt <- as.factor(paste0(fit_trim$pop, "_", fit_trim$trt))

# length and weight growth
# for the growth analyses, we want a df with no flowering individuals and no dead individuals
fit_growth <- fit_trim_env[is.na(fit_trim_env$Length_flower1),]
fit_growth <- fit_growth[fit_growth$isAlive == 1,]

# standardized growth variables
# length
fit_growth$length_growth <- rowSums(fit_growth[,c("Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5")], na.rm = T)

# percent length growth
fit_growth$length_std_change <- fit_growth$length_growth / fit_growth$length_longest_leaf_start

# weight
fit_growth$weight_change <- fit_growth$total_weight - fit_growth$weight_start

# percent weight change
fit_growth$weight_std_change <- fit_growth$weight_change / fit_growth$weight_start

# ratio of above vs below ground weight
fit_growth$above_below_weight_ratio <- fit_growth$above_ground_weight / fit_growth$below_ground_weight

# add treatment block to both data frames for testing potential random effects
# block changes every 8 IDs - note though that salinity was independently controlled in each bag, so block may not actually appropriate descriptor of design
# note two plants were removed
table(fit_trim$genet) # gen-13 & gen-35
table(fit_trim$pop) # VAT & HOG
fit_trim_ord <- fit_trim[order(fit_trim$bagnum),]
fit_trim_ord$block <- ifelse(fit_trim_ord$bagnum <= 384, ((fit_trim_ord$bagnum - 1) %/% 8) %% 24 + 1, NA)

# ensure correct coding of relevant variables
fit_trim_ord$pop <- as.factor(fit_trim_ord$pop)
fit_trim_ord$tank_ID <- as.factor(fit_trim_ord$tank_ID)
fit_trim_ord$trt <- as.factor(fit_trim_ord$trt)
fit_trim_ord$block <- as.factor(fit_trim_ord$block)

# combine with the exp env dat
fit_trim_env <- merge(fit_trim_ord, env[,c("DNA_label", "mean_sal", "median_sal", "min_sal", "max_sal", "sd_sal", "mean_temp", "median_temp", "min_temp", "max_temp", "sd_temp", "x_coord", "y_coord", "position")], by = c("DNA_label"))

# correlations between fitness metrics
# survival
fit_surv_cordf <- fit_trim[,c("IsBlackLesions_start","isAlive","nbr_leaves_start","weight_start","length_longest_leaf_start")]
pairs.panels(fit_surv_cordf)
fit_surv_cors <- cor(fit_surv_cordf)
corrplot(fit_surv_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)

# growth
fit_growth_cordf <- fit_growth[,c("IsBlackLesions_start","nbr_leaves_start","nbr_leaves_end","length_longest_leaf_start","length_growth","length_std_change","weight_change","weight_std_change","weight_start","total_weight","above_ground_weight","below_ground_weight","above_below_weight_ratio")]
pairs.panels(fit_growth_cordf)
fit_growth_cors <- cor(fit_growth_cordf, use = "complete.obs")
corrplot(fit_growth_cors, method = "circle",
         type = "lower",
         addCoef.col = "#333333",
         number.cex = 0.7)
```

## Fitness Difference Modeling

Here I will compare differences among groups. I will use tank, treatment, and population as explanatory factors in my analysis.

### Survival Modeling

I will start by modeling differences in survival. To do so, I will use a binomial GLMM with binary survival data as response variable, fixed explanatory factors treatment & population, and random factor tank.

```{r survival modeling trt}
# analysis of fixed effects
anova(lm(isAlive ~ pop + trt, data = fit_trim_env)) # both significant

# start by looking for evidence of random effects
# model with no random effect
surv_glm <- glm(isAlive ~ pop * trt,
               family = binomial(link = "logit"),
               data = fit_trim_env)
plot(surv_glm)
Anova(surv_glm, type = "III")
drop1(surv_glm) # keep everything
summary(surv_glm)

# now look at residuals
resid_surv <- data.frame(residuals = residuals(surv_glm), tank_ID = fit_trim_env$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_surv, main = "Residuals of Survival GLM by Tank") # tank doesn't seem to have any huge systematic effects, some differences in variance but not mean

# # try with random effects
# surv_glmm <- glmer(isAlive ~ pop * trt + (1 | tank_ID),
#                family = binomial(link = "logit"),
#                data = fit_trim_env)
# plot(surv_glmm)
# Anova(surv_glmm, type = "III")
# drop1(surv_glmm) # keep all
# summary(surv_glmm)

# check dispersion & multicollinearity
check_overdispersion(surv_glm) # looks good

# significance testing with tukey
surv_glm_poptrt <- glm(isAlive ~ pop_trt,
                       family = binomial,
                       data = fit_trim_env)
tuk_surv <- TukeyHSD(aov(surv_glm_poptrt), conf.level = 0.95)
cld_surv <- multcompLetters4(aov(surv_glm_poptrt), tuk_surv)

# get significance letters
letters_surv_df <- data.frame(letters = as.data.frame.list(cld_surv$pop_trt), popsite = names((cld_surv$pop_trt)$Letters))
letters_surv_df_red <- letters_surv_df[,c("letters.Letters","popsite")]
colnames(letters_surv_df_red) <- c("letters","popsite")
rownames(letters_surv_df_red) <- NULL
```

Rather than using treatment directly, I will try using salinity and temperature to represent temperature. Additionally, because tank_ID is confounded with temperature, I will try using x/y coordinates as a fixed effect and avoid having random effects altogether. If I do this, I can use a GLM. 

```{r survival modeling sal/temp}
# anova
anova(lm(isAlive ~ pop + median_sal + median_temp, data = fit_trim_env)) # pop, sal, temp all significant

# start with a glm
surv_glm_env <- glm(isAlive ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp,
                family = binomial(link = "logit"),
                data = fit_trim_env) # warning: glm.fit: fitted probabilities numerically 0 or 1 occurred

# check dispersion & multicollinearity
check_overdispersion(surv_glm_env) # looks good
vif(surv_glm_env) # pop is 21.827777, pop:median_temp is 26.376966

# inspect model
summary(surv_glm_env)
plot(surv_glm_env)
Anova(surv_glm_env, type = "III") # pop, sal, temp all signif
step(surv_glm_env)

# check for random effects
resid_surv_env <- data.frame(residuals = residuals(surv_glmm_env), tank_ID = fit_trim_env$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_surv_env, main = "Residuals of Survival GLM by Tank") # tank effects look minimal, but some amount of randomness - try a glmm

# try a glmm
surv_glmm_env <- glmer(isAlive ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp + (1|tank_ID),
                family = binomial(link = "logit"),
                data = fit_trim_env)
# Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model failed to converge with max|grad| = 0.0119148 (tol = 0.002, component 1)
# Warning in checkConv(attr(opt, "derivs"), opt$par, ctrl = control$checkConv,  :
#   Model is nearly unidentifiable: large eigenvalue ratio
#  - Rescale variables?

# check dispersion & multicollinearity
check_overdispersion(surv_glmm_env) # looks good
vif(surv_glmm_env) # similar numbers to the glm

# inspect model
summary(surv_glmm_env)
plot(surv_glmm_env)
Anova(surv_glmm_env, type = "III") # sal, temp signif - pop and any pop interactions not significant
drop1(surv_glmm_env) # remove pop:median_temp

# removing pop temp interaction
surv_glmm_env.1 <- glmer(isAlive ~ pop + median_sal + median_temp + pop:median_sal + (1|tank_ID),
                family = binomial(link = "logit"),
                data = fit_trim_env) 
# boundary (singular) fit: see help('isSingular')

# check dispersion & multicollinearity
check_overdispersion(surv_glmm_env.1) # looks good
vif(surv_glmm_env.1) # looks good

# inspect model
summary(surv_glmm_env.1)
plot(surv_glmm_env.1)
Anova(surv_glmm_env.1, type = "III") # intercept, sal, temp significant
drop1(surv_glmm_env.1) # remove pop:median_sal

# removing pop sal interaction, this is solely an additive model
surv_glmm_env.2 <- glmer(isAlive ~ pop + median_sal + median_temp + (1|tank_ID),
                family = binomial(link = "logit"),
                data = fit_trim_env) 
# boundary (singular) fit: see help('isSingular')

# check dispersion & multicollinearity
check_overdispersion(surv_glmm_env.2) # looks good
vif(surv_glmm_env.2) # looks good

# inspect model
summary(surv_glmm_env.2)
plot(surv_glmm_env.2)
Anova(surv_glmm_env.2, type = "III") # everything significant
drop1(surv_glmm_env.2) # final model
```

### Growth Modeling

I will now model differences in growth, namely changes in length and changes in weight. I no longer need to use a GLM, I can use an LMM since I am not working with a binary response variable. 

Notes on this dataset: The dataset used in this modeling approach excludes individuals that died. The reason for this is that there is no biologically relevant way to code this into the model - if I code the dead individuals as having no length or weight change, they will appear more fit than many other plants that lost length/biomass, and thus had negative length/weight changes. 

I will start by building models for length change (standardized) using treatment, then environmental variables, to explain differences in length change. 

```{r length modeling, standardized}
# quick check
anova(lm(length_std_change ~ pop + trt, data = fit_growth)) # both significant
hist(fit_growth$length_std_change) # left skewed, try log transformation
hist(log(fit_growth$length_std_change)) # yay it's normally distributed now

# start with lm since data is relatively normally distributed and not incorporating random effects yet
growlen_std_lm <- lm(log(length_std_change) ~ pop * trt, data = fit_growth)
plot(growlen_std_lm) # qq plot has some outliers deviating significantly from 1-1 line
anova(lm(length_std_change ~ pop * trt, data = fit_growth)) # both significant, interaction isn't
step(growlen_std_lm) # drop interaction
summary(growlen_std_lm)

# without interaction
growlen_std_lm.1 <- lm(log(length_std_change) ~ pop + trt, data = fit_growth)
plot(growlen_std_lm.1) # qq plot has some outliers deviating significantly from 1-1 line, but not as bad as before
anova(growlen_std_lm.1) # both significant
step(growlen_std_lm.1) # final model
summary(growlen_std_lm.1)

# check for random effects
resid_len <- data.frame(residuals = residuals(growlen_std_lm.1), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_len, main = "Residuals of Length LM by Tank") # some variation, very slight, try an lmm

# lmm
growlen_std_lmm <- lmer(log(length_std_change) ~ pop * trt + (1|tank_ID), data = fit_growth)
plot(growlen_std_lmm)
Anova(growlen_std_lmm, type = "III") # all significant
step(growlen_std_lmm) # keep all
summary(growlen_std_lmm)
```

I will look at separating out salinity and temperature for these models as well. I will only do this for the standardized data.

```{r length modeling sal/temp, standardized}
# quick check
anova(lm(length_std_change ~ pop + median_sal + median_temp, data = fit_growth)) # pop and temp signif, not sal

# we know we have skew, so log transform
len_saltemp_lm <- lm(log(length_std_change) ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp, data = fit_growth) # what family?
plot(len_saltemp_lm)
anova(len_saltemp_lm) # all signif except pop:median_temp interaction
step(len_saltemp_lm) # remove pop temp interaction
summary(len_saltemp_lm)

# without interaction
len_saltemp_lm.1 <- lm(log(length_std_change) ~ pop + median_sal + median_temp + pop:median_sal, data = fit_growth) # what family?
plot(len_saltemp_lm.1)
anova(len_saltemp_lm.1) # all signif
step(len_saltemp_lm.1) # final model
summary(len_saltemp_lm.1)

# check residuals for random effects
resid_len_saltemp <- data.frame(residuals = residuals(len_saltemp_lm.1), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_len_saltemp, main = "Residuals of Length LM by Tank") # some variation, very slight, try an lmm

# lmm
len_saltemp_lmm <- lmer(log(length_std_change) ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp + (1|tank_ID), data = fit_growth)
plot(len_saltemp_lmm)
Anova(len_saltemp_lmm, type = "III") # intercept, temp, pop:sal significant
step(len_saltemp_lmm) # drop temp/pop interaction
summary(len_saltemp_lmm)

len_saltemp_lmm.1 <- lmer(log(length_std_change) ~ pop + median_sal + median_temp + pop:median_sal + (1|tank_ID), data = fit_growth)
plot(len_saltemp_lmm.1)
Anova(len_saltemp_lmm.1, type = "III") # everything signif except sal
step(len_saltemp_lmm.1) # this is final model
summary(len_saltemp_lmm.1)

# check dispersion & multicollinearity
check_overdispersion(len_saltemp_lmm.1) # looks good
vif(len_saltemp_lmm.1) # suggests no collinearity

# # another option would be to do a glm with a gamma family, which is meant for skewed data
# len_saltemp_glm.1 <- glm(length_std_change ~ pop + median_temp, family = "gamma", data = fit_growth)
# plot(len_saltemp_glm.1)
# Anova(len_saltemp_glm.1, type = "III") # all signif
# step(len_saltemp_glm.1)
# summary(len_saltemp_glm.1)

# # check dispersion & multicollinearity
# check_overdispersion(len_saltemp_glm.1) # looks good
# vif(len_saltemp_glm.1) # suggests no collinearity, everything is around 1
```

Now look into weight. I am testing only standardized weight change.

```{r weight modeling, standardized}
# quick check
anova(lm(weight_std_change ~ pop + trt, data = fit_growth)) # pop and temp signif, not sal

# check distribution
hist(fit_growth$weight_std_change) # beautifully normally distributed omg

# linear model to start
growweightstd_lm <- lm(weight_std_change ~ pop * trt, data = fit_growth)
plot(growweightstd_lm)
anova(growweightstd_lm) # pop & trt signif
step(growweightstd_lm) # remove interaction

growweightstd_lm.1 <- lm(weight_std_change ~ pop + trt, data = fit_growth)
plot(growweightstd_lm.1)
anova(growweightstd_lm.1) # everything signif
step(growweightstd_lm.1) # final model
summary(growweightstd_lm.1)

# check for random effects
resid_weight <- data.frame(residuals = residuals(growweightstd_lm.1), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_weight, main = "Residuals of Weight GLM by Tank") # some randomness, try an lmm

# lmm weight model
growweightstd_lmm <- lmer(weight_std_change ~ pop * trt + (1|tank_ID), data = fit_growth)
plot(growweightstd_lmm)
Anova(growweightstd_lmm, type = "III") # nothing is significant
step(growweightstd_lmm) # drop the interaction
summary(growweightstd_lmm)

# dropping interaction
growweightstd_lmm.1 <- lmer(weight_std_change ~ pop + trt + (1|tank_ID), data = fit_growth)
plot(growweightstd_lmm.1)
Anova(growweightstd_lmm.1, type = "III") # both significant
step(growweightstd_lmm.1) # final model, although model without random effect has AIC slightly lower (but 144.76 vs 146.75 AIC, so can leave it)
summary(growweightstd_lmm.1)

# check dispersion & multicollinearity
check_overdispersion(growweightstd_lmm.1) # looks good
vif(growweightstd_lmm.1) # suggests no collinearity, everything is around 1
```

I will now look at this same data using salinity/temp in the place of treatment. Again, I will only do this for standardized data.

```{r weight modeling sal/temp, standardized}
# quick check
anova(lm(weight_std_change ~ pop + median_sal + median_temp, data = fit_growth)) # pop and temp signif, not sal

# linear model
weight_saltemp_lm <- lm(weight_std_change ~ pop * median_sal + median_temp, data = fit_growth)
plot(weight_saltemp_lm) # looks pretty good
anova(weight_saltemp_lm) # pop and temp signif
step(weight_saltemp_lm) # just keep pop and temp
summary(weight_saltemp_lm)

weight_saltemp_lm.1 <- lm(weight_std_change ~ pop + median_temp, data = fit_growth)
plot(weight_saltemp_lm.1) # looks pretty good
anova(weight_saltemp_lm.1) # both signif
step(weight_saltemp_lm.1) # final model
summary(weight_saltemp_lm.1)

# check for random effects
resid_weight_saltemp <- data.frame(residuals = residuals(weight_saltemp_lm.1), tank_ID = fit_growth$tank_ID)
boxplot(residuals ~ tank_ID, data = resid_weight_saltemp, main = "Residuals of Weight LM by Tank") # some randomness, try an lmm

# with random effect
growweight_saltemp_lmm <- lmer(weight_std_change ~ pop + median_sal + median_temp + pop:median_sal + pop:median_temp + (1|tank_ID), data = fit_growth)
plot(growweight_saltemp_lmm)
Anova(growweight_saltemp_lmm, type = "III") # median temp not significant
step(growweight_saltemp_lmm) # keep pop and temp, can keep random effect because AIC close to what i
summary(growweight_saltemp_lmm)

growweight_saltemp_lmm.1 <- lmer(weight_std_change ~ pop + median_temp + (1|tank_ID), data = fit_growth)
plot(growweight_saltemp_lmm.1)
Anova(growweight_saltemp_lmm.1, type = "III") # all significant
step(growweight_saltemp_lmm.1) # final model, although model without random effect has AIC slightly different (but 140.64 w vs 138.65 wo AIC, so can leave it)
summary(growweight_saltemp_lmm.1)

# check dispersion & multicollinearity
check_overdispersion(growweight_saltemp_lmm.1) # looks good
vif(growweight_saltemp_lmm.1) # suggests no collinearity, everything is around 1
```


### Modeling Outputs

Now I have my beautiful models! I'll write relevant results out for future reference:

```{r}
sink(paste0("results/eelgrass_fitness_models_",Sys.Date(),".txt"))

"Survival Model Pop/trt\n"
summary(surv_glm)
Anova(surv_glm, type = "III")

"Survival Model Env\n"
summary(surv_glmm_env.2)
Anova(surv_glmm_env.2, type = "III")


"Length Model Pop/trt\n"
summary(growlen_std_lm.1)
anova(growlen_std_lm.1)

"Length Model Env\n"
summary(len_saltemp_lm.1)
anova(len_saltemp_lm.1)

"Weight Model Pop/trt\n"
summary(growweightstd_lm.1)
anova(growweightstd_lm.1)

"Weight Model Env\n"
summary(weight_saltemp_lm.1)
anova(weight_saltemp_lm.1)

sink()
```