---
title: "eelgrass_fitness_modeling"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Eelgrass/BalticEelgrass")
```

```{r}
# libraries
library(dplyr)
library(tidyr)
library(ggplot2)
library(stringr)
library(lme4)
library(lmerTest)
library(multcompView)
library(car)
library(performance)

# data
fit <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
```

# Fitness Modeling

This script is to compare fitness across populations/treatments in the GO eelgrass mesocosm experiment. I may also use some site of origin environmental data to see if environmental factors explain performance differences, but this will come later (if at all).

## Setup

I need first to put together my data so that I have all the fitness and growth metrics possible.

```{r}
# data cleaning to just relevant columns
fit_trim <- fit[,c("DNA_label", "new_label", "pop", "genet", "bagnum", "tank_ID", "sal_full", "temp_full", "trt", "nbr_leaves_start", "sideShoot", "nbr_sideShootLeaves_start", "length_longest_leaf_start", "weight_start", "IsBlackLesions_start", "nbr_leaves_end", "nbr_new_leaves", "width_leaf", "Growth_leaf1", "Growth_leaf2", "Growth_leaf3", "Growth_leaf4", "Growth_leaf5", "Growth_leaf6", "Growth_leaf7", "Length_new_leaf1", "Length_new_leaf2", "Length_new_leaf3", "Length_new_leaf4", "Length_new_leaf5", "Length_new_leaf6", "Nbr_new_shoots", "Nbr_leaves_new_shoot1", "Length_new_shoot1_leaf1", "Length_new_shoot1_leaf2", "Length_new_shoot1_leaf3", "Length_new_shoot1_leaf4", "Length_new_shoot1_leaf5", "Length_new_shoot1_leaf6", "Nbr_leaves_new_shoot2", "Length_new_shoot2_leaf1", "Length_new_shoot2_leaf2", "Length_new_shoot2_leaf3", "Length_new_shoot2_leaf4", "Length_new_shoot2_leaf5", "below_ground_weight", "above_ground_weight", "weight_new_shoot1", "weight_new_shoot2", "total_weight", "lesions.", "isAlive", "Length_flower1", "Length_bract1")]

# binary survival column
fit_trim$isAlive <- ifelse(fit_trim$isAlive == TRUE, 1, 0)

# pop_trt factor for later significance testing
fit_trim$pop_trt <- as.factor(paste0(fit_trim$pop, "_", fit_trim$trt))

# ensure correct coding of relevant variables
fit_trim$pop <- as.factor(fit_trim$pop)
fit_trim$tank_ID <- as.factor(fit_trim$tank_ID)
fit_trim$trt <- as.factor(fit_trim$trt)

# length and weight growth
# for the growth analyses, we want a df with no flowering individuals and no dead individuals
fit_growth <- fit_trim[is.na(fit_trim$Length_flower1),]
fit_growth <- fit_growth[fit_growth$isAlive == 1,]

# length
fit_growth$length_growth <- rowSums(fit_growth[,c("Growth_leaf1","Growth_leaf2","Growth_leaf3","Growth_leaf4","Growth_leaf5","Growth_leaf6","Growth_leaf7","Length_new_leaf1","Length_new_leaf2","Length_new_leaf3","Length_new_leaf4","Length_new_leaf5","Length_new_leaf6","Length_new_shoot1_leaf1","Length_new_shoot1_leaf2","Length_new_shoot1_leaf3","Length_new_shoot1_leaf4","Length_new_shoot1_leaf5","Length_new_shoot1_leaf6","Length_new_shoot2_leaf1","Length_new_shoot2_leaf2","Length_new_shoot2_leaf3","Length_new_shoot2_leaf4","Length_new_shoot2_leaf5")], na.rm = T)

# weight
fit_growth$weight_growth <- fit_growth$total_weight - fit_growth$weight_start

# ratio of above vs below ground weight
fit_growth$above_below_weight_ratio <- fit_growth$above_ground_weight / fit_growth$below_ground_weight
```

## Fitness Difference Modeling

Here I will compare differences among groups. I will use tank, treatment, (???coast), and population as explanatory factors in my analysis.

### Survival Modeling

I will start by modeling differences in survival. To do so, I will use a binomial GLMM with binary survival data as response variable, fixed explanatory factors treatment & population, and random factor tank.

```{r}
# analysis of fixed effects
anova(lm(isAlive ~ pop + trt, data = fit_trim)) # both significant

# now try building a model
mod1 <- glmer(isAlive ~ pop * trt + (1 | tank_ID), 
               family = binomial(link = "logit"),
               data = fit_trim)
plot(mod1)
Anova(mod1, type = "III")
drop1(mod1) # keep all
summary(mod1)

# significance testing with tukey
mod1.1 <- glmer(isAlive ~ pop_trt + (1 | tank_ID), # getting boundary (singular) fit: see help('isSingular')
               family = binomial(link = "logit"),
               data = fit_trim)
mod1.2 <- glm(isAlive ~ pop_trt,
               family = binomial,
               data = fit_trim)

tuk_surv <- TukeyHSD(aov(mod1.2), conf.level = 0.95)
cld_surv <- multcompLetters4(aov(mod1.2), tuk_surv)

# multcomp letters to get letters
letters_surv_df <- data.frame(letters = as.data.frame.list(cld_surv$pop_trt), popsite = names((cld_surv$pop_trt)$Letters))
letters_surv_df_red <- letters_surv_df[,c("letters.Letters","popsite")]
colnames(letters_surv_df_red) <- c("letters","popsite")
rownames(letters_surv_df_red) <- NULL
```


### Growth Modeling

I will now model differences in growth, namely changes in length and changes in weight. I no longer need to use a GLMM, I can use an LMM since I am not working with a binary response variable.

```{r}
# start with length
# analysis of fixed effects
anova(lm(length_growth ~ pop + trt, data = fit_growth)) # both significant

# now try building a model
mod2 <- lmer(length_growth ~ pop * trt + (1 | tank_ID), data = fit_growth)
plot(mod2)
Anova(mod2, type = "III") # all signif except trt * pop
drop1(mod2) # drop interaction
summary(mod2)

mod2.1 <- lmer(length_growth ~ pop + trt + (1 | tank_ID), data = fit_growth)
plot(mod2.1)
Anova(mod2.1, type = "III") # all signif
drop1(mod2.1) # keep all
summary(mod2.1)

# significance testing with tukey
mod2.2 <- lmer(length_growth ~ pop_trt + (1 | tank_ID), data = fit_growth) # singular, throws error
tuk_len <- TukeyHSD(aov(mod2.2), conf.level = 0.95)
cld_len <- multcompLetters4(aov(mod2.2), tuk_len)

# try a linear model
mod2.3 <- lm(length_growth ~ pop * trt, data = fit_growth)
mod2.4 <- lm(length_growth ~ pop_trt, data = fit_growth)

```

