---
title: "eelgrass_fitness_offsets_clim"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Documents/Github/BalticEelgrass")

# setwd
setwd("/Users/camillerumberger/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse) # for data cleaning
library(stringr) # for data cleaning
library(dplyr) # for data cleaning
library(ggplot2) # for plotting
library(raster) # for environmental pixels
library(viridis) # plotting
library(sf) # mapping
library(s2) # mapping
library(rnaturalearth) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(stars) # mapping
library(ggpubr) # plotting
library(plotrix) # std error
library(dplyr)
library(gstat)
library(terra)
library(lwgeom)
library(officer)
library(rvg)

# data
cf_off <- read.csv("results/fit_results/climate_change_fitness_offsets_2026-05-28.csv")[,-1]
sites <- read.csv("data/experiment/eelgrass_exp_sites.csv")
poptrts_df <- read.csv("data/experiment/poptrts_df.csv")
surv_dat <- read.csv("data/experiment/eelgrass_mesocosm_start_end_complete.csv")[,-1]
```

# Plotting Fitness Offsets

This script plots current-future fitness offsets from W & E coast conditions to future W & E coast conditions, respectively. Fitness offsets are calculated in the script "eelgrass_fitness_offsets.Rmd"

## Data Prep

Combine data

```{r}
# check data
head(cf_off)
head(sites)

# rename
sites$pop <- sites$site_abbrev

# combine
clim_offs <- merge(sites, cf_off, by = "pop")

# survival data indiv level
# pop/trt info
w_pops <- c("VAT", "VIK", "BAR", "HOG")
e_pops <- c("KUR", "KAL", "HOR", "BJO")
w_trts <- c("TempControl-21psu", "TempWarm-16psu")
e_trts <- c("TempControl-7psu", "TempWarm-5psu")
curr_trts <- c("TempControl-21psu", "TempControl-7psu")
fut_trts <- c("TempWarm-16psu", "TempWarm-5psu")

# add pop/trt metadata
fit_full$trt <- gsub("_", "-", fit_full$trt)
fit_full$coast_pop <- ifelse(fit_full$pop %in% w_pops, "west", "east")
fit_full$coast_trt <- ifelse(fit_full$trt %in% w_trts, "west", "east")
fit_full$curr_fut <- ifelse(fit_full$trt %in% curr_trts, "current", "future")

# surv
surv_all <- fit_full[,c("pop", "trt", "bagnum", "tank_ID", "isAlive", "coast_pop", "coast_trt", "curr_fut")]
surv_all$isAlive <- ifelse(surv_all$isAlive == TRUE, 1, 0)

# pop avgs for plotting
surv_avgs <- surv_all %>% 
  group_by(pop, coast_trt, curr_fut) %>% 
  summarise(surv = mean(isAlive), surv_se = std.error(isAlive), .groups = "drop_last")

# just relevant west/east combos
surv_avgs_w <- surv_avgs %>% filter(pop %in% w_pops & coast_trt == "west")
surv_avgs_e <- surv_avgs %>% filter(pop %in% e_pops & coast_trt == "east")
surv_avgs_df <- rbind(surv_avgs_w, surv_avgs_e) %>% rename(coast = coast_trt) %>% as.data.frame()
```

## Plotting

### Maps

Start by visualizing fitness offsets as maps

I'm going to start off making "climate change" maps, which visualize the offset between current/future W and current/future E on the same plot.

First prep the plotting range.

```{r}
# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 8, xmax = 20, ymin = 54, ymax = 61))
```

#### Survival Maps

```{r}
# relevant data
surv_cf <- clim_offs %>% filter(metric == "surv")

# quick summaries
mean(surv_cf[surv_cf$coast == "West",]$CF_abs_offset)
mean(surv_cf[surv_cf$coast == "East",]$CF_abs_offset)
std.error(surv_cf[surv_cf$coast == "West",]$CF_abs_offset)
std.error(surv_cf[surv_cf$coast == "East",]$CF_abs_offset)

# mapping
# absolute survival
###################
surv_abs_map_curr <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = abs_current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", direction = 1, limits = c(0, 1)) +
  scale_color_distiller(palette = "RdYlGn", direction = 1, limits = c(0, 1)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(A) Survival in Current \nTreatment", x = "Longitude", y = "Latitude", fill = "Survival")
surv_abs_map_curr

surv_abs_map_fut <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = abs_future), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", direction = 1, limits = c(0, 1)) +
  scale_color_distiller(palette = "RdYlGn", direction = 1, limits = c(0, 1)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(B) Survival in Future \nTreatment", x = "Longitude", y = "Latitude", fill = "Survival")
surv_abs_map_fut
###################

# relative survival
###################
# surv_rel_map <- ggplot(data = world) +
#   theme_bw()+
#   geom_sf(data = world_crop, fill = 'antiquewhite1') +
#   geom_point(data = surv_cf, aes(x = long, y = lat, 
#                                  fill = CF_relative), 
#              shape = 21, size = 4) +
#   scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
#   scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
#   geom_label_repel(data = surv_cf,
#                   aes(x=long, y=lat, label = pop),
#                   fill = "white")+
#   theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
#         panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
#   labs(title = expression(paste("(B) Common Future Relative Survival \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nSurvival")
# surv_rel_map
###################

# abs surv offset
#################
surv_abs_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = surv_cf, aes(x = long, y = lat, 
                                 fill = CF_abs_offset), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", direction = 1, limits = c(-1, 0)) +
  scale_color_distiller(palette = "RdYlGn", direction = 1, limits = c(-1, 0)) +
  geom_label_repel(data = surv_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(C) Survival Offset in \nFuture Treatment", x = "Longitude", y = "Latitude", fill = "Fitness \nOffset")
surv_abs_off_map
#################

# abs surv offset
#################
# surv_rel_off_map <- ggplot(data = world) +
#   theme_bw()+
#   geom_sf(data = world_crop, fill = 'antiquewhite1') +
#   geom_point(data = surv_cf, aes(x = long, y = lat, 
#                                  fill = CF_rel_offset), 
#              shape = 21, size = 4) +
#   scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
#   scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
#   geom_label_repel(data = surv_cf,
#                   aes(x=long, y=lat, label = pop),
#                   fill = "white")+
#   theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
#         panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
#   labs(title = expression(paste("(D) Common Future Relative Survival \nOffset in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nFitness Offset")
# surv_rel_off_map
#################

# combine plots
###############
#surv_cf_maps <- ggarrange(plotlist = c(surv_abs_map, surv_rel_map, surv_abs_off_map, surv_rel_off_map))
surv_maps <- ggarrange(plotlist = c(surv_abs_map_curr, surv_abs_map_fut, surv_abs_off_map), nrow = 1)
surv_maps
###############
```

Also interpolate survival around coast

```{r}
## prep dat
###########
# check data
head(surv_cf)

# order columns
surv_interpolation_coasts <- surv_cf %>% dplyr::select(-CF_rel_offset, -CF_relative, -metric, -site_abbrev) %>% rename(order = pop_ord) %>% arrange(order) %>% as.data.frame()
###########

## prep geom
############
sf::sf_use_s2(FALSE)

# load and crop coastline
coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- st_crop(coast, c(xmin = 8.5, xmax = 20, ymin = 54, ymax = 61))

# turn multiline into single lines, id + inspect pieces
coast_parts <- st_cast(coast_crop, "MULTILINESTRING") |> st_cast("LINESTRING")
coast_parts$id <- seq_len(nrow(coast_parts))

plot(st_geometry(coast_parts))
cent <- st_coordinates(st_centroid(st_geometry(coast_parts)))
text(cent, labels = coast_parts$id, col = "red", cex = 1)

# keep only the relevant piece -- this single line covers BOTH coasts
keep_ids <- 24
coast_sweden <- coast_parts[coast_parts$id %in% keep_ids, ]
plot(st_geometry(coast_sweden), col = "lightblue")

# trim area beyond desired interpolation region to west of sweden
# eyeball the unwanted region approx lon/lat extent
rm_bbox <- st_as_sfc(st_bbox(c(xmin = 8.5, xmax = 10.5, ymin = 58, ymax = 60), crs = st_crs(coast_sweden)))
plot(st_geometry(coast_sweden), col = "lightblue")
plot(rm_bbox, add = TRUE, border = "red")

# identify main coast region
main_coast <- st_difference(coast_sweden, rm_bbox)
plot(st_geometry(main_coast), col = "darkblue")

# split coast
coast_seg <- st_segmentize(main_coast, units::set_units(200, "m")) # vertex every ~200 m

coast_pts <- coast_seg %>% st_cast("MULTIPOINT") %>% st_cast("POINT") %>% st_sf(geometry = .)

seg_len <- st_distance(coast_pts[-1, ], coast_pts[-nrow(coast_pts), ], by_element = TRUE)
coast_pts$dist_along <- c(0, cumsum(as.numeric(seg_len)))

kur_sf <- st_as_sf(surv_interpolation_coasts[surv_interpolation_coasts$pop == "KUR", ], coords = c("long", "lat"), crs = 4326)
kur_idx  <- st_nearest_feature(kur_sf, coast_pts)
kur_dist <- coast_pts$dist_along[kur_idx]

# sanity check
coast_pts$side <- ifelse(coast_pts$dist_along <= kur_dist, "West", "East")
plot(st_geometry(coast_pts), col = ifelse(coast_pts$dist_along <= kur_dist, "blue", "red"))

coast_w <- coast_pts[coast_pts$side == "West", ] %>% st_combine() %>% st_cast("LINESTRING")
coast_e <- coast_pts[coast_pts$side == "East", ] %>% st_combine() %>% st_cast("LINESTRING")

# buffers + grid
dist <- 0.6  # roughly miles

buffer_w <- st_buffer(st_union(coast_w), dist)
buffer_e <- st_buffer(st_union(coast_e), dist)
buffered_region <- st_union(buffer_w, buffer_e)

grid <- st_make_grid(buffered_region, cellsize = 0.05, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
############

## interpolate
##############
sf_surv_w <- st_as_sf(surv_interpolation_coasts[surv_interpolation_coasts$coast == "West", ], coords = c("long", "lat"), crs = 4326)
sf_surv_e <- st_as_sf(surv_interpolation_coasts[surv_interpolation_coasts$coast == "East", ], coords = c("long", "lat"), crs = 4326)

interpolated_curr_w <- idw(abs_current~1, sf_surv_w, grid_int)
interpolated_fut_w <- idw(abs_future~1,  sf_surv_w, grid_int)
interpolated_cf_w <- idw(CF_abs_offset~1, sf_surv_w, grid_int)

interpolated_curr_e <- idw(abs_current~1, sf_surv_e, grid_int)
interpolated_fut_e <- idw(abs_future~1,  sf_surv_e, grid_int)
interpolated_cf_e <- idw(CF_abs_offset~1, sf_surv_e, grid_int)
##############

## combine west + east blend southern tip
#########################################
grid_cent <- st_centroid(grid_int)
nearest_grid <- st_nearest_feature(grid_cent, coast_pts)
grid_side <- coast_pts$side[nearest_grid]

grid_lat <- st_coordinates(grid_cent)[, "Y"]
kur_lat <- st_coordinates(kur_sf)[, "Y"]

bandwidth_north_m <- 40000
bandwidth_north_deg <- bandwidth_north_m / 111320

d_lat <- grid_lat - kur_lat

w_west <- ifelse(grid_side == "West", pmin(1, pmax(0, d_lat / bandwidth_north_deg)), 0)
w_east <- 1 - w_west
interpolated_curr <- interpolated_curr_w
interpolated_curr$var1.pred <- w_west * interpolated_curr_w$var1.pred + w_east * interpolated_curr_e$var1.pred

interpolated_fut <- interpolated_fut_w
interpolated_fut$var1.pred <- w_west * interpolated_fut_w$var1.pred + w_east * interpolated_fut_e$var1.pred

interpolated_cf <- interpolated_cf_w
interpolated_cf$var1.pred <- w_west * interpolated_cf_w$var1.pred + w_east * interpolated_cf_e$var1.pred
#########################################

## rasterize
############
ext <- ext(st_bbox(interpolated_curr))
rast_empty <- rast(ext, resolution = 0.05, crs = "+proj=longlat +datum=WGS84")

sf_ls <- list(interpolated_curr, interpolated_fut, interpolated_cf)
rast_ls <- list()
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_curr", "interpolated_fut", "interpolated_cf")

# save
saveRDS(rast_all, "results/fit_results/survival_interpolated.RDS")
############

## plotting setup
#################
# color ramp
pal <- colorRamp(c("green", "red"))

# masking region for interpolation
interpolated_curr_masked <- interpolated_curr %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_fut_masked <- interpolated_fut %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_cf_masked <- interpolated_cf %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
#################

# plot curr
###########
surv_curr_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_curr_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = surv_interpolation_coasts, aes(x = long, y = lat, fill = abs_current), shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 1), direction = 1) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 1), direction = 1) +
  geom_label_repel(data = surv_interpolation_coasts,
                  aes(x=long, y=lat, label = pop),
                  fill = "white",
                  size = 2.5)+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(A) Absolute Survival", subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Survival", col = "Survival")
surv_curr_coast
###########

# plot fut
###########
surv_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_fut_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = surv_interpolation_coasts, aes(x = long, y = lat, fill = abs_future), shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 1), direction = 1) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 1), direction = 1) +
  geom_label_repel(data = surv_interpolation_coasts,
                  aes(x=long, y=lat, label = pop),
                  fill = "white",
                  size = 2.5)+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(B) Absolute Survival", subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Survival", col = "Survival")
surv_fut_coast
###########

# plot cf
###########
surv_cf_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_cf_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = surv_interpolation_coasts, aes(x = long, y = lat, fill = CF_abs_offset), shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(-1, 0), direction = 1) +
  scale_color_distiller(palette = "RdYlGn", limits = c(-1, 0), direction = 1) +
  geom_label_repel(data = surv_interpolation_coasts,
                  aes(x=long, y=lat, label = pop),
                  fill = "white",
                  size = 2.5)+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(C) Survival Offset", subtitle = "Current to Future Treatment", x = "Longitude", y = "Latitude", fill = "Survival \nOffset", col = "Survival \nOffset")
surv_cf_coast
###########

# export plots
##############
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_curr_coast), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_fut_coast), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_cf_coast), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))

print(doc, target = "results/figures/climchg_fit_maps.pptx")
##############
```

#### Length Plots

```{r}
# relevant data
len_cf <- clim_offs %>% filter(metric == "len")

# mapping
# absolute length
#################
len_abs_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = abs_future), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,2)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,2)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Length Change in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Length Change")
len_abs_map
###################

# relative length
#################
len_rel_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = CF_relative), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(0,1)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Common Future Relative Length Change \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nLength Change")
len_rel_map
#################

# abs len offset
################
len_abs_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = CF_abs_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-2,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-2,0)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Common Future Length Offset \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Fitness Offset")
len_abs_off_map
################

# rel len offset
################
len_rel_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = len_cf, aes(x = long, y = lat, 
                                 fill = CF_rel_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = len_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Common Future Relative Length \nOffset in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nFitness Offset")
len_rel_off_map
################

# combine plots
###############
len_cf_maps <- ggarrange(plotlist = c(len_abs_map, len_rel_map, len_abs_off_map, len_rel_off_map))
###############
```

#### Weight Plots

```{r}
# relevant data
weigh_cf <- clim_offs %>% filter(metric == "weigh")

# mapping
# absolute weight chg
###################
weigh_abs_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = abs_future), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Weight Change in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Weight Change")
weigh_abs_map
###################

# relative weight chg
###################
weigh_rel_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = CF_relative), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-12,6)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-12,6)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Common Future Relative Weight Change \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nWeight Change")
weigh_rel_map
###################

# abs weigh offset
#################
weigh_abs_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = CF_abs_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-1,0)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Common Future Weight Change Offset \nin Future Treatments")), x = "Longitude", y = "Latitude", fill = "Fitness Offset")
weigh_abs_off_map
#################

# abs weigh offset
#################
weigh_rel_off_map <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'white') +
  geom_point(data = weigh_cf, aes(x = long, y = lat, 
                                 fill = CF_rel_offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = 1, limits = c(-13,5)) +
  scale_fill_viridis_c(option = "magma", direction = 1, limits = c(-13,5)) +
  geom_label_repel(data = weigh_cf,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Common Future Relative Weight Change \nOffset in Future Treatments")), x = "Longitude", y = "Latitude", fill = "Relative \nFitness Offset")
weigh_rel_off_map
#################

# combine plots
###############
weigh_cf_maps <- ggarrange(plotlist = c(weigh_abs_map, weigh_rel_map, weigh_abs_off_map, weigh_rel_off_map))
###############
```

#### Export

```{r}
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_rel_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_rel_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_abs_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_rel_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_abs_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_rel_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_abs_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_rel_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_abs_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_rel_off_map), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))

print(doc, target = "results/figures/climchg_fit_offset_maps.pptx")

doc2 <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_cf_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = len_cf_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = weigh_cf_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) 

print(doc2, target = "results/figures/climchg_fit_offset_maps_combined.pptx")

c(surv_abs_map_curr, surv_abs_map_fut, surv_abs_off_map, surv_maps)

doc3 <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_map_curr), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_map_fut), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_abs_off_map), location = ph_location(left = 1, top = 1, width = 13, height = 8)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = surv_maps), location = ph_location(left = 1, top = 1, width = 13, height = 8)) 
print(doc3, target = "results/figures/climchg_fit_offset_maps_surv.pptx")
```
