---
title: "eelgrass_LEA_gea"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: pdf_document
---

# GEA Modeling with LEA

This script is for conducting a GEA study using LEA, specifically for the project "VR_Zostera_genomic_offset." It includes training of the models using seascape genomic data, as well as identification of loci putatively under selection.  

## Setup

```{r setup, include = F}
# markdown options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/BalticEelgrass/")
setwd("~/Documents/Github/BalticEelgrass")

# libraries
library(LEA) # for LEA gea
library(qvalue) # for determination of significance
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(forcats) # data vis
library(vcfR) # for vcf files

# read in data
## indiv dat
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
sea_pops <- read.csv("data/seascape_data/population_data.csv")[,-1]
exp_pops <- read.csv("data/experiment/eelgrass_exp_sites.csv")

## gen dat
lfmm_mll <- read.lfmm("data/seascape_data/gen_MLL_imp_mat.lfmm")
geno_mll <- read.geno("data/seascape_data/gen_MLL_imp_mat.geno")
lfmm_full <- read.lfmm("data/seascape_data/gen_full_rm_imp_mat.lfmm")
geno_full <- read.geno("data/seascape_data/gen_full_rm_imp_mat.geno")
lfmm_full_rm <- read.lfmm("data/seascape_data/eel_full_rm.lfmm") # this one has only snps found in both full and mll
eelgrass_vcf_MLL <- read.vcfR(
  "data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.vcf")
eelgrass_vcf_full <- read.vcfR(
  "data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.vcf")

## env dat
curr_full_env <- read.env("data/EnvDat/full_seascapeEnv_2025-12-03.env")
curr_mll_env <- read.env("data/EnvDat/mll_seascapeEnv_2025-12-03.env")
sea_scaled_mll <- read.csv("data/EnvDat/sea_scaled_envdat_mll_2025-12-03.csv")
sea_scaled_full <- read.csv("data/EnvDat/sea_scaled_envdat_full_2025-12-03.csv")
```

## Data Prep

### Individual Data Prep

I will start by processing the individual data to work with both MLL and full datasets. 

```{r}
# inspect data
sea_pops
indiv_list
exp_pops

# rename sea pop columns
colnames(sea_pops) <- c("region","site_name","pop","long","lat")

# rename exp pop columns
colnames(exp_pops) <- c("site_name", "pop", "long", "lat")

# add metadata to inds
indiv_list <- as.data.frame(indiv_list)
colnames(indiv_list) <- c("mll", "ind")
indiv_list$pop <- str_split_fixed(indiv_list$ind, "-", 2)[,1]
inds <- merge(sea_pops, indiv_list, by = c("pop"))

# reorder by mll
inds_ord <- inds[order(inds$mll),]

# unique mll inds
inds_mll <- inds_ord[unique(inds_ord$mll),]
```

### Environmental Data Prep

Environmental Data has been pre-processed using the script "EnvDatProcess.Rmd". This script scales all environmental data together, then splits it into the relevant data sets for seascape, experimental pops, and treatments before outputting the correct formats for each GO method. Here, I just ensure that column names are consistent and that the order of the .env files matches with our other metadata

```{r}
# inspect data
# seascape
curr_full_env
curr_mll_env
sea_scaled

# exp sites
exp_env
exp_scaled

# trt
trt_env
trt_scaled

# rownames for trt
rownames(trt_env) <- trt_scaled$trt

# colnames for env files
colnames(curr_full_env) <- colnames(curr_mll_env) <- colnames(exp_env) <- colnames(trt_env) <- c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")

# now check the order of data
exp_env == exp_scaled[,9:12] 
trt_env == trt_scaled[,6:9]
curr_mll_env == sea_scaled_mll[,15:18]
curr_full_env == sea_scaled_full[,15:18]

# add pops as rownames
rownames(exp_env) <- exp_scaled$pop
rownames(curr_mll_env) <- sea_scaled_mll$pop
rownames(curr_full_env) <- sea_scaled_full$pop
```

### Genomic Data Prep

Seascape genomic data comes from Steffi Ries and Marlene Jahnke at the University of Gothenburg, who had already processed the genomic dataset (check what part of bioinformatic pipeline had been done already). I further process this data to impute missing genomic data and remove non-variable genomic sites in the script "GenDatProcess.Rmd"

```{r}
# inspect
lfmm_mll[1:10,1:10]
lfmm_full[1:10,1:10]

# add pops as rownames
rownames(lfmm_mll) <- inds_mll$pop
rownames(lfmm_full) <- inds_ord$pop
```


## GEA Study

Now that I have prepped and cleaned my data, I will use it to train LFMM models used to identify outlier loci. Importantly, I do not need to trim or impute my genomic data - this has been done in the script "GenDatProcess.Rmd". I also do not determine ancestry, Steffi has done this and recovered K = 12. 

```{r}
# fitting model
mod_lfmm_full <- lfmm2(input = lfmm_full,
                       env = curr_full_env, 
                       K = 12,
                       effect.sizes = T)
mod_lfmm_MLL <- lfmm2(input = lfmm_mll,
                      env = curr_mll_env,
                      K = 12,
                      effect.sizes = T)

# perform test
pv_lfmm_full <- lfmm2.test(object = mod_lfmm_full,
                       input = lfmm_full,
                       env = curr_full_env, 
                       full = TRUE)
pv_lfmm_MLL <- lfmm2.test(object = mod_lfmm_MLL,
                          input = lfmm_mll,
                          env = curr_mll_env, 
                          full = TRUE)

# pvalues
pv_full <- pv_lfmm_full$pvalues
pv_MLL <- pv_lfmm_MLL$pvalues

# qvalues
qv_full <- qvalue(pv_full, fdr.level = 0.01)
qv_MLL <- qvalue(pv_MLL, fdr.level = 0.01)

# candidates
candidates_full <- which(qv_full$significant)
candidates_pvals_full <- (pv_full)[candidates_full]

candidates_MLL <- which(qv_MLL$significant)
candidates_pvals_MLL <- (pv_MLL)[candidates_MLL]

# using FDR to filter out false positives
length(which(qv_full$qvalues < 0.01)) # how many SNPs have an FDR < 1%? 358
length(which(qv_MLL$qvalues < 0.01)) # how many SNPs have an FDR < 1%? 58

# identify which SNPs have qval less than 1%
cands_full <- colnames(lfmm_full)[which(qv_full$qvalues < 0.01)]
cands_index_full <- which(qv_full$qvalues < 0.01)

cands_MLL <- colnames(lfmm_mll)[which(qv_MLL$qvalues < 0.01)]
cands_index_MLL <- which(qv_MLL$qvalues < 0.01)

# df for candidates
cand_df_full <- data.frame(index = colnames(lfmm_full),
                           pvals = pv_full, qvals = qv_full$qvalues)
cand_df_MLL <- data.frame(index = colnames(lfmm_mll),
                          pvals = pv_MLL, qvals = qv_MLL$qvalues)

# get the actual position info from vcf
# first we need chromosome assignments, should be in the vcf
mll_posinfo <- as.data.frame(eelgrass_vcf_MLL@fix[,1:2])
full_posinfo <- as.data.frame(eelgrass_vcf_full@fix[,1:2])
mll_posinfo$POS <- as.numeric(mll_posinfo$POS)
full_posinfo$POS <- as.numeric(full_posinfo$POS)

mll_gtinfo <- eelgrass_vcf_MLL@gt
full_gtinfo <- eelgrass_vcf_full@gt

names_mll <- c()
for (i in 1:nrow(mll_gtinfo)) {
  name = paste0("V",i)
  names_mll = c(names_mll, name)
}
names_full <- c()
for (i in 1:nrow(full_gtinfo)) {
  name = paste0("V",i)
  names_full = c(names_full, name)
}

rownames(mll_gtinfo) <- names_mll
rownames(full_gtinfo) <- names_full

# find the names that are not in the filtered version
nonvar_mll <- which(!rownames(mll_gtinfo) %in% colnames(gen_MLL_imp_trim))
nonvar_full <- which(!rownames(full_gtinfo) %in% colnames(gen_full_imp_trim))

# remove from pos df
mll_posinfo_rm <- mll_posinfo[-(nonvar_mll),]
full_posinfo_rm <- full_posinfo[-(nonvar_full),]

# now bind with candidate dfs
mll_poscands <- cbind(mll_posinfo_rm, cand_df_MLL)
full_poscands <- cbind(full_posinfo_rm, cand_df_full)

# overlap in candidates?
overlapping_mll_all <- which(mll_poscands$POS %in% full_poscands$POS)
overlapping_full_all <- which(full_poscands$POS %in% mll_poscands$POS)
length(overlapping_full_all) # 2395 SNPs in common overall
overlapping_mll <- which(mll_poscands[mll_poscands$qvals < 0.01,]$POS %in% full_poscands[full_poscands$qvals < 0.01,]$POS)
overlapping_full <- which(full_poscands[full_poscands$qvals < 0.01,]$POS %in% mll_poscands[mll_poscands$qvals < 0.01,]$POS)
length(overlapping_mll) # 5 overlapping candidates

overlap_cands_pos <- full_poscands[full_poscands$qvals < 0.01,][overlapping_full, c("CHROM","POS")]

# environmental associations of loci?
# pull out effect sizes
mll_effect <- as.data.frame(mod_lfmm_MLL@B)
full_effect <- as.data.frame(mod_lfmm_full@B)

# identify strongest effect size
mll_effect$pred <- colnames(mll_effect)[max.col(mll_effect, ties.method = "first")]
full_effect$pred <- colnames(full_effect)[max.col(full_effect, ties.method = "first")]

# name of locus + add predictor data
mll_effect$index <- rownames(mll_effect)
full_effect$index <- rownames(full_effect)

# combine with candidate data
mll_poscands_df <- merge(mll_poscands, mll_effect, by = "index")
full_poscands_df <- merge(full_poscands, full_effect, by = "index")

# explicitly designate candidates
mll_poscands_df$cand <- ifelse(mll_poscands_df$qvals < 0.01, TRUE, FALSE)
full_poscands_df$cand <- ifelse(full_poscands_df$qvals < 0.01, TRUE, FALSE)

# quick summary
table(mll_poscands_df[mll_poscands_df$cand == T,]$pred)
#   max_temp_scaled  median_sal_scaled median_temp_scaled     min_sal_scaled 
#                23                 17                 10                  8 

table(full_poscands_df[full_poscands_df$cand == T,]$pred)
#   max_temp_scaled  median_sal_scaled median_temp_scaled     min_sal_scaled 
#               104                 97                 37                120

# output these outlier loci
write.csv(full_poscands_df, "results/GO_results/GEA/gea_cands_lea_full.csv", row.names = F)
write.csv(mll_poscands_df, "results/GO_results/GEA/gea_cands_lea_mll.csv", row.names = F)
```

## Plotting

Now that I have my outliers, I want to create Manhattan plots. 

```{r}
# temporary position
cand_df_full$pos <- 1:length(cand_df_full$index)
cand_df_MLL$pos <- 1:length(cand_df_MLL$index)

# manhattan plots
plot(cand_df_full$pos, -log10(cand_df_full$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs Full Dataset",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df_full$pos[cand_df_full$qvals < 0.01], 
       -log10(cand_df_full$pvals[cand_df_full$qvals < 0.01]),
       pch = 19,
       col = "magenta4")

plot(cand_df_MLL$pos, -log10(cand_df_MLL$pvals), 
     pch = 19, col = "darkslategray4",
     main = "Candidate SNPs MLL Dataset",
     xlab = "Position",
     ylab = "-log10(Pvals)")
points(cand_df_MLL$pos[cand_df_MLL$qvals < 0.01], 
       -log10(cand_df_MLL$pvals[cand_df_MLL$qvals < 0.01]),
       pch = 19,
       col = "magenta4")

# nicer plots
mll_plot <- ggplot(mll_poscands, aes(x = as.numeric(POS), y = -log10(pvals)))+
  theme_bw()+
  geom_point(aes(color=as.factor(CHROM), alpha = 0.8)) +
  scale_color_manual(values = rep(c("cadetblue", "skyblue"), 68))+
  geom_point(data = mll_poscands[mll_poscands$qvals < .01,], aes(x = as.numeric(POS), y = -log10(pvals)), color = "magenta4")+
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5), legend.position = "none")+
  ggtitle("Candidate Loci (Clone Corrected)") +
  labs(x = "Position", y = expression(-log[10](p-value)))
mll_plot

full_plot <- ggplot(full_poscands, aes(x = as.numeric(POS), y = -log10(pvals)))+
  theme_bw()+
  geom_point(aes(color=as.factor(CHROM), alpha = 0.8)) +
  scale_color_manual(values = rep(c("cadetblue", "skyblue"), 68))+
  geom_point(data = full_poscands[full_poscands$qvals < .01,], aes(x = as.numeric(POS), y = -log10(pvals)), color = "magenta4")+
  theme(axis.text.x=element_text(angle = 90, vjust = 0.5), legend.position = "none")+
  ggtitle("Candidate Loci (Uncorrected)") +
  labs(x = "Position", y = expression(-log[10](p-value)))
full_plot
```
