---
title: "eelgrass_LEA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: pdf_document
---

# Offset Modeling with LEA

This script is for offset calculation using LEA, specifically for the project "VR_Zostera_genomic_offset." It includes training of the models using seascape genomic data, as well as calculation of genomic offset for experimental populations in each mesocosm treatment. 

## Setup

```{r setup, include = F}
# markdown options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/BalticEelgrass/")

# libraries
library(LEA) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(vegan) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(terra) # interpolation
library(stars) # interpolation
library(gstat) # interpolation

# read in data
## indiv dat
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
sea_pops <- read.csv("data/seascape_data/population_data.csv")[,-1]
exp_pops <- read.csv("data/experiment/eelgrass_exp_sites.csv")

## gen dat
eelgrass_lfmm_MLL <- read.lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.lfmm")
eelgrass_geno_MLL <- read.geno("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.geno")
eelgrass_lfmm_full <- read.lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.lfmm")
eelgrass_geno_full <- read.geno("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.geno")
eelgrass_lfmm_full_rm <- read.lfmm("data/seascape_data/eel_full_rm.lfmm")

## env dat
curr_full_env <- read.env("data/EnvDat/full_seascapeEnv_2025-09-12.env")
curr_mll_env <- read.env("data/EnvDat/mll_seascapeEnv_2025-09-12.env")
exp_env <- read.env("data/EnvDat/expEnv_2025-09-12.env")
trt_env <- read.env("data/EnvDat/trtEnv_2025-09-12.env")
trt_scaled <- read.csv("data/EnvDat/trt_scaled_envdat_2025-09-12.csv")
exp_scaled <- read.csv("data/EnvDat/exp_scaled_envdat_2025-09-12.csv")
sea_scaled <- read.csv("data/EnvDat/sea_scaled_envdat_2025-09-12.csv")
```

## Data Prep

### Individual Data Prep

I will start by processing the individual data to work with both MLL and full datasets. 

```{r}
# inspect data
sea_pops
indiv_list
exp_pops

# rename sea pop columns
colnames(sea_pops) <- c("region","site_name","pop","long","lat")

# rename exp pop columns
colnames(exp_pops) <- c("site_name", "pop", "long", "lat")

# add metadata to inds
indiv_list <- as.data.frame(indiv_list)
colnames(indiv_list) <- c("mll", "ind")
indiv_list$pop <- str_split_fixed(indiv_list$ind, "-", 2)[,1]
inds <- merge(sea_pops, indiv_list, by = c("pop"))
```

### Environmental Data Prep

Environmental Data has been pre-processed using the script "EnvDatProcess.Rmd". This script scales all environmental data together, then splits it into the relevant data sets for seascape, experimental pops, and treatments before outputting the correct formats for each GO method. Here, I just ensure that column names are consistent and that the order of the .env files matches with our other metadata

```{r}
# inspect data
# seascape
curr_full_env
curr_mll_env
sea_scaled

# exp sites
exp_env
exp_scaled

# trt
trt_env
trt_scaled

# colnames for env files
colnames(curr_full_env) <- colnames(curr_mll_env) <- colnames(exp_env) <- colnames(trt_env) <- c("temp_scaled", "max_temp_scaled", "sal_scaled", "min_sal_scaled")

# now check the order of data
# exp env
exp_env == exp_scaled[,9:12] # temperature is correct but salinity isn't, need to check this in env dat process script

# trt env
trt_env == trt_scaled[,6:9] # salinity is different here as well

# seascape
curr_mll_env == sea_scaled[,12:15] # these aren't in order at all
unique(curr_full_env) == sea_scaled[,12:15] # not even close and the dimensions are wrong too
```

### Genomic Data Prep

Seascape genomic data comes from Steffi Ries and Marlene Jahnke at the University of Gothenburg, who had already processed the genomic dataset to do XX (check what part of bioinformatic pipeline had been done already).

```{r}

```


## Seascape Model Training

Now that I have prepped and cleaned my data, I will use it to train LFMM models later used to predict genomic offset to some other (in this case experimental) climate. Importantly, I do not need to trim or impute my genomic data - this has been done in the script "GenDatProcess.Rmd". I also do not determine ancestry, Steffi has done this and recovered K = 12. 

