---
title: "eelgrass_GO_mapping"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# wd
setwd("~/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse) # for data cleaning
library(stringr) # for data cleaning
library(dplyr) # for data cleaning
library(ggplot2) # for plotting
library(raster) # for environmental pixels
library(viridis) # plotting
library(sf) # mapping
library(s2) # mapping
library(rnaturalearth) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(stars) # mapping

# data
# metadata
latlons <- read.csv("data/experiment/eelgrass_exp_sites.csv")
poptrts_df <- read.csv("data/experiment/poptrts_df.csv")

# experimental
lea_go <- read.csv("results/GO_results/LEA_offsets_exptrt_ind_long_2026-08-13.csv")[,-1]
rda_go <- read.csv("results/GO_results/RDA_offsets_exptrt_ind_long_2026-08-13.csv")[,-1]
gf_go <- read.csv("results/GO_results/GF_offsets_exptrt_ind_long_2026-08-13.csv")[,-1]
env_dist <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]

# experimental averaged env data
lea_go_avg <- read.csv("results/GO_results/LEA_offsets_exptrt_long_2025-12-03.csv")[,-1]
rda_go_avg <- read.csv("results/GO_results/RDA_offsets_exptrt_long_2025-12-03.csv")[,-1]
gf_go_avg <- read.csv("results/GO_results/GF_offsets_exptrt_long_2025-12-03.csv")[,-1]
env_dist_avg <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]

# seascape
lea_go_sea <- read.csv("results/GO_results/lea_offset_seascape_go.csv")[,-1]
rda_go_sea <- read.csv("results/GO_results/rda_offset_seascape_go.csv")[,-1]
gf_go_sea <- read.csv("results/GO_results/gf_offset_seascape_go.csv")[,-1]
```

# Eelgrass GO Mapping

This script is to map the GO of experimental populations to experimental treatments. 

## Data Prep

Start by combining all the data to prep for mapping.

```{r}
# prep latlon data
# vat and vik are switched in latlon metadata
latlons[latlons$site == "Inre Vattenholmen", c("site", "site_abbrev")] <- c("Vik", "VIK")
latlons[2, c("site", "site_abbrev")] <- c("Inre Vattenholmen", "VAT")
latlons$pop <- latlons$site_abbrev

# prep the offset dat (averaged from individual env dat)
lea_prep <- lea_go %>% group_by(pop, trt, cc) %>% summarize(offset = mean(offset)) %>% mutate(method = "lea_offset") %>% as.data.frame()
rda_prep <- rda_go %>% group_by(pop, trt, cc) %>% summarize(offset = mean(offset)) %>% mutate(method = "rda_offset") %>% as.data.frame()
gf_prep <- gf_go %>% group_by(pop, trt, cc) %>% summarize(offset = mean(offset)) %>% mutate(method = "gf_offset") %>% as.data.frame()

# and env dist
env_prep <- env_dist %>% filter(pops2 %in% c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu") & !(pops1 %in% c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu"))) %>% mutate(pop = pops1, trt = pops2, offset = distances, cc = "mll", method = "env") %>% dplyr::select(pop, trt, cc, offset, method)

# prep the offset dat (calculated from population averaged env data)
lea_prep_avg <- lea_go_avg %>% pivot_longer(cols = c(gap_mll, gap_full), names_to = "cc", values_to = "offset") %>% mutate(cc = str_split_fixed(cc, "_", 2)[,2], method = "lea") 
rda_prep_avg <- rda_go_avg %>% pivot_longer(cols = c(offset_mll, offset_full), names_to = "cc", values_to = "offset") %>% mutate(cc = str_split_fixed(cc, "_", 2)[,2], method = "rda")
gf_prep_avg <- gf_go_avg %>% pivot_longer(cols = c(offset_gf_geno_mll, offset_gf_geno_full), names_to = "cc", values_to = "offset") %>% mutate(cc = str_split_fixed(cc, "_", 4)[,4], method = "gf") %>% dplyr::select(-offset_gf_af_mll, -offset_gf_af_full)

env_prep_avg <- env_dist_avg %>% filter(pops2 %in% c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu") & !(pops1 %in% c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu"))) %>% mutate(pop = pops1, trt = pops2, offset = distances, cc = "mll", method = "env") %>% dplyr::select(pop, trt, cc, offset, method)

# combine
go <- rbind(lea_prep, rda_prep, gf_prep, env_prep)
go_avg <- rbind(lea_prep_avg, rda_prep_avg, gf_prep_avg, env_prep_avg)

# add the latlon
go_prep <- merge(latlons, go, by = "pop")
go_prep_avg <- merge(latlons, go_avg, by = "pop")

# coast info
pops <- poptrts_df %>% dplyr::select(poptrt, poptrt2, order, coast) %>% filter(!(coast == "trt")) %>% rename(pop = poptrt, pop_coast = coast, pop_order = order, poptrt_names = poptrt2)
go_prep1 <- merge(pops, go_prep, by = "pop")
go_prep1_avg <- merge(pops, go_prep_avg, by = "pop")

# trts
trtdf <- data.frame(trt = c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu"), trt_coast = c("West", "East", "West", "East"), temp = c("current", "current", "future", "future"), trt_order = c(1:4))

# add trt dat
go_plotting <- merge(trtdf, go_prep1, by = "trt")
go_plotting_avg <- merge(trtdf, go_prep1_avg, by = "trt")
```

Also interpolate the seascape data to prep for plotting

```{r}
## prep GOs
###########
# arrange gos by pop order and rm single outlier pop from seascape
lea_go_sea <- lea_go_sea %>% arrange(order) %>% filter(go_future_mll < 0.0055) %>% as.data.frame()
rda_go_sea <- rda_go_sea %>% arrange(order) %>% as.data.frame()
gf_go_sea <- gf_go_sea %>% arrange(order) %>% as.data.frame()

# add experimental gos into interpolation dataset
# first pull out experimental site values from dfs
lea_go_sea_rm <- lea_go_sea %>% filter(!(pop %in% lea_prep$pop))
rda_go_sea_rm <- rda_go_sea %>% filter(!(pop %in% rda_prep$pop))
gf_go_sea_rm <- gf_go_sea %>% filter(!(pop %in% gf_prep$pop))

# put together
go_sea_rm <- rbind(lea_go_sea_rm, rda_go_sea_rm, gf_go_sea_rm) %>% dplyr::select(pop, lat, long, order, coast, method, go_current_full, go_future_full, go_current_mll, go_future_mll)
col_order <- colnames(go_sea_rm)

# pull out relevant dat - pivot_wider gives lots of NAs so doing stepwise method
# lea
lea_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & method == "lea_offset")
lea_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & method == "lea_offset")
lea_future_coast_go <- rbind(lea_go_west_fut, lea_go_east_fut) %>% pivot_wider(names_from = "cc", values_from = "offset") %>% dplyr::select(-c(trt, trt_coast, trt_order)) %>% rename(go_future_full = full, go_future_mll = mll, coast = pop_coast) %>% as.data.frame()
lea_go_west_curr <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "current" & method == "lea_offset")
lea_go_east_curr <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "current" & method == "lea_offset")
lea_current_coast_go <- rbind(lea_go_west_curr, lea_go_east_curr) %>% pivot_wider(names_from = "cc", values_from = "offset") %>% dplyr::select(-c(trt, trt_coast, trt_order)) %>% rename(go_current_full = full, go_current_mll = mll, coast = pop_coast) %>% as.data.frame()
lea_coast_go <- merge(lea_future_coast_go, lea_current_coast_go, by = c("temp", "pop", "pop_order", "coast", "site", "lat", "long", "method"))

# rda
rda_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & method == "rda_offset")
rda_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & method == "rda_offset")
rda_future_coast_go <- rbind(rda_go_west_fut, rda_go_east_fut) %>% pivot_wider(names_from = "cc", values_from = "offset") %>% dplyr::select(-c(trt, trt_coast, trt_order)) %>% rename(go_future_full = full, go_future_mll = mll, coast = pop_coast) %>% as.data.frame()
rda_go_west_curr <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "current" & method == "rda_offset")
rda_go_east_curr <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "current" & method == "rda_offset")
rda_current_coast_go <- rbind(rda_go_west_curr, rda_go_east_curr) %>% pivot_wider(names_from = "cc", values_from = "offset") %>% dplyr::select(-c(trt, trt_coast, trt_order)) %>% rename(go_current_full = full, go_current_mll = mll, coast = pop_coast) %>% as.data.frame()
rda_coast_go <- merge(rda_future_coast_go, rda_current_coast_go, by = c("temp", "pop", "pop_order", "coast", "site", "lat", "long", "method"))

# gf
gf_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & method == "gf_offset")
gf_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & method == "gf_offset")
gf_future_coast_go <- rbind(gf_go_west_fut, gf_go_east_fut) %>% pivot_wider(names_from = "cc", values_from = "offset") %>% dplyr::select(-c(trt, trt_coast, trt_order)) %>% rename(go_future_full = full, go_future_mll = mll, coast = pop_coast) %>% as.data.frame()
gf_go_west_curr <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "current" & method == "gf_offset")
gf_go_east_curr <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "current" & method == "gf_offset")
gf_current_coast_go <- rbind(gf_go_west_curr, gf_go_east_curr) %>% pivot_wider(names_from = "cc", values_from = "offset") %>% dplyr::select(-c(trt, trt_coast, trt_order)) %>% rename(go_current_full = full, go_current_mll = mll, coast = pop_coast) %>% as.data.frame()
gf_coast_go <- merge(gf_future_coast_go, gf_current_coast_go, by = c("temp", "pop", "pop_order", "coast", "site", "lat", "long", "method"))

# put together & order columns
go_interpolation_coasts <- rbind(lea_coast_go, rda_coast_go, gf_coast_go) %>% dplyr::select(-temp, - site) %>% rename(order = pop_order) %>% relocate(col_order)

go_interpolation <- rbind(go_sea_rm, go_interpolation_coasts)
###########

## prep geom 
############
# set geom
sf::sf_use_s2(FALSE)

# load and crop layers
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- st_crop(world, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))

coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- st_crop(coast, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))

# turn multiline into single line
coast_parts <- st_cast(coast_crop, "MULTILINESTRING") |>
  st_cast("LINESTRING")

coast_parts$id <- seq_len(nrow(coast_parts))

# label and id pieces
plot(st_geometry(coast_parts))
cent <- st_coordinates(st_centroid(st_geometry(coast_parts)))
text(cent, labels = coast_parts$id, col = "red", cex = 1)

# keep only relevant ids
keep_ids <- c(1, 10, 13, 14, 15, 22)

coast_sweden <- coast_parts[coast_parts$id %in% keep_ids,]

plot(st_geometry(coast_sweden), col = "lightblue")

# eyeball the unwanted region approx lon/lat extent from above plot
rm_bbox <- st_as_sfc(st_bbox(c(xmin = 13.2, xmax = 14.1, ymin = 55.8, ymax = 57), crs = st_crs(coast_sweden)))

plot(st_geometry(coast_sweden), col = "lightblue")
plot(rm_bbox, add = TRUE, border = "red")

# split into main vs. everything else
main_coast <- st_difference(coast_sweden, rm_bbox)
plot(st_geometry(main_coast), col = "darkblue")

# buffer dist
dist <- 0.6  # roughly miles

buffered_region <- st_buffer(st_union(main_coast), dist)
plot(buffered_region, col = "lightblue2")

# crop ocean layer
oceans50 <- ne_download(scale = 50, type = "ocean", category = "physical")
oceans_crop <- st_crop(oceans50, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))

# intersect
int_region <- st_intersection(oceans_crop, buffered_region)

plot(st_geometry(int_region), col = "lightblue2")
plot(st_geometry(main_coast), col = "darkblue", add = TRUE)

# define grid across ocean area
grid <- st_make_grid(buffered_region, cellsize = 0.05, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
grid_int_crop <- sf::st_crop(grid_int, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))

# another version using just experimental sites for interpolation
# keep only relevant region
keep_ids1 <- 22
coast_sweden1 <- coast_parts[coast_parts$id %in% keep_ids1,]
buffered_region1 <- st_buffer(st_union(coast_sweden1), dist)

# intersect
int_region1 <- st_intersection(oceans_crop, buffered_region1)

grid1 <- st_make_grid(buffered_region1, cellsize = 0.05, what = "polygons")
grid_int1 <- st_intersection(grid1, buffered_region1)
grid_int_crop1 <- sf::st_crop(grid_int1, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))
############

## interpolate
##############
# go data as sf
sf_lea <- st_as_sf(go_interpolation[go_interpolation$method == "lea_offset",], 
                   coords = c("long","lat"),
                   crs = 4326) # for WGS84 projection
sf_rda <- st_as_sf(go_interpolation[go_interpolation$method == "rda_offset",], 
                   coords = c("long","lat"),
                   crs = 4326) # for WGS84 projection
sf_gf <- st_as_sf(go_interpolation[go_interpolation$method == "gf_offset",], 
                   coords = c("long","lat"),
                   crs = 4326) # for WGS84 projection

# doing mll only since only plotting mll
interpolated_lea_curr <- idw(go_current_mll~1, sf_lea, grid_int)
interpolated_lea_fut <- idw(go_future_mll~1, sf_lea, grid_int)
interpolated_rda_curr <- idw(go_current_mll~1, sf_rda, grid_int)
interpolated_rda_fut <- idw(go_future_mll~1, sf_rda, grid_int)
interpolated_gf_curr <- idw(go_current_mll~1, sf_gf, grid_int)
interpolated_gf_fut <- idw(go_future_mll~1, sf_gf, grid_int)
##############

## rasterize
############
# empty raster
ext <- ext(st_bbox(interpolated_lea_curr)) # get spatial extent
rast_empty <- rast(ext, resolution = 0.05, crs = "+proj=longlat +datum=WGS84")

# create a raster stack
# set up
sf_ls <- list(interpolated_lea_curr, interpolated_lea_fut, interpolated_rda_curr, interpolated_rda_fut, interpolated_gf_curr, interpolated_gf_fut)
rast_ls <- list()

# loop through sf files to rasterize
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}

# combine
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lea_curr", "interpolated_lea_fut", "interpolated_rda_curr", "interpolated_rda_fut", "interpolated_gf_curr", "interpolated_gf_fut")

# save
saveRDS(rast_all, "results/GO_results/seascape_interpolated_go.RDS")
############
```

And interpolate using just the experimental predictions

```{r}
## prep dat
###########
# check data
head(go_plotting_avg)

# pull relevant data
go_interpolation_w_avg <- go_plotting_avg %>% filter(pop_coast == "West" & trt_coast == "West") %>% dplyr::select(-trt, -trt_coast, -trt_order) %>% rename(coast = pop_coast) %>% filter(cc == "mll") %>% 
  pivot_wider(names_from = temp, values_from = offset) %>% as.data.frame()
go_interpolation_e_avg <- go_plotting_avg %>% filter(pop_coast == "East" & trt_coast == "East") %>% dplyr::select(-trt, -trt_coast, -trt_order) %>% rename(coast = pop_coast) %>% filter(cc == "mll") %>% 
  pivot_wider(names_from = temp, values_from = offset) %>% as.data.frame()
go_plotting_exp <- rbind(go_interpolation_w_avg, go_interpolation_e_avg)

# split by method and coast
go_lea_exp_w <- go_interpolation_w_avg %>% filter(method == "lea" & coast == "West")
go_lea_exp_e <- go_interpolation_e_avg %>% filter(method == "lea" & coast == "East")
go_rda_exp_w <- go_interpolation_w_avg %>% filter(method == "rda" & coast == "West")
go_rda_exp_e <- go_interpolation_e_avg %>% filter(method == "rda" & coast == "East")
go_gf_exp_w <- go_interpolation_w_avg %>% filter(method == "gf" & coast == "West")
go_gf_exp_e <- go_interpolation_e_avg %>% filter(method == "gf" & coast == "East")
go_env_exp_w <- go_interpolation_w_avg %>% filter(method == "env" & coast == "West")
go_env_exp_e <- go_interpolation_e_avg %>% filter(method == "env" & coast == "East")
###########

## prep geom
############
sf::sf_use_s2(FALSE)

# load and crop coastline
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 8, xmax = 20, ymin = 54, ymax = 61))

coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- st_crop(coast, c(xmin = 8.5, xmax = 20, ymin = 54, ymax = 61))

# turn multiline into single lines, id + inspect pieces
coast_parts <- st_cast(coast_crop, "MULTILINESTRING") |> st_cast("LINESTRING")
coast_parts$id <- seq_len(nrow(coast_parts))

plot(st_geometry(coast_parts))
cent <- st_coordinates(st_centroid(st_geometry(coast_parts)))
text(cent, labels = coast_parts$id, col = "red", cex = 1)

# keep only the relevant piece -- this single line covers BOTH coasts
keep_ids <- 24
coast_sweden <- coast_parts[coast_parts$id %in% keep_ids, ]
plot(st_geometry(coast_sweden), col = "lightblue")

# trim area beyond desired interpolation region to west of sweden
# eyeball the unwanted region approx lon/lat extent
rm_bbox <- st_as_sfc(st_bbox(c(xmin = 8.5, xmax = 10.5, ymin = 58, ymax = 60), crs = st_crs(coast_sweden)))
plot(st_geometry(coast_sweden), col = "lightblue")
plot(rm_bbox, add = TRUE, border = "red")

# identify main coast region
main_coast <- st_difference(coast_sweden, rm_bbox)
plot(st_geometry(main_coast), col = "darkblue")

# split coast
coast_seg <- st_segmentize(main_coast, units::set_units(200, "m")) # vertex every ~200 m

coast_pts <- coast_seg %>% st_cast("MULTIPOINT") %>% st_cast("POINT") %>% st_sf(geometry = .)

seg_len <- st_distance(coast_pts[-1, ], coast_pts[-nrow(coast_pts), ], by_element = TRUE)
coast_pts$dist_along <- c(0, cumsum(as.numeric(seg_len)))

kur_sf <- st_as_sf(surv_interpolation_coasts[surv_interpolation_coasts$pop == "KUR", ], coords = c("long", "lat"), crs = 4326)
kur_idx  <- st_nearest_feature(kur_sf, coast_pts)
kur_dist <- coast_pts$dist_along[kur_idx]

# sanity check
coast_pts$side <- ifelse(coast_pts$dist_along <= kur_dist, "West", "East")
plot(st_geometry(coast_pts), col = ifelse(coast_pts$dist_along <= kur_dist, "blue", "red"))

coast_w <- coast_pts[coast_pts$side == "West", ] %>% st_combine() %>% st_cast("LINESTRING")
coast_e <- coast_pts[coast_pts$side == "East", ] %>% st_combine() %>% st_cast("LINESTRING")

# buffers + grid
dist <- 0.6  # roughly 40 miles

buffer_w <- st_buffer(st_union(coast_w), dist)
buffer_e <- st_buffer(st_union(coast_e), dist)
buffered_region <- st_union(buffer_w, buffer_e)

grid <- st_make_grid(buffered_region, cellsize = 0.05, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
############

## interpolate
##############
lea_w_avg <- st_as_sf(go_lea_exp_w, coords = c("long", "lat"), crs = 4326)
lea_e_avg <- st_as_sf(go_lea_exp_e, coords = c("long", "lat"), crs = 4326)

rda_w_avg <- st_as_sf(go_rda_exp_w, coords = c("long", "lat"), crs = 4326)
rda_e_avg <- st_as_sf(go_rda_exp_e, coords = c("long", "lat"), crs = 4326)

gf_w_avg <- st_as_sf(go_gf_exp_w, coords = c("long", "lat"), crs = 4326)
gf_e_avg <- st_as_sf(go_gf_exp_e, coords = c("long", "lat"), crs = 4326)

env_w_avg <- st_as_sf(go_env_exp_w, coords = c("long", "lat"), crs = 4326)
env_e_avg <- st_as_sf(go_env_exp_e, coords = c("long", "lat"), crs = 4326)

# lea
int_lea_curr_w <- idw(current~1, lea_w_avg, grid_int)
int_lea_curr_e <- idw(current~1,  lea_e_avg, grid_int)
int_lea_fut_w <- idw(future~1, lea_w_avg, grid_int)
int_lea_fut_e <- idw(future~1,  lea_e_avg, grid_int)

# rda
int_rda_curr_w <- idw(current~1, rda_w_avg, grid_int)
int_rda_curr_e <- idw(current~1,  rda_e_avg, grid_int)
int_rda_fut_w <- idw(future~1, rda_w_avg, grid_int)
int_rda_fut_e <- idw(future~1,  rda_e_avg, grid_int)

# gf
int_gf_curr_w <- idw(current~1, gf_w_avg, grid_int)
int_gf_curr_e <- idw(current~1,  gf_e_avg, grid_int)
int_gf_fut_w <- idw(future~1, gf_w_avg, grid_int)
int_gf_fut_e <- idw(future~1,  gf_e_avg, grid_int)

# env dist
int_env_curr_w <- idw(current~1, env_w_avg, grid_int)
int_env_curr_e <- idw(current~1,  env_e_avg, grid_int)
int_env_fut_w <- idw(future~1, env_w_avg, grid_int)
int_env_fut_e <- idw(future~1,  env_e_avg, grid_int)
##############

## combine west + east blend southern tip
#########################################
grid_cent <- st_centroid(grid_int)
nearest_grid <- st_nearest_feature(grid_cent, coast_pts)
grid_side <- coast_pts$side[nearest_grid]

grid_lat <- st_coordinates(grid_cent)[, "Y"]
kur_lat <- st_coordinates(kur_sf)[, "Y"]

bandwidth_north_m <- 40000
bandwidth_north_deg <- bandwidth_north_m / 111320

d_lat <- grid_lat - kur_lat

w_west <- ifelse(grid_side == "West", pmin(1, pmax(0, d_lat / bandwidth_north_deg)), 0)
w_east <- 1 - w_west

# lea
interpolated_curr_lea_avg <- int_lea_curr_w
interpolated_curr_lea_avg$var1.pred <- w_west * int_lea_curr_w$var1.pred + w_east * int_lea_curr_e$var1.pred

interpolated_fut_lea_avg <- int_lea_fut_w
interpolated_fut_lea_avg$var1.pred <- w_west * int_lea_fut_w$var1.pred + w_east * int_lea_fut_e$var1.pred

# rda
interpolated_curr_rda_avg <- int_rda_curr_w
interpolated_curr_rda_avg$var1.pred <- w_west * int_rda_curr_w$var1.pred + w_east * int_rda_curr_e$var1.pred

interpolated_fut_rda_avg <- int_rda_fut_w
interpolated_fut_rda_avg$var1.pred <- w_west * int_rda_fut_w$var1.pred + w_east * int_rda_fut_e$var1.pred

# gf
interpolated_curr_gf_avg <- int_gf_curr_w
interpolated_curr_gf_avg$var1.pred <- w_west * int_gf_curr_w$var1.pred + w_east * int_gf_curr_e$var1.pred

interpolated_fut_gf_avg <- int_gf_fut_w
interpolated_fut_gf_avg$var1.pred <- w_west * int_gf_fut_w$var1.pred + w_east * int_gf_fut_e$var1.pred

# env
interpolated_curr_env_avg <- int_env_curr_w
interpolated_curr_env_avg$var1.pred <- w_west * int_env_curr_w$var1.pred + w_east * int_env_curr_e$var1.pred

interpolated_fut_env_avg <- int_env_fut_w
interpolated_fut_env_avg$var1.pred <- w_west * int_env_fut_w$var1.pred + w_east * int_env_fut_e$var1.pred
#########################################

## rasterize
############
ext <- ext(st_bbox(interpolated_curr_lea_avg))
rast_empty <- rast(ext, resolution = 0.05, crs = "+proj=longlat +datum=WGS84")

sf_ls <- list(interpolated_curr_lea_avg, interpolated_fut_lea_avg, interpolated_curr_rda_avg, interpolated_fut_rda_avg, interpolated_curr_gf_avg, interpolated_fut_gf_avg, interpolated_curr_env_avg, interpolated_fut_env_avg)
rast_ls <- list()
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lea_curr", "interpolated_lea_fut", "interpolated_rda_curr", "interpolated_rda_fut", "interpolated_gf_curr", "interpolated_gf_fut", "interpolated_env_curr", "interpolated_env_fut")

# save
saveRDS(rast_all, "results/GO_results/go_exp_interpolated_avg.RDS")
############

## plotting setup
#################
# color ramp
pal <- colorRamp(c("green", "red"))

# masking region for interpolation
interpolated_lea_curr_masked_avg <- interpolated_curr_lea_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_lea_fut_masked_avg <- interpolated_fut_lea_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_rda_curr_masked_avg <- interpolated_curr_rda_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_rda_fut_masked_avg <- interpolated_fut_rda_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_gf_curr_masked_avg <- interpolated_curr_gf_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_gf_fut_masked_avg <- interpolated_fut_gf_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_env_curr_masked_avg <- interpolated_curr_env_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_env_fut_masked_avg <- interpolated_fut_env_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
#################
```


## Mapping

Now make the plots. I'm going to start off making "climate change" maps, which visualize the genomic offset between current/future W and current/future E on the same plot.

### LEA

```{r}
# color ramp
pal <- colorRamp(c("green", "red"))

# masking region for interpolation
interpolated_lea_fut_masked <- interpolated_lea_fut %>%
  filter(y >= 54 & y <= 61, x >= 10 & x <= 20)
interpolated_lea_curr_masked <- interpolated_lea_curr %>%
  filter(y >= 54 & y <= 61, x >= 10 & x <= 20)

# map
## lea curr
###########
lea_curr_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_lea_curr_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = lea_go_sea, aes(x = long, y = lat, fill = go_current_mll), shape = 21, size = 2) +
  geom_point(data = lea_current_coast_go, aes(x = long, y = lat, 
                                              fill = go_current_mll), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.0055)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.0055)) +
  geom_label_repel(data = lea_current_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", LEA[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
lea_curr_coast
###########

## lea fut
##########
lea_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_lea_fut_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = lea_go_sea, aes(x = long, y = lat, fill = go_future_mll), shape = 21, size = 2) +
  geom_point(data = lea_future_coast_go, aes(x = long, y = lat, 
                                              fill = go_future_mll), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.0055)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.0055)) +
  geom_label_repel(data = lea_future_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Predicted ", LEA[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
lea_fut_coast
##########
```

### RDA

```{r}
# masking region for interpolation
interpolated_rda_fut_masked <- interpolated_rda_fut %>%
  filter(y >= 54 & y <= 61, x >= 10 & x <= 20)
interpolated_rda_curr_masked <- interpolated_rda_curr %>%
  filter(y >= 54 & y <= 61, x >= 10 & x <= 20)

# map
## rda curr
###########
rda_curr_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_rda_curr_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = rda_go_sea, aes(x = long, y = lat, fill = go_current_mll), shape = 21, size = 2) +
  geom_point(data = rda_current_coast_go, aes(x = long, y = lat, 
                                              fill = go_current_mll), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 3.5)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 3.5)) +
  geom_label_repel(data = rda_current_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", RDA[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
rda_curr_coast
###########

## rda fut
##########
rda_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_rda_fut_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = rda_go_sea, aes(x = long, y = lat, fill = go_future_mll), shape = 21, size = 2) +
  geom_point(data = rda_future_coast_go, aes(x = long, y = lat, 
                                              fill = go_future_mll), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 3.5)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 3.5)) +
  geom_label_repel(data = rda_future_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Predicted ", RDA[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
rda_fut_coast
##########
```

### GF

```{r}
# masking region for interpolation
interpolated_gf_fut_masked <- interpolated_gf_fut %>%
  filter(y >= 54 & y <= 61, x >= 10 & x <= 20)
interpolated_gf_curr_masked <- interpolated_gf_curr %>%
  filter(y >= 54 & y <= 61, x >= 10 & x <= 20)

# map
## gf curr
###########
gf_curr_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_gf_curr_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = gf_go_sea, aes(x = long, y = lat, fill = go_current_mll), shape = 21, size = 2) +
  geom_point(data = gf_current_coast_go, aes(x = long, y = lat, 
                                              fill = go_current_mll), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.12)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.12)) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", GF[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
gf_curr_coast
###########

## gf fut
##########
gf_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_gf_fut_masked), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = gf_go_sea, aes(x = long, y = lat, fill = go_future_mll), shape = 21, size = 2) +
  geom_point(data = gf_future_coast_go, aes(x = long, y = lat, 
                                              fill = go_future_mll), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.12)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.12)) +
  geom_label_repel(data = gf_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Predicted ", GF[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
gf_fut_coast
##########
```

### Env Dist

```{r}
# relevant data
env_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & cc == "mll" & method == "env")
env_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & cc == "mll" & method == "env")
env_future_coast_go <- rbind(env_go_west_fut, env_go_east_fut)

# map
env_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = env_future_coast_go, aes(x = long, y = lat, 
                                              fill = offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "rocket", direction = -1, limits = c(0,8)) +
  scale_fill_viridis_c(option = "rocket", direction = -1, limits = c(0,8)) +
  geom_label_repel(data = env_future_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(D) Environmental Distance to Future Climate", x = "Longitude", y = "Latitude", col = "Env. Dist.", fill = "Env. Dist.")
env_fut_coast
```

## Mapping with Averages

### LEA

```{r}
# reset world crop
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 8, xmax = 20, ymin = 54, ymax = 61))

# subset data
go_plotting_exp_lea <- go_plotting_exp %>% filter(method == "lea")

# map
## lea curr
###########
lea_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_lea_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_lea, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.06)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.06)) +
  geom_label_repel(data = lea_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", LEA[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
lea_curr_coast_avg
###########

## lea fut
##########
lea_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_lea_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_lea, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.06)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.06)) +
  geom_label_repel(data = lea_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Predicted ", LEA[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
lea_fut_coast_avg
##########
```

### RDA

```{r}
# subset data
go_plotting_exp_rda <- go_plotting_exp %>% filter(method == "rda")

# map
## rda curr
###########
rda_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_rda_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_rda, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 2)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 2)) +
  geom_label_repel(data = rda_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Predicted ", RDA[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
rda_curr_coast_avg
###########

## rda fut
##########
rda_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_rda_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_rda, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 2)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 2)) +
  geom_label_repel(data = rda_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Predicted ", RDA[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
rda_fut_coast_avg
##########
```

### GF

```{r}
# subset data
go_plotting_exp_gf <- go_plotting_exp %>% filter(method == "gf")

# map
## gf curr
###########
gf_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_gf_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_gf, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.1)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.1)) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(E) Predicted ", GF[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
gf_curr_coast_avg
###########

## gf fut
##########
gf_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_gf_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_gf, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 0.1)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 0.1)) +
  geom_label_repel(data = gf_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(F) Predicted ", GF[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
gf_fut_coast_avg
##########
```

### env dist

```{r}
# subset data
go_plotting_exp_env <- go_plotting_exp %>% filter(method == "env")

# map
## env curr
###########
env_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_env_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_env, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, 7)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 7)) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(G) Environmental Distance", subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
env_curr_coast_avg
###########

## env fut
##########
env_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_env_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "white") +
  geom_point(data = go_plotting_exp_env, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, 7)) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, 7)) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(H) Environmental Distance", subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
env_fut_coast_avg
##########
```

### Export Plots

```{r}
# export
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = lea_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = lea_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = rda_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = rda_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = gf_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = gf_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = env_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = env_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))
print(doc, target = "results/figures/climchg_genomic_offset_maps.pptx")
```
