---
title: "eelgrass_GO_mapping"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# wd
setwd("~/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse) # for data cleaning
library(stringr) # for data cleaning
library(dplyr) # for data cleaning
library(ggplot2) # for plotting
library(raster) # for environmental pixels
library(viridis) # plotting
library(sf) # mapping
library(s2) # mapping
library(rnaturalearth) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(stars) # mapping

# data
# metadata
latlons <- read.csv("data/experiment/eelgrass_exp_sites.csv")
poptrts_df <- read.csv("data/experiment/poptrts_df.csv")

# experimental
lea_go <- read.csv("results/GO_results/LEA_offsets_exptrt_long_2025-12-03.csv")[,-1]
rda_go <- read.csv("results/GO_results/RDA_offsets_exptrt_long_2025-12-03.csv")[,-1]
gf_go <- read.csv("results/GO_results/GF_offsets_exptrt_long_2025-12-03.csv")[,-1]
```

# Eelgrass GO Mapping

This script is to map the GO of experimental populations to experimental treatments. 

## Data Prep

Start by combining all the data to prep for mapping.

```{r}
# prep the offset dat
lea_prep <- lea_go %>% pivot_longer(cols = c(gap_mll, gap_full), names_to = "cc", values_to = "offset") %>% mutate(cc = str_split_fixed(cc, "_", 2)[,2], method = "lea") 
rda_prep <- rda_go %>% pivot_longer(cols = c(offset_mll, offset_full), names_to = "cc", values_to = "offset") %>% mutate(cc = str_split_fixed(cc, "_", 2)[,2], method = "rda")
gf_prep <- gf_go %>% pivot_longer(cols = c(offset_gf_geno_mll, offset_gf_geno_full), names_to = "cc", values_to = "offset") %>% mutate(cc = str_split_fixed(cc, "_", 4)[,4], method = "gf") %>% dplyr::select(-offset_gf_af_mll, -offset_gf_af_full)

# combine
go <- rbind(lea_prep, rda_prep, gf_prep)

# add the latlon
latlons <- latlons %>% rename(pop = site_abbrev)
go_prep <- merge(latlons, go, by = "pop")

# coast info
pops <- popsdf %>% dplyr::select(poptrt, order, coast) %>% filter(!(coast == "trt")) %>% rename(pop = poptrt, pop_coast = coast, pop_order = order)
go_prep1 <- merge(pops, go_prep, by = "pop")

# trts
trtdf <- data.frame(trt = c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu"), trt_coast = c("West", "East", "West", "East"), temp = c("current", "current", "future", "future"), trt_order = c(1:4))

# add trt dat
go_plotting <- merge(trtdf, go_prep1, by = "trt")
```

## Mapping

Now make the plots. I'm going to start off making "climate change" maps, which visualize the offset between current/future W and current/future E on the same plot.

First prep the plotting range.

```{r}
# set spherical geometry
sf::sf_use_s2(FALSE)

# set object
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 10, xmax = 20, ymin = 54, ymax = 61))
```

### LEA

```{r}
# relevant data
lea_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & cc == "full" & method == "lea")
lea_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & cc == "full" & method == "lea")
lea_future_coast_go <- rbind(lea_go_west_fut, lea_go_east_fut)

# map
lea_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = lea_future_coast_go, aes(x = long, y = lat, 
                                              fill = offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = -1, limits = c(0,0.05)) +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.05)) +
  geom_label_repel(data = lea_future_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", LEA[offset], " to Future Climate")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
lea_fut_coast
```

### RDA

```{r}
# relevant data
rda_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & cc == "full" & method == "rda")
rda_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & cc == "full" & method == "rda")
rda_future_coast_go <- rbind(rda_go_west_fut, rda_go_east_fut)

# map
rda_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = rda_future_coast_go, aes(x = long, y = lat, 
                                              fill = offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = -1, limits = c(0,1.5)) +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,1.5)) +
  geom_label_repel(data = rda_future_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Predicted ", RDA[offset], " to Future Climate")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
rda_fut_coast
```

### GF

```{r}
# relevant data
gf_go_west_fut <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West" & temp == "future" & cc == "full" & method == "gf")
gf_go_east_fut <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East" & temp == "future" & cc == "full" & method == "gf")
gf_future_coast_go <- rbind(gf_go_west_fut, gf_go_east_fut)

# map
gf_fut_coast <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = world_crop, fill = 'antiquewhite1') +
  geom_point(data = gf_future_coast_go, aes(x = long, y = lat, 
                                              fill = offset), 
             shape = 21, size = 4) +
  scale_colour_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  geom_label_repel(data = gf_future_coast_go,
                  aes(x=long, y=lat, label = pop),
               #   nudge_x = -3,
               #   nudge_y = 1,
                  fill = "white")+
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Predicted ", GF[offset], " to Future Climate")), x = "Longitude", y = "Latitude", col = "Genomic \nOffset", fill = "Genomic \nOffset")
gf_fut_coast
```
