---
title: "Eelgrass_RDA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Eelgrass/BalticEelgrass/")
```

# Offset Modeling with RDA

This script is for offset calculation using RDAdapt, specifically for the project "VR_Zostera_genomic_offset." It includes training of the models using seascape genomic data, as well as calculation of genomic offset for experimental populations in each mesocosm treatment. 

## Setup & Formatting

```{r Setup, echo = F}
# packages
library(rdadapt) # for offsets
library(LEA) # read.geno
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(vegan) # for env dat processing
library(terra) # for env dat processing
library(reshape2) # for env dat processing
library(raster) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(ggplot2) # for data vis
library(viridis) # for data vis
library(foreach) # for rdadapt functions
library(qvalue) # for rdadapt functions
library(robust) # for rdadapt functions
library(forcats) # for data vis

# data
## indiv dat
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
sea_pops <- read.csv("data/seascape_data/population_data.csv")[,-1]
exp_pops <- read.csv("data/experiment/eelgrass_exp_sites.csv")

## gen dat
geno_mll <- read.geno("data/seascape_data/gen_MLL_imp_mat.geno")
geno_full <- read.geno("data/seascape_data/gen_full_rm_imp_mat.geno")

## env dat
trt_scaled <- read.csv("data/EnvDat/trt_scaled_envdat_2025-12-03.csv")
exp_scaled <- read.csv("data/EnvDat/exp_scaled_envdat_2025-12-03.csv")
sea_scaled_mll <- read.csv("data/EnvDat/sea_scaled_envdat_mll_2025-12-03.csv")
sea_scaled_full <- read.csv("data/EnvDat/sea_scaled_envdat_full_2025-12-03.csv")
###############
```


## RDA Functions

The rdadapt package does not install consistently across platforms, so I am pulling the code for relevant functions directly from the [package github page](https://github.com/landscape-genomics/rdadapt): 

```{r}
## rdadapt isn't installing correctly, this is the function i need directly from the git repo

##' 
##' @export
##' 

## Class not exported by vegan, so initialized here
setClass("rda", slots = c("colsum", "tot.chi", "Ybar"
                          , "method", "call"
                          , "pCCA", "CCA", "CA", "inertia"
                          , "regularization", "terms", "terminfo"))

##########################################################################################
setGeneric("rdadapt", def = function(RDA, K, scores) { standardGeneric( "rdadapt") })

##'
##' @rdname rdadapt
##' @export
##'

setMethod('rdadapt', signature(RDA = "rda"), function(RDA, K)
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("v" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$v seems not to exist")
  }
  if (K > ncol(RDA$CCA$v)) {
    K = ncol(RDA$CCA$v)
    warning("\n Not enough RDA axis available, K is set to ncol(RDA$CCA$v)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  scores <- RDA$CCA$v[, 1:as.numeric(K)]
  return(rdadapt(scores = scores))
})

##'
##' @rdname rdadapt
##' @export
##'

setMethod('rdadapt', signature(scores = "data.frame", RDA = "missing", K = "missing"), function(scores)
{
  scores <- as.matrix(scores)
  return(rdadapt(scores = scores))
})

##'
##' @rdname rdadapt
##' @export
##'

setMethod('rdadapt', signature(scores = "matrix", RDA = "missing", K = "missing"), function(scores)
{
  K <- ncol(scores)
  resscale <- apply(scores, 2, scale)
  resmaha <- covRob(resscale, distance = TRUE, na.action = na.omit, estim = "pairwiseGK")$dist
  lambda <- median(resmaha) / qchisq(0.5, df = K)
  reschi2test <- pchisq(resmaha / lambda, K, lower.tail = FALSE)
  qval <- qvalue(reschi2test)
  q.values_rdadapt <- qval$qvalues
  return(data.frame(p.values = reschi2test, q.values = q.values_rdadapt))
})

library(qvalue)
library(robust)
##########################################################################################

###################################################################################################
##' @name adaptive_index
##' @author Thibaut Capblancq
##' 
##' @title Project adaptive variation across space onto current or future landscapes
##' 
##' @description The \code{adaptive_index} function allows the user to estimate the optimal 
##' adaptive genetic component for any combination of environmental conditions, and project it 
##' onto the landscape.
##' 
##' @param RDA a \code{RDA} model from which to extract loci and environmental variable scores
##' @param K an \code{integer} specifying the number of RDA axes to use for the projection
##' @param env a \code{RasterStack} object or a \code{data.frame} with the environmental 
##' conditions of the sites to predict
##' @param env_mask (\emph{optional, default} \code{NULL}) \cr a \code{Raster} object to 
##' limit the projection to a specific area
##' @param method  (\emph{default} \code{'loadings'}) \cr a \code{character} defining whether 
##' the function is to use weighted averages (scaling type 1, \code{loadings}) or linear 
##' combinations (scaling type 2, \code{predict}) of the projected environmental variables 
##' to predict site scores (i.e., adaptive index)
##' 
##' @return  
##' 
##' A \code{RasterStack} object or a \code{data.frame} containing the projections for the 
##' \code{K} first RDA axes
##' 
##' @details
##' 
##' The \emph{adaptive index} provides an estimate of adaptive genetic similarity or difference 
##' of all pixels on the landscape as a function of the values of the environmental predictors 
##' at that location. 
##' When projected on a map, it allows visualizing the different adaptive gradients across a 
##' species range.
##' 
##' 
##' @keywords RDA, adaptive index, genetic similarity, genetic difference, projection
## 
## @seealso 
## 
## @examples
##' 
##' 
##' @importFrom foreach foreach %do%
##' @importFrom terra crs rast as.data.frame mask
##' @include rda_class.R
##' 
##' @export
##' 
##'
###################################################################################################

setGeneric("adaptive_index", def = function(RDA, K, env, env_mask = NULL, method = "loadings") { standardGeneric("adaptive_index") })

##'
##' @rdname adaptive_index
##' @export
##'

setMethod('adaptive_index', signature(RDA = "rda", env = "SpatRaster"), function(RDA, K, env, env_mask = NULL, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("biplot" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$biplot seems not to exist")
  }
  if (K > ncol(RDA$CCA$biplot)) {
    K <- ncol(RDA$CCA$biplot)
    warning("\n Not enough RDA axis available, K is set to ncol(RDA$CCA$biplot)")
  }
  if (!inherits(env, c("SpatRaster", "RasterLayer", "RasterStack"))) {
    stop("\n env must be a 'SpatRaster', 'RasterLayer' or 'RasterStack' object")
  } else if (inherits(env, c("RasterLayer", "RasterStack"))) {
    env <- rast(env)
  }
  if (nlyr(env) != nrow(RDA$CCA$biplot) || any(! names(env) %in% rownames(RDA$CCA$biplot))) {
    stop("\n env must contain same variables as the ones used in RDA (see rownames(RDA$CCA$biplot))")
  }
  if (!is.null(env_mask)) {
    if (!inherits(env_mask, c("SpatRaster", "RasterLayer"))) {
      stop("\n env_mask must be a 'SpatRaster' or 'RasterLayer' object")
    } else if (inherits(env_mask, "RasterLayer")) {
      env_mask <- rast(env_mask)
    }
    if (nlyr(env_mask) > 1) {
      stop("\n env_mask must contain only one layer")
    }
  }
  if (!(method %in% c("loadings", "predict"))) {
    stop("\n method must be 'loadings' or 'predict'")
  }
  
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Get RDA information -----------------------------------------------------
  RDA_biplot <- RDA$CCA$biplot
  var_names <- rownames(RDA_biplot)
  
  ## Transform environmental raster for prediction
  if (!is.null(env_mask)) { ## Mask with env_mask
    env <- try(mask(env, env_mask))
    if (inherits(env, "try-error")) {
      stop("\n env and env_mask must match in terms of extent, origin, crs")
    }
  }
  env_crs <- crs(env)
  env_df <- as.data.frame(env[[var_names]], xy = TRUE)
  if (any(is.na(env_df))) {
    env_df <- na.exclude(env_df)
    warning("\n env contained NA values which have been removed")
  }
  env_xy <- env_df[, c("x", "y")]
  env_var <- env_df[, var_names]
  
  
  ## Make predictions ---------------------------------------------------------
  if (method == "predict") {
    pred <- predict(RDA, env_var, type = "lc")
  }
  Proj <- foreach(i = 1:K) %do%
    {
      if (method == "loadings") { ## Predict pixels genetic component based on RDA axes
        tmp_df <- data.frame(env_xy, z = as.vector(apply(env_var, 1, function(x) sum(x * RDA_biplot[, i]))))
      } else if (method == "predict") { ## Predict with RDA model and linear combinations
        tmp_df <- data.frame(env_xy, z = as.vector(pred[, i]))
      }
      ras <- rast(tmp_df, type = "xyz", crs = crs(env))
      names(ras) <- paste0("RDA", i)
      return(ras)
    }
  Proj <- rast(Proj)
  
  ## Return predictions for each RDA axis
  return(Proj)
})


##'
##' @rdname adaptive_index
##' @export
##'

setMethod('adaptive_index', signature(RDA = "rda", env = "data.frame", env_mask = "missing"), function(RDA, K, env, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("biplot" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$biplot seems not to exist")
  }
  if (K > ncol(RDA$CCA$biplot)) {
    K <- ncol(RDA$CCA$biplot)
    warning("\n Not enough RDA axis available, K is set to ncol(RDA$CCA$biplot)")
  }
  if (ncol(env) != nrow(RDA$CCA$biplot) || any(! colnames(env) %in% rownames(RDA$CCA$biplot))) {
    stop("\n env must contain same variables as the ones used in RDA (see rownames(RDA$CCA$biplot))")
  }
  if (!(method %in% c("loadings", "predict"))) {
    stop("\n method must be 'loadings' or 'predict'")
  }
  
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Get RDA information -----------------------------------------------------
  RDA_biplot <- RDA$CCA$biplot
  var_names <- rownames(RDA_biplot)
  env_var <- env[, var_names]
  
  ## Make predictions ---------------------------------------------------------
  if (method == "predict") {
    pred <- predict(RDA, env_var, type = "lc")
  }
  AI <- foreach(i = 1:K) %do%
    {
      if (method == "loadings") { ## Predict genetic component based on RDA axes
        tmp_vec <- as.vector(apply(env_var, 1, function(x) sum(x * RDA_biplot[, i])))
      } else if (method == "predict") { ## Predict with RDA model and linear combinations
        tmp_vec <- as.vector(pred[, i])
      }
      return(tmp_vec)
    }
  names(AI) <- paste0("RDA", 1:K)
  AI <- as.data.frame(do.call(cbind, AI))
  
  ## Return predictions for each RDA axis
  return(AI)
})

###################################################################################################
##' @name genomic_offset
##' @author Thibaut Capblancq
##' 
##' @title Predict temporal or geographic genomic offset
##' 
##' @description The \code{genomic_offset} function allows the user to predict genomic offset 
##' from a RDA model. The \code{genomic_offset} function can estimate both temporal or spatial 
##' genomic offset and accommodates raster data or discrete populations.
##' 
##' @param RDA a \code{RDA} model from which to extract loci and environmental variable scores
##' @param K an \code{integer} specifying the number of RDA axes to use for the projection
##' @param env_pres a \code{RasterStack} object or a \code{data.frame} with the environmental 
##' conditions in the present
##' @param env_fut a \code{RasterStack} object or a \code{data.frame} with the environmental 
##' conditions in the future
##' @param env_mask (\emph{optional, default} \code{NULL}) \cr a \code{Raster} object to limit 
##' the projection to a specific area
##' @param method  (\emph{default} \code{'loadings'}) \cr a \code{character} defining whether 
##' the function is to use weighted averages (scaling type 1, \code{loadings}) or linear 
##' combinations (scaling type 2, \code{predict}) of the projected environmental variables 
##' to predict site scores (i.e., adaptive index)
##' 
##' 
##' @return  
##' 
##' A \code{list} containing :
##' \itemize{
##'   \item \code{genomic_offset} : a \code{RasterStack} or a \code{data.frame} containing 
##'   the genomic offset predictions for the \code{K} first RDA axes, as well as the overall 
##'   genomic offset prediction
##'   \item \code{weights} : the weights associated with each RDA axis used for the predictions
##' }
##' 
##' 
##' @details
##' 
##' This RDA-based method to predict \emph{genomic offset} is relatively simple. 
##' RDA is first used to predict the \emph{optimal adaptive genetic composition} for each 
##' environmental pixel under consideration (see \code{\link{adaptive_index}} function), 
##' using both current and future environmental conditions. 
##' The \emph{euclidean distance} between these two predictions in the RDA space provides an 
##' estimate of the change in genetic composition that would be required to track climate change.
##' 
##' 
##' @keywords RDA, genomic offset, adaptive index, projection
## 
##' @seealso \code{\link{adaptive_index}}
## 
## @examples
##' 
##' 
##' @importFrom stats dist
##' @importFrom raster rasterToPoints rasterFromXYZ mask
##' @include rda_class.R
##'  
##' @export
##' 
##'
###################################################################################################

setGeneric("genomic_offset", def = function(RDA, K, env_pres, env_fut, env_mask = NULL, env_gar, method = "loadings") { standardGeneric("genomic_offset") })

##'
##' @rdname genomic_offset
##' @export
##'

setMethod('genomic_offset', signature(RDA = "rda", env_pres = "SpatRaster", env_gar = "missing"), function(RDA, K, env_pres, env_fut, env_mask = NULL, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("eig" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$eig seems not to exist")
  }
  if (K > length(RDA$CCA$eig)) {
    K <- length(RDA$CCA$eig)
    warning("\n Not enough RDA axis available, K is set to length(RDA$CCA$eig)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Make predictions
  AI_pres <- adaptive_index(RDA = RDA
                            , K = K
                            , env = env_pres
                            , env_mask = env_mask
                            , method = method)
  AI_fut <- adaptive_index(RDA = RDA
                           , K = K
                           , env = env_fut
                           , env_mask = env_mask
                           , method = method)
  
  ## Single axis genetic offset -----------------------------------------------
  offset <- foreach(i = 1:K) %do%
    {
      ras <- abs(AI_pres[[i]] - AI_fut[[i]])
      names(ras) <- paste0("RDA", i)
      return(ras)
    }
  offset <- rast(offset)
  
  ## Weight current and future adaptive indices based on eigen values of associated axes
  weights <- RDA$CCA$eig / sum(RDA$CCA$eig)
  weights <- weights[1:K]
  AI_pres_w <- AI_pres * weights
  AI_fut_w <- AI_fut * weights
  
  ## Predict a global genetic offset
  offset_global <- offset[[1]]
  offset_global[!is.na(offset_global)] <- sapply(1:ncell(AI_pres_w), function(x) {
    tmp_mat <- rbind(AI_pres_w[x], AI_fut_w[x])
    return(dist(tmp_mat, method = "euclidean"))
  })
  names(offset_global) <- "Global_offset"
  
  
  return(list(Proj_pres = AI_pres,
              Proj_fut = AI_fut,
              Proj_offset = offset,
              Proj_offset_global = offset_global,
              weights = weights))
})


##'
##' @rdname genomic_offset
##' @export
##'

setMethod('genomic_offset', signature(RDA = "rda", env_pres = "data.frame", env_mask = "missing", env_gar = "missing"), function(RDA, K, env_pres, env_fut, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("eig" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$eig seems not to exist")
  }
  if (K > length(RDA$CCA$eig)) {
    K <- length(RDA$CCA$eig)
    warning("\n Not enough RDA axis available, K is set to length(RDA$CCA$eig)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Make predictions
  AI_pres <- adaptive_index(RDA = RDA
                            , K = K
                            , env = env_pres
                            , method = method)
  AI_fut <- adaptive_index(RDA = RDA
                           , K = K
                           , env = env_fut
                           , method = method)
  
  ## Single axis genetic offset -----------------------------------------------
  offset <- foreach(i = 1:K) %do%
    {
      vec <- abs(AI_pres[,i] - AI_fut[,i])
      return(vec)
    }
  offset <- as.data.frame(do.call(cbind, offset))
  
  ## Weight offset based on eigen values of associated axes
  weights <- RDA$CCA$eig / sum(RDA$CCA$eig)
  weights <- weights[1:K]
  offset <- data.frame(do.call(cbind, lapply(1:ncol(offset), function(x) offset[,x]*weights[x])))
  colnames(offset) <- paste0("RDA", 1:K)
  
  ## Predict a global genetic offset
  offset$Global <- apply(offset, 1, function(x) sqrt(sum(x*x)))
  
  return(list(Proj_pres = AI_pres,
              Proj_fut = AI_fut,
              genomic_offset = offset,
              weights = weights))
})


##'
##' @rdname genomic_offset
##' @export
##'

setMethod('genomic_offset', signature(RDA = "rda", env_pres = "data.frame", env_mask = "missing", env_fut = "missing"), function(RDA, K, env_pres, env_gar, method = "loadings")
{
  ## CHECKS -------------------------------------------------------------------
  if (!inherits(RDA, "rda")) { stop("\n RDA must be a 'rda' object") }
  if (!("CCA" %in% names(RDA)) || !("eig" %in% names(RDA$CCA))) {
    stop("\n RDA$CCA$eig seems not to exist")
  }
  if (K > length(RDA$CCA$eig)) {
    K <- length(RDA$CCA$eig)
    warning("\n Not enough RDA axis available, K is set to length(RDA$CCA$eig)")
  }
  
  ## FUNCTION -----------------------------------------------------------------
  
  ## Make predictions
  AI_pres <- adaptive_index(RDA = RDA
                            , K = K
                            , env = env_pres
                            , method = method)
  
  AI_gar <- adaptive_index(RDA = RDA
                           , K = K
                           , env = env_gar
                           , method = method)
  
  ## Single axis genetic offset -----------------------------------------------
  offset <- foreach(i = 1:K) %do%
    {
      vec <- abs(AI_pres[,i] - AI_gar[,i])
      return(vec)
    }
  offset <- as.data.frame(do.call(cbind, offset))
  
  ## Weight offset based on eigen values of associated axes
  weights <- RDA$CCA$eig / sum(RDA$CCA$eig)
  weights <- weights[1:K]
  offset <- data.frame(do.call(cbind, lapply(1:ncol(offset), function(x) offset[,x]*weights[x])))
  colnames(offset) <- paste0("RDA", 1:K)
  
  ## Predict a global genetic offset
  offset$Global <- apply(offset, 1, function(x) sqrt(sum(x*x)))
  
  return(list(genomic_offset = offset,
              weights = weights))
})
```


## Data Checks

The RDA functions are ready, but I need to double check all of my data before I can train models and make GO predictions. 

### Individual Data

```{r}
# inspect data
indiv_list
sea_pops
exp_pops

# indiv list
colnames(indiv_list) <- c("mll", "indiv")
indiv_list$pop <- str_split_fixed(indiv_list$indiv, "-", 2)[,1]

# standardize colnames for pop dfs
colnames(sea_pops) <- c("region", "site_full", "pop", "long", "lat")
colnames(exp_pops) <- c("site_full", "pop", "long", "lat")

# merge inds with seascape pop metadata
inds_sea <- merge(sea_pops, indiv_list, by = c("pop"))

# reorder this by mll number
inds_sea_ord <- inds_sea[order(inds_sea$mll),]

# mll inds
inds_sea_ord_mll <- inds_sea_ord[!(duplicated(inds_sea_ord$mll)),]
```


### Environmental Data

Environmental data has been pre-processed in the script "EnvDatProcess.Rmd". Here I just double check that everything looks right and is in the correct order to match with genomic data.

```{r}
# inspect data
summary(sea_scaled_mll)
summary(sea_scaled_full)
summary(trt_scaled)
summary(exp_scaled)

# check seascape environmental data in same order as individuals
sea_scaled_mll$pop == inds_sea_ord_mll$pop # all good!
sea_scaled_full$pop == inds_sea_ord$pop # all good!

# reduced datasets for ease of use
sea_scaled_mll_red <- sea_scaled_mll[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(sea_scaled_mll_red) <- sea_scaled_mll$indv_no
sea_scaled_full_red <- sea_scaled_full[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(sea_scaled_full_red) <- sea_scaled_full$indv_no
trt_scaled_red <- trt_scaled[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(trt_scaled_red) <- trt_scaled$trt
exp_scaled_red <- exp_scaled[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(exp_scaled_red) <- exp_scaled$pop
```


### Genomic Data

Seascape genomic data comes from Steffi Ries and Marlene Jahnke at the University of Gothenburg, who had already processed the genomic dataset (check what part of bioinformatic pipeline had been done already). I further process this data to impute missing genomic data and remove non-variable genomic sites in the script "GenDatProcess.Rmd"

```{r}
# inspect data
geno_mll[1:10,1:10]
geno_full[1:10,1:10]

# set populations as rownames
rownames(geno_mll) <- inds_sea_ord_mll$pop
rownames(geno_full) <- inds_sea_ord$pop
```


## Seascape Model Training & Outlier Detection

While I do not intend to use an outlier dataset to calculate adaptive index/offset, I will still go through the process of outlier detection with RDA. This process is also part of training the RDA model for offsets later.

```{r}
# train model
#############
# mll
start <- (Sys.time())
rda_mll <- rda(geno_mll ~ ., as.data.frame(sea_scaled_mll_red))
plot(rda_mll)
screeplot(rda_mll, main = "Eigenvalues of constrained axes")
end <- (Sys.time())
print(end-start)

# full
start <- (Sys.time())
rda_full <- rda(geno_full ~ ., as.data.frame(sea_scaled_full_red))
plot(rda_full)
screeplot(rda_full, main = "Eigenvalues of constrained axes")
end <- (Sys.time())
print(end-start)
#############

# outlier detection
###################
# run rdadapt with K = 12 - error, not enough RDA axes available, so going back to K = 2 for first 2 axes
rdadapt_sea_mll <- rdadapt(RDA = rda_mll, K = 2)
rdadapt_sea_full <- rdadapt(RDA = rda_full, K = 2)

# bonferroni correction
# P-values threshold after Bonferroni correction
thres_env_mll <- 0.01 / length(rdadapt_sea_mll$p.values)
thres_env_full <- 0.01 / length(rdadapt_sea_full$p.values)

# Identifying the loci that are below the p-value threshold
outliers_mll <- (1:ncol(geno_mll))[which(rdadapt_sea_mll$p.values < thres_env_mll)]
length(outliers_mll) # 406 outliers

outliers_full <- (1:ncol(geno_full))[which(rdadapt_sea_full$p.values < thres_env_full)]
length(outliers_full) # 707 outliers
###################

# model selection
#################
# model sel
fwdsel_mll <- ordiR2step(rda(geno_mll ~ 1, data = as.data.frame(sea_scaled_mll_red)), # empty model
               scope = formula(rda_mll), # full model
               direction = "forward",
               R2scope = T, 
               pstep = 1000,
               trace = T)
fwdsel_full <- ordiR2step(rda(geno_full ~ 1, data = as.data.frame(sea_scaled_full_red)), # empty model
               scope = formula(rda_full), # full model
               direction = "forward",
               R2scope = T, 
               pstep = 1000,
               trace = T)

# variables retained?
(fwdsel_mll$call) # gets rid of median temp
(fwdsel_full$call)  # keep all

# signif testing
(signif_mll <- anova.cca(rda_mll, parallel=getOption("mc.cores"))) # signif
(signif_full <- anova.cca(rda_full, parallel=getOption("mc.cores"))) # signif
#################
```


## Offset Prediction

Now I predict offset for all of my experimental populations to each treatment.

```{r}
# put together dataset for exp pop origin with trt (pred from and pred to)
exp_expanded <- exp_scaled_red[rep(seq_len(nrow(exp_scaled_red)), each = nrow(trt_scaled_red)),]
trt_expanded <- as.matrix(do.call(rbind, replicate(8, trt_scaled_red, simplify = FALSE)))
trt_expanded_trts <- rep(trt_scaled$trt, 8)
exp_expanded_pops <- (rep(rownames(exp_scaled_red), 4))[order(rep(rownames(exp_scaled_red), 4))]

# estimate genomic offset
offsets_mll <- rdadapt::genomic_offset(RDA = rda_mll, 
                              K = 2, # because we only have 4 axes, we can only have max K = 4 - instead I will use just the first two axes (broken stick criterion)
                              env_pres = as.data.frame(exp_expanded), 
                              env_fut = as.data.frame(trt_expanded), 
                              method = "loadings")
offsets_full <- rdadapt::genomic_offset(RDA = rda_full,
                                 K = 2, # because we only have 4 axes, we can only have max K = 4 - instead I will use just the first two axes (broken stick criterion)
                                 env_pres = as.data.frame(exp_expanded), 
                                 env_fut = as.data.frame(trt_expanded), 
                                 method = "loadings")

# add metadata
offsets_rda <- data.frame(pop = exp_expanded_pops, trt = trt_expanded_trts, offset_mll = offsets_mll$genomic_offset$Global, offset_full = offsets_full$genomic_offset$Global)

# wide versions
mll_offsets_w <- pivot_wider(offsets_rda,
                             id_cols = "trt",
                             names_from = "pop", 
                             values_from = "offset_mll") %>% 
  as.data.frame()
mll_offsets_w$cc <- "mll"

full_offsets_w <- pivot_wider(offsets_rda,
                             id_cols = "trt",
                             names_from = "pop", 
                             values_from = "offset_full") %>% 
  as.data.frame()
full_offsets_w$cc <- "full"

offsets_rda_wide <- rbind(mll_offsets_w, full_offsets_w)

# save offsets
write.csv(offsets_rda, paste0("results/GO_results/RDA_offsets_exptrt_long_",Sys.Date(),".csv"))
write.csv(offsets_rda_wide, paste0("results/GO_results/RDA_offsets_exptrt_wide",Sys.Date(),".csv"))
```


## Plotting

Now that I have my offsets, I want to plot them in a grid where populations are columns and treatments are rows 

```{r}
# prep for plotting
exp_pops$order <- 1:8
offsets_rda_plotting <- merge(exp_pops, offsets_rda, by = "pop")

# plot
ggplot(offsets_rda_plotting, aes(x = fct_reorder(pop, order), y = trt, fill = (offset_mll))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,4)) +
  theme_classic() +
  labs(title = "Genomic Offsets (RDA, MLL)", x = "Population", y = "Treatment", fill = "Offset")

# and the full offsets
ggplot(offsets_rda_plotting, aes(x = fct_reorder(pop, order), y = trt, fill = (offset_full))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,4)) +
  theme_classic() +
  labs(title = "Genomic Offsets (RDA, Full)", x = "Population", y = "Treatment", fill = "Offset")

cor(offsets_rda$offset_mll, offsets_rda$offset_full) # 0.939 correlation
```

