---
title: "eelgrass_GF_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Offset Modeling with Gradient Forest

This script is for offset calculation using gradientForest, specifically for the project "VR_Zostera_genomic_offset." It includes training of the models using seascape genomic data, as well as calculation of genomic offset for experimental populations in each mesocosm treatment. 

## Setup & Formatting

```{r Setup, echo = F}
# packages
library(BiocManager) # needed to download specific packages
library(gradientForest) # for offsets
library(LEA) # read.geno
library(readxl) # reading xlsx files
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(ggplot2) # for data vis
library(viridis) # for data vis

# data
## indiv dat
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
sea_pops <- read.csv("data/seascape_data/population_data.csv")[,-1]
exp_pops <- read.csv("data/experiment/eelgrass_exp_sites.csv")

## gen dat
geno_mll <- read.geno("data/seascape_data/gen_MLL_imp_mat.geno")
geno_full <- read.geno("data/seascape_data/gen_full_rm_imp_mat.geno")

## env dat
trt_scaled <- read.csv("data/EnvDat/trt_scaled_envdat_2025-12-03.csv")
exp_scaled <- read.csv("data/EnvDat/exp_scaled_envdat_2025-12-03.csv")
sea_scaled_mll <- read.csv("data/EnvDat/sea_scaled_envdat_mll_2025-12-03.csv")
sea_scaled_full <- read.csv("data/EnvDat/sea_scaled_envdat_full_2025-12-03.csv")
###############
```


## Data Checks

## Data Checks

This data has largely been processed in other scripts, but I still need to double check everything before I can train models and make GO predictions. 

### Individual Data

```{r}
# inspect data
indiv_list
sea_pops
exp_pops

# indiv list
colnames(indiv_list) <- c("mll", "indiv")
indiv_list$pop <- str_split_fixed(indiv_list$indiv, "-", 2)[,1]

# standardize colnames for pop dfs
colnames(sea_pops) <- c("region", "site_full", "pop", "long", "lat")
colnames(exp_pops) <- c("site_full", "pop", "long", "lat")

# merge inds with seascape pop metadata
inds_sea <- merge(sea_pops, indiv_list, by = c("pop"))

# reorder this by mll number
inds_sea_ord <- inds_sea[order(inds_sea$mll),]

# mll inds
inds_sea_ord_mll <- inds_sea_ord[!(duplicated(inds_sea_ord$mll)),]
```


### Environmental Data

Environmental data has been pre-processed in the script "EnvDatProcess.Rmd". Here I just double check that everything looks right and is in the correct order to match with genomic data.

```{r}
# inspect data
summary(sea_scaled_mll)
summary(sea_scaled_full)
summary(trt_scaled)
summary(exp_scaled)

# check seascape environmental data in same order as individuals
sea_scaled_mll$pop == inds_sea_ord_mll$pop # all good!
sea_scaled_full$pop == inds_sea_ord$pop # all good!

# af actually only uses one obs per pop
sea_scaled_mll_trim <- sea_scaled_mll[!duplicated(sea_scaled_mll$pop),]

# reduced datasets for ease of use
sea_scaled_trim_red <- sea_scaled_mll_trim[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(sea_scaled_trim_red) <- sea_scaled_mll_trim$pop
sea_scaled_mll_red <- sea_scaled_mll[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(sea_scaled_mll_red) <- sea_scaled_mll$indv_no
sea_scaled_full_red <- sea_scaled_full[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(sea_scaled_full_red) <- sea_scaled_full$indv_no
trt_scaled_red <- trt_scaled[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(trt_scaled_red) <- trt_scaled$trt
exp_scaled_red <- exp_scaled[,c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")]
rownames(exp_scaled_red) <- exp_scaled$pop
```


### Genomic Data

Seascape genomic data comes from Steffi Ries and Marlene Jahnke at the University of Gothenburg, who had already processed the genomic dataset (check what part of bioinformatic pipeline had been done already). I further process this data to impute missing genomic data and remove non-variable genomic sites in the script "GenDatProcess.Rmd"

```{r}
# inspect data
geno_mll[1:10,1:10]
geno_full[1:10,1:10]

# set populations as rownames
rownames(geno_mll) <- inds_sea_ord_mll$pop
rownames(geno_full) <- inds_sea_ord$pop

# allele frequencies
# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
geno_mllT <- as.matrix(t(geno_mll))
geno_fullT <- as.matrix(t(geno_full))

# any missing data?
sum((geno_mllT == 9)) # 0 - good 
sum((geno_fullT == 9)) # 0 - good

# add pop data to genomic data
colnames(geno_mllT) <- inds_sea_ord_mll$pop
colnames(geno_fullT) <- inds_sea_ord$pop

# function to calculate allele frequency @ single locus
calc_freq <- function(x){
  a <- sum(x, na.rm=TRUE)
  b <- (2*length(na.omit(x)))
  a/b
}

# function extending this to calculate allele frequencies across whole pop
calcfreq <- function(a, pop){
  tapply(a, pop, calc_freq)
}

# allele freqs
freqs_mll <- apply((geno_mll), 2, calcfreq, inds_sea_ord_mll$pop)
str(freqs_mll)

freqs_full <- apply((geno_full), 2, calcfreq, inds_sea_ord$pop)
str(freqs_full)

# update colnames for gf
colnames(freqs_mll) <- make.names(as.character(1:ncol(geno_mll)))
colnames(freqs_full) <- make.names(as.character(1:ncol(geno_full)))

# do the same for the full geno matrix, just in case
colnames(geno_mll) <- make.names(as.character(1:ncol(geno_mll)))
colnames(geno_full) <- make.names(as.character(1:ncol(geno_full)))
```


## Seascape Model Training

Now that I know my data is ready to use, I can train my models!

```{r}
# start by defining a maximum number of splits
maxLevel_trim <- log2(0.368*nrow(sea_scaled_mll_trim)/2)
maxLevel_mll <- log2(0.368*nrow(sea_scaled_mll)/2)
maxLevel_full <- log2(0.368*nrow(sea_scaled_full)/2)
 
# run gradient forest
# allele freq models
# mll
start_time <- Sys.time() # time start
gf_af_mll <- gradientForest(cbind(sea_scaled_trim_red, freqs_mll), predictor.vars = colnames(sea_scaled_trim_red), response.vars = (colnames(freqs_mll)), ntree = 500, maxLevel = maxLevel_trim, trace = T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 30 sec

# full
start_time <- Sys.time() # time start
gf_af_full <- gradientForest(cbind(sea_scaled_trim_red, freqs_full), predictor.vars = colnames(sea_scaled_trim_red), response.vars = (colnames(freqs_full)), ntree = 500, maxLevel = maxLevel_trim, trace = T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 30 sec

# geno models
start_time <- Sys.time() # time start
gf_geno_mll <- gradientForest(cbind(sea_scaled_mll_red, geno_mll), predictor.vars = colnames(sea_scaled_mll_red), response.vars=colnames(geno_mll), ntree=500, maxLevel=maxLevel_mll, trace=T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 10 mins
start_time <- Sys.time() # time start
gf_geno_full <- gradientForest(cbind(sea_scaled_full_red, geno_full), predictor.vars = colnames(sea_scaled_full_red), response.vars=colnames(geno_full), ntree=500, maxLevel=maxLevel_full, trace=T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 10 mins

saveRDS(gf_af_mll, paste0("results/GO_results/gf_af_mll_",Sys.Date(),".RDS"))
saveRDS(gf_af_full, paste0("results/GO_results/gf_af_full_",Sys.Date(),".RDS"))

saveRDS(gf_geno_mll, paste0("results/GO_results/gf_geno_mll_",Sys.Date(),".RDS"))
saveRDS(gf_geno_full, paste0("results/GO_results/gf_geno_full_",Sys.Date(),".RDS"))

# read in models trained in trainGF.R / trainGF.sh
gf_af_mll <- readRDS("results/GO_results/gf_af_mll_2025-12-03.RDS")
gf_af_full <- readRDS("results/GO_results/gf_af_full_2025-12-03.RDS")
gf_geno_mll <- readRDS("results/GO_results/gf_geno_mll_2025-12-03.RDS")
gf_geno_full <- readRDS("results/GO_results/gf_geno_full_2025-12-03.RDS")

# plot bar graphs depicting importance of each spatial and climate variable
plot(gf_af_mll)
plot(gf_af_full)
plot(gf_geno_mll)
plot(gf_geno_full)

sink("results/GO_results/gf_importance.txt")
importance(gf_af_mll)
importance(gf_af_full)
importance(gf_geno_mll)
importance(gf_geno_full)
sink()
```


## Offset Prediction

Now that we have trained our gradient forest models, we need to use these to predict performance in the mesocosm treatments.

```{r}
# predictions
#############
# expanded env dat
exp_expanded <- exp_scaled_red[rep(seq_len(nrow(exp_scaled_red)), each = nrow(trt_scaled_red)),]
trt_expanded <- do.call(rbind, replicate(8, trt_scaled_red, simplify = FALSE))
trt_expanded_trts <- rep(trt_scaled$trt, 8)
exp_expanded_pops <- (rep(rownames(exp_scaled_red), 4))[order(rep(rownames(exp_scaled_red), 4))]

# predict to home site
pred_af_home_mll <- predict(gf_af_mll, exp_expanded)
pred_af_home_full <- predict(gf_af_full, exp_expanded)
pred_geno_home_mll <- predict(gf_geno_mll, exp_expanded)
pred_geno_home_full <- predict(gf_geno_full, exp_expanded)

# predict to treatments
pred_af_trt_mll <- predict(gf_af_mll, trt_expanded)
pred_af_trt_full <- predict(gf_af_full, trt_expanded)
pred_geno_trt_mll <- predict(gf_geno_mll, trt_expanded)
pred_geno_trt_full <- predict(gf_geno_full, trt_expanded)
#############

# calculate euclidean distance for offset
#########################################
# number preds
nb_pred <- nrow(trt_expanded)

# function
euclidean <- function(a, b) sqrt(sum((a - b)^2)) # Euclidean distance for genetic offset

# calculate offset
# af mll
df_af_mll <- data.frame() # df
for (i in 1:nb_pred){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_af_home_mll[i,], pred_af_trt_mll[i,]))
  df_af_mll <- rbind(df_af_mll, ec_dist)
}
af_offset_mll <- data.frame(pop = exp_expanded_pops, trt = trt_expanded_trts, offset_gf_af_mll = df_af_mll) # associate metadata with offset calculation
colnames(af_offset_mll) <- c("pop","trt","offset_gf_af_mll")

# af full
df_af_full <- data.frame() # df
for (i in 1:nb_pred){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_af_home_full[i,], pred_af_trt_full[i,]))
  df_af_full <- rbind(df_af_full, ec_dist)
}
af_offset_full <- data.frame(pop = exp_expanded_pops, trt = trt_expanded_trts, offset_gf_af_full = df_af_full) # associate metadata with offset calculation
colnames(af_offset_full) <- c("pop","trt","offset_gf_af_full")

# geno mll
df_geno_mll <- data.frame() # df
for (i in 1:nb_pred){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_mll[i,], pred_geno_trt_mll[i,]))
  df_geno_mll <- rbind(df_geno_mll, ec_dist)
}
geno_offset_mll <- data.frame(pop = exp_expanded_pops, trt = trt_expanded_trts, offset_gf_geno_mll = df_geno_mll) # associate metadata with offset calculation
colnames(geno_offset_mll) <- c("pop","trt","offset_gf_geno_mll")

# geno full
df_geno_full <- data.frame() # df
for (i in 1:nb_pred){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_full[i,], pred_geno_trt_full[i,]))
  df_geno_full <- rbind(df_geno_full, ec_dist)
}
geno_offset_full <- data.frame(pop = exp_expanded_pops, trt = trt_expanded_trts, offset_gf_geno_full = df_geno_full) # associate metadata with offset calculation
colnames(geno_offset_full) <- c("pop","trt","offset_gf_geno_full")
#########################################

# put it all together
#####################
# complete offset df
gfaf_offsets <- merge(af_offset_mll, af_offset_full, by = c("pop","trt"))
gfgeno_offsets <- merge(geno_offset_mll, geno_offset_full, by = c("pop","trt"))
gf_offsets <- merge(gfaf_offsets, gfgeno_offsets, by = c("pop","trt"))

# wide versions of each set preds
gfaf_offset_mll_wide <- pivot_wider(gf_offsets,
                                    id_cols = "trt",
                                    names_from = "pop", 
                                    values_from = "offset_gf_af_mll") %>% 
  as.data.frame()
gfaf_offset_mll_wide$cc <- "mll"
gfaf_offset_mll_wide$mod <- "af"

gfaf_offset_full_wide <- pivot_wider(gf_offsets,
                                    id_cols = "trt",
                                    names_from = "pop", 
                                    values_from = "offset_gf_af_full") %>% 
  as.data.frame()
gfaf_offset_full_wide$cc <- "full"
gfaf_offset_full_wide$mod <- "af"

gfgeno_offset_mll_wide <- pivot_wider(gf_offsets,
                                    id_cols = "trt",
                                    names_from = "pop", 
                                    values_from = "offset_gf_geno_mll") %>% 
  as.data.frame()
gfgeno_offset_mll_wide$cc <- "mll"
gfgeno_offset_mll_wide$mod <- "geno"

gfgeno_offset_full_wide <- pivot_wider(gf_offsets,
                                    id_cols = "trt",
                                    names_from = "pop", 
                                    values_from = "offset_gf_geno_full") %>% 
  as.data.frame()
gfgeno_offset_full_wide$cc <- "full"
gfgeno_offset_full_wide$mod <- "geno"

# put it together
gf_offsets_wide <- rbind(gfaf_offset_mll_wide, gfaf_offset_full_wide, gfgeno_offset_mll_wide, gfgeno_offset_full_wide)

# save outputs
write.csv(gf_offsets, paste0("results/GO_results/GF_offsets_exptrt_long_",Sys.Date(),".csv"))
write.csv(gf_offsets_wide, paste0("results/GO_results/GF_offsets_exptrt_wide",Sys.Date(),".csv"))
#####################
```

## Plotting

Now that I have my offsets, I want to plot them in a grid where populations are columns and treatments are rows 

```{r}
# plot
# af offsets
# mll
ggplot(gf_offsets, aes(x = pop, y = trt, fill = (offset_gf_af_mll))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  theme_classic() +
  labs(title = "Genomic Offsets (GFaf, MLL)", x = "Population", y = "Treatment", fill = "Offset")

# and the full offsets
ggplot(gf_offsets, aes(x = pop, y = trt, fill = (offset_gf_af_full))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  theme_classic() +
  labs(title = "Genomic Offsets (GFaf, Full)", x = "Population", y = "Treatment", fill = "Offset")

# geno offsets
# mll
ggplot(gf_offsets, aes(x = pop, y = trt, fill = (offset_gf_geno_mll))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  theme_classic() +
  labs(title = "Genomic Offsets (GFgeno, MLL)", x = "Population", y = "Treatment", fill = "Offset")

# and the full offsets
ggplot(gf_offsets, aes(x = pop, y = trt, fill = (offset_gf_geno_full))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  theme_classic() +
  labs(title = "Genomic Offsets (GFgeno, Full)", x = "Population", y = "Treatment", fill = "Offset")
```


