---
title: "eelgrass_GO_mapping"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# wd
setwd("~/Documents/GitHub/BalticEelgrass")

# libraries
library(tidyverse) # for data cleaning
library(stringr) # for data cleaning
library(dplyr) # for data cleaning
library(ggplot2) # for plotting
library(raster) # for environmental pixels
library(viridis) # plotting
library(sf) # mapping
library(s2) # mapping
library(rnaturalearth) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(stars) # mapping

# data
# metadata
latlons <- read.csv("data/experiment/eelgrass_exp_sites.csv")
poptrts_df <- read.csv("data/experiment/poptrts_df.csv")

# experimental
lea_go <- read.csv("results/GO_results/LEA_offsets_exptrt_ind_long_2026-09-07.csv")[,-1]
rda_go <- read.csv("results/GO_results/RDA_offsets_exptrt_ind_long_2026-09-07.csv")[,-1]
gf_go <- read.csv("results/GO_results/GF_offsets_exptrt_ind_long_2026-09-07.csv")[,-1]
env_dist <- read.csv("results/GO_results/euclidean_long_trt_exp.csv")[,-1]
```

# Eelgrass GO Mapping

This script is to map the GO of experimental populations to experimental treatments. 

## Data Prep

Start by combining all the data to prep for mapping.

```{r}
# prep latlon data
latlons$pop <- latlons$site_abbrev

# prep the offset dat (averaged from individual env dat)
lea_prep <- lea_go %>% group_by(pop, trt, cc) %>% summarize(offset = mean(offset)) %>% mutate(method = "lea_offset") %>% as.data.frame()
rda_prep <- rda_go %>% group_by(pop, trt, cc) %>% summarize(offset = mean(offset)) %>% mutate(method = "rda_offset") %>% as.data.frame()
gf_prep <- gf_go %>% group_by(pop, trt, cc) %>% summarize(offset = mean(offset)) %>% mutate(method = "gf_offset") %>% as.data.frame()

# and env dist
env_prep <- env_dist %>% filter(pops2 %in% c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu") & !(pops1 %in% c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu"))) %>% mutate(pop = pops1, trt = pops2, offset = distances, cc = "mll", method = "env") %>% dplyr::select(pop, trt, cc, offset, method)

# combine
go <- rbind(lea_prep, rda_prep, gf_prep, env_prep)

# add the latlon
go_prep <- merge(latlons, go, by = "pop")

# coast info
pops <- poptrts_df %>% dplyr::select(poptrt, poptrt2, order, coast) %>% filter(!(coast == "trt")) %>% rename(pop = poptrt, pop_coast = coast, pop_order = order, poptrt_names = poptrt2)
go_prep1 <- merge(pops, go_prep, by = "pop")

# trts
trtdf <- data.frame(trt = c("TempControl-21psu", "TempControl-7psu", "TempWarm-16psu", "TempWarm-5psu"), trt_coast = c("West", "East", "West", "East"), temp = c("current", "current", "future", "future"), trt_order = c(1:4))

# add trt dat
go_plotting <- merge(trtdf, go_prep1, by = "trt")
```

Interpolate using just the experimental predictions

```{r}
## prep dat
###########
# check data
head(go_plotting)

# pull relevant data
go_interpolation_w <- go_plotting %>% filter(pop_coast == "West" & trt_coast == "West") %>% dplyr::select(-trt, -trt_coast, -trt_order) %>% rename(coast = pop_coast) %>% filter(cc == "mll") %>% 
  pivot_wider(names_from = temp, values_from = offset) %>% as.data.frame()
go_interpolation_e <- go_plotting %>% filter(pop_coast == "East" & trt_coast == "East") %>% dplyr::select(-trt, -trt_coast, -trt_order) %>% rename(coast = pop_coast) %>% filter(cc == "mll") %>% 
  pivot_wider(names_from = temp, values_from = offset) %>% as.data.frame()
go_plotting_exp <- rbind(go_interpolation_w, go_interpolation_e)

# split by method and coast
go_lea_exp_w <- go_interpolation_w %>% filter(method == "lea_offset" & coast == "West")
go_lea_exp_e <- go_interpolation_e %>% filter(method == "lea_offset" & coast == "East")
go_rda_exp_w <- go_interpolation_w %>% filter(method == "rda_offset" & coast == "West")
go_rda_exp_e <- go_interpolation_e %>% filter(method == "rda_offset" & coast == "East")
go_gf_exp_w <- go_interpolation_w %>% filter(method == "gf_offset" & coast == "West")
go_gf_exp_e <- go_interpolation_e %>% filter(method == "gf_offset" & coast == "East")
go_env_exp_w <- go_interpolation_w %>% filter(method == "env" & coast == "West")
go_env_exp_e <- go_interpolation_e %>% filter(method == "env" & coast == "East")
###########

## prep geom
############
sf::sf_use_s2(FALSE)

# load and crop coastline
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 8, xmax = 20, ymin = 54, ymax = 61))

coast <- ne_coastline(scale = "medium", returnclass = "sf")
coast_crop <- st_crop(coast, c(xmin = 8.5, xmax = 20, ymin = 54, ymax = 61))

# turn multiline into single lines, id + inspect pieces
coast_parts <- st_cast(coast_crop, "MULTILINESTRING") |> st_cast("LINESTRING")
coast_parts$id <- seq_len(nrow(coast_parts))

#plot(st_geometry(coast_parts))
cent <- st_coordinates(st_centroid(st_geometry(coast_parts)))
#text(cent, labels = coast_parts$id, col = "red", cex = 1)

# keep only the relevant piece -- this single line covers BOTH coasts
keep_ids <- 24
coast_sweden <- coast_parts[coast_parts$id %in% keep_ids, ]
#plot(st_geometry(coast_sweden), col = "lightblue")

# trim area beyond desired interpolation region to west of sweden
# eyeball the unwanted region approx lon/lat extent
rm_bbox <- st_as_sfc(st_bbox(c(xmin = 8.5, xmax = 10.5, ymin = 58, ymax = 60), crs = st_crs(coast_sweden)))
#plot(st_geometry(coast_sweden), col = "lightblue")
#plot(rm_bbox, add = TRUE, border = "red")

# identify main coast region
main_coast <- st_difference(coast_sweden, rm_bbox)
#plot(st_geometry(main_coast), col = "darkblue")

# split coast
coast_seg <- st_segmentize(main_coast, units::set_units(200, "m")) # vertex every ~200 m

coast_pts <- coast_seg %>% st_cast("MULTIPOINT") %>% st_cast("POINT") %>% st_sf(geometry = .)

seg_len <- st_distance(coast_pts[-1, ], coast_pts[-nrow(coast_pts), ], by_element = TRUE)
coast_pts$dist_along <- c(0, cumsum(as.numeric(seg_len)))

kur_sf <- st_as_sf(surv_interpolation_coasts[surv_interpolation_coasts$pop == "KUR", ], coords = c("long", "lat"), crs = 4326)
kur_idx  <- st_nearest_feature(kur_sf, coast_pts)
kur_dist <- coast_pts$dist_along[kur_idx]

# sanity check
coast_pts$side <- ifelse(coast_pts$dist_along <= kur_dist, "West", "East")
#plot(st_geometry(coast_pts), col = ifelse(coast_pts$dist_along <= kur_dist, "blue", "red"))

coast_w <- coast_pts[coast_pts$side == "West", ] %>% st_combine() %>% st_cast("LINESTRING")
coast_e <- coast_pts[coast_pts$side == "East", ] %>% st_combine() %>% st_cast("LINESTRING")

# buffers + grid
dist <- 0.6  # roughly 40 miles

buffer_w <- st_buffer(st_union(coast_w), dist)
buffer_e <- st_buffer(st_union(coast_e), dist)
buffered_region <- st_union(buffer_w, buffer_e)

grid <- st_make_grid(buffered_region, cellsize = 0.05, what = "polygons")
grid_int <- st_intersection(grid, buffered_region)
############

## interpolate
##############
lea_w_avg <- st_as_sf(go_lea_exp_w, coords = c("long", "lat"), crs = 4326)
lea_e_avg <- st_as_sf(go_lea_exp_e, coords = c("long", "lat"), crs = 4326)

rda_w_avg <- st_as_sf(go_rda_exp_w, coords = c("long", "lat"), crs = 4326)
rda_e_avg <- st_as_sf(go_rda_exp_e, coords = c("long", "lat"), crs = 4326)

gf_w_avg <- st_as_sf(go_gf_exp_w, coords = c("long", "lat"), crs = 4326)
gf_e_avg <- st_as_sf(go_gf_exp_e, coords = c("long", "lat"), crs = 4326)

env_w_avg <- st_as_sf(go_env_exp_w, coords = c("long", "lat"), crs = 4326)
env_e_avg <- st_as_sf(go_env_exp_e, coords = c("long", "lat"), crs = 4326)

# lea
int_lea_curr_w <- idw(current~1, lea_w_avg, grid_int)
int_lea_curr_e <- idw(current~1,  lea_e_avg, grid_int)
int_lea_fut_w <- idw(future~1, lea_w_avg, grid_int)
int_lea_fut_e <- idw(future~1,  lea_e_avg, grid_int)

# rda
int_rda_curr_w <- idw(current~1, rda_w_avg, grid_int)
int_rda_curr_e <- idw(current~1,  rda_e_avg, grid_int)
int_rda_fut_w <- idw(future~1, rda_w_avg, grid_int)
int_rda_fut_e <- idw(future~1,  rda_e_avg, grid_int)

# gf
int_gf_curr_w <- idw(current~1, gf_w_avg, grid_int)
int_gf_curr_e <- idw(current~1,  gf_e_avg, grid_int)
int_gf_fut_w <- idw(future~1, gf_w_avg, grid_int)
int_gf_fut_e <- idw(future~1,  gf_e_avg, grid_int)

# env dist
int_env_curr_w <- idw(current~1, env_w_avg, grid_int)
int_env_curr_e <- idw(current~1,  env_e_avg, grid_int)
int_env_fut_w <- idw(future~1, env_w_avg, grid_int)
int_env_fut_e <- idw(future~1,  env_e_avg, grid_int)
##############

## combine west + east blend southern tip
#########################################
grid_cent <- st_centroid(grid_int)
nearest_grid <- st_nearest_feature(grid_cent, coast_pts)
grid_side <- coast_pts$side[nearest_grid]

grid_lat <- st_coordinates(grid_cent)[, "Y"]
kur_lat <- st_coordinates(kur_sf)[, "Y"]

bandwidth_north_m <- 40000
bandwidth_north_deg <- bandwidth_north_m / 111320

d_lat <- grid_lat - kur_lat

w_west <- ifelse(grid_side == "West", pmin(1, pmax(0, d_lat / bandwidth_north_deg)), 0)
w_east <- 1 - w_west

# lea
interpolated_curr_lea_avg <- int_lea_curr_w
interpolated_curr_lea_avg$var1.pred <- w_west * int_lea_curr_w$var1.pred + w_east * int_lea_curr_e$var1.pred

interpolated_fut_lea_avg <- int_lea_fut_w
interpolated_fut_lea_avg$var1.pred <- w_west * int_lea_fut_w$var1.pred + w_east * int_lea_fut_e$var1.pred

# rda
interpolated_curr_rda_avg <- int_rda_curr_w
interpolated_curr_rda_avg$var1.pred <- w_west * int_rda_curr_w$var1.pred + w_east * int_rda_curr_e$var1.pred

interpolated_fut_rda_avg <- int_rda_fut_w
interpolated_fut_rda_avg$var1.pred <- w_west * int_rda_fut_w$var1.pred + w_east * int_rda_fut_e$var1.pred

# gf
interpolated_curr_gf_avg <- int_gf_curr_w
interpolated_curr_gf_avg$var1.pred <- w_west * int_gf_curr_w$var1.pred + w_east * int_gf_curr_e$var1.pred

interpolated_fut_gf_avg <- int_gf_fut_w
interpolated_fut_gf_avg$var1.pred <- w_west * int_gf_fut_w$var1.pred + w_east * int_gf_fut_e$var1.pred

# env
interpolated_curr_env_avg <- int_env_curr_w
interpolated_curr_env_avg$var1.pred <- w_west * int_env_curr_w$var1.pred + w_east * int_env_curr_e$var1.pred

interpolated_fut_env_avg <- int_env_fut_w
interpolated_fut_env_avg$var1.pred <- w_west * int_env_fut_w$var1.pred + w_east * int_env_fut_e$var1.pred
#########################################

## rasterize
############
ext <- ext(st_bbox(interpolated_curr_lea_avg))
rast_empty <- rast(ext, resolution = 0.05, crs = "+proj=longlat +datum=WGS84")

sf_ls <- list(interpolated_curr_lea_avg, interpolated_fut_lea_avg, interpolated_curr_rda_avg, interpolated_fut_rda_avg, interpolated_curr_gf_avg, interpolated_fut_gf_avg, interpolated_curr_env_avg, interpolated_fut_env_avg)
rast_ls <- list()
for (i in seq_along(sf_ls)) {
  rast = st_rasterize(sf_ls[[i]] %>% dplyr::select(var1.pred, geometry))
  rast_ls[[i]] = rast
}
rast_all <- c(rast_ls)
names(rast_all) <- c("interpolated_lea_curr", "interpolated_lea_fut", "interpolated_rda_curr", "interpolated_rda_fut", "interpolated_gf_curr", "interpolated_gf_fut", "interpolated_env_curr", "interpolated_env_fut")

# save
#saveRDS(rast_all, "results/GO_results/go_exp_interpolated_avg.RDS")
############

## plotting setup
#################
# color ramp
pal <- colorRamp(c("green", "red"))

# masking region for interpolation
interpolated_lea_curr_masked_avg <- interpolated_curr_lea_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_lea_fut_masked_avg <- interpolated_fut_lea_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_rda_curr_masked_avg <- interpolated_curr_rda_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_rda_fut_masked_avg <- interpolated_fut_rda_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_gf_curr_masked_avg <- interpolated_curr_gf_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_gf_fut_masked_avg <- interpolated_fut_gf_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_env_curr_masked_avg <- interpolated_curr_env_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
interpolated_env_fut_masked_avg <- interpolated_fut_env_avg %>% filter(y >= 54 & y <= 61, x >= 8.5 & x <= 20)
#################
```


## Mapping

Now make the plots. I'm going to start off making "climate change" maps, which visualize the genomic offset between current/future W and current/future E on the same plot.

### LEA

```{r}
# reset world crop
world <- ne_countries(scale = "medium", returnclass = "sf")
world_crop <- sf::st_crop(world, c(xmin = 8, xmax = 20, ymin = 54, ymax = 61))

# subset data
go_plotting_exp_lea <- go_plotting_exp %>% filter(method == "lea_offset")

# map
## lea curr
###########
lea_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_lea_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_lea, aes(x = long, y = lat, 
                                             fill = current), 
             color = "black", shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_lea$future))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_lea$future))) +
  geom_label_repel(data = lea_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(A) Predicted ", LEA[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
lea_curr_coast_avg
###########

## lea fut
##########
lea_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_lea_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_lea, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_lea$future))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_lea$future))) +
  geom_label_repel(data = lea_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(B) Predicted ", LEA[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
lea_fut_coast_avg
##########
```

### RDA

```{r}
# subset data
go_plotting_exp_rda <- go_plotting_exp %>% filter(method == "rda_offset")

# map
## rda curr
###########
rda_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_rda_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_rda, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_rda$current))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_rda$current))) +
  geom_label_repel(data = rda_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(C) Predicted ", RDA[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
rda_curr_coast_avg
###########

## rda fut
##########
rda_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_rda_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_rda, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_rda$current))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_rda$current))) +
  geom_label_repel(data = rda_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(D) Predicted ", RDA[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
rda_fut_coast_avg
##########
```

### GF

```{r}
# subset data
go_plotting_exp_gf <- go_plotting_exp %>% filter(method == "gf_offset")

# map
## gf curr
###########
gf_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_gf_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_gf, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_gf$future))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_gf$future))) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(E) Predicted ", GF[offset])), subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
gf_curr_coast_avg
###########

## gf fut
##########
gf_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_gf_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_gf, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_gf$future))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_gf$future))) +
  geom_label_repel(data = gf_future_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = expression(paste("(F) Predicted ", GF[offset])), subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
gf_fut_coast_avg
##########
```

### env dist

```{r}
# subset data
go_plotting_exp_env <- go_plotting_exp %>% filter(method == "env")

# map
## env curr
###########
env_curr_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_env_curr_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_env, aes(x = long, y = lat, 
                                             fill = current), 
             shape = 21, size = 4) +
  scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_env$future))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_env$future))) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(G) Environmental Distance", subtitle = "Current Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
env_curr_coast_avg
###########

## env fut
##########
env_fut_coast_avg <- ggplot(data = world) +
  theme_bw()+
  geom_sf(data = st_as_sf(interpolated_env_fut_masked_avg), mapping = aes(fill = var1.pred, col = var1.pred)) +
  geom_sf(data = world_crop, fill = "gray90") +
  geom_point(data = go_plotting_exp_env, aes(x = long, y = lat, 
                                              fill = future), 
             shape = 21, size = 4) +
    scale_fill_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_env$future))) +
  scale_color_distiller(palette = "RdYlGn", limits = c(0, max(go_plotting_exp_env$future))) +
  geom_label_repel(data = gf_current_coast_go,
                  aes(x=long, y=lat, label = pop),
                  fill = "white")+
  geom_text(label = "North \nSea", 
             x = 9.3, y = 57.9,
             size = 3, fontface = "italic") +
  geom_text(label = "Baltic \nSea", 
             x = 18.7, y = 56.2,
             size = 3, fontface = "italic") +
  theme(plot.title = element_text(size = 18), panel.grid.major = element_line(color = "aliceblue"),
        panel.background = element_rect(fill = "aliceblue"), legend.position = 'right')+
  labs(title = "(H) Environmental Distance", subtitle = "Future Treatment", x = "Longitude", y = "Latitude", fill = "Genomic \nOffset", col = "Genomic \nOffset")
env_fut_coast_avg
##########
```

### Export Plots

```{r}
# export
doc <- read_pptx() %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = lea_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = lea_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = rda_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = rda_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = gf_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = gf_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>%
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = env_curr_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4)) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with(value = dml(ggobj = env_fut_coast_avg), location = ph_location(left = 1, top = 1, width = 6.5, height = 4))
print(doc, target = "results/figures/climchg_genomic_offset_maps.pptx")
```
