---
title: "eelgrass_GF_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Offset Modeling with Gradient Forest

This script is for offset calculation using gradientForest, specifically for the project "VR_Zostera_genomic_offset." It includes training of the models using seascape genomic data, as well as calculation of genomic offset for experimental populations in each mesocosm treatment. 

## Setup & Formatting

```{r Setup, echo = F}
# packages
library(BiocManager) # needed to download specific packages
library(gradientForest) # for offsets
library(LEA) # read.geno
library(readxl) # reading xlsx files
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(ggplot2) # for data vis
library(viridis) # for data vis
library(forcats) # for data vis

# data
## indiv dat
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
sea_pops <- read.csv("data/seascape_data/population_data.csv")[,-1]
exp_pops <- read.csv("data/experiment/eelgrass_exp_sites.csv")
inds_exp <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-12-31.csv")[,-1]

## gen dat
geno_mll <- read.geno("data/seascape_data/gen_MLL_imp_mat.geno")
geno_full <- read.geno("data/seascape_data/gen_full_rm_imp_mat.geno")

## env dat
trt_scaled <- read.csv("data/EnvDat/exp_trt_scaled_indiv_2026-08-13.csv")[,-1]
exp_scaled <- read.csv("data/EnvDat/exp_home_scaled_indiv_2026-08-13.csv")[,-1]
sea_scaled_mll <- read.csv("data/EnvDat/mll_seascape_scaled_indiv_2026-08-13.csv")[,-1]
sea_scaled_full <- read.csv("data/EnvDat/full_seascape_scaled_indiv_2026-08-13.csv")[,-1]

## seascape data
seascape_sites_trim <- read.csv("data/seascape_data/seascape_sites_forinterpolation.csv")[,-1]

full_sea_dat_curr <- read.csv("data/EnvDat/fullsea_curr_forint.csv")[,-1]
mll_sea_dat_curr <- read.csv("data/EnvDat/mllsea_curr_forint.csv")[,-1]
full_sea_dat_fut <- read.csv("data/EnvDat/fullsea_fut_forint.csv")[,-1]
mll_sea_dat_fut <- read.csv("data/EnvDat/mllsea_fut_forint.csv")[,-1]
###############
```


## Data Checks

## Data Checks

This data has largely been processed in other scripts, but I still need to double check everything before I can train models and make GO predictions. 

### Individual Data

```{r}
# inspect data
indiv_list
sea_pops
exp_pops

# indiv list
colnames(indiv_list) <- c("mll", "indiv")
indiv_list$pop <- str_split_fixed(indiv_list$indiv, "-", 2)[,1]

# exp inds
inds_exp1 <- inds_exp %>% mutate(bagnum = as.numeric(bagnum_text)) %>% dplyr::select(pop, trt, bagnum)

# standardize colnames for pop dfs
colnames(sea_pops) <- c("region", "site_full", "pop", "long", "lat")
colnames(exp_pops) <- c("site_full", "pop", "long", "lat")

# merge inds with seascape pop metadata
inds_sea <- merge(sea_pops, indiv_list, by = c("pop"))

# reorder this by mll number
inds_sea_ord <- inds_sea[order(inds_sea$mll),]

# mll inds
inds_sea_ord_mll <- inds_sea_ord[!(duplicated(inds_sea_ord$mll)),]
```


### Environmental Data

Environmental data has been pre-processed in the script "EnvDatProcess.Rmd". Here I just double check that everything looks right and is in the correct order to match with genomic data.

```{r}
# inspect data
summary(sea_scaled_mll)
summary(sea_scaled_full)
summary(trt_scaled)
summary(exp_scaled)

# check seascape environmental data in same order as individuals
sea_scaled_mll$pop == inds_sea_ord_mll$pop # needs reorder
sea_scaled_full$pop == inds_sea_ord$pop # needs reorder

sea_scaled_mll <- sea_scaled_mll %>% arrange(mll_no)
sea_scaled_full <- sea_scaled_full %>% arrange(mll_no)

# recheck
sea_scaled_mll$pop == inds_sea_ord_mll$pop # good!
sea_scaled_full$pop == inds_sea_ord$pop # good!

# pull just relevant columns
sea_scaled_mll_red <- sea_scaled_mll[,c("median_temp", "max_temp", "median_sal", "min_sal")]
rownames(sea_scaled_mll_red) <- sea_scaled_mll$indv_no
sea_scaled_full_red <- sea_scaled_full[,c("median_temp", "max_temp", "median_sal", "min_sal")]
rownames(sea_scaled_full_red) <- sea_scaled_full$order
trt_scaled_red <- trt_scaled[,c("median_temp", "max_temp", "median_sal", "min_sal")]
rownames(trt_scaled_red) <- trt_scaled$bagnum_text
exp_scaled_red <- exp_scaled[,c("median_temp", "max_temp", "median_sal", "min_sal")]
rownames(exp_scaled_red) <- exp_scaled$bagnum_text
```


### Genomic Data

Seascape genomic data comes from Steffi Ries and Marlene Jahnke at the University of Gothenburg, who had already processed the genomic dataset (check what part of bioinformatic pipeline had been done already). I further process this data to impute missing genomic data and remove non-variable genomic sites in the script "GenDatProcess.Rmd"

```{r}
# inspect data
geno_mll[1:10,1:10]
geno_full[1:10,1:10]

# set populations as rownames
rownames(geno_mll) <- inds_sea_ord_mll$pop
rownames(geno_full) <- inds_sea_ord$pop

# not using allele frequency based models

# check names
colnames(geno_mll) <- make.names(as.character(1:ncol(geno_mll)))
colnames(geno_full) <- make.names(as.character(1:ncol(geno_full)))
```


## Seascape Model Training

Now that I know my data is ready to use, I can train my models!

```{r}
# start by defining a maximum number of splits
maxLevel_mll <- log2(0.368*nrow(sea_scaled_mll)/2)
maxLevel_full <- log2(0.368*nrow(sea_scaled_full)/2)
 
# run gradient forest

# geno models
start_time <- Sys.time() # time start
gf_geno_mll <- gradientForest(cbind(sea_scaled_mll_red, geno_mll), predictor.vars = colnames(sea_scaled_mll_red), response.vars=colnames(geno_mll), ntree=500, maxLevel=maxLevel_mll, trace=T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 10 mins

start_time <- Sys.time() # time start
gf_geno_full <- gradientForest(cbind(sea_scaled_full_red, geno_full), predictor.vars = colnames(sea_scaled_full_red), response.vars=colnames(geno_full), ntree=500, maxLevel=maxLevel_full, trace=T, corr.threshold=0.50)
end_time <- Sys.time() # time end
(end_time - start_time) # about 10 mins

saveRDS(gf_geno_mll, paste0("results/GO_results/gf_geno_mll_indiv_",Sys.Date(),".RDS"))
saveRDS(gf_geno_full, paste0("results/GO_results/gf_geno_full_indiv_",Sys.Date(),".RDS"))

# read in models trained in trainGF.R / trainGF.sh
gf_geno_mll <- readRDS("results/GO_results/gf_geno_mll_indiv_2026-08-13.RDS")
gf_geno_full <- readRDS("results/GO_results/gf_geno_full_indiv_2026-08-13.RDS")

# plot bar graphs depicting importance of each spatial and climate variable
plot(gf_geno_mll)
plot(gf_geno_full)

sink("results/GO_results/gf_importance_ind.txt")
importance(gf_geno_mll)
importance(gf_geno_full)
sink()
```


## Offset Prediction

Now that we have trained our gradient forest models, we need to use these to predict performance in the mesocosm treatments.

```{r}
# predictions
#############
# predict to home site
pred_geno_home_mll <- predict(gf_geno_mll, exp_scaled_red)
pred_geno_home_full <- predict(gf_geno_full, exp_scaled_red)

# predict to treatments
pred_geno_trt_mll <- predict(gf_geno_mll, trt_scaled_red)
pred_geno_trt_full <- predict(gf_geno_full, trt_scaled_red)
#############

# calculate euclidean distance for offset
#########################################
# number preds
nb_pred <- nrow(trt_scaled_red)

# function
euclidean <- function(a, b) sqrt(sum((a - b)^2)) # Euclidean distance for genetic offset

# calculate offset
# geno mll
df_geno_mll <- data.frame() # df
for (i in 1:nb_pred){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_mll[i,], pred_geno_trt_mll[i,]))
  df_geno_mll <- rbind(df_geno_mll, ec_dist)
}
geno_offset_mll <- data.frame(pop = trt_scaled$pop, trt = trt_scaled$trt, bagnum = trt_scaled$bagnum_text, offset_gf_geno_mll = df_geno_mll, cc = "mll") # associate metadata with offset calculation
colnames(geno_offset_mll) <- c("pop", "trt", "bagnum", "offset", "cc")

# geno full
df_geno_full <- data.frame() # df
for (i in 1:nb_pred){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_full[i,], pred_geno_trt_full[i,]))
  df_geno_full <- rbind(df_geno_full, ec_dist)
}
geno_offset_full <- data.frame(pop = trt_scaled$pop, trt = trt_scaled$trt, bagnum = trt_scaled$bagnum_text, offset_gf_geno_full = df_geno_full, cc = "full") # associate metadata with offset calculation
colnames(geno_offset_full) <- c("pop", "trt", "bagnum", "offset", "cc")
#########################################

# put it all together
#####################
# complete offset df
gf_offsets <- rbind(geno_offset_mll, geno_offset_full)

# save output
write.csv(gf_offsets, paste0("results/GO_results/GF_offsets_exptrt_ind_long_",Sys.Date(),".csv"))
#####################
```

Also calculate GOs for seascape populations

```{r}
# seascape env data
full_sea_dat_curr_red <- full_sea_dat_curr %>% dplyr::select(median_temp, max_temp, median_sal, min_sal)
full_sea_dat_fut_red <- full_sea_dat_fut %>% dplyr::select(median_temp, max_temp, median_sal, min_sal)
mll_sea_dat_curr_red <- mll_sea_dat_curr %>% dplyr::select(median_temp, max_temp, median_sal, min_sal)
mll_sea_dat_fut_red <- mll_sea_dat_fut %>% dplyr::select(median_temp, max_temp, median_sal, min_sal)

# predictions
#############
# predict to home site
pred_geno_home_sea_mll <- predict(gf_geno_mll, sea_scaled_mll_red)
pred_geno_home_sea_full <- predict(gf_geno_full, sea_scaled_full_red)

# predict to current trt
pred_geno_curr_sea_mll <- predict(gf_geno_mll, mll_sea_dat_curr_red)
pred_geno_curr_sea_full <- predict(gf_geno_full, full_sea_dat_curr_red)

# predict to future trt
pred_geno_fut_sea_mll <- predict(gf_geno_mll, mll_sea_dat_fut_red)
pred_geno_fut_sea_full <- predict(gf_geno_full, full_sea_dat_fut_red)
#############

# calculate euclidean distance for offset
#########################################
# number preds
nb_pred_sea_mll <- nrow(mll_sea_dat_fut_red)
nb_pred_sea_full <- nrow(full_sea_dat_fut_red)

# calculate offset
# mll current
df_geno_mll_sea_curr <- data.frame() # df
for (i in 1:nb_pred_sea_mll){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_sea_mll[i,], pred_geno_curr_sea_mll[i,]))
  df_geno_mll_sea_curr <- rbind(df_geno_mll_sea_curr, ec_dist)
}
geno_offset_mll_sea_curr <- data.frame(pop = mll_sea_dat_curr$pop, offset_gf_geno_mll = df_geno_mll_sea_curr, cc = "mll") # associate metadata with offset calculation
colnames(geno_offset_mll_sea_curr) <- c("pop", "offset_curr", "cc")

# geno full
df_geno_full_sea_curr <- data.frame() # df
for (i in 1:nb_pred_sea_full){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_sea_full[i,], pred_geno_curr_sea_full[i,]))
  df_geno_full_sea_curr <- rbind(df_geno_full_sea_curr, ec_dist)
}
geno_offset_full_sea_curr <- data.frame(pop = full_sea_dat_curr$pop, offset_gf_geno_full = df_geno_full_sea_curr, cc = "full") # associate metadata with offset calculation
colnames(geno_offset_full_sea_curr) <- c("pop", "offset_curr", "cc")

# geno mll fut
df_geno_mll_sea_fut <- data.frame() # df
for (i in 1:nb_pred_sea_mll){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_sea_mll[i,], pred_geno_fut_sea_mll[i,]))
  df_geno_mll_sea_fut <- rbind(df_geno_mll_sea_fut, ec_dist)
}
geno_offset_mll_sea_fut <- data.frame(pop = mll_sea_dat_fut$pop, offset_gf_geno_mll = df_geno_mll_sea_fut, cc = "mll") # associate metadata with offset calculation
colnames(geno_offset_mll_sea_fut) <- c("pop", "offset_fut", "cc")

# geno full fut
df_geno_full_sea_fut <- data.frame() # df
for (i in 1:nb_pred_sea_full){ # calc offset
  ec_dist <- as.data.frame(euclidean(pred_geno_home_sea_full[i,], pred_geno_fut_sea_full[i,]))
  df_geno_full_sea_fut <- rbind(df_geno_full_sea_fut, ec_dist)
}
geno_offset_full_sea_fut <- data.frame(pop = full_sea_dat_fut$pop, offset_gf_geno_full = df_geno_full_sea_fut, cc = "full") # associate metadata with offset calculation
colnames(geno_offset_full_sea_fut) <- c("pop", "offset_fut", "cc")
#########################################

# put it all together
#####################
# combine & trim to relevant pops
offsets_mll_sea_go_gf <- merge(geno_offset_mll_sea_curr, geno_offset_mll_sea_fut, by = c("pop", "cc")) %>% filter(pop %in% seascape_sites_trim$pop) %>% as.data.frame()
offsets_full_sea_go_gf <- merge(geno_offset_full_sea_curr, geno_offset_full_sea_fut, by = c("pop", "cc")) %>% filter(pop %in% seascape_sites_trim$pop) %>% as.data.frame()

# summarize
offsets_mll_sea_go_gf_summary <- offsets_mll_sea_go_gf %>%
  group_by(pop) %>% 
  summarize(go_current_mll = mean(offset_curr), go_future_mll = mean(offset_fut))
offsets_full_sea_go_gf_summary <- offsets_full_sea_go_gf %>%
  group_by(pop) %>% 
  summarize(go_current_full = mean(offset_curr), go_future_full = mean(offset_fut))

# combine averaged info & add metadata
gf_offsets_sea <- merge(offsets_mll_sea_go_gf_summary, offsets_full_sea_go_gf_summary, by = "pop")
gf_offsets_sea_mg <- merge(seascape_sites_trim, gf_offsets_sea, by = "pop") %>% mutate(method = "gf_offset")

# save output
write.csv(gf_offsets_sea_mg, paste0("results/GO_results/gf_offset_seascape_go.csv"))
#####################
```

## Plotting

Now that I have my offsets, I want to plot them in a grid where populations are columns and treatments are rows 

```{r}
# prep for plotting
exp_pops$order <- 1:8
gf_offsets_plotting <- merge(exp_pops, gf_offsets, by = "pop")

# plot
# mll
ggplot(gf_offsets_plotting[gf_offsets_plotting$cc == "mll",], aes(x = fct_reorder(pop, order), y = trt, fill = (offset))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.1)) +
  theme_classic() +
  labs(title = "Genomic Offsets (GFgeno, MLL)", x = "Population", y = "Treatment", fill = "Offset")

# and the full offsets
ggplot(gf_offsets_plotting[gf_offsets_plotting$cc == "full",], aes(x = fct_reorder(pop, order), y = trt, fill = (offset))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,0.15)) +
  theme_classic() +
  labs(title = "Genomic Offsets (GFgeno, Full)", x = "Population", y = "Treatment", fill = "Offset")
```


