---
title: "eelgrass_LEA_offset"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: pdf_document
---

# Offset Modeling with LEA

This script is for offset calculation using LEA, specifically for the project "VR_Zostera_genomic_offset." It includes training of the models using seascape genomic data, as well as calculation of genomic offset for experimental populations in each mesocosm treatment. 

## Setup

```{r setup, include = F}
# markdown options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/BalticEelgrass/")

# libraries
library(LEA) # for offsets
library(BiocManager) # needed to download specific packages
library(readxl) # reading xlsx files
library(ggplot2) # for plotting
library(vegan) # for env dat processing
library(dplyr) # for data wrangling
library(stringr) # for data wrangling
library(s2) # mapping
library(rnaturalearth) # mapping
library(rnaturalearthdata) # mapping
library(maps) # mapping
library(ggspatial) # mapping
library(ggrepel) # mapping
library(terra) # interpolation
library(stars) # interpolation
library(gstat) # interpolation
library(forcats) # data vis

# read in data
## indiv dat
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
sea_pops <- read.csv("data/seascape_data/population_data.csv")[,-1]
exp_pops <- read.csv("data/experiment/eelgrass_exp_sites.csv")
inds_exp <- read.csv("data/EnvDat/indiv_trt_env_dat_2025-12-31.csv")[,-1]

## gen dat
lfmm_mll <- read.lfmm("data/seascape_data/gen_MLL_imp_mat.lfmm")
geno_mll <- read.geno("data/seascape_data/gen_MLL_imp_mat.geno")
lfmm_full <- read.lfmm("data/seascape_data/gen_full_rm_imp_mat.lfmm")
geno_full <- read.geno("data/seascape_data/gen_full_rm_imp_mat.geno")
lfmm_full_rm <- read.lfmm("data/seascape_data/eel_full_rm.lfmm") # this one has only snps found in both full and mll

## env dat
curr_full_env <- read.env("data/EnvDat/full_seascapeEnv_indlvl_only_2026-09-04.env")
curr_mll_env <- read.env("data/EnvDat/mll_seascapeEnv_indlvl_only_2026-09-04.env")
exp_env <- read.env("data/EnvDat/full_expEnv_indlvl_only_2026-09-04.env")
trt_env <- read.env("data/EnvDat/full_trtEnv_indlvl_only_2026-09-04.env")
trt_scaled <- read.csv("data/EnvDat/trt_scaled_ind_only_2026-09-04.csv")[,-1]
exp_scaled <- read.csv("data/EnvDat/exp_scaled_ind_only_2026-09-04.csv")[,-1]
sea_scaled_mll <- read.csv("data/EnvDat/mll_seascape_scaled_ind_only_2026-09-04.csv")[,-1]
sea_scaled_full <- read.csv("data/EnvDat/full_seascape_scaled_ind_only_2026-09-04.csv")[,-1]

## seascape data
# seascape_sites_trim <- read.csv("data/seascape_data/seascape_sites_forinterpolation.csv")[,-1]
# sea_scaled_full_curr <- read.env("data/EnvDat/fullsea_curr_forint.env")
# sea_scaled_mll_curr <- read.env("data/EnvDat/mllsea_curr_forint.env")
# sea_scaled_full_fut <- read.env("data/EnvDat/fullsea_fut_forint.env")
# sea_scaled_mll_fut <- read.env("data/EnvDat/mllsea_fut_forint.env")
# 
# full_sea_dat <- read.csv("data/EnvDat/fullsea_curr_forint.csv")[,c("coast", "cf", "pop", "mll_no", "indv_no", "site_full", "trt")]
# mll_sea_dat <- read.csv("data/EnvDat/mllsea_curr_forint.csv")[,c("coast", "cf", "pop", "mll_no", "indv_no", "site_full", "trt")]
```

## Data Prep

### Individual Data Prep

I will start by processing the individual data to work with both MLL and full datasets. 

```{r}
# inspect data
sea_pops
indiv_list
inds_exp
exp_pops

# rename sea pop columns
colnames(sea_pops) <- c("region","site_name","pop","long","lat")

# rename exp pop columns
colnames(exp_pops) <- c("site_name", "pop", "long", "lat")

# add metadata to inds
indiv_list <- as.data.frame(indiv_list)
colnames(indiv_list) <- c("mll", "ind")
indiv_list$pop <- str_split_fixed(indiv_list$ind, "-", 2)[,1]
inds <- merge(sea_pops, indiv_list, by = c("pop"))

# reorder by mll
inds_ord <- inds[order(inds$mll),]

# unique mll inds
inds_mll <- inds_ord[unique(inds_ord$mll),]

# experimental individuals
inds_exp1 <- inds_exp %>% dplyr::select(pop, trt, tank_ID, new_label, bagnum_text)
```

### Environmental Data Prep

Environmental Data has been pre-processed using the script "EnvDatProcess.Rmd". This script scales all environmental data together, then splits it into the relevant data sets for seascape, experimental pops, and treatments before outputting the correct formats for each GO method. Here, I just ensure that column names are consistent and that the order of the .env files matches with our other metadata

```{r}
# inspect data
# seascape
curr_full_env
curr_mll_env

# exp sites
exp_env
exp_scaled

# trt
trt_env
trt_scaled

# rownames for trt
rownames(trt_env) <- trt_scaled$trt

# colnames for env files
colnames(curr_full_env) <- colnames(curr_mll_env) <- colnames(exp_env) <- colnames(trt_env) <- c("median_temp_scaled", "max_temp_scaled", "median_sal_scaled", "min_sal_scaled")

# now check the order of data
exp_env == exp_scaled[,2:5] 
trt_env == trt_scaled[,5:8]
curr_mll_env == sea_scaled_mll[,6:9]
curr_full_env == sea_scaled_full[,6:9]

# add pop as rownames
rownames(exp_env) <- exp_scaled$pop
rownames(curr_mll_env) <- sea_scaled_mll$pop
rownames(curr_full_env) <- sea_scaled_full$pop
```

### Genomic Data Prep

Seascape genomic data comes from Steffi Ries and Marlene Jahnke at the University of Gothenburg, who had already processed the genomic dataset (check what part of bioinformatic pipeline had been done already). I further process this data to impute missing genomic data and remove non-variable genomic sites in the script "GenDatProcess.Rmd"

```{r}
# inspect
lfmm_mll[1:10,1:10]
lfmm_full[1:10,1:10]

# add pops as rownames
rownames(lfmm_mll) <- inds_mll$pop
rownames(lfmm_full) <- inds_ord$pop
```


## Seascape Model Training

Now that I have prepped and cleaned my data, I will use it to train LFMM models later used to predict genomic offset to some other (in this case experimental) climate. Importantly, I do not need to trim or impute my genomic data - this has been done in the script "GenDatProcess.Rmd". I also do not determine ancestry, Steffi has done this and recovered K = 12. 

```{r}
# fit model
mod_mll <- lfmm2(input = lfmm_mll,
                  env = curr_mll_env,
                  K = 12,
                  effect.sizes = T)
mod_full <- lfmm2(input = lfmm_full,
                      env = curr_full_env,
                      K = 12,
                      effect.sizes = T)
```

## Offset Prediction

Now I predict offset for all of my experimental populations to each treatment.

```{r}
# put together dataset for exp pop origin with trt (pred from and pred to)
exp_mat <- as.matrix(exp_env) 
trt_mat <- as.matrix(trt_env)

# genetic offset
gg_full_ind <- genetic.offset(input = lfmm_full,
                          env = curr_full_env,
                          new.env = exp_mat,
                          pred.env = trt_mat, 
                          K = 12)
gg_mll_ind <- genetic.offset(input = lfmm_mll,
                          env = curr_mll_env,
                          new.env = exp_mat,
                          pred.env = trt_mat, 
                          K = 12)

# check corr
cor(gg_mll_ind$offset, gg_full_ind$offset) # 0.9955437

# associate with metadata
inds_exp_trim <- inds_exp %>% dplyr::select(pop, trt, bagnum_text, tank_ID)
gap_df_full <- cbind(inds_exp_trim, gg_full_ind$offset) %>% mutate(cc = "full") %>% rename(offset = `gg_full_ind$offset`)
gap_df_mll <- cbind(inds_exp_trim, gg_mll_ind$offset) %>% mutate(cc = "mll") %>% rename(offset = `gg_mll_ind$offset`)
gap_df <- rbind(gap_df_full, gap_df_mll)
gap_df_wide <- cbind(inds_exp_trim, gg_mll_ind$offset, gg_full_ind$offset) %>% rename(gg_mll_ind = `gg_mll_ind$offset`, gg_full_ind = `gg_full_ind$offset`)
gap_df$method <- gap_df_wide$method <- "lea_offset"

# save offsets
write.csv(gap_df, paste0("results/GO_results/LEA_offsets_exptrt_ind_long_",Sys.Date(),".csv"))
write.csv(gap_df_wide, paste0("results/GO_results/LEA_offsets_exptrt_ind_wide_",Sys.Date(),".csv"))
```

I will also predict offsets for the seascape populations in order to interpolate across the study range. 

```{r}
# current
gg_full_sea_curr <- genetic.offset(input = lfmm_full,
                          env = curr_full_env,
                          pred.env = sea_scaled_full_curr, 
                          K = 12)
gg_mll_sea_curr <- genetic.offset(input = lfmm_mll,
                          env = curr_mll_env,
                          pred.env = sea_scaled_mll_curr, 
                          K = 12)

# future
gg_full_sea_fut <- genetic.offset(input = lfmm_full,
                          env = curr_full_env,
                          pred.env = sea_scaled_full_fut, 
                          K = 12)
gg_mll_sea_fut <- genetic.offset(input = lfmm_mll,
                          env = curr_mll_env,
                          pred.env = sea_scaled_mll_fut, 
                          K = 12)

# df
sea_full_go_df <- data.frame(pop = names(gg_full_sea_curr$offset), go_current = gg_full_sea_curr$offset, go_future = gg_full_sea_fut$offset, cc = "full")
sea_mll_go_df <- data.frame(pop = names(gg_mll_sea_curr$offset), go_current = gg_mll_sea_curr$offset, go_future = gg_mll_sea_fut$offset, cc = "mll")
sea_go_df <- rbind(sea_full_go_df, sea_mll_go_df)

# write this out
write.csv(sea_go_df, "results/GO_results/seascape_GO_all.csv")

# averages by population
sea_full_go_avg <- sea_full_go_df %>% 
  group_by(pop) %>% 
  summarize(go_current_full = mean(go_current), go_future_full = mean(go_future))
sea_mll_go_avg <- sea_mll_go_df %>% 
  group_by(pop) %>% 
  summarize(go_current_mll = mean(go_current), go_future_mll = mean(go_future))
sea_go_avg <- merge(sea_full_go_avg, sea_mll_go_avg, by = "pop")

# subset to relevant seascape populations, i.e. coastal swedish populations by merging with trimmed metadata
sea_go_trim <- merge(seascape_sites_trim, sea_go_avg, by = "pop") %>% arrange(order) %>% mutate(method = "lea_offset")

# write
write.csv(sea_go_trim, "results/GO_results/lea_offset_seascape_go.csv")
```

## Offset Plotting

Now that I have my offsets, I want to arrange them nicely by coast and plot. 

```{r}
# popsdf
exp_pops$order <- 1:8
gap_df_plotting <- merge(exp_pops, gap_df_wide, by = "pop")

# plot
ggplot(gap_df_plotting, aes(x = fct_reorder(pop, order), y = trt, fill = (gg_mll_ind))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,max(gap_df_plotting$gg_mll_ind))) +
  theme_classic() +
  labs(title = "Genomic Offsets (LEA, MLL)", x = "Population", y = "Treatment", fill = "Offset")

# and the full offsets
ggplot(gap_df_plotting, aes(x = fct_reorder(pop, order), y = trt, fill = (gg_full_ind))) +
  geom_tile(color = "white") +
  scale_fill_viridis_c(option = "magma", direction = -1, limits = c(0,max(gap_df_plotting$gg_full_ind))) +
  theme_classic() +
  labs(title = "Genomic Offsets (LEA, Full)", x = "Population", y = "Treatment", fill = "Offset")
```
