---
title: "GenDatProcess"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

# Genetic Data Processing

This script processes VCF files into the formats needed for downstream GO analysis, specifically for the project "VR_Zostera_genomic_offset." It includes data formatting + conversion to create .geno and .lfmm format files, as well as calculation of population level allele frequencies.  

```{r setup, include=FALSE}
# markdown options
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/BalticEelgrass/")

# libraries
library(tidyverse)
library(vcfR)
library(LEA)
library(readxl)

# read in data
# gen dat
eelgrass_vcf_MLL <- read.vcfR(
  "data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.vcf")
eelgrass_vcf_full <- read.vcfR(
  "data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.vcf")

# indiv/site dat
sites <- read.csv("data/seascape_data/sampling_sites_coordinates_Baltic_Sea.csv")
indiv_list <- read_excel("data/seascape_data/MLL_per_individual_list.xlsx")
gen_div <- read_excel("data/seascape_data/Gen_div_table.xlsx")
```

## Data Checks

```{r vcf processing}
# investigate vcf data - MLL
dim(eelgrass_vcf_MLL) # 3287 variants, 8 fix cols, 242 genotypes
fix_MLL <- as.data.frame(eelgrass_vcf_MLL@fix)
gt_MLL <- as.data.frame(eelgrass_vcf_MLL@gt)

# how many variants pass filtering?
table(fix_MLL$FILTER) # all pass

# quality summary
summary(as.numeric(fix_MLL$QUAL))

#investigate vcf data - FULL
dim(eelgrass_vcf_full) # 4149 variants, 8 fix cols, 696 genotypes
fix_full <- as.data.frame(eelgrass_vcf_full@fix)
gt_full <- as.data.frame(eelgrass_vcf_full@gt)

# how many variants pass filtering?
table(fix_full$FILTER) # all pass

# how many individuals in each population?
head(indiv_list)
pops_split <- stringr::str_split_fixed(indiv_list$`Indv #`, "-", 2) # split Indv # to have population in a seperate column
indiv_list_full <- cbind(indiv_list, pops_split) # add this info back to original file
colnames(indiv_list_full) <- c("MLL_no","Indv_no","Pop","Pop_no")
table(indiv_list_full$Pop) # between 4 and 20 individuals in each pop - most 17+ inds

# need indiv data for MLL dataset
# individuals are stored in column names of vcf - isolate these names
colnames(eelgrass_vcf_MLL@gt)
MLL_split <- stringr::str_split_fixed(colnames(eelgrass_vcf_MLL@gt), "-", 3)
MLL_inds <- MLL_split[-1,3]

# now get subsetted indiv dataframe from indiv list full
indiv_list_MLL <- indiv_list_full[indiv_list_full$Indv_no %in% MLL_inds,]
```

## File formatting

```{r lfmm & geno formats}
# convert vcf files to lfmm file format
eelgrass_lfmm_MLL <- vcf2lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.vcf")
eelgrass_lfmm_full <- vcf2lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.vcf")

# convert vcf files to lfmm file format
eelgrass_lfmm_MLL <- vcf2lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.vcf")
eelgrass_lfmm_full <- vcf2lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.vcf")

# read in geno and lfmm files
eelgrass_lfmm_MLL <- read.lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.lfmm")
eelgrass_geno_MLL <- read.geno("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_MLL.recode.geno")
eelgrass_lfmm_full <- read.lfmm("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.lfmm")
eelgrass_geno_full <- read.geno("data/seascape_data/zostera_230619_seascape_miss25_sorted_95_no_reps.recode.geno")
eelgrass_lfmm_full_rm <- read.lfmm("data/seascape_data/eel_full_rm.lfmm")

# check for missing data
sum((eelgrass_lfmm_MLL == 9)) # 16894/174269 missing data (~9.7% missing data)
sum((eelgrass_lfmm_full_rm == 9)) # 51254/615284 missing data (~8.3% missing data)

# impute missing data - this is a super naive imputation method, replace with more robust method using snmf later
gen_full_imp <- apply(eelgrass_lfmm_full_rm, 2, function(x) replace(x, x == 9, as.numeric(names(which.max(table(x))))))
sum((gen_full_imp == 9)) 

gen_MLL_imp <- apply(eelgrass_lfmm_MLL, 2, function(x) replace(x, x == 9, as.numeric(names(which.max(table(x))))))
sum((gen_MLL_imp == 9))

# remove non-variable genomic sites
gen_full_imp_df <- as.data.frame(gen_full_imp)
gen_full_imp_trim <- Filter(var, gen_full_imp_df)
gen_full_imp_mat <- as.matrix(gen_full_imp_trim)

gen_MLL_imp_df <- as.data.frame(gen_MLL_imp)
gen_MLL_imp_trim <- Filter(var, gen_MLL_imp_df)
gen_MLL_imp_mat <- as.matrix(gen_MLL_imp_trim)

# re-write lfmm files
write.lfmm(gen_full_imp_mat, "data/seascape_data/gen_full_rm_imp_mat.lfmm")
write.lfmm(gen_MLL_imp_mat, "data/seascape_data/gen_MLL_imp_mat.lfmm")

# also re-write geno files
write.geno(gen_full_imp_mat, "data/seascape_data/gen_full_rm_imp_mat.geno")
write.geno(gen_MLL_imp_mat, "data/seascape_data/gen_MLL_imp_mat.geno")

# read in lfmm files
gen_full <- read.lfmm("data/seascape_data/gen_full_rm_imp_mat.lfmm") # has 3264 SNPs after filtering
gen_MLL <- read.lfmm("data/seascape_data/gen_MLL_imp_mat.lfmm") # has 3036 SNPs after filtering
```


## Allele Frequency Calculation

```{r af prep}
# convert lfmm data to dataframe
eel_full_df <- as.data.frame(eelgrass_lfmm_full)
eel_MLL_df <- as.data.frame(eelgrass_lfmm_MLL)

# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
eel_full_df <- as.matrix(t(eel_full_df))
eel_MLL_df <- as.matrix(t(eel_MLL_df))

# replace 9/-9 with NA
eel_full_df[eel_full_df == 9] <- NA
eel_MLL_df[eel_MLL_df == 9] <- NA
# check it worked
sum((eel_full_df == 9)) # returns NA - good
sum((eel_MLL_df == 9)) # returns NA - good

# get vector of populations
pop_vector_full <- indiv_list_full$Pop
pop_vector_MLL <- indiv_list_MLL$Pop

# add pop data to genomic data
rownames(eel_full_df) <- pop_vector_full
rownames(eel_MLL_df) <- pop_vector_MLL

# convert lfmm data to dataframe
eel_full_df <- as.data.frame(eelgrass_lfmm_full)
eel_MLL_df <- as.data.frame(eelgrass_lfmm_MLL)

# rotate df so rows are now columns (we want individuals as columns, alleles as rows)
eel_full_df <- as.matrix(t(eel_full_df))
eel_MLL_df <- as.matrix(t(eel_MLL_df))

# replace 9/-9 with NA
eel_full_df[eel_full_df == 9] <- NA
eel_MLL_df[eel_MLL_df == 9] <- NA
# check it worked
sum((eel_full_df == 9)) # returns NA - good
sum((eel_MLL_df == 9)) # returns NA - good

# get vector of populations
pop_vector_full <- indiv_list_full$Pop
pop_vector_MLL <- indiv_list_MLL$Pop

# add pop data to genomic data
rownames(eel_full_df) <- pop_vector_full
rownames(eel_MLL_df) <- pop_vector_MLL
```


```{r genetic data positions}
# quick check of data
eelgrass_vcf_full
eelgrass_vcf_MLL

# use vcfR to extract positional information
full_fix <- as.data.frame(getFIX(eelgrass_vcf_full))
MLL_fix <- as.data.frame(getFIX(eelgrass_vcf_MLL))

# which SNPs are in both datasets?
dim(MLL_fix[MLL_fix$POS %in% full_fix$POS,]) # all MLL SNPs are in full

# filter out full SNPs that aren't in MLL dataset
full_fix_rm <- full_fix[full_fix$POS %in% MLL_fix$POS,]

# are these the same?
which(full_fix_rm != MLL_fix) # yes

# how will I exclude these SNPs from whole vcf file?
rm <- which(!(full_fix$POS %in% MLL_fix$POS)) # these are indices of the SNPs to remove

# reduce eel_full_df file to match eel_MLL_df file
eel_full_df_rm <- eel_full_df[-rm,]

# sanity check
dim(eel_MLL_df)
dim(eel_full_df_rm)
# looks good

# write the trimmed data back to lfmm format
write.lfmm(t(eel_full_df_rm), "seascape_data/eel_full_rm.lfmm")
```


```{r allele frequency calculation}
# function to calculate allele frequency @ single locus
calc_freq <- function(x){
  a <- sum(x, na.rm=TRUE)
  b <- (2*length(na.omit(x)))
  a/b
}

# function extending this to calculate allele frequencies across whole pop
calcfreq <- function(a, pop){
  tapply(a, pop, calc_freq)
}

# full dataset
str(eel_full_df_rm)
freqs_full <- apply(t(eel_full_df_rm), 2, calcfreq, pop_vector_full)
str(freqs_full)
colSums(freqs_full[,1:5])
View(freqs_full[,1:10])

# MLL dataset
freqs_MLL <- apply(t(eel_MLL_df), 2, calcfreq, pop_vector_MLL)
str(eel_MLL_df)
str(freqs_MLL)
colSums(freqs_MLL[,1:5])
View(freqs_MLL[,1:10])
```

