---
title: "eelgrass_fitness_offsets"
author: "Camille Rumberger"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = "~/Eelgrass/BalticEelgrass/data")
```

# Fitness Offsets

This script is intended to process raw fitness (survival and growth) data from the 2024 eelgrass mesocosm experiments in order to calculate absolute and relative fitness, as well as absolute and relative common-garden/current-future fitness offsets. 

A fitness offset is a change in fitness of an individual or population in a new environment compared to some reference. Fitness offsets can be delineated into “current-future” fitness offsets, which aim to forecast the change in fitness of a genotype in a future environment relative to its current environment, and “common garden” fitness offsets, which aim to forecast the difference between foreign genotypes and local genotypes in a common environment.

## Pre-processing

I need to first pre-process data to combine pop/trt/mort/growth data into a single workable data frame. 

```{r}
# read in data
setup <- read.csv("experiment/GOEEL-eelgrass-Sweden-exp - Setup.csv")
takedown <- read.csv("experiment/GO_eelgrass_mesocosm_takedown_data.csv")
head(setup)
head(takedown)

# remove practice rows (bagKey F054C074) & trim dataset
setup_rm <- setup[!setup$bagKey == "F054C074",]
setup_rm2 <- setup_rm[,c("setupKey","setupLabel","bagKey","setupTimestamp","nbr_leaves",
                         "sideShoot","nbr_sideShootLeaves","length_longest_leaf",
                         "leaf_width_hole","weight","IsBlackLesions")]

# get more meaningful labels in setup
dim(setup_rm2)
setup_rm2[,12:15] <- str_split_fixed(setup_rm$setupLabel, "_", 5)[,1:4]
colnames(setup_rm2) <- c("setupKey","setupLabel","bagKey","setupTimestamp","nbr_leaves_start","sideShoot","nbr_sideShootLeaves_start","length_longest_leaf_start","leaf_width_hole","weight_start","IsBlackLesions_start","bagnum","tank_ID","pop","genet")

# match setup and takedown
# can do this using "DNA_label" stuctured like HOR_gen-01
setup_rm2$DNA_label <- paste0(setup_rm2$pop,"_",setup_rm2$genet)

# check this 
head(setup_rm2)

# we also had to get rid of one individual because of planting error - ind 274
setup_rm3 <- setup_rm2[!setup_rm2$bagnum == 274,]

# now put together setup and takedown data
full_dat <- merge(setup_rm3, takedown, by = c("DNA_label", "tank_ID"))
dim(full_dat) # we've lost two individuals... one was the guy we had to remove

# which individual was not supposed to be lost?
setup_rm3[which(!(setup_rm3$DNA_label %in% takedown$DNA_label)),]
# VAT_gen-13 is missing from takedown data, confirmed by inspecting data sheet - we will have to do without this individual

takedown[which(!(takedown$DNA_label %in% setup_rm3$DNA_label)),]
# bagnum 274 is missing, which it should be since we removed it

# is there a different one that was duplicated in takedown? in setup?
which(duplicated(takedown$DNA_label)) # nope
which(duplicated(setup_rm3$DNA_label)) # nope

# back to our full dataset
full_dat # looks good, but still need treatment information
full_dat$trt <- paste0(full_dat$temp_full, "_", full_dat$sal_full)
write.csv(full_dat, "experiment/eelgrass_mesocosm_start_end_complete.csv")
```


## Absolute & Relative Fitness Data

I will start by processing raw fitness data to calculate survival and growth metrics. Survival is straightforward, but I will also calculate growth both by length (as the length of new leaves + growth in old leaves) and weight (as the difference in weight pre/post experiment). 

### Raw Survival Data - Absolute Fitness

To start, I will calculate the proportion of surviving individuals in each tank/treatment/population. This will be our "Absolute Fitness". For now, I won't bother calculating relative fitness for these data, as I will calculate relative fitness offsets later.

```{r}
# pops
pops <- levels(as.factor(full_dat$pop))

# trts
trts <- levels(as.factor(full_dat$trt))

# set up mortality matrix
# mortality by pop
# set up df
mort_pop <- matrix(data = NA, nrow = length(pops), ncol = 2)

# calc
for (i in 1:length(pops)) {
  # split pops
  split_mort = full_dat[full_dat$pop == pops[i],]
  mort_pop[i,1] = pops[i]
  # number that survived
  mort_pop[i,2] = length(split_mort[split_mort$isAlive == "TRUE",]$isAlive)/length(split_mort$isAlive)
}

# clean up
mort_pop <- as.data.frame(mort_pop)
colnames(mort_pop) <- c("pop","surv")
mort_pop$surv <- as.numeric(mort_pop$surv)

# mortality by treatment
mort_trt <- matrix(data = NA, nrow = length(trts), ncol = 2)
for (i in 1:length(trts)) {
  # split pops
  split_mort = full_dat[full_dat$trt == trts[i],]
  mort_trt[i,1] = trts[i]
  # number that survived
  mort_trt[i,2] = length(split_mort[split_mort$isAlive == "TRUE",]$isAlive)/length(split_mort$isAlive)
}
mort_trt <- as.data.frame(mort_trt)
colnames(mort_trt) <- c("trt","surv")
mort_trt$surv <- as.numeric(mort_trt$surv)

# mortality by population & treatment
# add trt_pop column
full_dat$trt_pop <- paste0(full_dat$trt,"-",full_dat$pop)
trt_pop <- levels(as.factor(full_dat$trt_pop))

# calc
mort_trt_pop <- matrix(data = NA, nrow = length(trt_pop), ncol = 2)
for (i in 1:length(trt_pop)) {
  # split pops
  split_mort = full_dat[full_dat$trt_pop == trt_pop[i],]
  mort_trt_pop[i,1] = trt_pop[i]
  # number that survived
  mort_trt_pop[i,2] = length(split_mort[split_mort$isAlive == "TRUE",]$isAlive)/length(split_mort$isAlive)
}
mort_trt_pop <- as.data.frame(mort_trt_pop)
colnames(mort_trt_pop) <- c("trt_pop","surv")
mort_trt_pop$surv <- as.numeric(mort_trt_pop$surv)

# add in trt and pop as separate columns
mort_trt_pop[,3:4] <- str_split_fixed(mort_trt_pop$trt_pop, "-", 2)
colnames(mort_trt_pop) <- c("trt_pop","surv","trt","pop")

# population order
pop_order <- data.frame(pop = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO"), 
                        order = c(1:8))

mort_trt_pop_mg <- merge(mort_trt_pop, pop_order, by = "pop")

# by coast?
mort_trt_pop_mg$coast <- ifelse(mort_trt_pop_mg$order >= 5, "East", "West")

# more informative trt names
trt <- levels(as.factor(mort_trt_pop_mg$trt))
names <- c("Current W Coast", "Current E Coast", "Future W Coast", "Future E Coast")
trt_nm <- data.frame(trt = trt, trt_name = names)
mort_trt_pop_mg <- merge(mort_trt_pop_mg, trt_nm, by = "trt")

# calc by coast
mortality_coasttrt_surv <- mort_trt_pop_mg %>% group_by(coast,trt) %>%
  summarise(mean_surv = mean(as.numeric(surv)), 
            sd_surv = sd(as.numeric(surv)),
            se_surv = std.error(as.numeric(surv)),
            .groups = 'drop')
mortality_coasttrt_surv <- merge(mortality_coasttrt_surv, trt_nm, by = "trt")
mortality_coasttrt_surv$CurrFut <- str_split_fixed(as.character(mortality_coasttrt_surv$trt_name)," ",2)[,1]

mort_pop
mort_trt
mort_trt_pop_mg
mortality_coasttrt_surv

write.csv(mort_trt_pop_mg, "experiment/survival_dat_trt_pop.csv")
```

### Raw Length Data - Absolute Fitness

Here I will calculate two growth metrics: length grown and weight grown for each tank/treatment/population. This will be the "Absolute Fitness". Again, I won't bother calculating relative fitness for these data, as I will calculate relative fitness offsets later.

Length grown will be determined by summing together the growth on old leaves with the length of any new leaves or shoots.

Weight grown will be determined by subtracting the weight at the beginning of the experiment from the weight at the end of the experiment. 

It's important to note that there are a LOT of NAs in this dataset, but they aren't errors or data that we missed - rather these reflect biological differences in the plants. 

```{r}
# again using the full_dat data frame
full_dat
summary(full_dat)

# we only want to work with plants that are still alive
full_dat_alive <- full_dat[full_dat$isAlive == T,]

# I'm also going to remove flowering plants from this analysis
full_dat_alive_nofl <- full_dat_alive[is.na(full_dat_alive$Length_flower1),]

# let's start by getting length growth for each plant
colnames(full_dat_alive_nofl) # check which columns are relevant for length growth
full_dat_alive_nofl$length_growth <- rowSums(full_dat_alive_nofl[,c("Growth_leaf1","Growth_leaf2","Growth_leaf3","Growth_leaf4","Growth_leaf5","Growth_leaf6","Growth_leaf7","Length_new_leaf1","Length_new_leaf2","Length_new_leaf3","Length_new_leaf4","Length_new_leaf5","Length_new_leaf6","Length_new_shoot1_leaf1","Length_new_shoot1_leaf2","Length_new_shoot1_leaf3","Length_new_shoot1_leaf4","Length_new_shoot1_leaf5","Length_new_shoot1_leaf6","Length_new_shoot2_leaf1","Length_new_shoot2_leaf2","Length_new_shoot2_leaf3","Length_new_shoot2_leaf4","Length_new_shoot2_leaf5")], na.rm = T)

# and weights
full_dat_alive_nofl$weight_growth <- full_dat_alive_nofl[,c("total_weight")] - full_dat_alive_nofl[,c("weight_start")]
# note that weight change can be negative if more biomass was lost than gained

# length growth vs weight change
ggplot(full_dat_alive_nofl, aes(x = length_growth, y = weight_growth)) +
  geom_point() +
  geom_smooth(method = "lm") + 
  theme_classic()
cor(full_dat_alive_nofl$length_growth, full_dat_alive_nofl$total_weight) # pearson corr = 0.6192

# now we can calculate averages across populations and treatments
growth_pop <- full_dat_alive_nofl %>% group_by(pop) %>%
  summarise(mean_len_growth = mean(as.numeric(length_growth)), 
            sd_len_growth = sd(as.numeric(length_growth)),
            se_len_growth = std.error(as.numeric(length_growth)),
            mean_weight = mean(as.numeric(weight_growth)), 
            sd_weight = sd(as.numeric(weight_growth)),
            se_weight = std.error(as.numeric(weight_growth)),
            .groups = 'drop')
growth_trt <- full_dat_alive_nofl %>% group_by(trt) %>%
  summarise(mean_len_growth = mean(as.numeric(length_growth)), 
            sd_len_growth = sd(as.numeric(length_growth)),
            se_len_growth = std.error(as.numeric(length_growth)),
            mean_weight = mean(as.numeric(weight_growth)), 
            sd_weight = sd(as.numeric(weight_growth)),
            se_weight = std.error(as.numeric(weight_growth)),
            .groups = 'drop')
growth_pop_trt <- full_dat_alive_nofl %>% group_by(pop,trt) %>%
  summarise(mean_len_growth = mean(as.numeric(length_growth)), 
            sd_len_growth = sd(as.numeric(length_growth)),
            se_len_growth = std.error(as.numeric(length_growth)),
            mean_weight = mean(as.numeric(weight_growth)), 
            sd_weight = sd(as.numeric(weight_growth)),
            se_weight = std.error(as.numeric(weight_growth)),
            .groups = 'drop')

# plot of length
len_plot_eel <- ggplot(full_dat_alive_nofl, aes(x = trt, y = length_growth, fill = pop)) +
  geom_boxplot(position=position_dodge(0.8),
               width = 0.6) +
  #  geom_dotplot(binaxis='y', stackdir='center', 
  #             position=position_dodge(0.8),
  #             dotsize = 0.5) +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  ggtitle("Length Growth vs Treatment") +
  theme_classic()
len_plot_eel

# plot of weight
wght_plot_eel <- ggplot(full_dat_alive_nofl, aes(x = trt, y = weight_growth, fill = pop)) +
  geom_boxplot(position = position_dodge(0.8, preserve = "single"),
               width = 0.6) +
#  geom_dotplot(binaxis='y', stackdir='center', 
#               position= position_dodge(0.8), # need to change smth here
#               dotsize = 0.6) +
  scale_fill_manual(name = "Population",
                    values = c("#155d27","#25a244","#4ad66d","#b7efc5", 
                               "#caf0f8","#90e0ef","#0096c7","#023e8a"),
                    labels = c("VIK","VAT","HOG","BAR","KUR","KAL","HOR","BJO")) +
  ggtitle("Weight Change vs Treatment") +
  theme_classic()
wght_plot_eel
```


## Fitness Offsets

Now that I have calculated absolute (and not yet relative) fitness using both survival and growth metrics, I can use this to calculate common garden and current-future fitness offsets. 

Note that for these fitness offsets, I have more populations than I have treatments - that is, populations are assigned a "home" or "local" treatment that is broadly representative of their true home/local conditions, but not an exact match. In order to appropriately account for this detail, I need to scale calculated fitness offsets by environmental distance. Thus, my first step here is actually to calculate environmental distance between experimental treatments and experimental populations' environment-of-origin.

```{r}
# put together env of origin and treatment data - COME BACK HERE
env_trt_site <- cbind()

# calculate environmental distance
library(vegan)
env_dist <- round(vegdist(env_trt_site[,3:8], method = "euclidian",
                          upper = FALSE, diag = TRUE), 4)
env_dist_mal <- round(vegdist(env_trt_site[,3:8], method = "mahalanobis",
                              upper = FALSE, diag = TRUE), 4)
env_dist_df <- as.data.frame(as.matrix(env_dist))
env_dist_mal_df <- as.data.frame(as.matrix(env_dist_mal))
colnames(env_dist_df) <- rownames(env_dist_df) <- colnames(env_dist_mal_df) <- rownames(env_dist_mal_df) <- env_trt_site$site_name
env_dist_df
env_dist_mal_df
```

### Common Garden Fitness Offset

Common garden fitness offset is the fitness difference between foreign genotypes and local genotypes in a common environment. I will calculate both absolute and relative common garden fitness offset here. Importantly, I will treat E. coast pops as local to the current E. coast treatment and W. coast pops as local to the current W. coast treatment - no populations will be "local" to either of the future treatments. Additionally, because no population is from the exact conditions represented in the mesocosm treatment, I will scale fitness offsets for each population by the environmental distance between its environment-of-origin and the actual treatment environment.  

```{r}
# fake data
fake_fit <- data.frame(population = rep(c("A","B","C","D"),2), trt = c(rep("trt-1", 4), rep("trt-2", 4)), fitness = c(1, 0.8, 0.3, 0.4, 0.2, 0.5, 0.9, 0.8), envOrg = c("lcl","lcl","frn","frn","frn","frn","lcl","lcl"), envdist = c(0.1, 1.3, 8.8, 9.1, 10, 9.4, 1, 0))
fake_fit

# # start by separating out "local" and "foreign"
# fit_frn <- data.frame(pop_trt = paste0(fake_fit[fake_fit$envOrg == "frn",]$population, "_", fake_fit[fake_fit$envOrg == "frn",]$trt), fit = fake_fit[fake_fit$envOrg == "frn",]$fitness)
# fit_lcl <- data.frame(pop_trt = paste0(fake_fit[fake_fit$envOrg == "lcl",]$population, "_", fake_fit[fake_fit$envOrg == "lcl",]$trt), fit = fake_fit[fake_fit$envOrg == "lcl",]$fitness)

# I need to scale these data now by environmental distance before calculating fitness offset


# from Katie's paper
# This is a more general function for real data
# At least a 3x3 reciprocal transplant is needed
ExploreCorrsFlex <- function(W, main){
  # `main` is the name or title of the situation modeled
  # W is a nxn reciprocal transplant matrix with genotypes in columns and common gardents in rows
  
  print("Absolute fitness")
  print(W)
  par(mar=c(4,0,0,1))

  LocalAdaptation_SA <- mean(diag(W)) - mean(W[lower.tri(W) | upper.tri(W)])
   print(c("Delta-SA is", LocalAdaptation_SA))
   
  colors = grey(seq(0.1,0.9, length=nrow(W)))
   
#  pdf(paste0("figs/", main, "_LApatterns.pdf"), height=4, width=5)
  # Make reciprocal transplant plot
    par(mar=c(4,4,1,1))
    barplot(t(W), ylim=c(0,max(W)*1.5), 
        lwd=4, bty="n", beside=TRUE,
         xlab="Site", ylab="Absolute fitness", col=colors)
    legend("topright",legend=c(colnames(W)), fill=colors)
  #dev.off()

  #Home-away fitness offset
    offset_HA <- t(t(W)-diag(W))
  #Home-away relative fitness
    w_HA <- t(t(W)/diag(W))
  #Home-away relative fitness offset
    offset_HA_relative <- t(t(offset_HA)/diag(W))
  
  #Local-foreign fitness offset
    offset_LF <- W - diag(W)
  #Local-foreign relative fitness
    w_LF <- W/diag(W)
  #Local-foreign relative fitness offset
    offset_LF_relative <- offset_LF/diag(W)
  # 
  df <- data.frame(as.numeric(W), 
                   as.numeric(offset_HA), 
                  # as.numeric(w_HA), 
                   as.numeric(offset_HA_relative),
                   as.numeric(offset_LF), 
                  # as.numeric(w_LF),
                   as.numeric(offset_LF_relative))
  
#  pdf(paste0("figs/", main, "_RelFitpatterns.pdf"), height=4, width=5)
    # Make reciprocal transplant plot
    par(mar=c(4,4,1,1))
    barplot(t(w_LF), ylim=c(min(c(w_LF,0)),max(w_LF)*1.5), 
        lwd=4, bty="n", beside=TRUE,
         xlab="Site", ylab="Local-foreign relative fitness",
        col=colors)
    legend("topright",legend=c(colnames(W)), fill=colors)
    
    barplot(w_HA, ylim=c(min(c(w_HA,0)),max(w_HA)*1.5), 
        lwd=4, bty="n", beside=TRUE,
         xlab="Site", ylab="Home-away relative fitness", col=colors)
    legend("topright",legend=c(rownames(W)), fill=colors)
    
 # dev.off()
  
  
  
  colnames(df)<-c("Absolute fitness",
                  "Home-away fitness offset",
                #  "Home-away relative fitness",
                  "Home-away relative fitness offset",
                  "Local-foreign fitness offset",
               #   "Local-foreign relative fitness",
                  "Local-foreign relative fitness offset")
  
  # Plot a heatmap of the correlations among the matrices
  col <- colorRampPalette(c("#BB4444", "#EE9988",
                            "white",  "#77AADD", "#4477AA"))
 # pdf(paste0("figs/", main, "_corr.pdf"), height=8 , width=8)
  par(mar=c(0,0,2,2))
  corrplot(round(cor(df),2), method="color", col=col(100),
           type="upper", 
           addCoef.col="black", tl.col="black", 
           tl.cex=1, main=main, tl.srt=45, mar=c(0,0,2,2))
 # dev.off()
  return(list(W=W, offset_HA=offset_HA,
              offset_HA_relative=offset_HA_relative, 
              w_HA = w_HA,
              offset_LF=offset_LF,
              offset_LF_relative=offset_LF_relative, 
              w_LF = w_LF,
              cor_mat = cor(df),
              Delta_SA = LocalAdaptation_SA))

} #end function

### I THINK TO USE THE ABOVE FUNCTION, I NEED TO COERCE MY FITNESS DATA TO FIT WITHIN THE RECIPROCAL GARDEN FRAMEWORK - I ALSO NEED TO SCALE MY FITNESS DATA BY ENVIRONMENTAL DISTANCE BEFORE USING THIS FUNCTION

# # need to start by making a function to calculate the fitness offset
# commGardFit <- function(fit_df) {
#   fit_frn = fit_df[fit_df$envOrg == "frn",]$fitness
#   fit_lcl = fit_df[fit_df$envOrg == "lcl",]$fitness
#   fit_matrix = matrix(nrow = (length(fit_frn) + length(fit_lcl)), ncol = 2)
#   for (i in (length(fit_frn) + length(fit_lcl))) { # change fit frn and fit lcl to be rows in some dataframe that labels things as local or foreign
#     env_dist = fit_df 
#     fit = fit_lcl[i,2] - sum(fit_frn[i,2])
#     fit_scale = fit * (1 + env_dist) # 1 + dist so that we don't run into problems when env dist is 0
#     fit_matrix[i,1] = 
#     fit_matrix[i,2] = fit_scale
#   }
#   return(fit_scale)
# }
# 
# commGardFit(fake_fit)
```

TO DISCUSS WITH KATIE - HOW CAN I CALCULATE FITNESS OFFSETS WHEN NO INDIVIDUALS ARE LOCAL TO FUTURE TREATMENTS??? DOES IT COME BACK TO SCALING BY ENV DIST AGAIN?

### Current-Future Fitness Offset

Current-future fitness offset is the fitness difference between a genotype in some future environment compared to its current environment. I will calculate both absolute and relative current-future fitness offset here. Importantly, I will treat E. coast pops as being home in the current E. coast treatment and W. coast pops as being home in the current W. coast treatment - no populations will be "home" in either of the future treatments. Additionally, because no population is from the exact conditions represented in the mesocosm treatment, I will scale fitness offsets for each population by the environmental distance between its environment-of-origin and the actual treatment environment.  

```{r}

```


